### 2023-05-23 00_dataPrepReport.Rmd
prepared report for Orr

---
title: "Data preparation report"
author: "Kaija Gahm"
date: "2023-05-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE, message = FALSE, cache = FALSE)
```

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
```

### Step 1: Download raw data from Movebank

```{r download-movebank-data}
# base::load("movebankCredentials/pw.Rda")
# MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
# rm(pw)
minDate <- "2020-09-01 00:00"
maxDate <- "2022-12-15 11:59"
#dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
#write_feather(dat, "data/dat.feather")
dat <- read_feather("data/dat.feather")
```

Downloaded data from Movebank from `r minDate` to `r maxDate`. The
initial data has **`r nrow(dat)`** rows and `r ncol(dat)` columns. There
are **`r length(unique(dat$trackId))`** unique individual ID's in the
data at this point.

### Step 2: Fix vulture IDs

```{r fix-trackId-Nili_id}
## fix trackId. Marta told me that E03w was sometimes coded as E03, so I fixed that right up front. 
dat2 <- dat %>%
  mutate(trackId = as.character(trackId),
         trackId = case_when(trackId == "E03" ~ "E03w",
                             TRUE ~ trackId))
# number of unique individuals
length(unique(dat2$trackId)) # 126--makes sense that we lost one because one individual had a typo.

# Read in the who's who table and prepare to join the Nili_ids
ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

# Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
ww %>% filter(Movebank_id %in% dat2$trackId) %>% 
  dplyr::select(Nili_id, Movebank_id) %>% 
  distinct() %>% group_by(Nili_id) %>% 
  summarize(n = length(unique(Movebank_id))) %>% 
  arrange(desc(n))
ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili_id's

all(dat2$trackId %in% ww$Movebank_id) # true

dat2 <- left_join(dat2, ww, by = c("trackId" = "Movebank_id"))
dim(dat2) # good, same number of rows.
length(unique(dat2$Nili_id))
```

Next, we join in the Who's Who table in order to fix the Movebank
trackId's and associate them with individual vultures (as defined by the
"Nili_id" column). After joining in the Nili_ids, we now have a data
frame with **`r nrow(dat2)`** rows and `r ncol(dat2)` columns. There are
now **`r length(unique(dat2$Nili_id))`** unique individual vultures
identified--the change from **`r length(unique(dat$trackId))`** is
because there were a couple cases of different trackId's being assigned
to the same Nili_id, such as "Y26" and "Y26b". No individuals are
actually lost here.

### Step 3: Remove invalid periods (cages etc.)

Next, we remove periods of time when specific vultures were in cages or
in the hospital, or when tags were not attached to the bird, etc. This
is based on the "periods_to_remove" sheet in the who's who file.

```{r remove-periods}
## annotate the data with periods to remove
toRemove <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")

toRemove <- toRemove %>%
  dplyr::select(Nili_id,
                remove_start,
                remove_end,
                reason) %>%
  mutate(across(c(remove_start, remove_end), .fns = function(x){
    lubridate::ymd(x)
  })) %>%
  dplyr::filter(!is.na(remove_end))

toRemove_long <- toRemove %>%
  group_by(Nili_id, reason) %>%
  # sequence of daily dates for each corresponding start, end elements
  dplyr::mutate(dateOnly = map2(remove_start, remove_end, seq, by = "1 day")) %>%
  # unnest the list column
  unnest(cols = c(dateOnly)) %>%
  # remove any duplicate rows
  distinct() %>%
  dplyr::select(-c(remove_start, remove_end)) %>%
  rename("status" = reason)

# Join to the original data
datAnnot <- dat2 %>%
  left_join(toRemove_long, by = c("Nili_id", "dateOnly")) %>%
  mutate(status = replace_na(status, "valid"))
nrow(datAnnot) == nrow(dat2) #T: good, same number of rows.

# Actually REMOVE the periods to remove...
datAnnot <- datAnnot %>%
  filter(status == "valid")
```

The resulting data has dropped from **`r nrow(dat2)`** to
**`r nrow(datAnnot)`** rows, a
**`r round((nrow(dat2)-nrow(datAnnot))*100/nrow(dat2), 3)`% decrease**.
It has `r ncol(datAnnot)` columns and
**`r length(unique(datAnnot$Nili_id))`** unique individuals.

### Step 4: Attach birth year and sex info from who's who

```{r join-age-sex}
# Attach age and sex information
as <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()

datAnnot2 <- datAnnot %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot)
length(unique(datAnnot2$Nili_id)) # 125
```

No change in the number of rows here: `r nrow(datAnnot2)` rows and
`r ncol(datAnnot2)` columns.

```{r}
datAnnot <- datAnnot2 # a hack so the rest of the code will work
```

### Step 5: Clean the data using the cleanData function.

This function includes several data cleaning steps: - Remove as invalid
data points that have zeroes in `ground_speed`, `barometric_height`, and
`external_temperature`. - Remove points that took to long to get a GPS
fix (\>89) - Remove as invalid points with heading data \<0 or \>360
degrees - Remove points with fewer than 3 satellites (because 3 are
needed for triangulation) - Remove points that "jump" unrealistically
across the map, as calculated by distance/time differences between
consecutive points - Set to NA any altitude values that jump
unrealistically between consecutive points - Remove points that "moved"
farther than expected during the night - Remove **individuals** where
less than 1/3 of their duration tracked fell inside the rough
Israel-region geographic mak - For any individuals remaining, remove
**points** that fell outside of the geographic mask (e.g. to remove
long-range forays).

```{r clean-data}
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
mask <- sf::st_read("data/CutOffRegion.kml")
# datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 0.33, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
# save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
load("data/datAnnotCleaned.Rda")
dim(datAnnotCleaned) # about 27 percent of data outside the israel region (roughly)
length(unique(datAnnotCleaned$Nili_id))
```

After cleaning, the new data has **`r nrow(datAnnotCleaned)`** rows (a
**`r round((nrow(datAnnot)-nrow(datAnnotCleaned))*100/nrow(datAnnot), 3)`%
decrease**) and `r ncol(datAnnotCleaned)` columns.
**`r length(unique(datAnnot$Nili_id))-length(unique(datAnnotCleaned$Nili_id))`**
individuals were removed, for a remaining total of
**`r length(unique(datAnnotCleaned$Nili_id))`** individuals.

### Step 6: Fix times and split into seasons

First, I changed the time zone to Israel time instead of UTC, since this
will make a difference to where the date line falls. Then I
re-calculated the `dateOnly` column based on the new date/times.

Next, I split the data into fall/breeding/summer seasons based on
Marta's cutoff dates: September 15th, December 15th, and May 15th. I
fixed the vultures' ages to adjust for the breeding season and then
classified them as adults or sub-adults/juveniles. Finally, I split the
data into a list, with one item per season.

```{r fix-dates}
# Fix time zone so dates make sense ---------------------------------------
## Overwrite the dateOnly column from the new times
datAnnotCleaned <- datAnnotCleaned %>%
  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
         dateOnly = lubridate::date(timestampIsrael),
         month = lubridate::month(dateOnly),
         year = lubridate::year(dateOnly))
```

```{r split-seasons}
# Split into Marta's 3 seasons --------------------------------------------
s1 <- datAnnotCleaned %>%
  mutate(start_breeding = ifelse(month %in% c(1:5),
                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
         start_summer = ifelse(month == 12,
                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
         start_breeding = as.Date(start_breeding),
         start_summer = as.Date(start_summer),
         season = case_when(
           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
  filter(dateOnly != "2022-12-15") # remove December 15 2022, because it turns out that's a non-inclusive boundary (oops)

s2 <- s1 %>%
  mutate(age = year - birth_year,
         sex = case_when(sex == "f/m" ~ "u",
                         TRUE ~ sex)) %>%
  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
  group_by(Nili_id, year, month) %>%
  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
  mutate(age_group = case_when(age == 0 ~ "juv",
                               age >= 1 & age <= 4 ~ "sub",
                               age >= 5 ~ "adult",
                               TRUE ~ NA)) %>%
  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
  # explicitly set the factor levels in temporal order
  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall")),
         season = factor(season, levels = c("breeding", "summer", "fall")))
table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read

# remove columns that were used in the season calculation
s3 <- s2 %>%
  dplyr::select(-c(month, start_breeding, start_summer, start_fall))

# Separate the seasons -----------------------------
seasons <- s3 %>%
  group_by(seasonUnique) %>%
  group_split(.keep = T)
# extract the season names in case we need a separate vector at some point.
seasonNames <- map_chr(seasons, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
```

```{r save-soc-data}
# Save copies for social analysis
# The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis. 
seasons_forSoc <- seasons
#save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
```

There are `r length(seasonNames)` seasons: `r seasonNames` (we will end
up removing Summer 2020 because the Ornitela tags started partway
through that season.) `seasons` is therefore a list of length
**`r length(seasons)`**, with the following numbers of rows in each
season: `r map_dbl(seasons, nrow)`. This data is saved as
`seasons_forSoc.Rda` so it can be used for the social network analysis
without any further changes made.

### Step 7: Get roosts

Now I use Marta's roost code, implemented in vultureUtils as
`get_roosts_df`, to get roost locations for each individual on each
night.

```{r get-roosts}
# Get roosts for each season ----------------------------------------------
# forRoosts <- map(seasons, ~.x %>% sf::st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE))
# roosts_seasons <- purrr::map(forRoosts, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
# roosts_seasons <- roosts_seasons %>%
#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat")))
# save(roosts_seasons, file = "data/roosts_seasons.Rda")
load("data/roosts_seasons.Rda")
```

### Step 8: Remove nighttime points

Just for the data I'm going to use for movement analyses, I want to
exclude any points that occur at night. This will also help deal with
the problem that there are some individuals that seem to have way too
many points per day

```{r remove-night}
# Remove nighttime points -------------------------------------------------
before <- map_dbl(seasons, nrow)
seasons <- map(seasons, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889,
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

  .x <- .x %>%
    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
    dplyr::left_join(times, by = "dateOnly") %>%
    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
    dplyr::select(-c(sunrise, sunset)) %>%
    dplyr::filter(daylight == "day")
})
after <- map_dbl(seasons, nrow)
change <- after-before
changeDF <- data.frame(seasonName = seasonNames, nBefore = before, nAfter = after) %>%
  mutate(propRowsRemoved = round((before-after)/before, 2))
```

```{r include = T, echo = F}
print(changeDF)
```

### Step 9: Include only "southern individuals" in the movement calculations

For each individual in each season, I calculated the centroid of its
coordinates. Then I looked at some histograms of the centroids in each
season and decided that 3550000 latitude (in ITM coordinates) would be a
cutoff point that would neatly divide individuals in the southern
population from the northern population (there was a large gap in the
histogram consistently around this value).

I filtered each season down to remove individuals classified as
"northern" for that season, leaving only those in the Negev population.
Keep in mind:

1)  A given individual could be "northern" in one season but "southern"
    in another season, theoretically. It would be removed only in the
    seasons when it falls in the north.

2)  Any removed individuals can still contribute to the social network
    that I will calculate--that is, they can contribute to the
    degree/strength/etc. of southern birds. I'm only removing them for
    the movement calculations, so they won't be focal individuals in the
    model.

```{r remove-northern}
# Restrict to southern individuals ----------------------------------------
# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (in ITM)
## Transform to SF object, so we can get centroids
seasonsSF <- map(seasons, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

## Get centroids, so we can see who's "southern" for that season.
centroids <- map(seasonsSF, ~.x %>%
                   group_by(Nili_id) %>%
                   summarize(geometry = st_union(geometry)) %>%
                   st_centroid()) # XXX what is the actual definition of the centroid?

## Examine a histogram of centroid latitudes 
# walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.

## Get southern individuals for each season, so we can filter the data
southernIndivs <- map(centroids, ~.x %>%
                        filter(st_coordinates(.)[,2] < 3550000) %>%
                        pull(Nili_id))

## Remove individuals not in the south
before <- seasons
seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
after <- seasons
beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
```

After doing this filtering, the following numbers of rows and
individuals remained in each season:

```{r include = T, echo = F}
print(df)
```

### Step 10: Remove vulture\*days based on number of points per day

```{r view-ppd}
# Include only individuals with enough points per day ------------------------
# How many points per day are we looking at?
ppd <- map_dfr(seasons, ~.x %>%
             st_drop_geometry() %>%
             group_by(seasonUnique, season, Nili_id, dateOnly) %>%
             summarize(n = n()))

## Theoretically, should only have 72 points per day--12 hours, 6 points per hour. That should vary a bit. I don't really understand why so many individuals/days have more than 100 points per day, but let's proceed for now: going to downsample that to 10 minue intervals anyway, so it won't matter.
```

We had initially decided to remove vulture\*days with \< 30 points per
day. Here's what that would look like:

```{r include = T, echo = F}
# 
thresh <- 30
ppd %>%
  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
  ggplot(aes(x = n))+
  geom_histogram(aes(fill = Status))+
  facet_wrap(~seasonUnique, scales ="free")+
  geom_vline(aes(xintercept = thresh), col = "red")+
  scale_fill_manual(values = c("darkgray", "red"))+
  theme_classic()+
  ggtitle(paste("n removed =", thresh, sep = " "))+
  ylab("# vulture*days")+
  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
```

Restricting to days with 30 points per day ends up removing a huge
number of days, especially in the breeding season (winter) because of
lack of sunlight etc. After talking with Marta and Noa, I decided to
remove just 10 points per day. That would look like this:

```{r include = T, echo = F}
# 
thresh <- 10
ppd %>%
  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
  ggplot(aes(x = n))+
  geom_histogram(aes(fill = Status))+
  facet_wrap(~seasonUnique, scales ="free")+
  geom_vline(aes(xintercept = thresh), col = "red")+
  scale_fill_manual(values = c("darkgray", "red"))+
  theme_classic()+
  ggtitle(paste("n removed =", thresh, sep = " "))+
  ylab("# vulture*days")+
  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
```

```{r}
# thresh <- 10
# ppd <- ppd %>%
#   ungroup() %>%
#   mutate(removed = ifelse(n < thresh, T, F))
# 
# # what proportion of vulture*days are removed overall?
# (propRemoved <- ppd %>%
#     group_by(seasonUnique, season) %>%
#     summarize(nTot = n(),
#               nRemoved = sum(removed),
#               propRemoved = nRemoved/nTot))
# propRemoved %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
#   geom_point(size = 10)+
#   theme_classic()+
#   scale_color_manual(values = c("blue", "red", "darkorange"))
# 
# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
# 
# # what about by individual over each season?
# ppd %>%
#   group_by(seasonUnique, season, Nili_id) %>%
#   summarize(propRemoved = sum(removed)/n()) %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved))+
#   geom_boxplot(aes(group = seasonUnique, fill = season))+
#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
#   theme_classic()+ # tells the same story
#   ylab("Proportion of days removed")+
#   xlab("")

## Remove days on which individuals don't have at least 10 points.
before <- seasons
seasons <- map(seasons, ~.x %>%
                 group_by(Nili_id, dateOnly) %>%
                 filter(n() >= 10) %>% # hard-coding the 10 in here so it doesn't accidentally get changed.
                 ungroup())
after <- seasons

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))
```

Here's how many points get removed when we do this filtering:

```{r include = T, echo = F}
df
```

### Step 11: Remove individuals based on number of days tracked

This filtering comes after the removal of some vulture\*days for having
too few points. So, considering the remaining days, we only want to
retain individuals that are tracked for at least 1/3 of a given season
(in terms of number of days).

Here are the number of days in each season. As you can see, breeding
season is the longest, followed by summer, then fall. (Ignore Summer
2020--it will not be included in the analysis.)

```{r include = TRUE, echo = F}
seasonNames
durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))
durs
```

```{r remove-daysTracked}
# Include only individuals with enough days tracked -----------------------
## Must be tracked for at least 1/3 of the number of days in the season.

# walk2(seasons, durs, ~.x %>%
#   st_drop_geometry() %>%
#   group_by(Nili_id) %>%
#   summarize(nDaysTracked = length(unique(dateOnly)),
#             propDaysTracked = nDaysTracked/.y) %>%
#   pull(propDaysTracked) %>%
#   hist())

indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
                       st_drop_geometry() %>%
                       group_by(Nili_id) %>%
                       summarize(nDaysTracked = length(unique(dateOnly)),
                                 propDaysTracked = nDaysTracked/.y) %>%
                       filter(propDaysTracked >= 0.33) %>% # XXX ask Orr if this should be percentage vs. number of days, given the different season lengths. (email)
                       pull(Nili_id))

## remove those individuals not tracked for enough days
before <- seasons
after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))

```

Here's what this does to the row counts:

```{r include = T, echo = F}
df
```

### Step 12: Calculate flight altitudes

Using the `elevatr` package, I get ground elevation values for each gps
point in each season, at zoom level 10 (which should come out to around
150 meter cell resolution in the resulting raster, according to )

```{r get-elevation}
# Get elevation rasters ---------------------------------------------------
elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x %>% st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE, crs = "WGS84"), z = 10))
# 
save(elevs_z10_seasons, file = "data/elevs_z10_seasons.Rda")
load("data/elevs_z10_seasons.Rda")

groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y %>% sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = FALSE))) # have to convert the crs because the raster from elevatr is in WGS84.

## use elevation rasters to calculate height above ground level
seasons <- map2(.x = seasons, .y = groundElev_z10, 
                ~.x %>% mutate(height_above_ground = .y,
                               height_above_ground = case_when(height_above_ground <0 ~ 0,
                                                               TRUE ~ height_above_ground)))

# Export the data
save(seasons, file = "data/seasons.Rda")
```

As expected, adding the elevation information doesn't change the number
of rows in the data.

### Step 13: Downsample the data

Because occasionally, the vultures' tags were set to read more
frequently than every 10 minutes, I want to downsample the data to a
10-minute interval.

```{r downsample}
# Downsample the data and save the downsampled data
subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
  sub <- df %>%
    arrange(idCol, timestampCol) %>%
    group_by(.data[[idCol]]) %>%
    mutate(tc = cut(.data[[timestampCol]], 
                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
    group_by(.data[[idCol]], 
             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(-c("d", "tc"))
  return(sub)
}

# create and save downsampled datasets for later use
seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
seasons_20min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 20), .progress = T)
seasons_30min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 30), .progress = T)
seasons_60min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 60), .progress = T)
seasons_120min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 120), .progress = T)

map_dbl(seasons, nrow)
map_dbl(seasons_10min, nrow)

# save(seasons_10min, file = "data/seasons_10min.Rda")
# save(seasons_20min, file = "data/seasons_20min.Rda")
# save(seasons_30min, file = "data/seasons_30min.Rda")
# save(seasons_60min, file = "data/seasons_60min.Rda")
# save(seasons_120min, file = "data/seasons_120min.Rda")

# no need to calculate roosts for the downsampled data because in theory there's only one point per night.

df <- data.frame(season = seasonNames, 
                 nBefore = map_dbl(seasons, nrow),
                 nAfter = map_dbl(seasons_10min, nrow)) %>%
  mutate(propChange = round((nBefore-nAfter)/nBefore, 2))
```

Here are the row counts after downsampling:

```{r include = TRUE, echo = F}
df
```

And here are the final counts of number of vultures:

```{r final-counts}
df <- seasons_10min %>% purrr::list_rbind() %>% 
  group_by(seasonUnique) %>% 
  summarize(vultures = length(unique(Nili_id)))
```

```{r include = TRUE, echo = F}
df
```

### 2023-05-23 00_dataPrep.R

@@ -13,20 +13,21 @@ library(readxl)
 base::load("movebankCredentials/pw.Rda")
 MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
 rm(pw)
-
-# dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = "2020-09-01 00:00", dateTimeEndUTC = "2022-11-30 11:59") # XXX needs to go all the way to December 15th based on Marta's new seasons
-# write_feather(dat, "data/dat.feather")
+minDate <- "2020-09-01 00:00"
+maxDate <- "2022-12-15 11:59"
+#dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+#write_feather(dat, "data/dat.feather")
 dat <- read_feather("data/dat.feather")
 # number of unique individuals
 length(unique(dat$trackId)) # 127
 
 ## fix trackId
-dat <- dat %>%
+dat2 <- dat %>%
   mutate(trackId = as.character(trackId),
          trackId = case_when(trackId == "E03" ~ "E03w",
                              TRUE ~ trackId))
 # number of unique individuals
-length(unique(dat$trackId)) # 126--makes sense that we lost one because one individual had a typo.
+length(unique(dat2$trackId)) # 126--makes sense that we lost one because one individual had a typo.
 
   # Add the Nili_id
 ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
@@ -34,12 +35,12 @@ ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tag
   distinct()
 
 # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
-ww %>% filter(Movebank_id %in% dat$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n))
+ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n))
[…]

### 2023-05-26 00_dataPrepReport.Rmd
Added code to remove in-cage dates

@@
 
-The resulting data has dropped from **`r nrow(dat2)`** to
+```{r remove-cages}
+# Remove cages using Marta's data
+# Identify and remove capture dates using Marta's code --------------------
+## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function).
+Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset.
+This has been validated.
+Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages.
+But we're just dealing with southern individuals, so that's okay.
+#roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+#save(roosts, file = "data/roosts.Rda")
+load("data/roosts.Rda")
+
+# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
+start.day <- 01
+start.month <- 08
+end.day <-  30
+end.month <- 11
+distance <- 500 # distance, in meters, to calculate from the cage
+
+# Load the information about the capture sites
+sites <- read.csv("data/capture_sites.csv")
+
+# Subset the roost dataset with the start and end dates for the capture period
+sub.roosts <- roosts %>%
+  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
+                              format="%d-%m-%Y"),
+         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
+                            format="%d-%m-%Y")) %>%
+  filter(date >= start_date & date <= end_date)
+
+unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.
+
+# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
+crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat), 
+               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates
+
+DistanceMat <- matrix(ncol = nrow(sites), nrow = nrow(crds)) 
+colnames(DistanceMat) <- unique(sites$name)
+
+for(i in 1:nrow(crds)){ # for each roost point...
+  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
+  print(i)
+}
+
+ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
+ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site
+
+sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist) 
+sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")
+
+sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage
+
+sub.captured.dates <- subset(sub.captured, 
+                             select = c("Nili_id",
+                                        "date", 
+                                        "ClosestCaptureSite",
+                                        "ClosestCaptureDist", 
+                                        "Captured"))
+
+## This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it.
+So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data
+
+sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
+sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")
+
+AllCarmelDates <- read.csv("data/all_captures_carmel_2010-2021.csv")
+AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")
+
+AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
+AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
+AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))
+
+AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)
+
+sub.captured.carmel <- sub.captured.carmel %>%
+  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
+         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
+  filter(Captured == "yes") %>%
+  dplyr::select(-c(known_capture))
+
+names(sub.captured.no.carmel)
+names(sub.captured.carmel)
+
+sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)
+
+# We also need to exclude the day after the bird was captured
+sub.captured.dates.1 <- sub.captured.dates
+sub.captured.dates.1$date <- sub.captured.dates.1$date+1
+
+sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
+sub.captured.dates <- sub.captured.dates %>% 
+  dplyr::distinct(Nili_id, date, .keep_all = T)
+
+# It all looks ok, so we can subset the dataset to exclude the capture dates
+datAnnot_noCaptures <- datAnnot %>%
+  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
+nrow(datAnnot) == nrow(datAnnot_noCaptures) # should be TRUE. NOW we can filter.
+datAnnot_noCaptures <- datAnnot_noCaptures %>%
+  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured
+
+datAnnot_noCaptures <- datAnnot_noCaptures %>%
+  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
+
+# How many rows and individuals did we remove?
+before <- nrow(datAnnot)
+after <- nrow(datAnnot_noCaptures)
+(propChange <- (before-after)/before) # approx. 1% of data removed.
+length(unique(datAnnot$Nili_id))
+length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
+```
+
+The resulting data has dropped from **`r nrow(datAnnot_noCaptures)`** to
 **`r nrow(datAnnot)`** rows, a
-**`r round((nrow(dat2)-nrow(datAnnot))*100/nrow(dat2), 3)`% decrease**.
+**`r round((nrow(datAnnot_noCaptures)-nrow(datAnnot))*100/nrow(datAnnot_noCaptures), 3)`% decrease**.
 It has `r ncol(datAnnot)` columns and
@@
 
-datAnnot2 <- datAnnot %>%
+datAnnot2 <- datAnnot_noCaptures %>%
   dplyr::select(-c("sex")) %>%

### 2023-05-26 00_dataPrep.R

@@ -82,12 +82,116 @@ length(unique(datAnnot$Nili_id))
 nrow(dat2)
 nrow(datAnnot)
 
+# Identify and remove capture dates using Marta's code --------------------
+## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function). Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages. But we're just dealing with southern individuals, so that's okay.
+#roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+#save(roosts, file = "data/roosts.Rda")
+load("data/roosts.Rda")
+
+# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
+start.day <- 01
+start.month <- 08
+end.day <-  30
+end.month <- 11
+distance <- 500 # distance, in meters, to calculate from the cage
+
+# Load the information about the capture sites
+sites <- read.csv("data/capture_sites.csv")
+
+# Subset the roost dataset with the start and end dates for the capture period
+sub.roosts <- roosts %>%
+  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
+                              format="%d-%m-%Y"),
+         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
+                            format="%d-%m-%Y")) %>%
+  filter(date >= start_date & date <= end_date)
+
+unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.
+
+# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
+crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat), 
+               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates
+
[…]

### 2023-06-09 00_dataPrepReport.Rmd
changed data cleaning threshold to 30 days instead of 1/3 of time

@@
 But we're just dealing with southern individuals, so that's okay.
-#roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-#save(roosts, file = "data/roosts.Rda")
+roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+save(roosts, file = "data/roosts.Rda")
 load("data/roosts.Rda")
@@
   DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
-  print(i)
+  #print(i)
 }
@@
 
-The resulting data has dropped from **`r nrow(datAnnot_noCaptures)`** to
-**`r nrow(datAnnot)`** rows, a
-**`r round((nrow(datAnnot_noCaptures)-nrow(datAnnot))*100/nrow(datAnnot_noCaptures), 3)`% decrease**.
-It has `r ncol(datAnnot)` columns and
-**`r length(unique(datAnnot$Nili_id))`** unique individuals.
+The resulting data has dropped from **`r nrow(dat2)`** to
+**`r nrow(datAnnot_noCaptures)`** rows, a
+**`r round((nrow(dat2)-nrow(datAnnot_noCaptures))*100/nrow(dat2), 3)`% decrease**.
+It has `r ncol(dat2)` columns and
+**`r length(unique(datAnnot_noCaptures$Nili_id))`** unique individuals.
 
@@
   left_join(as, by = "Nili_id")
-nrow(datAnnot2) == nrow(datAnnot)
+nrow(datAnnot2) == nrow(datAnnot_noCaptures)
 length(unique(datAnnot2$Nili_id)) # 125
@@
 
-```{r clean-data}
+```{r clean-data, echo = T, include = T}
 # Clean the data
@@
 mask <- sf::st_read("data/CutOffRegion.kml")
-# datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 0.33, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
-# save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
+datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
+save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
 load("data/datAnnotCleaned.Rda")
-dim(datAnnotCleaned) # about 27 percent of data outside the israel region (roughly)
-length(unique(datAnnotCleaned$Nili_id))
 ```
@@
 # roosts_seasons <- roosts_seasons %>%
-#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat")))
+#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
 # save(roosts_seasons, file = "data/roosts_seasons.Rda")

### 2023-06-09 00_dataPrep.R

@@ -202,8 +202,8 @@ datAnnot <- datAnnot2 # a hack so the rest of the code will work
 # Clean the data
 ## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
 mask <- sf::st_read("data/CutOffRegion.kml")
-#datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 0.33, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T)
-#save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
+datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 0.33, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T)
+save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
 load("data/datAnnotCleaned.Rda")
 dim(datAnnotCleaned) # about 27 percent of data outside the israel region (roughly)
 length(unique(datAnnotCleaned$Nili_id)) # 120 individuals left
@@ -271,7 +271,7 @@ save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
 # Get roosts for each season ----------------------------------------------
 roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
 roosts_seasons <- roosts_seasons %>%
-  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat")))
+  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
 save(roosts_seasons, file = "data/roosts_seasons.Rda")
 
 # Remove nighttime points -------------------------------------------------

### 2023-06-09 00_dataPrepReport.Rmd
clarified UTM zone

@@
 coordinates. Then I looked at some histograms of the centroids in each
-season and decided that 3550000 latitude (in ITM coordinates) would be a
+season and decided that 3550000 latitude (in UTM zone 36N coordinates--https://epsg.io/32636) would be a
 cutoff point that would neatly divide individuals in the southern
@@
 # Restrict to southern individuals ----------------------------------------
-# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (in ITM)
+# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636)
 ## Transform to SF object, so we can get centroids

### 2023-06-09 00_dataPrep.R

@@ -202,7 +202,7 @@ datAnnot <- datAnnot2 # a hack so the rest of the code will work
 # Clean the data
 ## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
 mask <- sf::st_read("data/CutOffRegion.kml")
-datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 0.33, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T)
+datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall. If we want 30 days per season, will have to do that differently (e.g. split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
 save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
 load("data/datAnnotCleaned.Rda")
 dim(datAnnotCleaned) # about 27 percent of data outside the israel region (roughly)
@@ -272,7 +272,7 @@ save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
 roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
 roosts_seasons <- roosts_seasons %>%
   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
-save(roosts_seasons, file = "data/roosts_seasons.Rda")
+save(roosts_seasons, file = "data/roosts_seasons.Rda") # XXXXXX start here
 
 # Remove nighttime points -------------------------------------------------
 before <- map_dbl(seasons, nrow)
@@ -296,7 +296,7 @@ changeDF <- data.frame(seasonName = seasonNames, nBefore = before, nAfter = afte
 changeDF
 
 # Restrict to southern individuals ----------------------------------------
-# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (in ITM)
+# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636)
 ## Transform to SF object, so we can get centroids
 seasonsSF <- map(seasons, ~.x %>%
[…]

### 2023-06-09 00_dataPrepReport.Rmd
fixed battery/ppd filter, and removed summer 2020

@@
 
-## Remove days on which individuals don't have at least 10 points.
+## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
 before <- seasons
@@
                  group_by(Nili_id, dateOnly) %>%
-                 filter(n() >= 10) %>% # hard-coding the 10 in here so it doesn't accidentally get changed.
+                 filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>%
                  ungroup())
@@
                                  propDaysTracked = nDaysTracked/.y) %>%
-                       filter(propDaysTracked >= 0.33) %>% # XXX ask Orr if this should be percentage vs. number of days, given the different season lengths. (email)
+                       filter(nDaysTracked >= 30) %>%
                        pull(Nili_id))
@@
 df
+
+## Now I notice that season 1, Summer 2020, doesn't have enough individuals. Let's remove it--we're going to have to anyway.
+seasons <- seasons[-1]
 ```

### 2023-06-09 00_dataPrep.R

@@ -272,7 +272,7 @@ save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
 roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
 roosts_seasons <- roosts_seasons %>%
   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
-save(roosts_seasons, file = "data/roosts_seasons.Rda") # XXXXXX start here
+save(roosts_seasons, file = "data/roosts_seasons.Rda")
 
 # Remove nighttime points -------------------------------------------------
 before <- map_dbl(seasons, nrow)
@@ -331,13 +331,84 @@ df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
 df
 
 # Include only individuals with enough points per day ------------------------
-# How many points per day are we looking at?
-ppd <- map_dfr(seasons, ~.x %>%
-             st_drop_geometry() %>%
-             group_by(seasonUnique, season, Nili_id, dateOnly) %>%
-             summarize(n = n()))
+# Investigate battery percentage vs points per day ------------------------
+# Is there some way we can relate battery charge percentage to number of points per day?
+test <- seasons[[5]]
+names(test)
 
-## Theoretically, should only have 72 points per day--12 hours, 6 points per hour. That should vary a bit. I don't really understand why so many individuals/days have more than 100 points per day, but let's proceed for now: going to downsample that to 10 minue intervals anyway, so it won't matter.
+ppd <- map_dfr(seasons, ~.x %>%
+                 st_drop_geometry() %>%
+                 group_by(seasonUnique, season, Nili_id, dateOnly) %>%
+                 summarize(n = n(),
+                           minBatt = min(battery_charge_percent, na.rm = T),
+                           meanBatt = mean(battery_charge_percent, na.rm = T),
+                           medBatt = mean(battery_charge_percent, na.rm = T)))
+
+excl_batt <- 50
+excl_pts <- 10
+## min battery charge
+ppd %>%
+  filter(n < 100) %>% # only interested in the lower numbers
+  mutate(remove = case_when(minBatt <= excl_batt & n <= excl_pts ~ T,
+                            TRUE ~ F)) %>%
+  ggplot(aes(x = minBatt, y = n))+
[…]

### 2023-06-15 02_calcSocial.R
save social network data and load back in

@@ -14,9 +14,17 @@ load("data/roosts_seasons.Rda")
 seasonNames <- map_chr(seasons_forSoc, ~as.character(.x$seasonUnique[1]))
 
 # Social Networks ---------------------------------------------------------
-flightSeasons <- map(seasons_forSoc, ~vultureUtils::getFlightEdges(.x, roostPolygons = roostPolygons, distThreshold = 1000, idCol = "Nili_id", return = "sri"))
-feedingSeasons <- map(seasons_forSoc, ~vultureUtils::getFeedingEdges(.x, roostPolygons = roostPolygons, distThreshold = 50, return = "sri")) # XXX start here
-roostSeasons <- map(roosts_seasons, ~vultureUtils::getRoostEdges(.x, mode = "polygon", roostPolygons = roostPolygons, return = "sri", latCol = "location_lat", longCol = "location_long", idCol = "Nili_id", dateCol = "roost_date"))
+# flightSeasons <- map(seasons_forSoc, ~vultureUtils::getFlightEdges(.x, roostPolygons = roostPolygons, distThreshold = 1000, idCol = "Nili_id", return = "sri"))
+# feedingSeasons <- map(seasons_forSoc, ~vultureUtils::getFeedingEdges(.x, roostPolygons = roostPolygons, distThreshold = 50, return = "sri")) # XXX start here
+# roostSeasons <- map(roosts_seasons, ~vultureUtils::getRoostEdges(.x, mode = "polygon", roostPolygons = roostPolygons, return = "sri", latCol = "location_lat", longCol = "location_long", idCol = "Nili_id", dateCol = "roost_date"))
+# 
+# save(flightSeasons, file = "data/flightSeasons.Rda")
+# save(feedingSeasons, file = "data/feedingSeasons.Rda")
+# save(roostSeasons, file = "data/roostSeasons.Rda")
+
+load("data/flightSeasons.Rda")
+load("data/feedingSeasons.Rda")
+load("data/roostSeasons.Rda")
 
 flightSeasons_g <- map(flightSeasons, ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
 feedingSeasons_g <- map(feedingSeasons,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
@@ -58,8 +66,8 @@ networkMetrics <- map2_dfr(flightSeasons_g, seasonNames, ~{
 
 # Add info about number of indivs and relative measures
[…]

### 2023-06-15 03_PCA.R
random lines at the end?

@@ -145,3 +145,6 @@ linked %>%
   theme_classic()+
   facet_wrap(~season) # makes sense! As the DDT/DMD ratio increases, so does tortuosity. As expected. Good thing we fixed the tortuosity.
 
+
+
+

### 2023-06-15 05_visualize.R
More viz, and pc maps

@@ -2,15 +2,20 @@ library(tidyverse)
 library(jtools)
 library(sjPlot)
 library(ggplot2)
+library(ggeffects)
 
 load("data/mods.Rda")
 load("data/effects.Rda")
 load("data/contrib.Rda")
 load("data/linked.Rda")
+load("data/seasons_10min.Rda")
 
 # Set season colors
 seasonColors <- c("#2FF8CA", "#CA2FF8", "#F8CA2F")
-situationColors <- c("dodgerblue2", "olivedrab3", "gold")
+situationColors <- c("dodgerblue2", "olivedrab3", "gold") # flight, roosting, feeding
+flightColor <- "dodgerblue2"
+roostingColor <- "olivedrab3"
+feedingColor <- "gold"
 
 # Effect size plots
 # Here are some plots made with jtools, following the vignette here: https://cran.r-project.org/web/packages/jtools/vignettes/summ.html
@@ -18,21 +23,21 @@ situationColors <- c("dodgerblue2", "olivedrab3", "gold")
 # Degree
 deg.jt <- plot_summs(mods[["fd"]], mods[["ed"]], mods[["rd"]], inner_ci_level = .9,
            model.names = c("Flight", "Feeding", "Roosting"), coefs = c("PC1", "PC2"))+
-  scale_color_manual(values = situationColors)+
+  scale_color_manual(values = c("Flight" = flightColor, "Feeding" = feedingColor, "Roosting" = roostingColor))+
   ggtitle("Degree (normalized)")
 deg.jt
 
 # Strength
 str.jt <- plot_summs(mods[["fs"]], mods[["es"]], mods[["rs"]], inner_ci_level = .9,
            model.names = c("Flight", "Feeding", "Roosting"), coefs = c("PC1", "PC2"))+
-  scale_color_manual(values = situationColors)+
+  scale_color_manual(values = c("Flight" = flightColor, "Feeding" = feedingColor, "Roosting" = roostingColor))+
   ggtitle("Strength (normalized)")
 str.jt
 
 # Evenness
 eve.jt <- plot_summs(mods[["fe"]], mods[["ee"]], mods[["re"]], inner_ci_level = .9,
            model.names = c("Flight", "Feeding", "Roosting"), coefs = c("PC1", "PC2"))+
-  scale_color_manual(values = situationColors)+
[…]

### 2023-06-15 04_analysis.R
Model selection status before 3/14 VulTour meeting

@@ -18,6 +18,9 @@ library(sjPlot)
 library(broom.mixed)
 library(jtools) # similar to sjplot, for forest plots
 
+load("data/linked.Rda")
+load("data/mnPPD.Rda")
+
 # Set ggplot theme to classic
 theme_set(theme_classic())
 
@@ -25,12 +28,39 @@ theme_set(theme_classic())
 seasonColors <- c("#2FF8CA", "#CA2FF8", "#F8CA2F")
 situationColors <- c("dodgerblue2", "olivedrab3", "gold")
 
+# Check that the response variable distributions look normal--nothing too weird here.
+linked %>%
+  mutate(season = factor(season, levels = c("breeding", "summer", "fall"))) %>%
+  ggplot(aes(x = evenness))+
+  geom_density(aes(col = season), linewidth = 1.5)+
+  facet_wrap(~type, scales = "free_y")+
+  scale_color_manual(name = "Season", values = seasonColors)+
+  xlab("Evenness")+
+  ylab("")+
+  theme(text = element_text(size = 16))
+
+linked %>%
+  mutate(season = factor(season, levels = c("breeding", "summer", "fall"))) %>%
+  ggplot(aes(x = degreeRelative))+
+  geom_density(aes(col = season), linewidth = 1.5)+
+  facet_wrap(~type, scales = "free_y")+
+  scale_color_manual(name = "Season", values = seasonColors)+
+  xlab("Degree (normalized)")+
+  ylab("")+
+  theme(text = element_text(size = 16))
+
+linked %>%
+  mutate(season = factor(season, levels = c("breeding", "summer", "fall"))) %>%
+  ggplot(aes(x = strengthRelative))+
+  geom_density(aes(col = season), linewidth = 1.5)+
+  facet_wrap(~type, scales = "free_y")+
+  scale_color_manual(name = "Season", values = seasonColors)+
+  xlab("Strength (normalized)")+
+  ylab("")+
+  theme(text = element_text(size = 16))
+
 # Here's an interesting reference article for some modeling stuff: https://biol607.github.io/lab/12_gzlm.html
 
-# arbitrary changes to test for slowness
-
-load("data/linked.Rda")
-load("data/mnPPD.Rda")
 mnPPD <- mnPPD %>%
   mutate(year = factor(as.numeric(str_extract(seasonUnique, "[0-9]+"))),
[…]

### 2023-06-15 00_dataPrepReport.Rmd
Re-downloaded data including the 2023 breeding season

@@
 minDate <- "2020-09-01 00:00"
-maxDate <- "2022-12-15 11:59"
-#dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-#write_feather(dat, "data/dat.feather")
+maxDate <- "2023-05-15 11:59"
+dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+write_feather(dat, "data/dat.feather")
 dat <- read_feather("data/dat.feather")

### 2023-06-15 00_dataPrep.R

@@ -14,9 +14,9 @@ base::load("movebankCredentials/pw.Rda")
 MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
 rm(pw)
 minDate <- "2020-09-01 00:00"
-maxDate <- "2022-12-15 11:59"
-#dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-#write_feather(dat, "data/dat.feather")
+maxDate <- "2023-05-15 11:59"
+dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+write_feather(dat, "data/dat.feather")
 dat <- read_feather("data/dat.feather")
 # number of unique individuals
 length(unique(dat$trackId)) # 127

### 2023-08-17 00_dataPrepReport.Rmd
progress on the data prep report as rmarkdown doc

--- 2023-06-09 00_dataPrepReport.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = FALSE, message = FALSE, cache = FALSE)
+knitr::opts_chunk$set(include = FALSE, echo = FALSE, message = FALSE, cache = FALSE)
 ```
+
+This is a condensed and simplified data preparation report. It excludes some exploratory plotting and analyses. There are records of those in 00_dataPrep.R.
 
@@
 library(readxl)
+library(elevatr)
 ```
 
-### Step 1: Download raw data from Movebank
-
-```{r download-movebank-data}
+## 1. Download data from Movebank
+```{r}
+# ## Download data from movebank
 # base::load("movebankCredentials/pw.Rda")
@@
 # rm(pw)
-minDate <- "2020-09-01 00:00"
-maxDate <- "2022-12-15 11:59"
-#dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-#write_feather(dat, "data/dat.feather")
+# 
+# # Hatzofe INPA data
+# # inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
+# # inpadf <- methods::as(inpa, "data.frame")
+# # inpadf_touse <- inpadf %>%
+# #   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
+# #          year = as.numeric(lubridate::year(timestamp))) %>%
+# #   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
+# # write_feather(inpadf_touse, "data/inpadf_touse.feather")
+inpadf_touse <- read_feather("data/inpadf_touse.feather")
+# # Ornitela: download data from movebank (just a subset of the times for now)
+# minDate <- "2020-09-01 00:00"
+# maxDate <- "2023-05-15 11:59"
+# # dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+# # write_feather(dat, "data/dat.feather")
 dat <- read_feather("data/dat.feather")
@@
 
-Downloaded data from Movebank from `r minDate` to `r maxDate`. The
-initial data has **`r nrow(dat)`** rows and `r ncol(dat)` columns. There
-are **`r length(unique(dat$trackId))`** unique individual ID's in the
-data at this point.
+When we download the data, there are `r (a1 <- nrow(dat))` GPS fixes from the Ornitela dataset and `r (a2 <- nrow(inpadf_touse))` fixes from the INPA dataset.
+Their date ranges are from `r (a3 <- min(lubridate::date(dat$timestamp), na.rm = T))` to `r (a4 <- max(lubridate::date(dat$timestamp), na.rm = T))` (Ornitela) and `r (a5 <- min(lubridate::date(inpadf_touse$timestamp), na.rm = T))` to `r (a6 <- max(lubridate::date(inpadf_touse$timestamp), na.rm = T))` (INPA).
+The Ornitela data contains `r (a7 <- length(unique(dat$trackId)))` individuals, and the INPA dataset contains `r (a8 <- length(unique(inpadf_touse$trackId)))`, for a total of `r (a9 <- length(unique(c(dat$trackId, inpadf_touse$trackId))))` unique birds.
 
-### Step 2: Fix vulture IDs
+Products: **dat** (ornitela data); **inpadf_touse** (inpa data, prepared to join)
+Metrics: how many individuals in each, how many rows in each
 
-```{r fix-trackId-Nili_id}
-## fix trackId. Marta told me that E03w was sometimes coded as E03, so I fixed that right up front. 
-dat2 <- dat %>%
+## 2. Join INPA to Ornitela data. Clean ID's.
+```{r}
+# Will there be a problem joining these?
+names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
+names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
+
+# Add the dataset names so we can keep track of where the data comes from
+inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
+  mutate(dataset = "inpa")
+dat <- dat %>%
+  mutate(dataset = "ornitela")
+
+# Join the two datasets
+joined <- bind_rows(inpadf_tojoin, dat)
+dim(joined)
+
+## fix trackId
+dat2 <- joined %>%
   mutate(trackId = as.character(trackId),
@@
 # number of unique individuals
-length(unique(dat2$trackId)) # 126--makes sense that we lost one because one individual had a typo.
+length(unique(dat2$trackId)) # 206
 
-# Read in the who's who table and prepare to join the Nili_ids
+# Add the Nili_id
 ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
@@
 
-# Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
-ww %>% filter(Movebank_id %in% dat2$trackId) %>% 
-  dplyr::select(Nili_id, Movebank_id) %>% 
-  distinct() %>% group_by(Nili_id) %>% 
-  summarize(n = length(unique(Movebank_id))) %>% 
-  arrange(desc(n))
-ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili_id's
+# Are there any other individuals who have two different associated trackId's?
+ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
+ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
 
-all(dat2$trackId %in% ww$Movebank_id) # true
+all(dat2$trackId %in% ww$Movebank_id) # false.
 
-dat2 <- left_join(dat2, ww, by = c("trackId" = "Movebank_id"))
-dim(dat2) # good, same number of rows.
-length(unique(dat2$Nili_id))
+# Okay, let's investigate the ones that aren't included
+dat2 %>%
+  filter(!(trackId %in% ww$Movebank_id)) %>%
+  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
+
+dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
+ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
+
+sort(unique(ww$mb_id))
+unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
+dat2$trackId[dat2$trackId == "e86white"] <- "e86"
+unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+# Investigate the remaining ones:
+dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
+# I guess for now let's just make their trackId's into Nili_id's
+
+dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
+                  by = c("trackId" = "mb_id"))
+nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
+length(unique(dat3$Nili_id))
+# For any that didn't get Nili_ids assigned, make it the trackId
+dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
+dat3 <- dat3 %>%
+  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
+                             TRUE ~ Nili_id))
+unique(dat3$Nili_id)
 ```
 
-Next, we join in the Who's Who table in order to fix the Movebank
-trackId's and associate them with individual vultures (as defined by the
-"Nili_id" column). After joining in the Nili_ids, we now have a data
-frame with **`r nrow(dat2)`** rows and `r ncol(dat2)` columns. There are
-now **`r length(unique(dat2$Nili_id))`** unique individual vultures
-identified--the change from **`r length(unique(dat$trackId))`** is
-because there were a couple cases of different trackId's being assigned
-to the same Nili_id, such as "Y26" and "Y26b". No individuals are
-actually lost here.
+After joining the two datasets together and cleaning up the names, we end up with a single dataset with `r (b1 <- length(unique(dat3$Nili_id)))` individuals and `r (b2 <- nrow(dat3))` fixes, and with a date range from `r (b3 <- min(lubridate::date(dat3$timestamp), na.rm = T))` to `r (b4 <- max(lubridate::date(dat3$timestamp), na.rm = T))`.
+Product: **dat3**
+Metrics: number of rows; number of individuals; date range of individuals
 
-### Step 3: Remove invalid periods (cages etc.)
-
-Next, we remove periods of time when specific vultures were in cages or
-in the hospital, or when tags were not attached to the bird, etc. This
-is based on the "periods_to_remove" sheet in the who's who file.
-
-```{r remove-periods}
-## annotate the data with periods to remove
+## 3. Remove periods in hospital/cages
+```{r}
+# ## annotate the data with periods to remove
 toRemove <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
@@
 # Join to the original data
-datAnnot <- dat2 %>%
+datAnnot <- dat3 %>%
   left_join(toRemove_long, by = c("Nili_id", "dateOnly")) %>%
   mutate(status = replace_na(status, "valid"))
-nrow(datAnnot) == nrow(dat2) #T: good, same number of rows.
+nrow(datAnnot) == nrow(dat3) #T: good, same number of rows.
 
@@
   filter(status == "valid")
-```
+nrow(datAnnot)
+length(unique(datAnnot$Nili_id))
+nrow(dat3)
+nrow(datAnnot)
 
-```{r remove-cages}
-# Remove cages using Marta's data
+#
 # Identify and remove capture dates using Marta's code --------------------
@@
 But we're just dealing with southern individuals, so that's okay.
-roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-save(roosts, file = "data/roosts.Rda")
+# roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+# save(roosts, file = "data/roosts.Rda")
 load("data/roosts.Rda")
@@
 
-datAnnot_noCaptures <- datAnnot_noCaptures %>%
-  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
-
 # How many rows and individuals did we remove?
@@
 (propChange <- (before-after)/before) # approx. 1% of data removed.
-length(unique(datAnnot$Nili_id))
-length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
+length(unique(datAnnot$Nili_id)) == length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
+
+datAnnot_noCaptures <- datAnnot_noCaptures %>%
+  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
 ```
 
-The resulting data has dropped from **`r nrow(dat2)`** to
-**`r nrow(datAnnot_noCaptures)`** rows, a
-**`r round((nrow(dat2)-nrow(datAnnot_noCaptures))*100/nrow(dat2), 3)`% decrease**.
-It has `r ncol(dat2)` columns and
-**`r length(unique(datAnnot_noCaptures$Nili_id))`** unique individuals.
+After removing periods when individuals were in the hospital, the number of fixes has been reduced to `r (c1 <- nrow(datAnnot))`, which is a `r `(round(100*(b2-c1)/b2, 3))`% decrease.
+The number of individuals has not changed: it still has `r (c2 <- length(unique(datAnnot$Nili_id)))` unique vultures.
 
-### Step 4: Attach birth year and sex info from who's who
+Then we remove periods of time when individuals were in the capture cages.
+We are left with `r (c3 <- nrow(datAnnot_noCaptures))` fixes (a further `r `(round(100*(c1-c3)/c1, 3))`% decrease) and `r (c4 <- length(unique(datAnnot_noCaptures$Nili_id)))` individuals.
 
-```{r join-age-sex}
+Product: **datAnnot** (after removing periodsToRemove/hospital stays), datAnnot_noCaptures (after removing capture cage periods)
+Metrics: number of rows
+
+## 4. Attach age and sex
+```{r}
+# Age and sex info --------------------------------------------------------
 # Attach age and sex information
@@
   left_join(as, by = "Nili_id")
-nrow(datAnnot2) == nrow(datAnnot_noCaptures)
-length(unique(datAnnot2$Nili_id)) # 125
-```
+nrow(datAnnot2) == nrow(datAnnot)
+length(unique(datAnnot2$Nili_id)) # 196
 
-No change in the number of rows here: `r nrow(datAnnot2)` rows and
-`r ncol(datAnnot2)` columns.
-
-```{r}
 datAnnot <- datAnnot2 # a hack so the rest of the code will work
@@
 
-### Step 5: Clean the data using the cleanData function.
+After attaching the age and sex information, the number of rows and individuals has not changed (`r (d1 <- nrow(datAnnot2))` rows, `r (d2 <- length(unique(datAnnot2$Nili_id)))` individuals).
 
-This function includes several data cleaning steps: - Remove as invalid
-data points that have zeroes in `ground_speed`, `barometric_height`, and
-`external_temperature`. - Remove points that took to long to get a GPS
-fix (\>89) - Remove as invalid points with heading data \<0 or \>360
-degrees - Remove points with fewer than 3 satellites (because 3 are
-needed for triangulation) - Remove points that "jump" unrealistically
-across the map, as calculated by distance/time differences between
-consecutive points - Set to NA any altitude values that jump
-unrealistically between consecutive points - Remove points that "moved"
-farther than expected during the night - Remove **individuals** where
-less than 1/3 of their duration tracked fell inside the rough
-Israel-region geographic mak - For any individuals remaining, remove
-**points** that fell outside of the geographic mask (e.g. to remove
-long-range forays).
+Product: **datAnnot2** (then it gets renamed to datAnnot so the rest of the code will run, but datAnnot2 is a direct parallel of the state of datAnnot at the end of this).
+Metrics: number of columns before and after; number of rows (to show that none were added)
 
-```{r clean-data, echo = T, include = T}
+## 5. Clean data (remove invalid points; keep individuals with at least 30 days total in the entire dataset)
+```{r}
 # Clean the data
@@
 mask <- sf::st_read("data/CutOffRegion.kml")
-datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
-save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
+#datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall.
+If we want 30 days per season, will have to do that differently (e.g.
+split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
+#save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
 load("data/datAnnotCleaned.Rda")
-```
-
-After cleaning, the new data has **`r nrow(datAnnotCleaned)`** rows (a
-**`r round((nrow(datAnnot)-nrow(datAnnotCleaned))*100/nrow(datAnnot), 3)`%
-decrease**) and `r ncol(datAnnotCleaned)` columns.
-**`r length(unique(datAnnot$Nili_id))-length(unique(datAnnotCleaned$Nili_id))`**
-individuals were removed, for a remaining total of
-**`r length(unique(datAnnotCleaned$Nili_id))`** individuals.
-
-### Step 6: Fix times and split into seasons
-
-First, I changed the time zone to Israel time instead of UTC, since this
-will make a difference to where the date line falls. Then I
-re-calculated the `dateOnly` column based on the new date/times.
-
-Next, I split the data into fall/breeding/summer seasons based on
-Marta's cutoff dates: September 15th, December 15th, and May 15th. I
-fixed the vultures' ages to adjust for the breeding season and then
-classified them as adults or sub-adults/juveniles. Finally, I split the
-data into a list, with one item per season.
-
-```{r fix-dates}
+length(unique(datAnnotCleaned$Nili_id)) # 172 individuals
 # Fix time zone so dates make sense ---------------------------------------
@@
 
-```{r split-seasons}
-# Split into Marta's 3 seasons --------------------------------------------
-s1 <- datAnnotCleaned %>%
-  mutate(start_breeding = ifelse(month %in% c(1:5),
-                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
-                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
-         start_summer = ifelse(month == 12,
-                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
-                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
-         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
-         start_breeding = as.Date(start_breeding),
-         start_summer = as.Date(start_summer),
-         season = case_when(
-           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
-           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
-           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
-  filter(dateOnly != "2022-12-15") # remove December 15 2022, because it turns out that's a non-inclusive boundary (oops)
+Next, we clean the data with the `vultureUtils::cleanData()` function, removing invalid points.
+We retain only vultures that are tracked for at least 30 days in the entire dataset.
+We are left with `r (e1 <- nrow(datAnnotCleaned))` fixes from `r (e2 <- length(unique(datAnnotCleaned$Nili_id)))` unique vultures.
+This is a `r round(100*((d1-e1)/d1), 3)`% decrease in fixes and a `r (round(100*((d2-e2)/d2), 3))`% decrease in individuals.
 
-s2 <- s1 %>%
-  mutate(age = year - birth_year,
-         sex = case_when(sex == "f/m" ~ "u",
-                         TRUE ~ sex)) %>%
-  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
-  group_by(Nili_id, year, month) %>%
-  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
-  mutate(age_group = case_when(age == 0 ~ "juv",
-                               age >= 1 & age <= 4 ~ "sub",
-                               age >= 5 ~ "adult",
-                               TRUE ~ NA)) %>%
-  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
-                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
-                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
-  # explicitly set the factor levels in temporal order
-  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall")),
-         season = factor(season, levels = c("breeding", "summer", "fall")))
-table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read
+Product: **datAnnotCleaned**
+Metrics: number of individuals, number of rows
 
-# remove columns that were used in the season calculation
-s3 <- s2 %>%
-  dplyr::select(-c(month, start_breeding, start_summer, start_fall))
-
-# Separate the seasons -----------------------------
-seasons <- s3 %>%
-  group_by(seasonUnique) %>%
-  group_split(.keep = T)
-# extract the season names in case we need a separate vector at some point.
-seasonNames <- map_chr(seasons, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
+## 6. Split into seasons
+```{r}
+# # Split into Marta's 3 seasons --------------------------------------------
+# s1 <- datAnnotCleaned %>%
+#   mutate(start_breeding = ifelse(month %in% c(1:5),
+#                                  paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
+#                                  paste(as.character(year), "-", "12", "-", "15", sep = "")),
+#          start_summer = ifelse(month == 12,
+#                                paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
+#                                paste(as.character(year), "-", "05", "-", "15", sep = "")),
+#          start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
+#          start_breeding = as.Date(start_breeding),
+#          start_summer = as.Date(start_summer),
+#          season = case_when(
+#            dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
+#            dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
+#            dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
+#   filter(dateOnly != "2023-05-15") # remove May 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
+# 
+# s2 <- s1 %>%
+#   mutate(age = year - birth_year,
+#          sex = case_when(sex == "f/m" ~ "u",
+#                          TRUE ~ sex)) %>%
+#   mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
+#   group_by(Nili_id, year, month) %>%
+#   mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
+#   mutate(age_group = case_when(age == 0 ~ "juv",
+#                                age >= 1 & age <= 4 ~ "sub",
+#                                age >= 5 ~ "adult",
+#                                TRUE ~ NA)) %>%
+#   mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
+#                                   season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
+#                                   season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
+#   # explicitly set the factor levels in temporal order
+#   mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding")),
+#          season = factor(season, levels = c("breeding", "summer", "fall")))
+# table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read
+# 
+# # remove columns that were used in the season calculation
+# s3 <- s2 %>%
+#   dplyr::select(-c(month, start_breeding, start_summer, start_fall))
+# 
+# # Separate the seasons -----------------------------
+# seasons_orig <- s3 %>%
+#   group_by(seasonUnique) %>%
+#   group_split(.keep = T)
+# # extract the season names in case we need a separate vector at some point.
+# seasonNames_orig <- map_chr(seasons_orig, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
+# save(seasons_orig, file = "data/seasons_orig.Rda")
+# save(seasonNames_orig, file = "data/seasonNames_orig.Rda")
+load("data/seasons_orig.Rda")
+load("data/seasonNames_orig.Rda")
 ```
 
-```{r save-soc-data}
+We've split the data into seasons now.
+The season names are: `r (f1 <- seasonNames_orig)`.
+We have a total of `r (f2 <- length(f1))` seasons.
+Here are the date ranges, numbers of fixes, and numbers of individuals for each season:
+```{r echo = F, include = T}
+mins <- map_chr(seasons_orig, ~as.character(min(lubridate::date(.x$timestamp), na.rm = T)))
+maxes <- map_chr(seasons_orig, ~as.character(max(lubridate::date(.x$timestamp), na.rm = T)))
+fixes <- map_dbl(seasons_orig, nrow)
+inds <- map_dbl(seasons_orig, ~length(unique(.x$Nili_id)))
+f3 <- data.frame(season = seasonNames_orig, startDate = mins, endDate = maxes, nFixes = fixes, nVultures = inds)
+f3
+```
+
+Products: seasons_orig, seasonNames_orig
+Metrics: number of seasons; number of individuals and rows per season; which seasons are included; date ranges for each season.
+
+## 7. Remove individuals with too low a fix rate: Exclude individuals that never have a daily mode of 10 minutes
+```{r}
+# Optionally remove individuals with a low fix rate -----------------------
+Mode <- function(x) { # function to calculate mode fix rate
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
+# Calculate daily modes and add them to the dataset
+seasons <- map(seasons_orig, ~.x %>%
+                 group_by(dateOnly, Nili_id) %>%
+                 mutate(diff = as.numeric(difftime(lead(timestamp), timestamp, units = "mins"))) %>%
+                 group_by(dateOnly, Nili_id) %>%
+                 mutate(mode = Mode(round(diff))) %>%
+                 ungroup())
+
+# Exclude individuals that never have a daily mode of 10 minutes
+toKeep_fixrate <- map(seasons, ~.x %>%
+                sf::st_drop_geometry() %>%
+                group_by(Nili_id) %>%
+                summarize(minmode = min(mode, na.rm = T)) %>%
+                filter(minmode <= 10) %>%
+                pull(Nili_id) %>%
+                unique())
+save(toKeep_fixrate, file = "data/toKeep_fixrate.Rda")
+load("data/toKeep_fixrate.Rda")
+seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
+```
+Product: seasons, seasons_mode10
+Metrics: number of individuals per season, number of fixes per season
+
+After removing individuals that never have a 10-minute fix rate, here are the numbers of fixes and numbers of individuals for each season:
+```{r echo = F, include = T}
+fixes <- map_dbl(seasons_mode10, nrow)
+inds <- map_dbl(seasons_mode10, ~length(unique(.x$Nili_id)))
+g1 <- data.frame(season = seasonNames_orig, nFixes = fixes, nVultures = inds)
+g1
+```
+
+At this point, we save the data for social analysis, saving both seasons_forSoc (which contains all individuals, regardless of fix rate) and seasons_forSoc_mode10 (which contains only individuals with a 10-minute fix rate).
+
+## 8. Save SOCIAL data
+```{r}
 # Save copies for social analysis
@@
 seasons_forSoc <- seasons
-#save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
+seasons_forSoc_mode10 <- seasons_mode10
+save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
+save(seasons_forSoc_mode10, file = "data/seasons_forSoc_mode10.Rda")
+load("data/seasons_forSoc.Rda")
+load("data/seasons_forSoc_mode10.Rda")
+# rename, preparing to carry on
+seasons_mode10 <- seasons_forSoc_mode10
+seasons <- seasons_forSoc
+
+# Investigate: How many individuals are retained in each of these cases? What proportion were removed?
+inds_all <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
+inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
+propRemoved <- round((inds_all-inds_mode10)/inds_all, 3) # For all seasons, getting rid of individuals that don't have
+propRemoved # yikes that's a lot of birds. Alas.
 ```
 
-There are `r length(seasonNames)` seasons: `r seasonNames` (we will end
-up removing Summer 2020 because the Ornitela tags started partway
-through that season.) `seasons` is therefore a list of length
-**`r length(seasons)`**, with the following numbers of rows in each
-season: `r map_dbl(seasons, nrow)`. This data is saved as
-`seasons_forSoc.Rda` so it can be used for the social network analysis
-without any further changes made.
+Here are the numbers of individuals for each of these products:
+```{r include = T, echo = F}
+inds_reg <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
+inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
+h1 <- data.frame(season = seasonNames_orig, allIndivs = inds_reg, highFixRate = inds_mode10) %>% mutate(propHighFixRate = round(highFixRate/allIndivs, 2))
+h1
+```
 
-### Step 7: Get roosts
+Products: seasons_forSoc, seasons_forSoc_mode10. These shouldn't be any different from seasons and seasons_mode10, there's just some renaming that happens.
 
-Now I use Marta's roost code, implemented in vultureUtils as
-`get_roosts_df`, to get roost locations for each individual on each
-night.
-
-```{r get-roosts}
+## 9. Get ROOSTS
+```{r}
 # Get roosts for each season ----------------------------------------------
-# forRoosts <- map(seasons, ~.x %>% sf::st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE))
-# roosts_seasons <- purrr::map(forRoosts, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
+# roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
 # roosts_seasons <- roosts_seasons %>%
@@
 load("data/roosts_seasons.Rda")
+
+# roosts_seasons_mode10 <- purrr::map(seasons_mode10, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+# roosts_seasons_mode10 <- roosts_seasons_mode10 %>%
+#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
+# save(roosts_seasons_mode10, file = "data/roosts_seasons_mode10.Rda")
+load("data/roosts_seasons_mode10.Rda")
+```
+Products: roosts_seasons and roosts_seasons_mode10
+Metrics: number of individuals per season, number of rows per season
+
+Now, we use the full dataset (not restricted to high fix rate individuals) to obtain a locations for each individual on each night. This gives us:
+```{r echo = F, include = T}
+cat("All individuals:\n")
+(i1 <- data.frame(season = seasonNames_orig, 
+                 nIndivs = map_dbl(roosts_seasons, ~length(unique(.x$Nili_id))),
+                 vultureNights = map_dbl(roosts_seasons, nrow)))
 ```
 
-### Step 8: Remove nighttime points
+And for comparison, the vultures and vultureNights for roosts including only the high fix rate individuals:
 
-Just for the data I'm going to use for movement analyses, I want to
-exclude any points that occur at night. This will also help deal with
-the problem that there are some individuals that seem to have way too
-many points per day
+```{r echo = F, include = T}
+cat("\nHigh fix rate individuals:\n")
+(i2 <- data.frame(season = seasonNames_orig, 
+                 nIndivs = map_dbl(roosts_seasons_mode10, ~length(unique(.x$Nili_id))),
+                 vultureNights = map_dbl(roosts_seasons_mode10, nrow)))
+```
 
-```{r remove-night}
+## 10. Remove night points
+```{r}
 # Remove nighttime points -------------------------------------------------
 before <- map_dbl(seasons, nrow)
+beforeindivs <- map_dbl(seasons, ~length(unique(.x$Nili_id)))
 seasons <- map(seasons, ~{
   times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
-                                     lat = 31.434306, lon = 34.991889,
+                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                      keep = c("sunrise", "sunset")) %>%
@@
 })
-after <- map_dbl(seasons, nrow)
-change <- after-before
-changeDF <- data.frame(seasonName = seasonNames, nBefore = before, nAfter = after) %>%
-  mutate(propRowsRemoved = round((before-after)/before, 2))
+# after <- map_dbl(seasons, nrow)
+# change <- after-before
+# changeDF <- data.frame(seasonName = seasonNames_orig, nBefore = before, nAfter = after) %>%
+#   mutate(propRowsRemoved = round((before-after)/before, 2))
+# changeDF
+```
+Product: seasons/"after" (maybe need a unique name here?)
+Metrics: number of individuals per season (should be the same), number of rows per season
+
+After removing the night points, the number of individuals per season should not have changed, but some rows have been removed.
+```{r}
+j1 <- data.frame(season = seasonNames_orig, 
+                 nIndivsBefore = beforeindivs,
+                 nIndivsNow = map_dbl(seasons, ~length(unique(.x$Nili_id))),
+                 nIndivs = map_dbl(roosts_seasons, ~length(unique(.x$Nili_id))),
+                 vultureNights = map_dbl(roosts_seasons, nrow))
 ```
 
-```{r include = T, echo = F}
-print(changeDF)
-```
+<!-- ## 11. Restrict to southern individuals -->
+<!-- ```{r} -->
+<!-- # Restrict to southern individuals ---------------------------------------- -->
+<!-- # Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636) -->
+<!-- ## Transform to SF object, so we can get centroids -->
+<!-- # seasonsSF <- map(seasons_orig, ~.x %>% -->
+<!-- #                    sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>% -->
+<!-- #                    sf::st_set_crs("WGS84") %>% -->
+<!-- #                    sf::st_transform(32636)) -->
+<!-- #  -->
+<!-- # save(seasonsSF, file = "data/seasonsSF.Rda") -->
+<!-- load("data/seasonsSF.Rda") -->
 
-### Step 9: Include only "southern individuals" in the movement calculations
+<!-- ## Get centroids, so we can see who's "southern" for that season. -->
+<!-- centroids <- map(seasonsSF, ~.x %>% -->
+<!--                    group_by(Nili_id) %>% -->
+<!--                    summarize(geometry = st_union(geometry)) %>% -->
+<!--                    st_centroid()) -->
 
-For each individual in each season, I calculated the centroid of its
-coordinates. Then I looked at some histograms of the centroids in each
-season and decided that 3550000 latitude (in UTM zone 36N coordinates--https://epsg.io/32636) would be a
-cutoff point that would neatly divide individuals in the southern
-population from the northern population (there was a large gap in the
-histogram consistently around this value).
+<!-- ## Examine a histogram of centroid latitudes  -->
+<!-- # walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here. -->
 
-I filtered each season down to remove individuals classified as
-"northern" for that season, leaving only those in the Negev population.
-Keep in mind:
+<!-- ## Get southern individuals for each season, so we can filter the data -->
+<!-- southernIndivs <- map(centroids, ~.x %>% -->
+<!--                         filter(st_coordinates(.)[,2] < 3550000) %>% -->
+<!--                         pull(Nili_id)) -->
 
-1)  A given individual could be "northern" in one season but "southern"
-    in another season, theoretically. It would be removed only in the
-    seasons when it falls in the north.
+<!-- ## Remove individuals not in the south -->
+<!-- before <- seasons -->
+<!-- seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y)) -->
+<!-- after <- seasons -->
 
-2)  Any removed individuals can still contribute to the social network
-    that I will calculate--that is, they can contribute to the
-    degree/strength/etc. of southern birds. I'm only removing them for
-    the movement calculations, so they won't be focal individuals in the
-    model.
+<!-- beforeRows <- map_dbl(before, nrow) -->
+<!-- afterRows <- map_dbl(after, nrow) -->
+<!-- beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id))) -->
+<!-- afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id))) -->
+<!-- df <- data.frame(season = seasonNames_orig, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs) -->
+<!-- df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2) -->
+<!-- df -->
+<!-- rm(centroids) -->
+<!-- rm(seasonsSF) -->
+<!-- ``` -->
+<!-- Product: seasons/after (unique name?) -->
+<!-- Metrics: number of individuals per season, number of rows per season, boxplot of days per individual per season (so we can compare?) -->
 
-```{r remove-northern}
-# Restrict to southern individuals ----------------------------------------
-# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636)
-## Transform to SF object, so we can get centroids
-seasonsSF <- map(seasons, ~.x %>%
-                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
-                   sf::st_set_crs("WGS84") %>%
-                   sf::st_transform(32636))
+<!-- ## 12. Include only days with > 10ppd unless the battery never dropped below 50% on that day -->
+<!-- ```{r} -->
+<!-- ## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50% -->
+<!-- seasons <- map(seasons, ~.x %>% -->
+<!--                  group_by(Nili_id, dateOnly) %>% -->
+<!--                  filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>% -->
+<!--                  ungroup()) -->
+<!-- ``` -->
+<!-- Product: seasons (unique name?) -->
+<!-- Metrics: number of individuals per season (should not have changed since previous), number of rows per season, boxplot of days per individual per season (compare to previous) -->
 
-## Get centroids, so we can see who's "southern" for that season.
-centroids <- map(seasonsSF, ~.x %>%
-                   group_by(Nili_id) %>%
-                   summarize(geometry = st_union(geometry)) %>%
-                   st_centroid()) # XXX what is the actual definition of the centroid?
+<!-- ## 13. For each season, exclude individuals with < 30 valid days (after removing days with too few points) -->
+<!-- ```{r} -->
+<!-- # Include only individuals with enough days tracked ----------------------- -->
+<!-- ## Must be tracked for at least 30 days per season -->
+<!-- durs <- map_dbl(seasons, ~length(unique(.x$dateOnly))) -->
 
-## Examine a histogram of centroid latitudes 
-# walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
+<!-- # walk2(seasons, durs, ~.x %>% -->
+<!-- #         st_drop_geometry() %>% -->
+<!-- #         group_by(Nili_id) %>% -->
+<!-- #         summarize(nDaysTracked = length(unique(dateOnly)), -->
+<!-- #                   propDaysTracked = nDaysTracked/.y) %>% -->
+<!-- #         pull(propDaysTracked) %>% -->
+<!-- #         hist()) -->
 
-## Get southern individuals for each season, so we can filter the data
-southernIndivs <- map(centroids, ~.x %>%
-                        filter(st_coordinates(.)[,2] < 3550000) %>%
-                        pull(Nili_id))
+<!-- indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>% -->
+<!--                        st_drop_geometry() %>% -->
+<!--                        group_by(Nili_id) %>% -->
+<!--                        summarize(nDaysTracked = length(unique(dateOnly)), -->
+<!--                                  propDaysTracked = nDaysTracked/.y) %>% -->
+<!--                        filter(nDaysTracked >= 30) %>% -->
+<!--                        pull(Nili_id)) -->
 
-## Remove individuals not in the south
-before <- seasons
-seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
-after <- seasons
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
-afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
-df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
-df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
-```
+<!-- ## remove those individuals not tracked for enough days -->
+<!-- before <- seasons -->
+<!-- after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y)) -->
 
-After doing this filtering, the following numbers of rows and
-individuals remained in each season:
+<!-- beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id))) -->
+<!-- afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id))) -->
+<!-- seasons <- after -->
+<!-- ``` -->
+<!-- Product: seasons/after (unique name) -->
+<!-- Metric: number of individuals per season (should have decreased), number of points per season, boxplot of days per individual per season -->
 
-```{r include = T, echo = F}
-print(df)
-```
+<!-- ## 14. Get elevation information (and remove summer 2020 because it has too little data) -->
+<!-- ```{r} -->
+<!-- # Get elevation rasters --------------------------------------------------- -->
+<!-- # Remove the first element of `seasons` because it doesn't have any data -->
+<!-- # seasons <- seasons[-1] # removing summer 2020 -->
+<!-- # # XXXXXXX start here -->
+<!-- # seasons <- map(seasons, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84")) -->
+<!-- # elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x , z = 10)) -->
+<!-- # save(elevs_z10_seasons, file = "data/elevs_z10_seasons.Rda") -->
+<!-- load("data/elevs_z10_seasons.Rda") -->
 
-### Step 10: Remove vulture\*days based on number of points per day
+<!-- # groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84. -->
+<!-- # save(groundElev_z10, file = "data/groundElev_z10.Rda") -->
+<!-- load("data/groundElev_z10.Rda") -->
 
-```{r view-ppd}
-# Include only individuals with enough points per day ------------------------
-# How many points per day are we looking at?
-ppd <- map_dfr(seasons, ~.x %>%
-             st_drop_geometry() %>%
-             group_by(seasonUnique, season, Nili_id, dateOnly) %>%
-             summarize(n = n()))
+<!-- ## use elevation rasters to calculate height above ground level -->
+<!-- before <- seasons -->
+<!-- seasons <- map2(.x = seasons, .y = groundElev_z10,  -->
+<!--                 ~.x %>% mutate(height_above_ground = .y, -->
+<!--                                height_above_ground = case_when(height_above_ground < 0 ~ 0, -->
+<!--                                                                TRUE ~ height_above_ground))) -->
+<!-- after <- seasons -->
 
-## Theoretically, should only have 72 points per day--12 hours, 6 points per hour.
-That should vary a bit.
-I don't really understand why so many individuals/days have more than 100 points per day, but let's proceed for now: going to downsample that to 10 minue intervals anyway, so it won't matter.
-```
+<!-- # There should be no change in the number of rows -->
+<!-- all(map2_lgl(before, after, ~nrow(.x) == nrow(.y))) -->
+<!-- ``` -->
+<!-- Product: seasons/after -->
+<!-- Metrics: number of seasons (should be one fewer than before), number of rows per season (expect no change), number of columns, histograms of elevation by season (?) -->
 
-We had initially decided to remove vulture\*days with \< 30 points per
-day. Here's what that would look like:
+<!-- ## 15. Fix rate again (why?) -->
+<!-- ```{r} -->
+<!-- # Now restrict to the mode10 individuals ---------------------------------- -->
+<!-- load("data/toKeep_fixrate.Rda") -->
+<!-- toKeep_fixrate <- toKeep_fixrate[-1] -->
+<!-- seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y)) -->
 
-```{r include = T, echo = F}
-# 
-thresh <- 30
-ppd %>%
-  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
-  ggplot(aes(x = n))+
-  geom_histogram(aes(fill = Status))+
-  facet_wrap(~seasonUnique, scales ="free")+
-  geom_vline(aes(xintercept = thresh), col = "red")+
-  scale_fill_manual(values = c("darkgray", "red"))+
-  theme_classic()+
-  ggtitle(paste("n removed =", thresh, sep = " "))+
-  ylab("# vulture*days")+
-  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
-```
+<!-- #save(seasons, file = "data/seasons.Rda") -->
+<!-- #save(seasons_mode10, file = "data/seasons_mode10.Rda") -->
+<!-- load("data/seasons.Rda") -->
+<!-- load("data/seasons_mode10.Rda") -->
+<!-- ``` -->
+<!-- Products: seasons, seasons_mode10 -->
+<!-- Metrics: Number of individuals per season (I don't think this should have changed from the step before? I'm not clear on why we are doing this again) -->
 
-Restricting to days with 30 points per day ends up removing a huge
-number of days, especially in the breeding season (winter) because of
-lack of sunlight etc. After talking with Marta and Noa, I decided to
-remove just 10 points per day. That would look like this:
+<!-- ## 16. Downsample to 10 minutes to deal with the more frequent fixes -->
+<!-- ```{r} -->
+<!-- # Downsample the data and save the downsampled data -->
+<!-- subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){ -->
+<!--   sub <- df %>% -->
+<!--     arrange(idCol, timestampCol) %>% -->
+<!--     group_by(.data[[idCol]]) %>% -->
+<!--     mutate(tc = cut(.data[[timestampCol]],  -->
+<!--                     breaks = paste(as.character(mins), "min", sep = " "))) %>% -->
+<!--     group_by(.data[[idCol]],  -->
+<!--              "d" = lubridate::date(.data[[timestampCol]]), tc) %>% -->
+<!--     slice(1) %>% -->
+<!--     ungroup() %>% -->
+<!--     dplyr::select(-c("d", "tc")) -->
+<!--   return(sub) -->
+<!-- } -->
 
-```{r include = T, echo = F}
-# 
-thresh <- 10
-ppd %>%
-  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
-  ggplot(aes(x = n))+
-  geom_histogram(aes(fill = Status))+
-  facet_wrap(~seasonUnique, scales ="free")+
-  geom_vline(aes(xintercept = thresh), col = "red")+
-  scale_fill_manual(values = c("darkgray", "red"))+
-  theme_classic()+
-  ggtitle(paste("n removed =", thresh, sep = " "))+
-  ylab("# vulture*days")+
-  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
-```
+<!-- # create and save downsampled datasets for later use -->
+<!-- seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
+-->
+<!-- seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) -->
 
-```{r}
-# thresh <- 10
-# ppd <- ppd %>%
-#   ungroup() %>%
-#   mutate(removed = ifelse(n < thresh, T, F))
-# 
-# # what proportion of vulture*days are removed overall?
-# (propRemoved <- ppd %>%
-#     group_by(seasonUnique, season) %>%
-#     summarize(nTot = n(),
-#               nRemoved = sum(removed),
-#               propRemoved = nRemoved/nTot))
-# propRemoved %>%
-#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
-#   geom_point(size = 10)+
-#   theme_classic()+
-#   scale_color_manual(values = c("blue", "red", "darkorange"))
-# 
-# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
-# 
-# # what about by individual over each season?
-# ppd %>%
-#   group_by(seasonUnique, season, Nili_id) %>%
-#   summarize(propRemoved = sum(removed)/n()) %>%
-#   ggplot(aes(x = seasonUnique, y = propRemoved))+
-#   geom_boxplot(aes(group = seasonUnique, fill = season))+
-#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
-#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
-#   theme_classic()+ # tells the same story
-#   ylab("Proportion of days removed")+
-#   xlab("")
+<!-- map_dbl(seasons, nrow) -->
+<!-- map_dbl(seasons_10min, nrow) -->
+<!-- map_dbl(seasons_mode10, nrow) -->
 
-## Remove days on which individuals don't have at least 10 points.
-before <- seasons
-seasons <- map(seasons, ~.x %>%
-                 group_by(Nili_id, dateOnly) %>%
-                 filter(n() >= 10) %>% # hard-coding the 10 in here so it doesn't accidentally get changed.
-                 ungroup())
-after <- seasons
+<!-- seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly)))) -->
+<!-- seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly)))) -->
 
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
-  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))
-```
-
-Here's how many points get removed when we do this filtering:
-
-```{r include = T, echo = F}
-df
-```
-
-### Step 11: Remove individuals based on number of days tracked
-
-This filtering comes after the removal of some vulture\*days for having
-too few points. So, considering the remaining days, we only want to
-retain individuals that are tracked for at least 1/3 of a given season
-(in terms of number of days).
-
-Here are the number of days in each season. As you can see, breeding
-season is the longest, followed by summer, then fall. (Ignore Summer
-2020--it will not be included in the analysis.)
-
-```{r include = TRUE, echo = F}
-seasonNames
-durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))
-durs
-```
-
-```{r remove-daysTracked}
-# Include only individuals with enough days tracked -----------------------
-## Must be tracked for at least 1/3 of the number of days in the season.
-
-# walk2(seasons, durs, ~.x %>%
-#   st_drop_geometry() %>%
-#   group_by(Nili_id) %>%
-#   summarize(nDaysTracked = length(unique(dateOnly)),
-#             propDaysTracked = nDaysTracked/.y) %>%
-#   pull(propDaysTracked) %>%
-#   hist())
-
-indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
-                       st_drop_geometry() %>%
-                       group_by(Nili_id) %>%
-                       summarize(nDaysTracked = length(unique(dateOnly)),
-                                 propDaysTracked = nDaysTracked/.y) %>%
-                       filter(propDaysTracked >= 0.33) %>% # XXX ask Orr if this should be percentage vs. number of days, given the different season lengths. (email)
-                       pull(Nili_id))
-
-## remove those individuals not tracked for enough days
-before <- seasons
-after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))
-
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
-  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))
-
-```
-
-Here's what this does to the row counts:
-
-```{r include = T, echo = F}
-df
-```
-
-### Step 12: Calculate flight altitudes
-
-Using the `elevatr` package, I get ground elevation values for each gps
-point in each season, at zoom level 10 (which should come out to around
-150 meter cell resolution in the resulting raster, according to )
-
-```{r get-elevation}
-# Get elevation rasters ---------------------------------------------------
-elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x %>% st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE, crs = "WGS84"), z = 10))
-# 
-save(elevs_z10_seasons, file = "data/elevs_z10_seasons.Rda")
-load("data/elevs_z10_seasons.Rda")
-
-groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y %>% sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = FALSE))) # have to convert the crs because the raster from elevatr is in WGS84.
-
-## use elevation rasters to calculate height above ground level
-seasons <- map2(.x = seasons, .y = groundElev_z10, 
-                ~.x %>% mutate(height_above_ground = .y,
-                               height_above_ground = case_when(height_above_ground <0 ~ 0,
-                                                               TRUE ~ height_above_ground)))
-
-# Export the data
-save(seasons, file = "data/seasons.Rda")
-```
-
-As expected, adding the elevation information doesn't change the number
-of rows in the data.
-
-### Step 13: Downsample the data
-
-Because occasionally, the vultures' tags were set to read more
-frequently than every 10 minutes, I want to downsample the data to a
-10-minute interval.
-
-```{r downsample}
-# Downsample the data and save the downsampled data
-subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
-  sub <- df %>%
-    arrange(idCol, timestampCol) %>%
-    group_by(.data[[idCol]]) %>%
-    mutate(tc = cut(.data[[timestampCol]], 
-                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
-    group_by(.data[[idCol]], 
-             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
-    slice(1) %>%
-    ungroup() %>%
-    dplyr::select(-c("d", "tc"))
-  return(sub)
-}
-
-# create and save downsampled datasets for later use
-seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
-seasons_20min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 20), .progress = T)
-seasons_30min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 30), .progress = T)
-seasons_60min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 60), .progress = T)
-seasons_120min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 120), .progress = T)
-
-map_dbl(seasons, nrow)
-map_dbl(seasons_10min, nrow)
-
-# save(seasons_10min, file = "data/seasons_10min.Rda")
-# save(seasons_20min, file = "data/seasons_20min.Rda")
-# save(seasons_30min, file = "data/seasons_30min.Rda")
-# save(seasons_60min, file = "data/seasons_60min.Rda")
-# save(seasons_120min, file = "data/seasons_120min.Rda")
-
-# no need to calculate roosts for the downsampled data because in theory there's only one point per night.
-
-df <- data.frame(season = seasonNames, 
-                 nBefore = map_dbl(seasons, nrow),
-                 nAfter = map_dbl(seasons_10min, nrow)) %>%
-  mutate(propChange = round((nBefore-nAfter)/nBefore, 2))
-```
-
-Here are the row counts after downsampling:
-
-```{r include = TRUE, echo = F}
-df
-```
-
-And here are the final counts of number of vultures:
-
-```{r final-counts}
-df <- seasons_10min %>% purrr::list_rbind() %>% 
-  group_by(seasonUnique) %>% 
-  summarize(vultures = length(unique(Nili_id)))
-```
-
-```{r include = TRUE, echo = F}
-df
-```
+<!-- save(seasons_10min, file = "data/seasons_10min.Rda") -->
+<!-- save(seasons_mode10_10min, file = "data/seasons_mode10_10min.Rda") -->
+<!-- ``` -->

### 2023-08-17 00_dataPrep.R

@@ -3,7 +3,6 @@
 ## Load packages -----------------------------------------------------------
 library(vultureUtils)
 library(sf)
-# In lieu of tidyverse
 library(tidyverse)
 library(move)
 library(feather)
@@ -647,12 +646,6 @@ subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 
 # create and save downsampled datasets for later use
 seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
 seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
-# XXX will eventually have to come back to this and bring the mode10 dataset all the way through to the end here.
-# seasons_15min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 15), .progress = T)
-# seasons_20min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 20), .progress = T)
-# seasons_30min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 30), .progress = T)
-# seasons_60min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 60), .progress = T)
-# seasons_120min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 120), .progress = T)
 
 map_dbl(seasons, nrow)
 map_dbl(seasons_10min, nrow)
@@ -662,12 +655,7 @@ seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% 
 seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
 
 save(seasons_10min, file = "data/seasons_10min.Rda")
-# save(seasons_15min, file = "data/seasons_15min.Rda")
 save(seasons_mode10_10min, file = "data/seasons_mode10_10min.Rda")
-# save(seasons_20min, file = "data/seasons_20min.Rda")
-# save(seasons_30min, file = "data/seasons_30min.Rda")
[…]

### 2023-08-17 05.0_viz_descriptive.R
changed colors from seasons to situations

@@ -13,9 +13,10 @@ linked %>%
   ggplot(aes(x = evenness))+
   geom_density(aes(col = season), linewidth = 1.5)+
   facet_wrap(~type, scales = "free_y")+
-  scale_color_manual(name = "Season", values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
+  scale_color_manual(name = "Season", values = c(cc[["breedingColor"]], cc[["summerColor"]], cc[["fallColor"]]))+
   xlab("Evenness")+
   ylab("")+
+  theme_classic()+
   theme(text = element_text(size = 16))
 
 linked %>%
@@ -23,9 +24,10 @@ linked %>%
   ggplot(aes(x = degreeRelative))+
   geom_density(aes(col = season), linewidth = 1.5)+
   facet_wrap(~type, scales = "free_y")+
-  scale_color_manual(name = "Season", values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
+  scale_color_manual(name = "Season", values = c(cc[["breedingColor"]], cc[["summerColor"]], cc[["fallColor"]]))+
   xlab("Degree (normalized)")+
-  ylab("")+
+  ylab("")+ 
+  theme_classic()+
   theme(text = element_text(size = 16))
 
 linked %>%
@@ -33,7 +35,7 @@ linked %>%
   ggplot(aes(x = degree))+
   geom_density(aes(col = season), linewidth = 1.5)+
   facet_wrap(~type*year, scales = "free_y")+
-  scale_color_manual(name = "Season", values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
+  scale_color_manual(name = "Season", values = c(cc[["breedingColor"]], cc[["summerColor"]], cc[["fallColor"]]))+
   xlab("Degree")+
   ylab("")+
   theme(text = element_text(size = 16))
@@ -43,7 +45,7 @@ linked %>%
   ggplot(aes(x = strengthRelative))+
   geom_density(aes(col = season), linewidth = 1.5)+
   facet_wrap(~type, scales = "free_y")+
-  scale_color_manual(name = "Season", values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
+  scale_color_manual(name = "Season", values = c(cc[["breedingColor"]], cc[["summerColor"]], cc[["fallColor"]]))+
[…]

### 2023-08-17 05.1_viz_plotModels.R
changed point sizes

@@ -107,7 +107,7 @@ ggplot(d_eff_pc2_situ, aes(x, predicted)) +
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
               alpha = 0.2, linewidth = 0.6, show.legend = F)+
   geom_line(aes(col = group), linewidth = 1)+
-  geom_point(data = forModeling, aes(x = PC1, y = degree, col = situ), alpha = 0.5, size = 0.7)+
+  geom_point(data = forModeling, aes(x = PC1, y = degree, col = situ), alpha = 0.5)+
   scale_color_manual(name = "Situation", values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
   scale_fill_manual(name = "Situation", values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
   ylab("Degree")+
@@ -185,6 +185,7 @@ ggplot(s_eff_pc1_situ_season, aes(x, predicted)) +
 
 pc1_situ_season_emt <- emtrends(s, specs = c("situ", "season"), var = "PC1")
 pc1_situ_season_emt # There is no significant trend in any season for co-flight. For both co-roosting and co-feeding, there is a significant negative trend in the breeding season and the fall, but not in the summer. So, in summer, there is no significant relationship between movement and strength at all.
+pairs(pc1_situ_season_emt)
 
 # in terms of percent decreases:
 as.data.frame(pc1_situ_season_emt) %>% mutate(across(c("PC1.trend", "lower.CL", "upper.CL"), 

### 2023-08-17 05.2_viz_plotPCExtremes.R
to allow st_drop_geometry to work

@@ -1,6 +1,7 @@
 # Script for creating maps showing individuals' trajectories on either end of the principal component spectrum.
 library(tidyverse)
 library(ggspatial)
+library(sf)
 
 load("data/linked.Rda")
 load("data/seasons_10min.Rda")

### 2023-08-17 ggeffects_ggpredict_plotTesting_2023-08-04.R
testing ggeffects and ggpredict plotting with raw data

@@ -1,0 +1,27 @@
+library(tidyverse)
+library(sjPlot)
+library(ggeffects)
+load("d.Rda")
+
+# Using plot_model
+# Conditional (predict) plot with plot_model
+plot_model(d, type = "pred", terms = c("PC2", "situ"), show.data = T)+
+  theme_classic()
+
+# Marginal effects plot with plot_model
+plot_model(d, type = "eff", terms = c("PC2", "situ"), show.data = T)+
+  theme_classic()
+
+# Using ggeffects and then ggplot
+pred <- as.data.frame(ggpredict(d, terms = c("PC2", "situ")))
+eff <- as.data.frame(ggeffect(d, terms = c("PC2", "situ")))
+ggplot(pred, aes(x, predicted)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_line(aes(col = group), linewidth = 1.5)
+
+ggplot(eff, aes(x, predicted)) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_line(aes(col = group), linewidth = 1.5)+
+  geom_point(data = forModeling, aes(x = PC2, y = degree_scl, col = situ))

### 2023-08-17 05.2_viz_plotPCExtremes.R
Maps showing PC1 and PC2 extremes on a map, zoom level 9

@@ -2,10 +2,14 @@
 library(tidyverse)
 library(ggspatial)
 library(sf)
+library(grid)
+
+theme_set(theme_classic(base_size = 9))
 
 load("data/linked.Rda")
 load("data/seasons_10min.Rda")
 seasonNames <- map_chr(seasons_10min, ~as.character(.x$seasonUnique[1]))
+names(seasons_10min) <- seasonNames
 
 daysTracked <- map(seasons_10min, ~.x %>% st_drop_geometry() %>% select(Nili_id, seasonUnique, daysTracked) %>% distinct()) %>% purrr::list_rbind()
 
@@ -19,7 +23,7 @@ forplots <- linked %>%
   mutate(seasonUnique = paste(year, season, sep = "_")) %>%
   left_join(daysTracked, by = c("seasonUnique", "Nili_id"))
 
-# Take a look at the biplot again
+# Take a look at the biplot again, this time facetted by season
 forplots %>%
   ggplot(aes(x = PC1, y = PC2, col = daysTracked))+
   geom_point(size = 2)+
@@ -28,6 +32,80 @@ forplots %>%
   theme_bw()+
   scale_color_viridis_c()+
   facet_wrap(~seasonUnique)
+
+# not facetted by season:
+forplots %>%
+  ggplot(aes(x = PC1, y = PC2, col = daysTracked))+
+  geom_point(size = 2)+
+  geom_vline(aes(xintercept = 0))+
+  geom_hline(aes(yintercept = 0))+
+  theme_bw()+
+  scale_color_viridis_c()
+
+# colored and ellipsed by season:
+# XXX
+
+# Automatic plotting: top and bottom 5 for each PC
+pc1_top5 <- forplots %>%
+  select(-c(type, n, degree, degreeRelative, strength, strengthRelative, sbd, pageRank, pageRankRelative, evenness, dataset)) %>%
+  distinct() %>%
+  arrange(desc(PC1)) %>%
+  head(5) %>% select(Nili_id, seasonUnique, PC1, PC2, daysTracked)
+
+pc1_bottom5 <- forplots %>%
+  select(-c(type, n, degree, degreeRelative, strength, strengthRelative, sbd, pageRank, pageRankRelative, evenness, dataset)) %>%
+  distinct() %>%
+  arrange(PC1) %>%
+  head(5) %>% select(Nili_id, seasonUnique, PC1, PC2, daysTracked)
+
+pc2_top5 <- forplots %>%
[…]

### 2024-01-10 scripts/04.2_bayesianAnalysis.R
First try with Bayesian

@@ -1,0 +1,53 @@
+# Testing out Bayesian analysis in parallel
+library(tidyverse)
+library(rstan)
+library(cmdstanr)
+library(rstan)
+library(brms)
+library(posterior)
+library(bayesplot)
+library(tidybayes)
+library(MCMCvis)
+theme_set(theme_tidybayes())
+load("data/derived/forModeling.Rda")
+forModeling$type <- factor(forModeling$type)
+
+modelingSubset <- forModeling %>%
+  sample_n(400)
+
+# Degree ------------------------------------------------------------------
+# Define a named list of data to pass to stan
+data_degree <- list(
+  N = nrow(modelingSubset), # the number of data points
+  PC1 = modelingSubset$PC1, # independent variable
+  #type = as.integer(modelingSubset$type), # converting the factor to integer
+  degree = modelingSubset$degree, # dependent variable  
+  popsize = modelingSubset$n # pop size per season (part of dependent variable)
+)
+
+mod_degree <- stan(
+  file = "scripts/degree.stan", # name of the Stan program file
+  data = data_degree, # named list of data (defined above)
+  chains = 4, # number of Markov chains to run
+  warmup = 1000, # number of warmup iterations per chain
+  iter = 4000, # total number of iterations per chain
+  cores = 4 # number of cores (should use 1 per chain)
+)
+
+launch_shinystan(mod_degree)
+
+data_degreeRel <- list(
+  N = nrow(modelingSubset),
+  PC1 = modelingSubset$PC2,
+  type = as.integer(modelingSubset$type),
+  degreeRel = modelingSubset$degreeRelative
+)
+
+mod_degreeRel <- stan(
+  file = "scripts/degreeRel.stan",
+  data = data_degreeRel,
+  chains = 4,
+  warmup = 1000,
+  iter = 4000,
+  cores = 4
+)

### 2024-01-10 scripts/muxLib DF.R

@@ -1,0 +1,1285 @@
+library(Matrix)
+library(RSpectra)
+library(igraph) 
+
+####################################################
+# MuxNetLib: Library for Multilayer Network Analysis in muxViz
+#
+# Version: 0.1
+# Last update: Dec 2017
+# Authors: Manlio De Domenico
+#
+# History:
+#
+# Mar 2017: From Matlab to R!
+# May 2014: First release, including part of muxNet
+####################################################
+
+#Good refs, to check in general:
+#https://cran.r-project.org/doc/contrib/Hiebeler-matlabR.pdf
+#http://mathesaurus.sourceforge.net/octave-r.html
+#https://cran.r-project.org/web/packages/Matrix/Matrix.pdf
+
+
+###################################################################
+## MATLAB-like function
+###################################################################
+
+inv <- function(A){
+  return( solve(A) )
+}
+
+kron <- function(A, B){
+  #return the kronecker product between two matrices
+  return( Matrix::kronecker(A, B) )
+}
+
+speye <- function(n){
+  #return the identity matrix in sparse format
+  return( Matrix::Diagonal(n,1) )
+}
+
+zeros <- function(n,m){
+  #return a matrix full of zeros
+  #return( Matrix::Matrix(Matrix::kronecker( cbind(rep(0,n)), rbind(rep(1,m)), sparse=T), sparse=T) )
+  return(Matrix::Matrix(0, n, m, sparse=T))
+}
+
+ones <- function(n,m){
+  #return a matrix full of ones
+  #return( Matrix::Matrix(Matrix::kronecker( cbind(rep(1,n)), rbind(rep(1,m)) )) )
+  return(Matrix::Matrix(1, n, m, sparse=T))
+}
+
+spcan <- function(n, i, j){
+  #return a canonical matrix
+  A <- zeros(n,n)
+  A[i,j] <- 1
+  return(A)
+}
+
+rand <- function(n,m){
+  #return a matrix full of random numbers (uniformly distributed in [0,1])
+  return( Matrix::Matrix( runif(n*m), n, m, sparse=T ) )
+}
+
+blkdiag <- function(...){
+  #return a list of matrices in block diagonal form
+  return( Matrix::bdiag(...) )
+}
+
+reshapeR <- function(A, n, m){
+  #cannot call this function just "reshape", because it's an R builtin command
[…]

### 2024-01-10 scripts/temp_multilayer_test.R

@@ -1,0 +1,115 @@
+# Setup for temporal multilayer network analysis
+library(tidyverse)
+library(lubridate)
+library(tnet)
+library(sna)
+
+load("data/derived/flightSeasons_mode10_edges.Rda")
+fs <- flightSeasons_mode10_edges
+testseason <- fs[[5]] # just doing this for one arbitrary season
+
+# Make a binary edgelist
+bin <- testseason %>%
+  mutate(day = lubridate::date(minTimestamp)) %>%
+  mutate(hour = floor_date(minTimestamp, unit = "hours")) %>%
+  select(ID1, ID2, day, hour) %>%
+  distinct() %>%
+  mutate(weight = 1)
+
+# Plot associations
+bin %>%
+  group_by(hour) %>% 
+  summarize(n = n()) %>%
+  ggplot(aes(x = hour, y = n))+
+  geom_point()+
+  geom_path(alpha = 0.5)+
+  geom_smooth(method = "lm") # generally more in-flight interactions over the course of the season, which makes sense because it's warming up.
+
+opp_edges <- bin
+names(opp_edges)[1:2] <- c("ID2", "ID1")
+
+all_bin <- bind_rows(bin, opp_edges) # attach 
+
+# Add layer numbers (days) to make the next step easier
+all_bin <- all_bin %>%
+  select(-day) %>%
+  arrange(hour, ID1, ID2) %>%
+  mutate(layer = as.integer(as.factor(hour))) %>%
+  select(-hour) %>%
+  rename("layer1" = "layer") %>%
+  mutate(layer2 = layer1) %>%
+  relocate(layer1, .after = "ID1") %>%
+  relocate(layer2, .after = "ID2")
+
+self_edges <- expand.grid(as.character(unique(all_bin$ID1)), 1:(max(all_bin$layer1)-1)) # create all possible combinations of indiv ID and layer
+colnames(self_edges) <- c("ID1", "layer1")
+
+self_edges <- self_edges %>%
+  mutate(ID2 = ID1, layer2 = layer1 + 1, weight = 1)
+self_edges2 <- self_edges %>%
+  mutate(layer1 = layer1 + 1, layer2 = layer2-1)
+self_edges_all <- bind_rows(self_edges, self_edges2)
+
+# Combine with symmetrical edgelist
+all_edges <- bind_rows(all_bin, self_edges)
+
[…]

### 2024-01-10 scripts/test.R

@@ -1,0 +1,25 @@
+# Frequentist models parallel to Bayesian models
+# 1. degree ~ PC1
+# 2. degree ~ situ + PC1
+# 3. degree ~ situ + PC1 + PC2
+# 4. degree ~ situ + PC1 + PC2 + age_group
+
+library(tidyverse)
+library(lme4)
+load("data/derived/forModeling.Rda")
+
+# 1.
+mod1 <- lm(degree_scl ~ PC1, data = forModeling)
+summary(mod1)
+
+# 2.
+mod2 <- lm(degree_scl ~ situ + PC1, data = forModeling)
+summary(mod2)
+
+# 3.
+mod3 <- lm(degree_scl ~ situ + PC1 + PC2, data = forModeling)
+summary(mod3)
+
+# 4.
+mod4 <- lm(degree_scl ~ situ + PC1 + PC2 + age_group, data = forModeling)
+summary(mod4)

### 2024-01-10 manuscript.qmd
setting up manuscript

---
title: "Manuscript draft"
format: html
editor: visual
author: Kaija Gahm
---

# Introduction

The social structure of animal groups has consequences for disease transmission, information sharing, and population dynamics. At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration

# Research Questions

# Methods

# Results

# Discussion

### 2024-01-10 manuscript.qmd
beginning to copy over the intro and add citations

@@
 author: Kaija Gahm
+bibliography: references.bib
+csl: the-american-naturalist.csl
 ---
@@
 The social structure of animal groups has consequences for disease transmission, information sharing, and population dynamics.
-At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration
+At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration (@sah2017, @silk2017), govern the flow of information about food resources \[REF\], and contribute to population persistence through Allee effects \[REF\].
+Individuals differ in their encounter rates, number of contacts, and centrality in the population social network \[REF\].
+For individual animals, which make up the social environment of the group.
+social centrality in different behavioral situations contributes to their likelihood of contracting disease, finding mates, or accessing resources \[REFS\].
+For example, in networks with high levels of degree heterogeneity, the most connected individuals may function as disease superspreaders (@lloyd-smith2005).
+For example, possums with higher betweenness and closeness in a social network had higher infection rates (@corner2003), and higher connectivity in lizards was associated with a greater parasite load (@godfrey2009).
+In general, individuals that are more central in the social network both have a higher infection risk and transmit disease more quickly than those that are more peripheral (@romano2016, @balasubramaniam2019).
+
+An individual's social centrality may be determined by many factors.
+Differences in social position have been attributed to individual phenotypic characteristics, such as age, sex, or personality \[REFS\].
+Of course, how an animal uses space is also an important driver of how many conspecifics it interacts with, and how frequently.
+The relationship between movement, space use, and social interaction patterns has been an area of active research in recent years.
+Webber et al -@webber2023 have characterized the variation in patterns of animal movement and space use as the "spatial phenotype," as part of a larger framework for understanding the relationships between social and spatial drivers of animal behavior.
+In this light, the "spatial phenotype" may be considered in addition to more conventional phenotypic measures to explain why some individuals are more central than others in their social networks.
+
+Additionally, sociality has many dimensions.
+Individuals interact in different social and behavioral contexts, defined by the types of activities being performed (e.g.
+feeding, playing, sleeping) and/or the intent and consequences of the interaction itself (e.g.
+agonistic vs.
+affiliative interactions) \[REFS\].
+The overall social environment experienced by an individual is made up of its interactions in all of these different contexts @silk2013.
+Individuals differ in their centrality in these different contexts @sharma2023.
+Different types of interactions are also localized differently in space.
+For example, social foraging occurs, by definition, while individuals move concurrently across a landscape, so which individuals forage together and co-foraging rates are likely to be directly tied to movement behavior and spatial phenotypes of the individuals involved.
+By contrast, mating or fighting interactions might be more localized to specific places where these behaviors occur, such as nest or den sites, leks, territorial boundaries, or scarce food resources.
+An informed understanding of the relationship between the spatial and social phenotypes, then, should allow for the possibility that the spatial phenotype relates differently to an individual's position in each situation-specific social network.
+
+Eurasian griffon vultures (*Gyps fulvus*) provide a study system well suited to ascertaining the relationship between individuals' spatial phenotypes and their social centrality.
+
+-   obligate scavengers, large soaring fliers, lots of movement
+
+-   Unusual GPS tracking coverage of the population–high proportion of individuals tracked, so we can get a population-level sense of movement behavior and social interaction patterns
+
+-   Three clear social situations in which they interact, each with a different significance/relevance.
+
+    -   Communal roosts: many individuals together at once, variability in size/density of roosts.
+Used for information exchange \[REFS\] and also important for disease transmission \[or not–cf Elvira Nili\].
+Breeding/nesting also occurs at/near roost sites, on cliffs.
+
+    -   Flight: interactions in flight are important for information transfer @harel2017, finding locations of thermal uplift @sassi2023 and feeding sites @spiegel2013, @cortés-avizanda2014
+
+    -   Feeding: Communal feeding at carcasses.
+Also sometimes standing around the carcass site.
+Sometimes associate for a while before beginning to feed, e.g.
+waiting for critical mass to be reached (Orr Spiegel, pers.
+comm.) Potentially also important for disease transmission \[but not, cf.
+ref Elvira Nili\].
 

### 2024-01-10 manuscript.qmd
finished transferring intro draft

@@
 
+While some griffon vultures make long-range forays \[REFS\], there is a mostly resident population in a relatively small area in southern Israel \[REF\].
+The population is small and restricted enough that individuals' home ranges are highly overlapping, with most individuals using a large fraction of available space in the region.
+This makes habitat features likely to affect most individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement @harel2016.
+GPS fixes are taken once every 10 minutes, frequently enough to measure flight movements with good accuracy, and most individuals are tracked for several months to several years.
+Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population @sharma2023.\
+\
+\[General introduction to the importance of time scale in social network analysis\]
+
+-   @blonder2012, @cantor2012, @hobson2013, @rasmussen2013, @holme2015, @farine2018, @fisher2021, @pasquaretta2021
+
+Social situations occur over different time scales, and this is an important consideration when studying the causes and consequences of their structure.
+In defining co-roosting, co-flight, and co-feeding interactions based on proximity in each of these behavioral situations, we have created social networks in which edges already conceptually represent interactions over different amounts of time.
+Because individuals have one roosting location per night, edge weights in the co-roosting network are measured in units of days, while pairs of individuals can interact many times in flight or at a carcass on a given day.\
+\
+The biological processes driving the execution of these behaviors and the formation of these networks are likewise happening over different time scales.
+An individual must decide once per day which roost to go to @harel2016.
+But flight is a continuous process driven by many decisions resulting from individual and social information @sassi2023, and feeding events are concentrated in time on a given day.
+Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every \[NUMBER\] days and persist for hours to a few days (Orr Spiegel, pers.
+comm.).
+Individuals may return to the same carcass on successive days.
+While at a carcass, individuals may fight and interact many times.
+
 # Research Questions

### 2024-01-10 scripts/02.0_calcSocial.R
added network density calculation

@@ -66,6 +66,20 @@ flightSeasons_mode10_g <- map(flightSeasons_mode10, ~vultureUtils::makeGraph(mod
 feedingSeasons_mode10_g <- map(feedingSeasons_mode10,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
 roostSeasons_g <- map(roostSeasons,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
 
+# Describing the networks -------------------------------------------------
+# Network-level measures
+fldens <- map(flightSeasons_mode10_g, igraph::edge_density) %>% unlist()
+fedens <- map(feedingSeasons_mode10_g, igraph::edge_density) %>% unlist()
+rodens <- map(roostSeasons_g, igraph::edge_density) %>% unlist()
+dens <- data.frame(season = factor(seasonNames, levels = seasonNames), flight = fldens, feeding = fedens, roosting = rodens) %>%
+  pivot_longer(cols = c("flight", "feeding", "roosting"), names_to = "type", values_to = "density")
+
+dens %>%
+  ggplot(aes(x = season, y = density, col = type, group = type))+
+  geom_point()+
+  geom_line()
+
+# Extracting individual-level network measures ----------------------------
 networkMetrics <- map2_dfr(flightSeasons_mode10_g, seasonNames, ~{
   df <- data.frame(degree = igraph::degree(.x),
                    strength = igraph::strength(.x),

### 2024-01-10 manuscript.qmd
Added descriptive results and a few figures to manuscript

@@
 editor: visual
-author: Kaija Gahm
+date: last-modified
+author: Kaija Gahm, [others]
 bibliography: references.bib
@@
 # Introduction
+
+```{r}
+#| include: false
+library(tidyverse)
+```
 
@@
 
+Q1: What is the relationship between spatial phenotype and social interactions?
+
+-   H1A: Individuals that move more will have more, and stronger, encounters with other unique individuals.
+
+    -   P1A: Vultures that move more will encounter more others (higher degree centrality in the social network) and have more interactions (higher strength in the social network) than vultures who move less.
+
+-   H1B: Individuals that are more exploratory will have more but weaker encounters with other unique individuals.
+
+    -   P1B: Exploratory vultures will encounter more unique individuals (higher degree centrality), but will have fewer interactions (lower strength) compared to less exploratory vultures.
+
+Q2: How is the relationship between spatial phenotype and social interactions impacted by social situation and season?
+
+-   H2A: Social situations in which animals move more will have a greater impact on social interactions
+
+    -   P2A: Social centrality in co-flight networks will be more strongly associated with spatial phenotype than in co-roosting and co-feeding behavioral situations.
+
+-   H2B: Seasons in which animals move more will have a stronger impact on social interactions
+
+    -   P2B: The relationship between movement and social position will be stronger in the non-breeding seasons (summer and fall) than during the breeding season.
+
+Q3: How does the time scale of movement predict social interactions?
+
+-   H3A: The relationship between movement and social centrality will be strongest within the time scales at which change occurs in the social network.
+
+    -   P3A: Co-flight networks will change hourly to daily, co-feeding networks will change every few days, and co-roosting networks will change over longer periods of time (weeks to months).
+
+    -   P3B: Social position in each of the social situations will be best predicted by individuals' movement patterns aggregated over the corresponding time scale.
+
 # Methods
+
+## Study system
+
+To answer these questions, we drew on data from a free-flying population of Eurasian griffon vultures (Gyps fulvus) fitted with GPS tags.
+\[tag specs\].
+Over the period 2020-09-01 to 2023-05-15, the dataset contained \[number\] of tagged vultures.
+Tags recorded GPS fixes, as well as speed and heading readings, every 10 minutes when battery conditions allowed.
+The GPS data are stored on Movebank @kays2022 and were loaded into R using the `move` package @kranstauber2023.
+
+## Tagging and aging
+
+\[Tagging methods\] For each vulture, age was determined using \[methods\]
+
+## Data cleaning
+
+We subsetted the dataset to exclude periods of time that vultures spent in capture cages, in the wildlife hospital following poisonings or injuries, or in captivity prior to wild release (in the case of captive-born birds).
+We cleaned the remaining data to remove invalid GPS fixes \[more details\].
+Then we split data from each into three seasons.
+The fall, or "pre-breeding," season includes dates from 15 September through 14 December; the breeding season lasts from 15 December through 14 May of the following year; and the summer lasts from 15 May through 14 September.
+These season boundaries were chosen based on biological knowledge of the vultures' breeding timelines (see more details in the Study System section).
+Therefore, the analysis contained a total of 8 seasons, from the fall of 2020 through the 2023 breeding season (Table X).
+All of these data was reserved for use in proximity-based social network analysis.
+Since the GPS tags give few to no fixes overnight, and since reception is often very poor on the cliffs where vultures roost, we combined information from individuals' last locations of the evening and first locations of the morning to assign a roost site for each individual on each night \[more details\].
+Full details about the data cleaning and roost assignment procedure can be found in the supplementary material.
+
+To characterize individual movement behavior and spatial phenotype, we cleaned the data further.
+We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package suncalc @thieurmel2022.
+We restricted movement calculations to a set of focal individuals with activity regions centered in the Negev region of Israel, the site of most of our tagging effort.
+For each vulture, we also removed days with too few GPS fixes.
+Fixes were ideally collected every 10 minutes, but sometimes they were less frequent due to bad cellular reception in the cliff areas where vultures often sit or due to low battery charge.
+For each individual, only days with 10 or more GPS fixes per day were retained, unless the battery charge never dipped below 50% during that day.
+In those cases, the low number of points was attributed to the vultures spending most of their time sitting on a cliff without much movement, rather than to missed points due to low battery.
+Since these sedentary days represent an important component of vulture movement behavior, we retained this information (Orr Spiegel, pers.
+obs.).
+
+\[Geography\]
+
+For each season, we calculated movement measures only for individuals tracked for at least 30 days (after removing low-point days as described above).
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package @hollister2022, at a zoom level of \[level–more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
+Finally, since some tags were occasionally set to a much higher fix rate (e.g.
+every 1 minute) for short periods of time, we downsampled the data to a 10 minute fix rate to make sure that all individuals had comparable fix rates.
+
+## Movement calculations
+
+For each focal individual, we characterized a pattern-based spatial phenotype (Webber et al.
+2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
+Using the `adehabitatHR` package @calenge2023, we calculated home range as the 95% kernel utilization distribution (KUD) of a vulture's GPS fixes in a season; core area as the 50% KUD, and core area use intensity as the ratio of the 50% to the 95% KUD.
+
+To quantify roosting behavior, we assigned each individual's night roost site to a pre-defined roost site polygon determined from clustering of \[years\] of vulture roost locations \[more details?\].
+Then, for each vulture in each season, we calculated the number of unique roosts used, the proportion of nights switching roosts (versus returning to the same roost as the previous night), and the diversity of roosts used (Shannon diversity index).
+
+We identified in-flight fixes (ground speed \>= 5 m/s) and calculated measures of daily flight behavior, including maximum and median daily flight altitude and daily flight time (the sum of time intervals between adjacent in-flight fixes).
+Finally, we used all fixes to calculate daily distance traveled (the sum of all distances between successive fixes); daily maximum displacement (the straight-line distance between the beginning of the daily movement track and the farthest point reached on that day); and movement tortuosity (cumulative distance to the point of maximum displacement divided by the straight-line distance to the point of maximum displacement @benhamou2004).
+We took seasonal means of each of these individual daily measures to get a single value per individual vulture, per season.
+
+## Summarizing movement measures
+
+To gain a complete picture of individual vultures' spatial phenotype, we summarized the above-described movement measures (centered and scaled) in a principal component analysis.
+The first two principal components together explained 65% of the variance in spatial phenotype (Table X).
+
+PC1 explained 47.6% of the variance in spatial phenotype.
+All movement variables except tortuosity loaded positively on PC1, with the strongest contributors being mean daily maximum displacement, roost diversity, and proportion of nights roost-switching (Figure X).
+Overall, we can understand PC1 as an aggregate measure of individual movement.
+Vultures with high PC1 values cover more area, use more different roosts, travel farther, spend more time flying, and have straighter trajectories.
+
+PC2 explained 17.4% of the variance in spatial phenotype.
+The space use variables loaded positively on this principal component, while flight variables (both altitude measures, daily distance traveled, and daily flight time) contributed strongly negatively (Figure X).
+We can understand PC2 as a rough measure of exploratory behavior.
+Vultures with high PC2 values use a larger spatial area, while those with low PC2 values fly higher and spend more time flying (in a more concentrated area).
+
+## Construction of proximity-based social networks
+
+Using GPS fixes from the tagged vultures, we created proximity-based social networks in three different social situations, following @sharma2023: co-flight, co-feeding (diurnal ground interactions) and co-roosting (nocturnal ground interactions).
+Networks were created separately for each of the eight seasons.
+For each social situation in each season, we created a weighted, undirected network.
+Because not all individuals were tagged for exactly the same period of time, and to account for missing data due to low tag battery or poor reception, we calculated the simple ratio index (SRI) to quantify the strength of the association between each pair of vultures in each social situation.
+We calculated SRI by dividing the number of 10-minute time periods during which two vultures were observed interacting by the number of time periods during which both vultures had a recorded GPS fix (whether or not they were interacting during that period) @ginsberg1992.
+Thus, two vultures who never interacted in a given social situation would have an SRI of 0, while two individuals who were interacting in every time period during which a GPS fix was collected for both of them would have an SRI of 1.
+
+Non-focal vultures (e.g.
+those that did not always have enough points per day, or those that were not tracked for at least 30 days per season) were included in our calculations of social interactions.
+Even if an individual did not have enough data to calculate meaningful movement measures, it could still contribute to a focal individual's position in the social network.
+Defining edge weights with SRI ensured that missing data for these non-focal birds did not bias edge weight estimates for their social relationships.
+
+### Co-flight
+
+Vultures were considered to be flying if their ground speed (collected by a sensor on the GPS tags) was \> 5 m/s @spiegel2013 .
+Because vultures can see each other over long distances @pennycuick1972, @spiegel2013a , and in accordance with @sharma2023, we defined a co-flight interaction as two individuals flying within 1000m of each other for at least two consecutive 10-minute time periods.
+
+Co-feeding
+
+Vultures that were not flying (ground speed \< 5 m/s, @spiegel2013) who came within 50m of each other for at least two consecutive 10-minute time periods, were considered to be having a ground interaction @sharma2023.
+Any ground interactions that occurred at night or within the boundaries of a known roost site were excluded.
+Therefore, the remaining diurnal ground interactions should represent vultures' interactions while feeding or while present at feeding sites (Orr Spiegel and Nili Anglister, pers.
+comm.).
+Co-roosting Roost sites were assigned using \[methods\].
+Unlike flight and diurnal ground GPS fixes, vultures were only assigned one roosting location per night.
+Two vultures that roosted within the boundaries of the same roost site polygon on a given night were considered to be co-roosting.
+
+## Calculation of social network measures
+
+\[Explain the reasons for choosing the social network measures in the earlier hypotheses/predictions/introduction section\]
+
+We used the R package `igraph` @csardi2005 to calculate each individual's degree and strength for each social situation in each season.
+Degree is defined as the number of other individuals that a vulture is connected to (i.e.
+ever interacted with).
+Strength measures how much an individual interacts, and is calculated as the sum of the weights of all edges that the focal vulture is involved in.
+
+## Model fitting and selection
+
+We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the lme4 and lmerTest packages @bates2015, @kuznetsova2017.
+We fit separate models for each response variable (degree and strength).
+The predictors included were movement (PC1), exploration (PC2), situation (feeding, flight, roosting), age group (juvenile/subadult, adult), and season (breeding, summer, fall).
+To account for the different numbers of individuals in the network in each unique season, we included a random effect for the interaction of season and year, e.g.
+"2022_fall" (8 levels).
+To account for consistent individual differences in network position or behavior, and because individuals varied in how many seasons they were included in, we also included an individual-level random effect.
+
+Degree values were scaled and standardized before modeling.
+Because the distribution of strength values was extremely right-skewed, the strength values were incremented by one-half the minimum non-zero strength value (0.018) and then log-transformed.
+
+For each response variable, we began by fitting a model without any interaction terms to check for multicollinearity between the predictors and check model assumptions, using the performance and DHARMa packages @lüdecke2021, @hartig2022.
+Next, we fitted a model containing two three-way interaction terms (between season, situation, and each of the spatial phenotype variables), all of the included two-way interaction terms, age as a main effect, and the two random effects (individual ID; season by year).
+Examining the model summaries and p values produced by lmerTest, we then removed non-significant higher-order interaction terms, one by one, stopping when all remaining interaction terms were significant.
+All main and random effects were retained in the final model.
+
+For each model, we calculated marginal effects using the emmeans package @lenth2023 and created visualizations using ggeffects and ggplot2 @wickham2016, @lüdecke2018.
 
@@
 
+```{r}
+#| echo: false
+#| output: true
+#| label: fig-nvultures
+#| fig-cap: "Number of vultures in different social situations over the study period.
+Lines for the 'feeding' situation have been jittered upward because they overlap completely with the 'flight' situation.
+Dashed lines and open circles represent the total number of vultures tracked in the population.
+Solid lines and filled circles represent the number of focal vultures."
+
+load("data/derived/forModeling.Rda")
+load("data/derived/cc.Rda")
+load("data/derived/seasonNames_orig.Rda")
+seasonNames <- seasonNames_orig[-1]
+focals <- forModeling %>%
+  filter(!is.na(PC1)) %>%
+  group_by(season, year, type) %>%
+  summarize(total = n[1],
+            focal = length(unique(Nili_id)), .groups = "drop") %>%
+  arrange(year, season, type) %>%
+  mutate(seasonUnique = paste(year, season, sep = "_")) %>%
+  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding"))) %>%
+  mutate(seasonU = str_replace(seasonUnique, "breeding", "B")) %>%
+  mutate(seasonU = str_replace(seasonU, "fall", "F")) %>%
+  mutate(seasonU = str_replace(seasonU, "summer", "S")) %>%
+  mutate(seasonU = str_remove(seasonU, "_")) %>%
+  mutate(seasonU = str_remove(seasonU, "(?<=^)20")) %>%
+  mutate(seasonU = factor(seasonU, levels = c("20F", "21B", "21S", "21F", "22B", "22S", "22F", "23B")))  %>%
+  pivot_longer(cols = c(total, focal), names_to = "Vultures", values_to = "n") %>%
+  rename("Situation" = "type") %>%
+  ungroup()
+
+focals %>%
+  mutate(n = case_when(Situation == "feeding" ~ n + 1, # just adding a tiny bit of jitter so as to more easily see the different lines 
+                       TRUE ~ n)) %>%
+  ggplot(aes(x = seasonU, y = n, col = Situation))+
+  geom_line(aes(group = interaction(Situation, Vultures), linetype = Vultures), linewidth = 1)+
+  geom_point(size = 4, aes(pch = Vultures))+
+  scale_shape_manual(values = c(19, 1))+
+  theme_classic()+
+  ylab("# vultures")+
+  xlab("Season")+
+  scale_color_manual(values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
+  theme(text = element_text(size = 16))
+
+wide <- focals %>%
+  pivot_wider(id_cols = c("season", "year", "Situation", "seasonUnique", "seasonU"), names_from = "Vultures", values_from = "n")
+```
+
+Networks varied in the total and focal number of individuals, from a minimum of `r min(wide$total)` vultures (`r wide$focal[wide$total == min(wide$total)][1]` focal) in the `r wide$year[wide$total == min(wide$total)][1]` `r wide$season[wide$total == min(wide$total)][1]` season to a maximum of `r max(wide$total)` vultures (`r wide$focal[wide$total == max(wide$total)]` focal) in the `r wide$year[wide$total == max(wide$total)]` `r wide$season[wide$total == max(wide$total)]` season in the co-roosting network (@fig-nvultures) .
+The number of vultures tagged generally increased over time as more individuals were fitted with tags over the course of the study period.
+
+```{r}
+#| echo: false
+#| output: true
+#| label: fig-density
+#| fig-cap: "Density of the proximity-based social networks created for each season and social situation."
+
+load("data/derived/flightSeasons_mode10.Rda")
+load("data/derived/feedingSeasons_mode10.Rda")
+load("data/derived/roostSeasons.Rda")
+
+flightSeasons_mode10_g <- map(flightSeasons_mode10, ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
+feedingSeasons_mode10_g <- map(feedingSeasons_mode10,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
+roostSeasons_g <- map(roostSeasons,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
+
+# Describing the networks -------------------------------------------------
+# Network-level measures
+fldens <- map(flightSeasons_mode10_g, igraph::edge_density) %>% unlist()
+fedens <- map(feedingSeasons_mode10_g, igraph::edge_density) %>% unlist()
+rodens <- map(roostSeasons_g, igraph::edge_density) %>% unlist()
+dens <- data.frame(season = factor(seasonNames, levels = seasonNames), flight = fldens, feeding = fedens, roosting = rodens) %>%
+  pivot_longer(cols = c("flight", "feeding", "roosting"), names_to = "Situation", values_to = "Density") %>%
+  mutate(seasonU = str_replace(season, "breeding", "B")) %>%
+  mutate(seasonU = str_replace(seasonU, "fall", "F")) %>%
+  mutate(seasonU = str_replace(seasonU, "summer", "S")) %>%
+  mutate(seasonU = str_remove(seasonU, "_")) %>%
+  mutate(seasonU = str_remove(seasonU, "(?<=^)20")) %>%
+  mutate(seasonU = factor(seasonU, levels = c("20F", "21B", "21S", "21F", "22B", "22S", "22F", "23B")))
+
+dens %>%
+  ggplot(aes(x = seasonU, y = Density, col = Situation))+
+  geom_line(aes(group = Situation), linewidth = 1)+
+  geom_point(size = 4)+
+  theme_classic()+
+  ylab("Network density")+
+  xlab("Season")+
+  scale_color_manual(values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
+  theme(text = element_text(size = 16))
+```
+
+Seasonal social networks varied in density from `r min(dens$Density)` (`r stringr::str_extract(dens$season[dens$Density == min(dens$Density)], "[a-z]+")` `r str_extract(dens$season[dens$Density == min(dens$Density)], "[0-9]+")` ) to `r max(dens$Density)` (`r stringr::str_extract(dens$season[dens$Density == max(dens$Density)], "[a-z]+")` `r str_extract(dens$season[dens$Density == max(dens$Density)], "[0-9]+")`) (@fig-density).
+Co-roosting networks were generally denser than co-feeding or co-flight networks.
+
+Two-way interactions between situation and both movement (PC1) and exploration (PC2) were retained in the degree model.
+No interaction effects including season were retained, and the main effect of season was not significant (Table XX).
+
 # Discussion

### 2024-01-10 manuscript.qmd
switched to self-contained html for the manuscript

@@
 title: "Manuscript draft"
-format: html
+format:
+  html:
+    embed-resources: true
 editor: visual

### 2024-01-10 00.0_dataPrep.qmd
starting to transfer code over to qmd files

--- 2023-06-15 00_dataPrepReport.Rmd
@@
 ---
-title: "Data preparation report"
-author: "Kaija Gahm"
-date: "2023-05-22"
-output: html_document
-editor_options: 
-  markdown: 
-    wrap: 72
+title: "Data Prep"
+format:
+  html:
+    embed-resources: true
+editor: visual
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = FALSE, message = FALSE, cache = FALSE)
-```
+## Goals
+
+Prepare data for the movement/social analysis
+
+## Setup
 
@@
 library(readxl)
+library(elevatr)
 ```
 
-### Step 1: Download raw data from Movebank
+## Download data from Movebank
 
-```{r download-movebank-data}
-# base::load("movebankCredentials/pw.Rda")
-# MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
-# rm(pw)
+### Authenticate to Movebank
+
+```{r}
+base::load("movebankCredentials/pw.Rda")
+MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
+rm(pw)
+```
+
+### INPA data
+
+Data collected by Ohad Hatzofe and team with GPS loggers deployed for their study.
+
+Inputs:
+
+-   `MB.LoginObject`
+
+Outputs:
+
+-   `inpadf_touse`
+
+-   `data/derived/inpadf_touse.feather`
+
+```{r download-inpa}
+#| eval: false
+# inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
+# inpadf <- methods::as(inpa, "data.frame")
+# inpadf_touse <- inpadf %>%
+#   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
+#          year = as.numeric(lubridate::year(timestamp))) %>%
+#   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
+# write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")
+
+length(unique(inpadf_touse$trackId)) # 77
+```
+
+### Ornitela data
+
+Inputs:
+
+-   `MB.LoginObject`
+
+Outputs:
+
+-   `dat`
+
+-   `data/derived/dat.feather`
+
+```{r download-ornitela}
+#| eval: false
+# Ornitela: download data from movebank (just a subset of the times for now)
 minDate <- "2020-09-01 00:00"
 maxDate <- "2023-05-15 11:59"
-dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-write_feather(dat, "data/dat.feather")
-dat <- read_feather("data/dat.feather")
+# dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+# write_feather(ornitela, "data/derived/dat.feather")
+
+length(unique(dat$trackId)) # 127
 ```
 
-Downloaded data from Movebank from `r minDate` to `r maxDate`. The
-initial data has **`r nrow(dat)`** rows and `r ncol(dat)` columns. There
-are **`r length(unique(dat$trackId))`** unique individual ID's in the
-data at this point.
+## Join INPA and Ornitela data
 
-### Step 2: Fix vulture IDs
+*Inputs:*
 
-```{r fix-trackId-Nili_id}
-## fix trackId. Marta told me that E03w was sometimes coded as E03, so I fixed that right up front. 
-dat2 <- dat %>%
-  mutate(trackId = as.character(trackId),
-         trackId = case_when(trackId == "E03" ~ "E03w",
-                             TRUE ~ trackId))
-# number of unique individuals
-length(unique(dat2$trackId)) # 126--makes sense that we lost one because one individual had a typo.
+-   `data/derived/inpadf_touse.feather`
 
-# Read in the who's who table and prepare to join the Nili_ids
-ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
-  dplyr::select(Nili_id, Movebank_id) %>%
-  distinct()
+-   `data/derived/dat.feather`
 
-# Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
-ww %>% filter(Movebank_id %in% dat2$trackId) %>% 
-  dplyr::select(Nili_id, Movebank_id) %>% 
-  distinct() %>% group_by(Nili_id) %>% 
-  summarize(n = length(unique(Movebank_id))) %>% 
-  arrange(desc(n))
-ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili_id's
+*Outputs:*
 
-all(dat2$trackId %in% ww$Movebank_id) # true
+-   `dat3`
 
-dat2 <- left_join(dat2, ww, by = c("trackId" = "Movebank_id"))
-dim(dat2) # good, same number of rows.
-length(unique(dat2$Nili_id))
+```{r join-inpa-ornitela}
+#| eval: false
+
+inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
+dat <- read_feather("data/derived/dat.feather")
+# # Will there be a problem joining these?
+# names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
+# names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
+# 
+# # Add the dataset names so we can keep track of where the data comes from
+# inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
+#   mutate(dataset = "inpa")
+# dat <- dat %>%
+#   mutate(dataset = "ornitela")
+# 
+# # Join the two datasets
+# joined <- bind_rows(inpadf_tojoin, dat)
+# dim(joined)
+# 
+# ## fix trackId
+# dat2 <- joined %>%
+#   mutate(trackId = as.character(trackId),
+#          trackId = case_when(trackId == "E03" ~ "E03w",
+#                              TRUE ~ trackId))
+# # number of unique individuals
+# length(unique(dat2$trackId)) # 206
+# 
+# # Add the Nili_id
+# ww <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+#   dplyr::select(Nili_id, Movebank_id) %>%
+#   distinct()
+# 
+# # Are there any other individuals who have two different associated trackId's?
+# ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
+# ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
+# 
+# all(dat2$trackId %in% ww$Movebank_id) # false.
+# 
+# # Okay, let's investigate the ones that aren't included
+# dat2 %>%
+#   filter(!(trackId %in% ww$Movebank_id)) %>%
+#   pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
+# 
+# dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
+# ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
+# 
+# sort(unique(ww$mb_id))
+# unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+# dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
+# dat2$trackId[dat2$trackId == "e86white"] <- "e86"
+# unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+# # Investigate the remaining ones:
+# dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
+# # I guess for now let's just make their trackId's into Nili_id's
+# 
+# dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
+#                   by = c("trackId" = "mb_id"))
+# nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
+# length(unique(dat3$Nili_id))
+# # For any that didn't get Nili_ids assigned, make it the trackId
+# dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
+# dat3 <- dat3 %>%
+#   mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
+#                              TRUE ~ Nili_id))
+# unique(dat3$Nili_id)
 ```
 
-Next, we join in the Who's Who table in order to fix the Movebank
-trackId's and associate them with individual vultures (as defined by the
-"Nili_id" column). After joining in the Nili_ids, we now have a data
-frame with **`r nrow(dat2)`** rows and `r ncol(dat2)` columns. There are
-now **`r length(unique(dat2$Nili_id))`** unique individual vultures
-identified--the change from **`r length(unique(dat$trackId))`** is
-because there were a couple cases of different trackId's being assigned
-to the same Nili_id, such as "Y26" and "Y26b". No individuals are
-actually lost here.
+## Remove invalid periods
 
-### Step 3: Remove invalid periods (cages etc.)
+Removing known invalid periods, such as times when the vultures were in the hospital or the tags were known not to be attached to birds. These times are derived from a separate Excel sheet.
 
-Next, we remove periods of time when specific vultures were in cages or
-in the hospital, or when tags were not attached to the bird, etc. This
-is based on the "periods_to_remove" sheet in the who's who file.
+*Inputs:*
 
-```{r remove-periods}
+-   `data/raw/whoswho_vultures_20230315_new.xlsx`
+
+-   `dat3`
+
+*Outputs:*
+
+-   `datAnnot`
+
+```{r remove-hospital}
+#| eval: false
+
 ## annotate the data with periods to remove
-toRemove <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
+toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
 
@@
   filter(status == "valid")
+nrow(datAnnot)
+length(unique(datAnnot$Nili_id))
+nrow(dat3)
+nrow(datAnnot)
 ```
 
-```{r remove-cages}
-# Remove cages using Marta's data
+## Identify and remove capture cage periods
+
+This is Marta's code for identifying periods when the vultures were in the capture cages and removing days of data around those occasions.
+
+*Inputs:*
+
+-   `datAnnot`
+
+*Outputs:*
+
+-   `roosts`
+
+-   `data/derived/roosts.Rda`
+
+```{r}
+#| eval: false
+
 # Identify and remove capture dates using Marta's code --------------------
@@
 But we're just dealing with southern individuals, so that's okay.
-roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-save(roosts, file = "data/roosts.Rda")
-load("data/roosts.Rda")
+# roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+# save(roosts, file = "data/derived/roosts.Rda")
+load("data/derived/roosts.Rda")
 
@@
 # Load the information about the capture sites
-sites <- read.csv("data/capture_sites.csv")
+sites <- read.csv("data/raw/capture_sites.csv")
 
@@
 
-AllCarmelDates <- read.csv("data/all_captures_carmel_2010-2021.csv")
+AllCarmelDates <- read.csv("data/raw/all_captures_carmel_2010-2021.csv")
 AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")
@@
 (propChange <- (before-after)/before) # approx. 1% of data removed.
-length(unique(datAnnot$Nili_id))
-length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
+length(unique(datAnnot$Nili_id)) == length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
+
+datAnnot_noCaptures <- datAnnot_noCaptures %>%
+  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
 ```
-
-The resulting data has dropped from **`r nrow(dat2)`** to
-**`r nrow(datAnnot_noCaptures)`** rows, a
-**`r round((nrow(dat2)-nrow(datAnnot_noCaptures))*100/nrow(dat2), 3)`% decrease**.
-It has `r ncol(dat2)` columns and
-**`r length(unique(datAnnot_noCaptures$Nili_id))`** unique individuals.
-
-### Step 4: Attach birth year and sex info from who's who
-
-```{r join-age-sex}
-# Attach age and sex information
-as <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
-  dplyr::select(Nili_id, birth_year, sex) %>%
-  distinct()
-
-datAnnot2 <- datAnnot_noCaptures %>%
-  dplyr::select(-c("sex")) %>%
-  left_join(as, by = "Nili_id")
-nrow(datAnnot2) == nrow(datAnnot_noCaptures)
-length(unique(datAnnot2$Nili_id)) # 125
-```
-
-No change in the number of rows here: `r nrow(datAnnot2)` rows and
-`r ncol(datAnnot2)` columns.
-
-```{r}
-datAnnot <- datAnnot2 # a hack so the rest of the code will work
-```
-
-### Step 5: Clean the data using the cleanData function.
-
-This function includes several data cleaning steps: - Remove as invalid
-data points that have zeroes in `ground_speed`, `barometric_height`, and
-`external_temperature`. - Remove points that took to long to get a GPS
-fix (\>89) - Remove as invalid points with heading data \<0 or \>360
-degrees - Remove points with fewer than 3 satellites (because 3 are
-needed for triangulation) - Remove points that "jump" unrealistically
-across the map, as calculated by distance/time differences between
-consecutive points - Set to NA any altitude values that jump
-unrealistically between consecutive points - Remove points that "moved"
-farther than expected during the night - Remove **individuals** where
-less than 1/3 of their duration tracked fell inside the rough
-Israel-region geographic mak - For any individuals remaining, remove
-**points** that fell outside of the geographic mask (e.g. to remove
-long-range forays).
-
-```{r clean-data, echo = T, include = T}
-# Clean the data
-## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
-mask <- sf::st_read("data/CutOffRegion.kml")
-datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
-save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
-load("data/datAnnotCleaned.Rda")
-```
-
-After cleaning, the new data has **`r nrow(datAnnotCleaned)`** rows (a
-**`r round((nrow(datAnnot)-nrow(datAnnotCleaned))*100/nrow(datAnnot), 3)`%
-decrease**) and `r ncol(datAnnotCleaned)` columns.
-**`r length(unique(datAnnot$Nili_id))-length(unique(datAnnotCleaned$Nili_id))`**
-individuals were removed, for a remaining total of
-**`r length(unique(datAnnotCleaned$Nili_id))`** individuals.
-
-### Step 6: Fix times and split into seasons
-
-First, I changed the time zone to Israel time instead of UTC, since this
-will make a difference to where the date line falls. Then I
-re-calculated the `dateOnly` column based on the new date/times.
-
-Next, I split the data into fall/breeding/summer seasons based on
-Marta's cutoff dates: September 15th, December 15th, and May 15th. I
-fixed the vultures' ages to adjust for the breeding season and then
-classified them as adults or sub-adults/juveniles. Finally, I split the
-data into a list, with one item per season.
-
-```{r fix-dates}
-# Fix time zone so dates make sense ---------------------------------------
-## Overwrite the dateOnly column from the new times
-datAnnotCleaned <- datAnnotCleaned %>%
-  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
-         dateOnly = lubridate::date(timestampIsrael),
-         month = lubridate::month(dateOnly),
-         year = lubridate::year(dateOnly))
-```
-
-```{r split-seasons}
-# Split into Marta's 3 seasons --------------------------------------------
-s1 <- datAnnotCleaned %>%
-  mutate(start_breeding = ifelse(month %in% c(1:5),
-                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
-                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
-         start_summer = ifelse(month == 12,
-                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
-                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
-         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
-         start_breeding = as.Date(start_breeding),
-         start_summer = as.Date(start_summer),
-         season = case_when(
-           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
-           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
-           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
-  filter(dateOnly != "2022-12-15") # remove December 15 2022, because it turns out that's a non-inclusive boundary (oops)
-
-s2 <- s1 %>%
-  mutate(age = year - birth_year,
-         sex = case_when(sex == "f/m" ~ "u",
-                         TRUE ~ sex)) %>%
-  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
-  group_by(Nili_id, year, month) %>%
-  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
-  mutate(age_group = case_when(age == 0 ~ "juv",
-                               age >= 1 & age <= 4 ~ "sub",
-                               age >= 5 ~ "adult",
-                               TRUE ~ NA)) %>%
-  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
-                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
-                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
-  # explicitly set the factor levels in temporal order
-  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall")),
-         season = factor(season, levels = c("breeding", "summer", "fall")))
-table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read
-
-# remove columns that were used in the season calculation
-s3 <- s2 %>%
-  dplyr::select(-c(month, start_breeding, start_summer, start_fall))
-
-# Separate the seasons -----------------------------
-seasons <- s3 %>%
-  group_by(seasonUnique) %>%
-  group_split(.keep = T)
-# extract the season names in case we need a separate vector at some point.
-seasonNames <- map_chr(seasons, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
-```
-
-```{r save-soc-data}
-# Save copies for social analysis
-# The next step will be to remove individuals with too few points per day or too few days tracked.
-But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric.
-So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis.
-seasons_forSoc <- seasons
-#save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
-```
-
-There are `r length(seasonNames)` seasons: `r seasonNames` (we will end
-up removing Summer 2020 because the Ornitela tags started partway
-through that season.) `seasons` is therefore a list of length
-**`r length(seasons)`**, with the following numbers of rows in each
-season: `r map_dbl(seasons, nrow)`. This data is saved as
-`seasons_forSoc.Rda` so it can be used for the social network analysis
-without any further changes made.
-
-### Step 7: Get roosts
-
-Now I use Marta's roost code, implemented in vultureUtils as
-`get_roosts_df`, to get roost locations for each individual on each
-night.
-
-```{r get-roosts}
-# Get roosts for each season ----------------------------------------------
-# forRoosts <- map(seasons, ~.x %>% sf::st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE))
-# roosts_seasons <- purrr::map(forRoosts, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
-# roosts_seasons <- roosts_seasons %>%
-#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
-# save(roosts_seasons, file = "data/roosts_seasons.Rda")
-load("data/roosts_seasons.Rda")
-```
-
-### Step 8: Remove nighttime points
-
-Just for the data I'm going to use for movement analyses, I want to
-exclude any points that occur at night. This will also help deal with
-the problem that there are some individuals that seem to have way too
-many points per day
-
-```{r remove-night}
-# Remove nighttime points -------------------------------------------------
-before <- map_dbl(seasons, nrow)
-seasons <- map(seasons, ~{
-  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
-                                     lat = 31.434306, lon = 34.991889,
-                                     keep = c("sunrise", "sunset")) %>%
-    dplyr::select("dateOnly" = date, sunrise, sunset)
-
-  .x <- .x %>%
-    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
-    dplyr::left_join(times, by = "dateOnly") %>%
-    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
-    dplyr::select(-c(sunrise, sunset)) %>%
-    dplyr::filter(daylight == "day")
-})
-after <- map_dbl(seasons, nrow)
-change <- after-before
-changeDF <- data.frame(seasonName = seasonNames, nBefore = before, nAfter = after) %>%
-  mutate(propRowsRemoved = round((before-after)/before, 2))
-```
-
-```{r include = T, echo = F}
-print(changeDF)
-```
-
-### Step 9: Include only "southern individuals" in the movement calculations
-
-For each individual in each season, I calculated the centroid of its
-coordinates. Then I looked at some histograms of the centroids in each
-season and decided that 3550000 latitude (in UTM zone 36N coordinates--https://epsg.io/32636) would be a
-cutoff point that would neatly divide individuals in the southern
-population from the northern population (there was a large gap in the
-histogram consistently around this value).
-
-I filtered each season down to remove individuals classified as
-"northern" for that season, leaving only those in the Negev population.
-Keep in mind:
-
-1)  A given individual could be "northern" in one season but "southern"
-    in another season, theoretically. It would be removed only in the
-    seasons when it falls in the north.
-
-2)  Any removed individuals can still contribute to the social network
-    that I will calculate--that is, they can contribute to the
-    degree/strength/etc. of southern birds. I'm only removing them for
-    the movement calculations, so they won't be focal individuals in the
-    model.
-
-```{r remove-northern}
-# Restrict to southern individuals ----------------------------------------
-# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636)
-## Transform to SF object, so we can get centroids
-seasonsSF <- map(seasons, ~.x %>%
-                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
-                   sf::st_set_crs("WGS84") %>%
-                   sf::st_transform(32636))
-
-## Get centroids, so we can see who's "southern" for that season.
-centroids <- map(seasonsSF, ~.x %>%
-                   group_by(Nili_id) %>%
-                   summarize(geometry = st_union(geometry)) %>%
-                   st_centroid()) # XXX what is the actual definition of the centroid?
-
-## Examine a histogram of centroid latitudes 
-# walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
-
-## Get southern individuals for each season, so we can filter the data
-southernIndivs <- map(centroids, ~.x %>%
-                        filter(st_coordinates(.)[,2] < 3550000) %>%
-                        pull(Nili_id))
-
-## Remove individuals not in the south
-before <- seasons
-seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
-after <- seasons
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
-afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
-df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
-df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
-```
-
-After doing this filtering, the following numbers of rows and
-individuals remained in each season:
-
-```{r include = T, echo = F}
-print(df)
-```
-
-### Step 10: Remove vulture\*days based on number of points per day
-
-```{r view-ppd}
-# Include only individuals with enough points per day ------------------------
-# How many points per day are we looking at?
-ppd <- map_dfr(seasons, ~.x %>%
-             st_drop_geometry() %>%
-             group_by(seasonUnique, season, Nili_id, dateOnly) %>%
-             summarize(n = n()))
-
-## Theoretically, should only have 72 points per day--12 hours, 6 points per hour.
-That should vary a bit.
-I don't really understand why so many individuals/days have more than 100 points per day, but let's proceed for now: going to downsample that to 10 minue intervals anyway, so it won't matter.
-```
-
-We had initially decided to remove vulture\*days with \< 30 points per
-day. Here's what that would look like:
-
-```{r include = T, echo = F}
-# 
-thresh <- 30
-ppd %>%
-  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
-  ggplot(aes(x = n))+
-  geom_histogram(aes(fill = Status))+
-  facet_wrap(~seasonUnique, scales ="free")+
-  geom_vline(aes(xintercept = thresh), col = "red")+
-  scale_fill_manual(values = c("darkgray", "red"))+
-  theme_classic()+
-  ggtitle(paste("n removed =", thresh, sep = " "))+
-  ylab("# vulture*days")+
-  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
-```
-
-Restricting to days with 30 points per day ends up removing a huge
-number of days, especially in the breeding season (winter) because of
-lack of sunlight etc. After talking with Marta and Noa, I decided to
-remove just 10 points per day. That would look like this:
-
-```{r include = T, echo = F}
-# 
-thresh <- 10
-ppd %>%
-  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
-  ggplot(aes(x = n))+
-  geom_histogram(aes(fill = Status))+
-  facet_wrap(~seasonUnique, scales ="free")+
-  geom_vline(aes(xintercept = thresh), col = "red")+
-  scale_fill_manual(values = c("darkgray", "red"))+
-  theme_classic()+
-  ggtitle(paste("n removed =", thresh, sep = " "))+
-  ylab("# vulture*days")+
-  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
-```
-
-```{r}
-# thresh <- 10
-# ppd <- ppd %>%
-#   ungroup() %>%
-#   mutate(removed = ifelse(n < thresh, T, F))
-# 
-# # what proportion of vulture*days are removed overall?
-# (propRemoved <- ppd %>%
-#     group_by(seasonUnique, season) %>%
-#     summarize(nTot = n(),
-#               nRemoved = sum(removed),
-#               propRemoved = nRemoved/nTot))
-# propRemoved %>%
-#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
-#   geom_point(size = 10)+
-#   theme_classic()+
-#   scale_color_manual(values = c("blue", "red", "darkorange"))
-# 
-# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
-# 
-# # what about by individual over each season?
-# ppd %>%
-#   group_by(seasonUnique, season, Nili_id) %>%
-#   summarize(propRemoved = sum(removed)/n()) %>%
-#   ggplot(aes(x = seasonUnique, y = propRemoved))+
-#   geom_boxplot(aes(group = seasonUnique, fill = season))+
-#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
-#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
-#   theme_classic()+ # tells the same story
-#   ylab("Proportion of days removed")+
-#   xlab("")
-
-## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
-before <- seasons
-seasons <- map(seasons, ~.x %>%
-                 group_by(Nili_id, dateOnly) %>%
-                 filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>%
-                 ungroup())
-after <- seasons
-
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
-  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))
-```
-
-Here's how many points get removed when we do this filtering:
-
-```{r include = T, echo = F}
-df
-```
-
-### Step 11: Remove individuals based on number of days tracked
-
-This filtering comes after the removal of some vulture\*days for having
-too few points. So, considering the remaining days, we only want to
-retain individuals that are tracked for at least 1/3 of a given season
-(in terms of number of days).
-
-Here are the number of days in each season. As you can see, breeding
-season is the longest, followed by summer, then fall. (Ignore Summer
-2020--it will not be included in the analysis.)
-
-```{r include = TRUE, echo = F}
-seasonNames
-durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))
-durs
-```
-
-```{r remove-daysTracked}
-# Include only individuals with enough days tracked -----------------------
-## Must be tracked for at least 1/3 of the number of days in the season.
-
-# walk2(seasons, durs, ~.x %>%
-#   st_drop_geometry() %>%
-#   group_by(Nili_id) %>%
-#   summarize(nDaysTracked = length(unique(dateOnly)),
-#             propDaysTracked = nDaysTracked/.y) %>%
-#   pull(propDaysTracked) %>%
-#   hist())
-
-indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
-                       st_drop_geometry() %>%
-                       group_by(Nili_id) %>%
-                       summarize(nDaysTracked = length(unique(dateOnly)),
-                                 propDaysTracked = nDaysTracked/.y) %>%
-                       filter(nDaysTracked >= 30) %>%
-                       pull(Nili_id))
-
-## remove those individuals not tracked for enough days
-before <- seasons
-after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))
-
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
-  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))
-
-```
-
-Here's what this does to the row counts:
-
-```{r include = T, echo = F}
-df
-
-## Now I notice that season 1, Summer 2020, doesn't have enough individuals. Let's remove it--we're going to have to anyway.
-seasons <- seasons[-1]
-```
-
-### Step 12: Calculate flight altitudes
-
-Using the `elevatr` package, I get ground elevation values for each gps
-point in each season, at zoom level 10 (which should come out to around
-150 meter cell resolution in the resulting raster, according to )
-
-```{r get-elevation}
-# Get elevation rasters ---------------------------------------------------
-elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x %>% st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE, crs = "WGS84"), z = 10))
-# 
-save(elevs_z10_seasons, file = "data/elevs_z10_seasons.Rda")
-load("data/elevs_z10_seasons.Rda")
-
-groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y %>% sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = FALSE))) # have to convert the crs because the raster from elevatr is in WGS84.
-
-## use elevation rasters to calculate height above ground level
-seasons <- map2(.x = seasons, .y = groundElev_z10, 
-                ~.x %>% mutate(height_above_ground = .y,
-                               height_above_ground = case_when(height_above_ground <0 ~ 0,
-                                                               TRUE ~ height_above_ground)))
-
-# Export the data
-save(seasons, file = "data/seasons.Rda")
-```
-
-As expected, adding the elevation information doesn't change the number
-of rows in the data.
-
-### Step 13: Downsample the data
-
-Because occasionally, the vultures' tags were set to read more
-frequently than every 10 minutes, I want to downsample the data to a
-10-minute interval.
-
-```{r downsample}
-# Downsample the data and save the downsampled data
-subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
-  sub <- df %>%
-    arrange(idCol, timestampCol) %>%
-    group_by(.data[[idCol]]) %>%
-    mutate(tc = cut(.data[[timestampCol]], 
-                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
-    group_by(.data[[idCol]], 
-             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
-    slice(1) %>%
-    ungroup() %>%
-    dplyr::select(-c("d", "tc"))
-  return(sub)
-}
-
-# create and save downsampled datasets for later use
-seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
-seasons_20min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 20), .progress = T)
-seasons_30min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 30), .progress = T)
-seasons_60min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 60), .progress = T)
-seasons_120min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 120), .progress = T)
-
-map_dbl(seasons, nrow)
-map_dbl(seasons_10min, nrow)
-
-# save(seasons_10min, file = "data/seasons_10min.Rda")
-# save(seasons_20min, file = "data/seasons_20min.Rda")
-# save(seasons_30min, file = "data/seasons_30min.Rda")
-# save(seasons_60min, file = "data/seasons_60min.Rda")
-# save(seasons_120min, file = "data/seasons_120min.Rda")
-
-# no need to calculate roosts for the downsampled data because in theory there's only one point per night.
-
-df <- data.frame(season = seasonNames, 
-                 nBefore = map_dbl(seasons, nrow),
-                 nAfter = map_dbl(seasons_10min, nrow)) %>%
-  mutate(propChange = round((nBefore-nAfter)/nBefore, 2))
-```
-
-Here are the row counts after downsampling:
-
-```{r include = TRUE, echo = F}
-df
-```
-
-And here are the final counts of number of vultures:
-
-```{r final-counts}
-df <- seasons_10min %>% purrr::list_rbind() %>% 
-  group_by(seasonUnique) %>% 
-  summarize(vultures = length(unique(Nili_id)))
-```
-
-```{r include = TRUE, echo = F}
-df
-```

### 2024-01-11 00.0_dataPrep.qmd
further improvements to dataPrep

@@
 
--   `inpadf_touse`
-
 -   `data/derived/inpadf_touse.feather`
@@
 #| eval: false
-# inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
-# inpadf <- methods::as(inpa, "data.frame")
-# inpadf_touse <- inpadf %>%
-#   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
-#          year = as.numeric(lubridate::year(timestamp))) %>%
-#   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
-# write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")
+inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
+inpadf <- methods::as(inpa, "data.frame")
+inpadf_touse <- inpadf %>%
+  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
+         year = as.numeric(lubridate::year(timestamp))) %>%
+  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
+write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")
 
@@
 
--   `dat`
-
 -   `data/derived/dat.feather`
@@
 maxDate <- "2023-05-15 11:59"
-# dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-# write_feather(ornitela, "data/derived/dat.feather")
+dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+write_feather(ornitela, "data/derived/dat.feather")
 
@@
 
--   `dat3`
+-   `data/derived/dat3.feather`
+
+```{r load-inpa-ornitela}
+inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
+dat <- read_feather("data/derived/dat.feather")
+```
 
@@
 
-inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
-dat <- read_feather("data/derived/dat.feather")
-# # Will there be a problem joining these?
-# names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
-# names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
-# 
-# # Add the dataset names so we can keep track of where the data comes from
-# inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
-#   mutate(dataset = "inpa")
-# dat <- dat %>%
-#   mutate(dataset = "ornitela")
-# 
-# # Join the two datasets
-# joined <- bind_rows(inpadf_tojoin, dat)
-# dim(joined)
-# 
-# ## fix trackId
-# dat2 <- joined %>%
-#   mutate(trackId = as.character(trackId),
-#          trackId = case_when(trackId == "E03" ~ "E03w",
-#                              TRUE ~ trackId))
-# # number of unique individuals
-# length(unique(dat2$trackId)) # 206
-# 
-# # Add the Nili_id
-# ww <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
-#   dplyr::select(Nili_id, Movebank_id) %>%
-#   distinct()
-# 
-# # Are there any other individuals who have two different associated trackId's?
-# ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
-# ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
-# 
-# all(dat2$trackId %in% ww$Movebank_id) # false.
-# 
-# # Okay, let's investigate the ones that aren't included
-# dat2 %>%
-#   filter(!(trackId %in% ww$Movebank_id)) %>%
-#   pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
-# 
-# dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
-# ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
-# 
-# sort(unique(ww$mb_id))
-# unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
-# dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
-# dat2$trackId[dat2$trackId == "e86white"] <- "e86"
-# unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
-# # Investigate the remaining ones:
-# dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
-# # I guess for now let's just make their trackId's into Nili_id's
-# 
-# dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
-#                   by = c("trackId" = "mb_id"))
-# nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
-# length(unique(dat3$Nili_id))
-# # For any that didn't get Nili_ids assigned, make it the trackId
-# dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
-# dat3 <- dat3 %>%
-#   mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
-#                              TRUE ~ Nili_id))
-# unique(dat3$Nili_id)
+# Will there be a problem joining these?
+names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
+names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
+
+# Add the dataset names so we can keep track of where the data comes from
+inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
+  mutate(dataset = "inpa")
+dat <- dat %>%
+  mutate(dataset = "ornitela")
+
+# Join the two datasets
+joined <- bind_rows(inpadf_tojoin, dat)
+dim(joined)
+
+## fix trackId
+dat2 <- joined %>%
+  mutate(trackId = as.character(trackId),
+         trackId = case_when(trackId == "E03" ~ "E03w",
+                             TRUE ~ trackId))
+# number of unique individuals
+length(unique(dat2$trackId)) # 206
+
+# Add the Nili_id
+ww <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+  dplyr::select(Nili_id, Movebank_id) %>%
+  distinct()
+
+# Are there any other individuals who have two different associated trackId's?
+ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
+ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
+
+all(dat2$trackId %in% ww$Movebank_id) # false.
+
+# Okay, let's investigate the ones that aren't included
+dat2 %>%
+  filter(!(trackId %in% ww$Movebank_id)) %>%
+  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
+
+dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
+ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
+
+sort(unique(ww$mb_id))
+unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
+dat2$trackId[dat2$trackId == "e86white"] <- "e86"
+unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+# Investigate the remaining ones:
+dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
+# I guess for now let's just make their trackId's into Nili_id's
+
+dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
+                  by = c("trackId" = "mb_id"))
+nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
+length(unique(dat3$Nili_id))
+# For any that didn't get Nili_ids assigned, make it the trackId
+dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
+dat3 <- dat3 %>%
+  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
+                             TRUE ~ Nili_id))
+unique(dat3$Nili_id)
+write_feather(dat3, "data/derived/dat3.feather")
 ```
@@
 
--   `dat3`
+-   `data/derived/dat3.feather`
 
@@
 
+```{r load-hospital}
+dat3 <- read_feather("data/derived/dat3.feather")
+toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
+```
+
 ```{r remove-hospital}
@@
 ## annotate the data with periods to remove
-toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
-
 toRemove <- toRemove %>%
@@
   dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
+
+write_feather(datAnnot_noCaptures, "data/derived/datAnnot_noCaptures.feather")
 ```
+
+## Attach age and sex information
+
+*Inputs:*
+
+-   `data/raw/whoswho_vultures_20230315_new.xlsx`
+
+-   `data/derived/datAnnot_noCaptures.feather`
+
+*Outputs:*
+
+-   `datAnnot` (`datAnnot2`)
+
+```{r}
+datAnnot_noCaptures <- read_feather("data/derived/datAnnot_noCaptures.feather")
+```
+
+```{r}
+#| eval: false
+# Age and sex info --------------------------------------------------------
+# Attach age and sex information
+as <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+  dplyr::select(Nili_id, birth_year, sex) %>%
+  distinct()
+
+datAnnot2 <- datAnnot_noCaptures %>%
+  dplyr::select(-c("sex")) %>%
+  left_join(as, by = "Nili_id")
+nrow(datAnnot2) == nrow(datAnnot)
+length(unique(datAnnot2$Nili_id)) # 196
+
+datAnnot <- datAnnot2 # a hack so the rest of the code will work
+```
+
+## Clean data
+
+*Inputs:*
+
+-   `data/raw/CutOffRegion.kml`
+
+-   `datAnnot`
+
+*Outputs:*
+
+-   `data/derived/datAnnotCleaned.Rda`
+
+```{r}
+#| eval: false
+# Clean the data
+## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
+mask <- sf::st_read("data/raw/CutOffRegion.kml")
+datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall.
+If we want 30 days per season, will have to do that differently (e.g.
+split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
+length(unique(datAnnotCleaned$Nili_id)) # 172 individuals
+# Fix time zone so dates make sense ---------------------------------------
+## Overwrite the dateOnly column from the new times
+datAnnotCleaned <- datAnnotCleaned %>%
+  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
+         dateOnly = lubridate::date(timestampIsrael),
+         month = lubridate::month(dateOnly),
+         year = lubridate::year(dateOnly))
+
+save(datAnnotCleaned, file = "data/derived/datAnnotCleaned.Rda")
+load("data/derived/datAnnotCleaned.Rda")
+```
+
+## Split into seasons
+
+Season definitions, and code to split the data into seasons, come from Marta.
+
+*Inputs:*
+
+-   `datAnnotCleaned`
+
+*Outputs:*
+
+-   `data/derived/seasons_orig.Rda`
+
+-   `data/derived/seasonNames_orig.Rda`
+
+```{r}
+#| eval: false
+# Split into Marta's 3 seasons --------------------------------------------
+s1 <- datAnnotCleaned %>%
+  mutate(start_breeding = ifelse(month %in% c(1:5),
+                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
+                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
+         start_summer = ifelse(month == 12,
+                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
+                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
+         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
+         start_breeding = as.Date(start_breeding),
+         start_summer = as.Date(start_summer),
+         season = case_when(
+           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
+           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
+           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
+  filter(dateOnly != "2023-05-15") # remove May 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
+
+s2 <- s1 %>%
+  mutate(age = year - birth_year,
+         sex = case_when(sex == "f/m" ~ "u",
+                         TRUE ~ sex)) %>%
+  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
+  group_by(Nili_id, year, month) %>%
+  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
+  mutate(age_group = case_when(age == 0 ~ "juv",
+                               age >= 1 & age <= 4 ~ "sub",
+                               age >= 5 ~ "adult",
+                               TRUE ~ NA)) %>%
+  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
+                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
+                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
+  # explicitly set the factor levels in temporal order
+  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding")),
+         season = factor(season, levels = c("breeding", "summer", "fall")))
+table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read
+
+# remove columns that were used in the season calculation
+s3 <- s2 %>%
+  dplyr::select(-c(month, start_breeding, start_summer, start_fall))
+
+# Separate the seasons -----------------------------
+seasons_orig <- s3 %>%
+  group_by(seasonUnique) %>%
+  group_split(.keep = T)
+# extract the season names in case we need a separate vector at some point.
+seasonNames_orig <- map_chr(seasons_orig, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
+save(seasons_orig, file = "data/derived/seasons_orig.Rda")
+save(seasonNames_orig, file = "data/derived/seasonNames_orig.Rda")
+load("data/derived/seasons_orig.Rda")
+load("data/derived/seasonNames_orig.Rda")
+```
+
+## Remove individuals with low fix rate
+
+Find the mode (most common) observation frequency for each individual on each day. Then exclude individuals who never have observations at 10-minute frequencies.
+
+We keep the ones that theoretically have a 10-minute fix rate, even if in practice they often have much less frequent observations.
+
+*Inputs:*
+
+-   `seasons_orig`
+
+*Outputs:*
+
+-   `data/derived/toKeep_fixRate.Rda`
+
+-   `seasons_mode10`
+
+```{r}
+#| eval: false
+# Optionally remove individuals with a low fix rate -----------------------
+Mode <- function(x) { # function to calculate mode fix rate
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
+# Calculate daily modes and add them to the dataset
+seasons <- map(seasons_orig, ~.x %>%
+                 group_by(dateOnly, Nili_id) %>%
+                 mutate(diff = as.numeric(difftime(lead(timestamp), timestamp, units = "mins"))) %>%
+                 group_by(dateOnly, Nili_id) %>%
+                 mutate(mode = Mode(round(diff))) %>%
+                 ungroup())
+
+# Exclude individuals that never have a daily mode of 10 minutes
+toKeep_fixrate <- map(seasons, ~.x %>%
+                        sf::st_drop_geometry() %>%
+                        group_by(Nili_id) %>%
+                        summarize(minmode = min(mode, na.rm = T)) %>%
+                        filter(minmode <= 10) %>%
+                        pull(Nili_id) %>%
+                        unique())
+save(toKeep_fixrate, file = "data/derived/toKeep_fixrate.Rda")
+load("data/derived/toKeep_fixrate.Rda")
+seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
+```
+
+## Save dataset for social analysis
+
+The next step will be to remove individuals with too few points per day or too few days tracked.
+But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric.
+So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis.
+
+*Inputs:*
+
+-   `seasons`
+
+-   `seasons_mode10`
+
+*Outputs:*
+
+-   `data/derived/seasons_forSoc.Rda`
+
+-   `data/derived/seasons_forSoc_mode10.Rda`
+
+```{r}
+# Save copies for social analysis
+# 
+seasons_forSoc <- seasons
+seasons_forSoc_mode10 <- seasons_mode10
+save(seasons_forSoc, file = "data/derived/seasons_forSoc.Rda")
+save(seasons_forSoc_mode10, file = "data/derived/seasons_forSoc_mode10.Rda")
+load("data/derived/seasons_forSoc.Rda")
+load("data/derived/seasons_forSoc_mode10.Rda")
+seasons_mode10 <- seasons_forSoc_mode10
+seasons <- seasons_forSoc
+
+# Investigate: How many individuals are retained in each of these cases? What proportion were removed?
+inds_all <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
+inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
+propRemoved <- round((inds_all-inds_mode10)/inds_all, 3) # For all seasons, getting rid of individuals that don't have 
+propRemoved # yikes that's a lot of birds. Alas.
+```
+
+## Get roosts
+
+*Inputs:*
+
+-   `seasons`
+
+-   `seasons_mode10`
+
+*Outputs:*
+
+-   `data/derived/roosts_seasons.Rda`
+
+-   `data/derived/roosts_seasons_mode10.Rda`
+
+```{r}
+#| eval: false
+# Get roosts for each season ----------------------------------------------
+roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts_seasons <- roosts_seasons %>%
+  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
+save(roosts_seasons, file = "data/derived/roosts_seasons.Rda")
+load("data/derived/roosts_seasons.Rda")
+
+roosts_seasons_mode10 <- purrr::map(seasons_mode10, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts_seasons_mode10 <- roosts_seasons_mode10 %>%
+  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
+save(roosts_seasons_mode10, file = "data/derived/roosts_seasons_mode10.Rda")
+load("data/derived/roosts_seasons_mode10.Rda")
+```
+
+## Remove nighttime points
+
+XXX do we not also remove nighttime points for seasons_mode10?
+
+*Inputs:*
+
+-   `seasons`
+
+*Outputs:*
+
+-   `seasons`
+
+```{r}
+#| eval: false
+# Remove nighttime points -------------------------------------------------
+before <- map_dbl(seasons, nrow)
+seasons <- map(seasons, ~{
+  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
+                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
+                                     keep = c("sunrise", "sunset")) %>%
+    dplyr::select("dateOnly" = date, sunrise, sunset)
+
+  .x <- .x %>%
+    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
+    dplyr::left_join(times, by = "dateOnly") %>%
+    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
+    dplyr::select(-c(sunrise, sunset)) %>%
+    dplyr::filter(daylight == "day")
+})
+after <- map_dbl(seasons, nrow)
+change <- after-before
+changeDF <- data.frame(seasonName = seasonNames_orig, nBefore = before, nAfter = after) %>%
+  mutate(propRowsRemoved = round((before-after)/before, 2))
+changeDF
+```
+
+## Restrict to Southern individuals
+
+Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, <https://epsg.io/32636)>
+
+XXX why are we using `seasons_orig` here? shouldn't it be `seasons`?
+
+*Inputs:*
+
+-   `seasons_orig`
+
+*Outputs:*
+
+-   `data/derived/seasonsSF.Rda`
+
+-   `seasons` XXX seems wrong!
+
+```{r}
+#| eval: false
+# Restrict to southern individuals ----------------------------------------
+# 
+# Transform to SF object, so we can get centroids
+seasonsSF <- map(seasons_orig, ~.x %>%
+                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
+                   sf::st_set_crs("WGS84") %>%
+                   sf::st_transform(32636))
+
+save(seasonsSF, file = "data/derived/seasonsSF.Rda")
+load("data/derived/seasonsSF.Rda")
+
+## Get centroids, so we can see who's "southern" for that season.
+centroids <- map(seasonsSF, ~.x %>%
+                   group_by(Nili_id) %>%
+                   summarize(geometry = st_union(geometry)) %>%
+                   st_centroid())
+
+## Examine a histogram of centroid latitudes 
+walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
+
+## Get southern individuals for each season, so we can filter the data
+southernIndivs <- map(centroids, ~.x %>%
+                        filter(st_coordinates(.)[,2] < 3550000) %>%
+                        pull(Nili_id))
+
+## Remove individuals not in the south
+before <- seasons
+seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
+after <- seasons
+
+beforeRows <- map_dbl(before, nrow)
+afterRows <- map_dbl(after, nrow)
+beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
+afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
+df <- data.frame(season = seasonNames_orig, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
+df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
+df
+rm(centroids)
+rm(seasonsSF)
+```
+
+## Remove individuals with too few points per day
+
+*Inputs:*
+
+-   `seasons`
+
+*Outputs:*
+
+-   `seasons`
+
+```{r}
+#| eval: false
+# Include only individuals with enough points per day ------------------------
+# Investigate battery percentage vs points per day ------------------------
+# Is there some way we can relate battery charge percentage to number of points per day?
+test <- seasons[[5]]
+names(test)
+
+ppd <- map_dfr(seasons, ~.x %>%
+                 st_drop_geometry() %>%
+                 group_by(seasonUnique, season, Nili_id, dateOnly) %>%
+                 summarize(n = n(),
+                           minBatt = min(battery_charge_percent, na.rm = T),
+                           meanBatt = mean(battery_charge_percent, na.rm = T),
+                           medBatt = mean(battery_charge_percent, na.rm = T)))
+
+excl_batt <- 50
+excl_pts <- 10
+## min battery charge
+ppd %>%
+  filter(n < 100) %>% # only interested in the lower numbers
+  mutate(remove = case_when(minBatt <= excl_batt & n <= excl_pts ~ T,
+                            TRUE ~ F)) %>%
+  ggplot(aes(x = minBatt, y = n))+
+  geom_point(aes(col = remove))+
+  facet_wrap(~seasonUnique) +
+  scale_color_manual(values = c("black", "red")) +
+  theme_classic()
+
+# # cumulative histogram to look at the results of removing days based only on ppd
+# ppd %>%
+#   filter(n < 100) %>%
+#   ggplot(aes(x = n)) +
+#   stat_ecdf(aes(y = after_stat(y)*100))+
+#   ylab("Percentage of vulture-days included")+
+#   xlab("Points per day")+
+#   theme_classic()+
+#   facet_wrap(~seasonUnique)
+# 
+# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
+# ppd %>%
+#   filter(n < 100, minBatt > 50) %>%
+#   ggplot(aes(x = n)) +
+#   stat_ecdf(aes(y = after_stat(y)*100))+
+#   ylab("Percentage of vulture-days included")+
+#   xlab("Points per day")+
+#   theme_classic()+
+#   facet_wrap(~seasonUnique)
+# 
+# # same thing, but with batery set at 75
+# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
+# ppd %>%
+#   filter(n < 100, minBatt > 75) %>%
+#   ggplot(aes(x = n)) +
+#   stat_ecdf(aes(y = after_stat(y)*100))+
+#   ylab("Percentage of vulture-days included")+
+#   xlab("Points per day")+
+#   theme_classic()+
+#   facet_wrap(~seasonUnique)
+# 
+# # What if instead we decide we want a certain number of ppd--battery cuttoff may be sharper.
+Say, 30 ppd: is there a discernible battery cutoff that distinguishes "real" from "battery artifact" days?
+# ppd %>%
+#   filter(n < 30) %>%
+#   ggplot(aes(x = minBatt))+
+#   stat_ecdf(aes(y = after_stat(y)*100))+
+#   ylim(c(0, 100))+
+#   facet_wrap(~seasonUnique)+
+#   ylab("Percentage of vulture-days included")+
+#   xlab("Minimum battery percentage") # we do seem to see crooks here around 75%!
+# 
+# # Same thing but with 10 points per day? Predict much sharper cutoffs.
+# ppd %>%
+#   filter(n < 10) %>%
+#   ggplot(aes(x = minBatt))+
+#   stat_ecdf(aes(y = ..y..*100))+
+#   ylim(c(0, 100))+
+#   facet_wrap(~seasonUnique)+
+#   ylab("Percentage of vulture-days included")+
+#   xlab("Minimum battery percentage") # crooks around 50% battery minimum for 10ppd. This seems like a pretty good way to go!
+
+# Let's throw out anything with < 10ppd unless the battery has a min charge > 50%.
+
+# Orr suggests to just restrict to vulture*days with >30 points per day. Let's see what that would look like.
+thresh <- 30
+# ppd %>%
+#   mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
+#   ggplot(aes(x = n))+
+#   geom_histogram(aes(fill = Status))+
+#   facet_wrap(~seasonUnique, scales ="free")+
+#   geom_vline(aes(xintercept = thresh), col = "red")+
+#   scale_fill_manual(values = c("darkgray", "red"))+
+#   theme_classic()+
+#   ggtitle(paste("n removed =", thresh, sep = " "))+
+#   ylab("# vulture*days")+
+#   xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
+# 
+# ppd <- ppd %>%
+#   ungroup() %>%
+#   mutate(removed = ifelse(n < thresh, T, F))
+# 
+# # what proportion of vulture*days are removed overall?
+# (propRemoved <- ppd %>%
+#     group_by(seasonUnique, season) %>%
+#     summarize(nTot = n(),
+#               nRemoved = sum(removed),
+#               propRemoved = nRemoved/nTot))
+# propRemoved %>%
+#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
+#   geom_point(size = 10)+
+#   theme_classic()+
+#   scale_color_manual(values = c("blue", "red", "darkorange"))
+# 
+# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
+# 
+# # what about by individual over each season?
+# ppd %>%
+#   group_by(seasonUnique, season, Nili_id) %>%
+#   summarize(propRemoved = sum(removed)/n()) %>%
+#   ggplot(aes(x = seasonUnique, y = propRemoved))+
+#   geom_boxplot(aes(group = seasonUnique, fill = season))+
+#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
+#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
+#   theme_classic()+ # tells the same story
+#   ylab("Proportion of days removed")+
+#   xlab("")
+
+# Based on the battery charge analysis above, going to remove individuals that have 10 or fewer points per day, but only if the minimum battery charge is less than 50%.
+
+## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
+seasons <- map(seasons, ~.x %>%
+                 group_by(Nili_id, dateOnly) %>%
+                 filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>%
+                 ungroup())
+```
+
+## Remove individuals with too few days tracked
+
+*Inputs:*
+
+-   `seasons`
+
+*Outputs:*
+
+-   `seasons`
+
+```{r}
+#| eval: false
+# Include only individuals with enough days tracked -----------------------
+## Must be tracked for at least 30 days per season
+durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))
+
+# walk2(seasons, durs, ~.x %>%
+#         st_drop_geometry() %>%
+#         group_by(Nili_id) %>%
+#         summarize(nDaysTracked = length(unique(dateOnly)),
+#                   propDaysTracked = nDaysTracked/.y) %>%
+#         pull(propDaysTracked) %>%
+#         hist())
+
+indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
+                       st_drop_geometry() %>%
+                       group_by(Nili_id) %>%
+                       summarize(nDaysTracked = length(unique(dateOnly)),
+                                 propDaysTracked = nDaysTracked/.y) %>%
+                       filter(nDaysTracked >= 30) %>%
+                       pull(Nili_id))
+
+## remove those individuals not tracked for enough days
+before <- seasons
+after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))
+
+beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
+afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
+seasons <- after
+```
+
+## Calculate flight altitude
+
+*Inputs:*
+
+-   `seasons`
+
+*Outputs:*
+
+-   `data/derived/elevs_z10_seasons.Rda`
+
+-   `data/derived/groundElev_z10.Rda`
+
+-   `seasons`
+
+```{r}
+#| eval: false
+# Get elevation rasters ---------------------------------------------------
+# Remove the first element of `seasons` because it doesn't have any data
+seasons <- seasons[-1] # removing summer 2020
+# # XXXXXXX start here
+#seasons <- map(seasons, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
+#elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x , z = 10))
+#save(elevs_z10_seasons, file = "data/derived/elevs_z10_seasons.Rda")
+load("data/derived/elevs_z10_seasons.Rda")
+
+# groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.
+# save(groundElev_z10, file = "data/derived/groundElev_z10.Rda")
+load("data/derived/groundElev_z10.Rda")
+
+## use elevation rasters to calculate height above ground level
+before <- seasons
+seasons <- map2(.x = seasons, .y = groundElev_z10, 
+                ~.x %>% mutate(groundElev = .y) %>%
+                  mutate(height_above_ground = height_above_msl-groundElev,
+                         height_above_ground = case_when(height_above_ground < 0 ~ 0,
+                                                         TRUE ~ height_above_ground)))
+after <- seasons
+
+# There should be no change in the number of rows
+all(map2_lgl(before, after, ~nrow(.x) == nrow(.y)))
+```
+
+## Restrict to mode10 individuals
+
+XXX what is happening here
+
+*Inputs:*
+
+-   `data/derived/toKeep_fixrate.Rda`
+
+-   `seasons`
+
+*Outputs:*
+
+-   `data/derived/seasons.Rda`
+
+-   `data/derived/seasons_mode10.Rda`
+
+```{r}
+#| eval: false
+# Now restrict to the mode10 individuals ----------------------------------
+load("data/derived/toKeep_fixrate.Rda")
+toKeep_fixrate <- toKeep_fixrate[-1]
+seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
+
+save(seasons, file = "data/derived/seasons.Rda")
+save(seasons_mode10, file = "data/derived/seasons_mode10.Rda")
+load("data/derived/seasons.Rda")
+load("data/derived/seasons_mode10.Rda")
+```
+
+## Remove high-frequency fixes
+
+Downsampling data to 10-minute intervals exclusively
+
+*Inputs:*
+
+-   `seasons`
+
+-   `seasons_mode10`
+
+*Outputs:*
+
+-   `data/derived/seasons_10min.Rda`
+
+-   `data/derived/seasons_mode10_10min.Rda`
+
+XXX this is confusing. Need to rename after concepts, not minutes.
+
+```{r}
+#| eval: false
+# Downsample the data and save the downsampled data
+subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
+  sub <- df %>%
+    arrange(idCol, timestampCol) %>%
+    group_by(.data[[idCol]]) %>%
+    mutate(tc = cut(.data[[timestampCol]], 
+                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
+    group_by(.data[[idCol]], 
+             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
+    slice(1) %>%
+    ungroup() %>%
+    dplyr::select(-c("d", "tc"))
+  return(sub)
+}
+
+# create and save downsampled datasets for later use
+seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
+seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
+
+map_dbl(seasons, nrow)
+map_dbl(seasons_10min, nrow)
+map_dbl(seasons_mode10, nrow)
+
+seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
+seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
+
+save(seasons_10min, file = "data/derived/seasons_10min.Rda")
+save(seasons_mode10_10min, file = "data/derived/seasons_mode10_10min.Rda")
+```

### 2024-01-11 scripts/00.0_dataPrep.R

@@ -8,30 +8,30 @@ library(move)
 library(feather)
 library(readxl)
 library(elevatr)
-# 
-# ## Download data from movebank
-# base::load("movebankCredentials/pw.Rda")
-# MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
-# rm(pw)
-# 
-# # Hatzofe INPA data
-# # inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
-# # inpadf <- methods::as(inpa, "data.frame")
-# # inpadf_touse <- inpadf %>%
-# #   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
-# #          year = as.numeric(lubridate::year(timestamp))) %>%
-# #   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
-# # write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")
-# inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
-# # Ornitela: download data from movebank (just a subset of the times for now)
-# minDate <- "2020-09-01 00:00"
-# maxDate <- "2023-05-15 11:59"
-# # dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-# # write_feather(dat, "data/derived/dat.feather")
-# dat <- read_feather("data/derived/dat.feather")
-# # number of unique individuals
-# length(unique(dat$trackId)) # 127
-# length(unique(inpadf_touse$trackId)) # 77
+
+## Download data from movebank
+base::load("movebankCredentials/pw.Rda")
+MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
+rm(pw)
+
+# Hatzofe INPA data
+# inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
+# inpadf <- methods::as(inpa, "data.frame")
+# inpadf_touse <- inpadf %>%
+#   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
[…]

### 2024-01-11 manuscript.qmd
Fixed citations by adding brackets

--- 2024-01-10 manuscript.qmd
@@
 The social structure of animal groups has consequences for disease transmission, information sharing, and population dynamics.
-At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration (@sah2017, @silk2017), govern the flow of information about food resources \[REF\], and contribute to population persistence through Allee effects \[REF\].
+At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration [@sah2017; @silk2017], govern the flow of information about food resources \[REF\], and contribute to population persistence through Allee effects \[REF\].
 Individuals differ in their encounter rates, number of contacts, and centrality in the population social network \[REF\].
@@
 social centrality in different behavioral situations contributes to their likelihood of contracting disease, finding mates, or accessing resources \[REFS\].
-For example, in networks with high levels of degree heterogeneity, the most connected individuals may function as disease superspreaders (@lloyd-smith2005).
-For example, possums with higher betweenness and closeness in a social network had higher infection rates (@corner2003), and higher connectivity in lizards was associated with a greater parasite load (@godfrey2009).
-In general, individuals that are more central in the social network both have a higher infection risk and transmit disease more quickly than those that are more peripheral (@romano2016, @balasubramaniam2019).
+For example, in networks with high levels of degree heterogeneity, the most connected individuals may function as disease superspreaders [@lloyd-smith2005].
+For example, possums with higher betweenness and closeness in a social network had higher infection rates [@corner2003], and higher connectivity in lizards was associated with a greater parasite load [@godfrey2009].
+In general, individuals that are more central in the social network both have a higher infection risk and transmit disease more quickly than those that are more peripheral [@romano2016; @balasubramaniam2019].
 
@@
 The relationship between movement, space use, and social interaction patterns has been an area of active research in recent years.
-Webber et al -@webber2023 have characterized the variation in patterns of animal movement and space use as the "spatial phenotype," as part of a larger framework for understanding the relationships between social and spatial drivers of animal behavior.
+Webber et al [-@webber2023] have characterized the variation in patterns of animal movement and space use as the "spatial phenotype," as part of a larger framework for understanding the relationships between social and spatial drivers of animal behavior.
 In this light, the "spatial phenotype" may be considered in addition to more conventional phenotypic measures to explain why some individuals are more central than others in their social networks.
@@
 affiliative interactions) \[REFS\].
-The overall social environment experienced by an individual is made up of its interactions in all of these different contexts @silk2013.
-Individuals differ in their centrality in these different contexts @sharma2023.
+The overall social environment experienced by an individual is made up of its interactions in all of these different contexts [@silk2013].
+Individuals differ in their centrality in these different contexts [@sharma2023].
 Different types of interactions are also localized differently in space.
@@
 
-    -   Flight: interactions in flight are important for information transfer @harel2017, finding locations of thermal uplift @sassi2023 and feeding sites @spiegel2013, @cortés-avizanda2014
+    -   Flight: interactions in flight are important for information transfer [@harel2017], finding locations of thermal uplift [@sassi2023] and feeding sites [@spiegel2013; @cortés-avizanda2014].
 
@@
 The population is small and restricted enough that individuals' home ranges are highly overlapping, with most individuals using a large fraction of available space in the region.
-This makes habitat features likely to affect most individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement @harel2016.
+This makes habitat features likely to affect most individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement [@harel2016].
 GPS fixes are taken once every 10 minutes, frequently enough to measure flight movements with good accuracy, and most individuals are tracked for several months to several years.
-Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population @sharma2023.\
+Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population [@sharma2023].\
 \
@@
 
--   @blonder2012, @cantor2012, @hobson2013, @rasmussen2013, @holme2015, @farine2018, @fisher2021, @pasquaretta2021
+-   [@blonder2012; @cantor2012; @hobson2013; @rasmussen2013; @holme2015; @farine2018; @fisher2021; @pasquaretta2021]
 
@@
 The biological processes driving the execution of these behaviors and the formation of these networks are likewise happening over different time scales.
-An individual must decide once per day which roost to go to @harel2016.
-But flight is a continuous process driven by many decisions resulting from individual and social information @sassi2023, and feeding events are concentrated in time on a given day.
-Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every \[NUMBER\] days and persist for hours to a few days (Orr Spiegel, pers.
+An individual must decide once per day which roost to go to [@harel2016].
+But flight is a continuous process driven by many decisions resulting from individual and social information [@sassi2023], and feeding events are concentrated in time on a given day.
+Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every [NUMBER] days and persist for hours to a few days (Orr Spiegel, pers.
 comm.).
@@
 Tags recorded GPS fixes, as well as speed and heading readings, every 10 minutes when battery conditions allowed.
-The GPS data are stored on Movebank @kays2022 and were loaded into R using the `move` package @kranstauber2023.
+The GPS data are stored on Movebank [@kays2022] and were loaded into R using the `move` package\[ @kranstauber2023\].
 
@@
 To characterize individual movement behavior and spatial phenotype, we cleaned the data further.
-We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package suncalc @thieurmel2022.
+We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package suncalc [@thieurmel2022].
 We restricted movement calculations to a set of focal individuals with activity regions centered in the Negev region of Israel, the site of most of our tagging effort.
@@
 For each season, we calculated movement measures only for individuals tracked for at least 30 days (after removing low-point days as described above).
-We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package @hollister2022, at a zoom level of \[level–more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of [level–more details about resolution?] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
 Finally, since some tags were occasionally set to a much higher fix rate (e.g.
@@
 2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
-Using the `adehabitatHR` package @calenge2023, we calculated home range as the 95% kernel utilization distribution (KUD) of a vulture's GPS fixes in a season; core area as the 50% KUD, and core area use intensity as the ratio of the 50% to the 95% KUD.
+Using the `adehabitatHR` package [@calenge2023], we calculated home range as the 95% kernel utilization distribution (KUD) of a vulture's GPS fixes in a season; core area as the 50% KUD, and core area use intensity as the ratio of the 50% to the 95% KUD.
 
@@
 We identified in-flight fixes (ground speed \>= 5 m/s) and calculated measures of daily flight behavior, including maximum and median daily flight altitude and daily flight time (the sum of time intervals between adjacent in-flight fixes).
-Finally, we used all fixes to calculate daily distance traveled (the sum of all distances between successive fixes); daily maximum displacement (the straight-line distance between the beginning of the daily movement track and the farthest point reached on that day); and movement tortuosity (cumulative distance to the point of maximum displacement divided by the straight-line distance to the point of maximum displacement @benhamou2004).
+Finally, we used all fixes to calculate daily distance traveled (the sum of all distances between successive fixes); daily maximum displacement (the straight-line distance between the beginning of the daily movement track and the farthest point reached on that day); and movement tortuosity (cumulative distance to the point of maximum displacement divided by the straight-line distance to the point of maximum displacement [@benhamou2004]).
 We took seasonal means of each of these individual daily measures to get a single value per individual vulture, per season.
@@
 
-Using GPS fixes from the tagged vultures, we created proximity-based social networks in three different social situations, following @sharma2023: co-flight, co-feeding (diurnal ground interactions) and co-roosting (nocturnal ground interactions).
+Using GPS fixes from the tagged vultures, we created proximity-based social networks in three different social situations, following Sharma [-@sharma2023]: co-flight, co-feeding (diurnal ground interactions) and co-roosting (nocturnal ground interactions).
 Networks were created separately for each of the eight seasons.
@@
 Because not all individuals were tagged for exactly the same period of time, and to account for missing data due to low tag battery or poor reception, we calculated the simple ratio index (SRI) to quantify the strength of the association between each pair of vultures in each social situation.
-We calculated SRI by dividing the number of 10-minute time periods during which two vultures were observed interacting by the number of time periods during which both vultures had a recorded GPS fix (whether or not they were interacting during that period) @ginsberg1992.
+We calculated SRI by dividing the number of 10-minute time periods during which two vultures were observed interacting by the number of time periods during which both vultures had a recorded GPS fix (whether or not they were interacting during that period) [@ginsberg1992].
 Thus, two vultures who never interacted in a given social situation would have an SRI of 0, while two individuals who were interacting in every time period during which a GPS fix was collected for both of them would have an SRI of 1.
@@
 
-Vultures were considered to be flying if their ground speed (collected by a sensor on the GPS tags) was \> 5 m/s @spiegel2013 .
-Because vultures can see each other over long distances @pennycuick1972, @spiegel2013a , and in accordance with @sharma2023, we defined a co-flight interaction as two individuals flying within 1000m of each other for at least two consecutive 10-minute time periods.
+Vultures were considered to be flying if their ground speed (collected by a sensor on the GPS tags) was \> 5 m/s [@spiegel2013] .
+Because vultures can see each other over long distances [@pennycuick1972; @spiegel2013a] , and in accordance with [@sharma2023], we defined a co-flight interaction as two individuals flying within 1000m of each other for at least two consecutive 10-minute time periods.
 
@@
 
-Vultures that were not flying (ground speed \< 5 m/s, @spiegel2013) who came within 50m of each other for at least two consecutive 10-minute time periods, were considered to be having a ground interaction @sharma2023.
+Vultures that were not flying (ground speed \< 5 m/s, [@spiegel2013]) who came within 50m of each other for at least two consecutive 10-minute time periods, were considered to be having a ground interaction [@sharma2023].
 Any ground interactions that occurred at night or within the boundaries of a known roost site were excluded.
@@
 comm.).
-Co-roosting Roost sites were assigned using \[methods\].
+Co-roosting Roost sites were assigned using [methods].
 Unlike flight and diurnal ground GPS fixes, vultures were only assigned one roosting location per night.
@@
 
-We used the R package `igraph` @csardi2005 to calculate each individual's degree and strength for each social situation in each season.
+We used the R package `igraph` [@csardi2005] to calculate each individual's degree and strength for each social situation in each season.
 Degree is defined as the number of other individuals that a vulture is connected to (i.e.
@@
 
-We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the lme4 and lmerTest packages @bates2015, @kuznetsova2017.
+We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the lme4 and lmerTest packages [@bates2015; @kuznetsova2017].
 We fit separate models for each response variable (degree and strength).
@@
 
-For each response variable, we began by fitting a model without any interaction terms to check for multicollinearity between the predictors and check model assumptions, using the performance and DHARMa packages @lüdecke2021, @hartig2022.
+For each response variable, we began by fitting a model without any interaction terms to check for multicollinearity between the predictors and check model assumptions, using the performance and DHARMa packages [@lüdecke2021; @hartig2022].
 Next, we fitted a model containing two three-way interaction terms (between season, situation, and each of the spatial phenotype variables), all of the included two-way interaction terms, age as a main effect, and the two random effects (individual ID; season by year).
@@
 
-For each model, we calculated marginal effects using the emmeans package @lenth2023 and created visualizations using ggeffects and ggplot2 @wickham2016, @lüdecke2018.
+For each model, we calculated marginal effects using the emmeans package [@lenth2023] and created visualizations using ggeffects and ggplot2 [@wickham2016; @lüdecke2018].
 
@@
 # Discussion
+
+\[text here\]

### 2024-01-11 00.0_dataPrep.qmd
began changing variable names (through data cleaning step)

@@
 
-## Goals
+## Goals and introduction
 
-Prepare data for the movement/social analysis
+Prepare data for the movement/social analysis. This involves loading the data from Movebank and cleaning it.
 
@@
 
--   `data/derived/inpadf_touse.feather`
+-   `data/derived/inpa.feather`
 
@@
 inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
-inpadf <- methods::as(inpa, "data.frame")
-inpadf_touse <- inpadf %>%
+inpa <- methods::as(inpa, "data.frame")
+inpa <- inpa %>%
   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
@@
   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
-write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")
+write_feather(inpa, "data/derived/inpa.feather")
 
-length(unique(inpadf_touse$trackId)) # 77
+length(unique(inpa$trackId)) # 77
 ```
@@
 
--   `data/derived/dat.feather`
+-   `data/derived/ornitela.feather`
 
@@
 maxDate <- "2023-05-15 11:59"
-dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-write_feather(ornitela, "data/derived/dat.feather")
+ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
+write_feather(ornitela, "data/derived/ornitela.feather")
 
-length(unique(dat$trackId)) # 127
+length(unique(ornitela$trackId)) # 127
 ```
@@
 
--   `data/derived/inpadf_touse.feather`
+-   `data/derived/inpa.feather`
 
--   `data/derived/dat.feather`
+-   `data/derived/ornitela.feather`
 
@@
 
--   `data/derived/dat3.feather`
+-   `data/derived/joined.feather`
 
 ```{r load-inpa-ornitela}
-inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
-dat <- read_feather("data/derived/dat.feather")
+inpa <- read_feather("data/derived/inpa.feather")
+ornitela <- read_feather("data/derived/ornitela.feather")
 ```
@@
 # Will there be a problem joining these?
-names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
-names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
+names(ornitela)[!(names(ornitela) %in% names(inpa))] # all names in "ornitela" are found in inpa
+names(inpa)[!(names(inpa) %in% names(ornitela))] # many are not found in "ornitela" that are found in inpa
 
 # Add the dataset names so we can keep track of where the data comes from
-inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
+inpa_tojoin <- inpa[,names(ornitela)] %>%
   mutate(dataset = "inpa")
-dat <- dat %>%
+ornitela <- ornitela %>%
   mutate(dataset = "ornitela")
@@
 # Join the two datasets
-joined <- bind_rows(inpadf_tojoin, dat)
-dim(joined)
+joined0 <- bind_rows(inpa_tojoin, ornitela)
+dim(joined0)
 
 ## fix trackId
-dat2 <- joined %>%
+joined0 <- joined0 %>%
   mutate(trackId = as.character(trackId),
@@
                              TRUE ~ trackId))
+
 # number of unique individuals
-length(unique(dat2$trackId)) # 206
+length(unique(joined0$trackId)) # 206
 
@@
 # Are there any other individuals who have two different associated trackId's?
-ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
+ww %>% filter(Movebank_id %in% joined$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
 ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
 
-all(dat2$trackId %in% ww$Movebank_id) # false.
+all(joined0$trackId %in% ww$Movebank_id) # false.
 
 # Okay, let's investigate the ones that aren't included
-dat2 %>%
+joined0 %>%
   filter(!(trackId %in% ww$Movebank_id)) %>%
@@
 
-dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
+joined0$trackId <- str_replace_all(tolower(joined0$trackId), "\\.", "")
 ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
@@
 sort(unique(ww$mb_id))
-unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
-dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
-dat2$trackId[dat2$trackId == "e86white"] <- "e86"
-unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
+unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
+joined0$trackId[joined0$trackId == "y01t60w"] <- "t60white"
+joined0$trackId[joined0$trackId == "e86white"] <- "e86"
+unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
 # Investigate the remaining ones:
-dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
+joined0 %>% filter(!(trackId %in% ww$mb_id)) %>% 
+  dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>%
+  group_by(trackId, local_identifier, ring_id, dataset) %>% 
+  summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
 # I guess for now let's just make their trackId's into Nili_id's
 
-dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
+joined <- left_join(joined0, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
                   by = c("trackId" = "mb_id"))
-nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
-length(unique(dat3$Nili_id))
+nrow(joined) == nrow(joined0) # good, same number of rows. (check that the join worked)
+length(unique(joined$Nili_id))
 # For any that didn't get Nili_ids assigned, make it the trackId
-dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
-dat3 <- dat3 %>%
+joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
+joined <- joined %>%
   mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
                              TRUE ~ Nili_id))
-unique(dat3$Nili_id)
-write_feather(dat3, "data/derived/dat3.feather")
+unique(joined$Nili_id)
+write_feather(joined, "data/derived/joined.feather")
 ```
@@
 
--   `data/derived/dat3.feather`
+-   `data/derived/joined.feather`
 
@@
 
--   `datAnnot`
+-   `removed`
 
 ```{r load-hospital}
-dat3 <- read_feather("data/derived/dat3.feather")
+joined <- read_feather("data/derived/joined.feather")
 toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
@@
   mutate(across(c(remove_start, remove_end), .fns = function(x){
-    lubridate::ymd(x)
-  })) %>%
-  dplyr::filter(!is.na(remove_end))
-
-toRemove_long <- toRemove %>%
+    lubridate::ymd(x)})) %>%
+  dplyr::filter(!is.na(remove_end)) %>%
   group_by(Nili_id, reason) %>%
   # sequence of daily dates for each corresponding start, end elements
-  dplyr::mutate(dateOnly = map2(remove_start, remove_end, seq, by = "1 day")) %>%
+  dplyr::mutate(dateOnly = map2(remove_start, remove_end, 
+                                seq, by = "1 day")) %>%
   # unnest the list column
@@
 # Join to the original data
-datAnnot <- dat3 %>%
-  left_join(toRemove_long, by = c("Nili_id", "dateOnly")) %>%
+removal_annotated <- joined %>%
+  left_join(toRemove, by = c("Nili_id", "dateOnly")) %>%
   mutate(status = replace_na(status, "valid"))
-nrow(datAnnot) == nrow(dat3) #T: good, same number of rows.
+nrow(removal_annotated) == nrow(joined) #T: good, same number of rows.
 
 # Actually REMOVE the periods to remove...
-datAnnot <- datAnnot %>%
+removed <- removal_annotated %>%
   filter(status == "valid")
-nrow(datAnnot)
-length(unique(datAnnot$Nili_id))
-nrow(dat3)
-nrow(datAnnot)
+nrow(removed)
+length(unique(removal_annotated$Nili_id))
+nrow(joined)
+nrow(removal_annotated)
 ```
@@
 
+To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function).
+Then, if the bird roosted within 500m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset.
+This has been validated.
+This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it.
+So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data
+
 *Inputs:*
 
--   `datAnnot`
+-   `removed`
+
+-   `data/raw/capture_sites.csv`
 
@@
 
-```{r}
+```{r remove-capture-cage}
 #| eval: false
@@
 # Identify and remove capture dates using Marta's code --------------------
-## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function).
-Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset.
-This has been validated.
-Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages.
-But we're just dealing with southern individuals, so that's okay.
-# roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-# save(roosts, file = "data/derived/roosts.Rda")
+roosts <- get_roosts_df(df = removed, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+save(roosts, file = "data/derived/roosts.Rda")
 load("data/derived/roosts.Rda")
@@
 
-# Load the information about the capture sites
+# Load information about the capture sites
 sites <- read.csv("data/raw/capture_sites.csv")
@@
 
-## This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it.
-So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data
+## For Carmel--different protocol.
 
@@
 # It all looks ok, so we can subset the dataset to exclude the capture dates
-datAnnot_noCaptures <- datAnnot %>%
+removed_captures <- removed %>%
   left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
-nrow(datAnnot) == nrow(datAnnot_noCaptures) # should be TRUE. NOW we can filter.
-datAnnot_noCaptures <- datAnnot_noCaptures %>%
+nrow(removed) == nrow(removed_captures) # should be TRUE. NOW we can filter.
+removed_captures <- removed_captures %>%
   dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured
@@
 # How many rows and individuals did we remove?
-before <- nrow(datAnnot)
-after <- nrow(datAnnot_noCaptures)
+before <- nrow(removed)
+after <- nrow(removed_captures)
 (propChange <- (before-after)/before) # approx. 1% of data removed.
-length(unique(datAnnot$Nili_id)) == length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
+length(unique(removed$Nili_id)) == length(unique(removed_captures$Nili_id)) # no change in number of individuals! As expected.
 
-datAnnot_noCaptures <- datAnnot_noCaptures %>%
+removed_captures <- removed_captures %>%
   dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
 
-write_feather(datAnnot_noCaptures, "data/derived/datAnnot_noCaptures.feather")
+write_feather(removed_captures, "data/derived/removed_captures.feather")
 ```
@@
 
--   `data/derived/datAnnot_noCaptures.feather`
+-   `data/derived/removed_captures.feather`
 
@@
 
-```{r}
-datAnnot_noCaptures <- read_feather("data/derived/datAnnot_noCaptures.feather")
+```{r attach-age-sex}
+removed_captures <- read_feather("data/derived/removed_captures.feather")
 ```
@@
 
-datAnnot2 <- datAnnot_noCaptures %>%
+# Whenever I do left_joins, I save the result as something different before reassigning it to the variable name so I can make sure the join worked before committing to it.
+Hence the temporary "test" object.
+test <- removed_captures %>%
   dplyr::select(-c("sex")) %>%
   left_join(as, by = "Nili_id")
-nrow(datAnnot2) == nrow(datAnnot)
-length(unique(datAnnot2$Nili_id)) # 196
-
-datAnnot <- datAnnot2 # a hack so the rest of the code will work
+nrow(test) == nrow(removed_captures)
+length(unique(removed_captures$Nili_id)) # 196
+removed_captures <- test
+rm(test) # don't need this anymore after we know the join worked.
 ```
@@
 
--   `datAnnot`
+-   `removed_captures`
 
@@
 
--   `data/derived/datAnnotCleaned.Rda`
+-   `data/derived/cleaned.Rda`
 
-```{r}
+```{r clean-data}
 #| eval: false
@@
 mask <- sf::st_read("data/raw/CutOffRegion.kml")
-datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall.
+cleaned <- vultureUtils::cleanData(dataset = removed_captures, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall.
 If we want 30 days per season, will have to do that differently (e.g.
 split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
-length(unique(datAnnotCleaned$Nili_id)) # 172 individuals
+length(unique(cleaned$Nili_id)) # 172 individuals
 # Fix time zone so dates make sense ---------------------------------------
 ## Overwrite the dateOnly column from the new times
-datAnnotCleaned <- datAnnotCleaned %>%
+cleaned <- cleaned %>%
   mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
@@
 
-save(datAnnotCleaned, file = "data/derived/datAnnotCleaned.Rda")
-load("data/derived/datAnnotCleaned.Rda")
+save(cleaned, file = "data/derived/cleaned.Rda")
+load("data/derived/cleaned.Rda")
 ```
@@
 
--   `datAnnotCleaned`
+-   `cleaned`
 
@@
 
-```{r}
+```{r split-seasons}
 #| eval: false

### 2024-01-12 00.0_dataPrep.qmd
More work on changing variable names

@@
 length(unique(inpa$trackId)) # 77
+
+# tidy up
+rm(inpa)
+gc()
 ```
@@
 length(unique(ornitela$trackId)) # 127
+
+# tidy up
+rm(ornitela)
+gc()
 ```
@@
 
+rm(inpa)
+rm(inpa_tojoin)
+rm(ornitela)
+
 ## fix trackId
@@
 # number of unique individuals
-length(unique(joined0$trackId)) # 206
+length(unique(joined0$trackId))
 
@@
 # Are there any other individuals who have two different associated trackId's?
-ww %>% filter(Movebank_id %in% joined$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
+ww %>% filter(Movebank_id %in% joined0$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
 ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
@@
 # For any that didn't get Nili_ids assigned, make it the trackId
-joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
+joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expected.
 joined <- joined %>%
@@
                              TRUE ~ Nili_id))
-unique(joined$Nili_id)
+length(unique(joined$Nili_id))
 write_feather(joined, "data/derived/joined.feather")
+
+# tidy up
+rm(joined0)
+rm(joined)
+gc()
 ```
@@
 joined <- read_feather("data/derived/joined.feather")
-toRemove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
-```
-
-```{r remove-hospital}
-#| eval: false
-
+periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
+```
+
+```{r remove-periods}
 ## annotate the data with periods to remove
-toRemove <- toRemove %>%
+periods_to_remove <- periods_to_remove %>%
   dplyr::select(Nili_id,
@@
 # Actually REMOVE the periods to remove...
-removed <- removal_annotated %>%
+removed_periods <- removal_annotated %>%
   filter(status == "valid")
-nrow(removed)
+nrow(removed_periods)
 length(unique(removal_annotated$Nili_id))
-nrow(joined)
-nrow(removal_annotated)
-```
-
-## Identify and remove capture cage periods
+nrow(joined) == nrow(removal_annotated) # true
+nrow(removed_periods) != nrow(removal_annotated) # true
+
+write_feather(removed_periods, "data/derived/removed_periods.feather")
+
+# tidy up
+rm(periods_to_remove)
+rm(joined)
+rm(removal_annotated)
+rm(removed_periods)
+gc()
+```
+
+## Remove capture cage periods
 
@@
 
-To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function).
+To identify the capture dates, first we need to classify the roosts (now using the `get_roosts()` function).
 Then, if the bird roosted within 500m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset.
@@
 This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it.
-So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data
-
-*Inputs:*
-
--   `removed`
+So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data.
+
+Note: I'm calling the roosts data frame generated here `roosts0.Rda` because I want to distinguish it from the more final roost location *list* that we will generate further along in the cleaning process.
+
+*Inputs:*
+
+-   `data/derived/removed_periods.feather`
 
@@
 
--   `roosts`
-
--   `data/derived/roosts.Rda`
+-   `data/derived/roosts0.Rda`
+
+```{r}
+removed_periods <- read_feather("data/derived/removed_periods.feather")
+```
+
+```{r remove-capture-cage-getroosts}
+#| eval: false
+# Identify and remove capture dates using Marta's code --------------------
+roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
+save(roosts0, file = "data/derived/roosts0.Rda")
+```
+
+```{r}
+# Load information about the capture sites
+capture_sites <- read.csv("data/raw/capture_sites.csv")
+load("data/derived/roosts0.Rda")
+AllCarmelDates <- read.csv("data/raw/all_captures_carmel_2010-2021.csv")
+```
 
 ```{r remove-capture-cage}
-#| eval: false
-
-# Identify and remove capture dates using Marta's code --------------------
-roosts <- get_roosts_df(df = removed, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-save(roosts, file = "data/derived/roosts.Rda")
-load("data/derived/roosts.Rda")
-
 # Identify the period of time during which the capture sites are open (when we need to do this exclusion)
@@
 
-# Load information about the capture sites
-sites <- read.csv("data/raw/capture_sites.csv")
-
 # Subset the roost dataset with the start and end dates for the capture period
-sub.roosts <- roosts %>%
+sub.roosts <- roosts0 %>%
   mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
@@
 
-DistanceMat <- matrix(ncol = nrow(sites), nrow = nrow(crds))
-colnames(DistanceMat) <- unique(sites$name)
+DistanceMat <- matrix(ncol = nrow(capture_sites), nrow = nrow(crds))
+colnames(DistanceMat) <- unique(capture_sites$name)
 
 for(i in 1:nrow(crds)){ # for each roost point...
-  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
+  DistanceMat[i,] <- round(geosphere::distm(crds[i,], capture_sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
   #print(i)
@@
 ## For Carmel--different protocol.
-
 sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
@@
 
-AllCarmelDates <- read.csv("data/raw/all_captures_carmel_2010-2021.csv")
 AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")
@@
 # It all looks ok, so we can subset the dataset to exclude the capture dates
-removed_captures <- removed %>%
+removed_captures <- removed_periods %>%
   left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
-nrow(removed) == nrow(removed_captures) # should be TRUE. NOW we can filter.
+nrow(removed_periods) == nrow(removed_captures) # should be TRUE. NOW we can filter.
 removed_captures <- removed_captures %>%
@@
 # How many rows and individuals did we remove?
-before <- nrow(removed)
+before <- nrow(removed_periods)
 after <- nrow(removed_captures)
 (propChange <- (before-after)/before) # approx. 1% of data removed.
-length(unique(removed$Nili_id)) == length(unique(removed_captures$Nili_id)) # no change in number of individuals! As expected.
+length(unique(removed_periods$Nili_id)) == length(unique(removed_captures$Nili_id)) # no change in number of individuals! As expected.
 
@@
 write_feather(removed_captures, "data/derived/removed_captures.feather")
-```
-
-## Attach age and sex information
+
+# tidy up
+rm(AllCarmelDates)
+rm(AllCarmelDates.1)
+rm(AllCarmelDates.2)
+rm(AllCarmelDates.3)
+rm(AllCarmelDates.all)
+rm(capture_sites)
+rm(crds)
+rm(DistanceMat)
+rm(removed)
+rm(removed_captures)
+rm(roosts0)
+rm(sub.captured)
+rm(sub.captured.carmel)
+rm(sub.captured.dates)
+rm(sub.captured.dates.1)
+rm(sub.captured.no.carmel)
+rm(sub.roosts)
+rm(removed_periods)
+rm(after)
+rm(before)
+rm(ClosestCaptureDist)
+rm(ClosestCaptureSite)
+rm(distance)
+gc()
+```
+
+## Attach birth year and sex information
+
+Note: the actual calculation of age happens later on, after cleaning the data, for some reason.
 
@@
 
--   `datAnnot` (`datAnnot2`)
+-   `data/derived/with_age_sex.feather`
 
@@
 removed_captures <- read_feather("data/derived/removed_captures.feather")
-```
-
-```{r}
-#| eval: false
-# Age and sex info --------------------------------------------------------
-# Attach age and sex information
-as <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+age_sex <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
   dplyr::select(Nili_id, birth_year, sex) %>%
   distinct()
-
-# Whenever I do left_joins, I save the result as something different before reassigning it to the variable name so I can make sure the join worked before committing to it.
-Hence the temporary "test" object.
-test <- removed_captures %>%
+```
+
+```{r}
+#| eval: false
+with_age_sex <- removed_captures %>%
   dplyr::select(-c("sex")) %>%
-  left_join(as, by = "Nili_id")
-nrow(test) == nrow(removed_captures)
-length(unique(removed_captures$Nili_id)) # 196
-removed_captures <- test
-rm(test) # don't need this anymore after we know the join worked.
+  left_join(age_sex, by = "Nili_id")
+nrow(with_age_sex) == nrow(removed_captures)
+length(unique(with_age_sex$Nili_id)) # 196
+
+write_feather(with_age_sex, "data/derived/with_age_sex.feather")
+
+# tidy up
+rm(removed_captures)
+rm(with_age_sex)
+rm(age_sex)
+gc()
 ```
@@
 
+Here, we set the `inMaskThreshold` to 30 days overall.
+If we wanted 30 days per season, we would have to do that differently (e.g.
+split the seasons before this step, not after).
+But I think that gets covered later in the data cleaning anyway.
+
 *Inputs:*
@@
 
--   `removed_captures`
-
-*Outputs:*
-
--   `data/derived/cleaned.Rda`
+-   `data/derived/with_age_sex.feather`
+
+*Outputs:*
+
+-   `data/derived/cleaned.feather`
+
+```{r}
+mask <- sf::st_read("data/raw/CutOffRegion.kml")
+with_age_sex <- read_feather("data/derived/with_age_sex.feather")
+```
 
@@
 ## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
-mask <- sf::st_read("data/raw/CutOffRegion.kml")
-cleaned <- vultureUtils::cleanData(dataset = removed_captures, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall.
-If we want 30 days per season, will have to do that differently (e.g.
-split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
+cleaned <- vultureUtils::cleanData(dataset = with_age_sex, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) 
 length(unique(cleaned$Nili_id)) # 172 individuals
@@
 
+# This produces a huge file because it's an SF file. We don't really need it to be sf at this stage. Let's remove that.
+cleaned <- sf::st_drop_geometry(cleaned)
+
 save(cleaned, file = "data/derived/cleaned.Rda")
+
+# tidy up
+rm(mask)
+rm(with_age_sex)
+rm(cleaned)
+gc()
+```
+
+## Split into seasons
+
+Season definitions, and code to split the data into seasons, come from Marta.
+
+*Inputs:*
+
+-   `data/derived/cleaned.Rda`
+
+*Outputs:*
+
+-   `data/derived/seasons_list.Rda`
+
+-   `data/derived/season_names.Rda`
+
+```{r}
 load("data/derived/cleaned.Rda")
@@
 
-## Split into seasons
-
-Season definitions, and code to split the data into seasons, come from Marta.
-
-*Inputs:*
-
--   `cleaned`
-
-*Outputs:*
-
--   `data/derived/seasons_orig.Rda`
-
--   `data/derived/seasonNames_orig.Rda`
-
 ```{r split-seasons}
@@
 # Split into Marta's 3 seasons --------------------------------------------
-s1 <- datAnnotCleaned %>%
+cleaned <- cleaned %>%
   mutate(start_breeding = ifelse(month %in% c(1:5),
@@
 
-s2 <- s1 %>%
-  mutate(age = year - birth_year,
+with_seasons <- cleaned %>%
+  mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
          sex = case_when(sex == "f/m" ~ "u",
@@
          season = factor(season, levels = c("breeding", "summer", "fall")))
-table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read
 
 # remove columns that were used in the season calculation
-s3 <- s2 %>%
+with_seasons <- with_seasons %>%
   dplyr::select(-c(month, start_breeding, start_summer, start_fall))
 
+# REMOVE SUMMER 2020--we don't need to use it
+with_seasons <- with_seasons %>%
+  filter(seasonUnique != "2020_summer")
+length(unique(with_seasons$seasonUnique))
+
 # Separate the seasons -----------------------------
-seasons_orig <- s3 %>%
+seasons_list <- with_seasons %>%
   group_by(seasonUnique) %>%
   group_split(.keep = T)
+length(seasons_list) # should be 8
 # extract the season names in case we need a separate vector at some point.
-seasonNames_orig <- map_chr(seasons_orig, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
-save(seasons_orig, file = "data/derived/seasons_orig.Rda")
-save(seasonNames_orig, file = "data/derived/seasonNames_orig.Rda")
-load("data/derived/seasons_orig.Rda")
-load("data/derived/seasonNames_orig.Rda")
+season_names <- map_chr(seasons_list, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
+save(seasons_list, file = "data/derived/seasons_list.Rda")
+save(season_names, file = "data/derived/season_names.Rda")
+
+# tidy up
+rm(cleaned)
+rm(seasons_list)
+rm(with_seasons)
+rm(season_names)
+gc()
 ```
@@
 
-Find the mode (most common) observation frequency for each individual on each day. Then exclude individuals who never have observations at 10-minute frequencies.
-
-We keep the ones that theoretically have a 10-minute fix rate, even if in practice they often have much less frequent observations.
-
-*Inputs:*
-
--   `seasons_orig`
-
-*Outputs:*
-
--   `data/derived/toKeep_fixRate.Rda`
-
--   `seasons_mode10`
+START HERE!!!
+
+Some of the INPA vultures were tracked only once an hour, or some other low fix rate.
+However, determining which birds these are is tricky because even the vultures that should have a 10-minute fix rate do not always have points every 10 minutes, e.g.
+due to lack of reception or low tag battery.\
+\
+Therefore, we have to identify **low fix rate (lfr) individuals**, which are those that *never* have a 10-minute fix rate.
+To do this, we will find the mode (most common) fix frequency for each individual on each day.
+Then, we will exclude individuals who never have observations at 10-minute intervals.
+We will keep any that theoretically have a 10-minute fix rate (i.e.
+their observation frequency was 10 minutes at some point), even if in practice they often have much less frequent observations.
+
+*Inputs:*
+
+-   `data/derived/seasons_list.Rda`
+
+*Outputs:*
+
+-   `data/derived/toRemove_lfr.Rda`
+
+-   `data/derived/seasons_list_removed_lfr.Rda`
+
+```{r}
+load("data/derived/seasons_list.Rda")
+```
 
@@
 # Optionally remove individuals with a low fix rate -----------------------
-Mode <- function(x) { # function to calculate mode fix rate
+# function to calculate mode (most common) fix rate
+Mode <- function(x) { 
   ux <- unique(x)
@@
 
-# Calculate daily modes and add them to the dataset
-seasons <- map(seasons_orig, ~.x %>%
+# Calculate daily modes and add them to the dataset (using the function defined above)
+with_modes <- map(seasons_list, ~.x %>%
                  group_by(dateOnly, Nili_id) %>%
@@
 # Exclude individuals that never have a daily mode of 10 minutes
-toKeep_fixrate <- map(seasons, ~.x %>%
+toRemove_lfr <- map(with_modes, ~.x %>%
                         sf::st_drop_geometry() %>%
@@
                         summarize(minmode = min(mode, na.rm = T)) %>%
-                        filter(minmode <= 10) %>%
+                        filter(minmode > 10) %>%
                         pull(Nili_id) %>%
                         unique())
-save(toKeep_fixrate, file = "data/derived/toKeep_fixrate.Rda")
-load("data/derived/toKeep_fixrate.Rda")
-seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
-```
-
-## Save dataset for social analysis
+save(toRemove_lfr, file = "data/derived/toRemove_lfr.Rda")
+seasons_list_removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% filter(!(Nili_id %in% .y)))
+save(seasons_list_removed_lfr, file = "data/derived/seasons_list_removed_lfr.Rda")
+
+# Tidy up
+rm(seasons_list)
+rm(seasons_list_removed_lfr)
+rm(with_modes)
+rm(toRemove_lfr)
+gc()
+```
+
+### Data for social analysis
 
@@
 But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric.
-So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis.
-
-*Inputs:*
-
--   `seasons`
-
--   `seasons_mode10`
-
-*Outputs:*
-
--   `data/derived/seasons_forSoc.Rda`
-
--   `data/derived/seasons_forSoc_mode10.Rda`
-
-```{r}
-# Save copies for social analysis
-# 
-seasons_forSoc <- seasons
-seasons_forSoc_mode10 <- seasons_mode10
-save(seasons_forSoc, file = "data/derived/seasons_forSoc.Rda")
-save(seasons_forSoc_mode10, file = "data/derived/seasons_forSoc_mode10.Rda")
-load("data/derived/seasons_forSoc.Rda")
-load("data/derived/seasons_forSoc_mode10.Rda")
-seasons_mode10 <- seasons_forSoc_mode10
-seasons <- seasons_forSoc
-
-# Investigate: How many individuals are retained in each of these cases? What proportion were removed?
-inds_all <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
-inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
-propRemoved <- round((inds_all-inds_mode10)/inds_all, 3) # For all seasons, getting rid of individuals that don't have 
-propRemoved # yikes that's a lot of birds. Alas.
-```
-
-## Get roosts
-
-*Inputs:*
-
--   `seasons`
-
--   `seasons_mode10`
-
-*Outputs:*
-
--   `data/derived/roosts_seasons.Rda`
-
--   `data/derived/roosts_seasons_mode10.Rda`
+So when we do the social analysis, we will use `seasons_list_removed_lfr` and not anything further along.
+
+## Get roosts before removing night points
+
+We are going to need to get nightly roost locations for all of the remaining vultures.
+We have to do this now because getting roost locations requires using points that occur at night, which we will eventually want to remove in order to do the movement analysis.
+
+*Inputs:*
+
+-   `data/derived/seasons_list_removed_lfr`
+
+*Outputs:*
+
+-   \`data/derived/roosts_seasons_list.
+
+```{r}
+load("data/derived/seasons_list_removed_lfr.Rda")
+```
 
@@
 # Get roosts for each season ----------------------------------------------
-roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
-roosts_seasons <- roosts_seasons %>%
-  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
-save(roosts_seasons, file = "data/derived/roosts_seasons.Rda")
-load("data/derived/roosts_seasons.Rda")
-
-roosts_seasons_mode10 <- purrr::map(seasons_mode10, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
-roosts_seasons_mode10 <- roosts_seasons_mode10 %>%
-  map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
-save(roosts_seasons_mode10, file = "data/derived/roosts_seasons_mode10.Rda")
-load("data/derived/roosts_seasons_mode10.Rda")
+roosts <- purrr::map(seasons_list_removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts <- roosts %>%
+  map(., ~st_as_sf(.x, crs = "WGS84", 
+                   coords = c("location_long", "location_lat"), 
+                   remove = F))
+save(roosts, file = "data/derived/roosts.Rda")
+
+# Tidy up
+rm(seasons_list_removed_lfr)
+rm(roosts)
+gc()
 ```
@@
 ## Remove nighttime points
+
+START HERE
 

### 2024-01-13 00.0_dataPrep.qmd
finalized cleaning up the data pipeline

@@
 
-## Goals and introduction
+# Goals and introduction
 
@@
 
-## Setup
+# Setup
 
@@
 
+# Processing for all data
+
 ## Download data from Movebank
@@
 rm(ornitela)
+rm(minDate)
+rm(maxDate)
 gc()
@@
 
+### Hospital/pre-determined invalid
+
 Removing known invalid periods, such as times when the vultures were in the hospital or the tags were known not to be attached to birds. These times are derived from a separate Excel sheet.
@@
 
--   `removed`
+-   `data/derived/removed_periods.feather`
 
@@
 
-## Remove capture cage periods
+### Capture cages
 
@@
 
+-   `data/raw/all_captures_carmel_2010-2021.csv`
+
 *Outputs:*
@@
 -   `data/derived/roosts0.Rda`
+
+-   `data/derived/removed_captures.feather`
 
@@
 rm(distance)
+rm(start.day)
+rm(start.month)
+rm(end.day)
+rm(end.month)
+rm(propChange)
+
 gc()
@@
 
-START HERE!!!
-
 Some of the INPA vultures were tracked only once an hour, or some other low fix rate.
@@
 
--   `data/derived/toRemove_lfr.Rda`
-
--   `data/derived/seasons_list_removed_lfr.Rda`
+-   `data/derived/removed_lfr.Rda`
 
@@
                         unique())
-save(toRemove_lfr, file = "data/derived/toRemove_lfr.Rda")
-seasons_list_removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% filter(!(Nili_id %in% .y)))
-save(seasons_list_removed_lfr, file = "data/derived/seasons_list_removed_lfr.Rda")
+
+removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% 
+                                   filter(!(Nili_id %in% .y)))
+
+save(removed_lfr, 
+     file = "data/derived/removed_lfr.Rda")
 
@@
 rm(seasons_list)
-rm(seasons_list_removed_lfr)
+rm(removed_lfr)
 rm(with_modes)
@@
 But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric.
-So when we do the social analysis, we will use `seasons_list_removed_lfr` and not anything further along.
+So when we do the social analysis, we will use `removed_lfr` and not anything further along.
 
@@
 
--   `data/derived/seasons_list_removed_lfr`
-
-*Outputs:*
-
--   \`data/derived/roosts_seasons_list.
-
-```{r}
-load("data/derived/seasons_list_removed_lfr.Rda")
+-   `data/derived/removed_lfr`
+
+*Outputs:*
+
+-   `data/derived/roosts.Rda`
+
+```{r}
+load("data/derived/removed_lfr.Rda")
 ```
@@
 # Get roosts for each season ----------------------------------------------
-roosts <- purrr::map(seasons_list_removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
 roosts <- roosts %>%
@@
 # Tidy up
-rm(seasons_list_removed_lfr)
+rm(removed_lfr)
 rm(roosts)
@@
 
+# Further processing for movement data
+
 ## Remove nighttime points
 
-START HERE
-
-XXX do we not also remove nighttime points for seasons_mode10?
-
-*Inputs:*
-
--   `seasons`
-
-*Outputs:*
-
--   `seasons`
+*Inputs:*
+
+-   `data/derived/removed_lfr.Rda`
+-   `data/derived/season_names.Rda`
+
+*Outputs:*
+
+-   `data/derived/removed_nightime.Rda`
+
+```{r}
+load("data/derived/removed_lfr.Rda")
+load("data/derived/season_names.Rda")
+```
 
@@
 # Remove nighttime points -------------------------------------------------
-before <- map_dbl(seasons, nrow)
-seasons <- map(seasons, ~{
+removed_nighttime <- map(removed_lfr, ~{
   times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
@@
 })
-after <- map_dbl(seasons, nrow)
-change <- after-before
-changeDF <- data.frame(seasonName = seasonNames_orig, nBefore = before, nAfter = after) %>%
-  mutate(propRowsRemoved = round((before-after)/before, 2))
-changeDF
-```
-
-## Restrict to Southern individuals
+
+save(removed_nighttime, file = "data/derived/removed_nighttime.Rda")
+
+# tidy up
+rm(removed_nighttime)
+rm(removed_lfr)
+rm(season_names)
+gc()
+```
+
+## Restrict to southern population
 
@@
 
-XXX why are we using `seasons_orig` here? shouldn't it be `seasons`?
-
-*Inputs:*
-
--   `seasons_orig`
-
-*Outputs:*
-
--   `data/derived/seasonsSF.Rda`
-
--   `seasons` XXX seems wrong!
-
-```{r}
-#| eval: false
-# Restrict to southern individuals ----------------------------------------
-# 
+*Inputs:*
+
+-   `data/derived/removed_nighttime.Rda`
+-   `data/derived/season_names.Rda`
+
+*Outputs:*
+
+-   `data/derived/removed_northern.Rda`
+
+```{r}
+load("data/derived/removed_nighttime.Rda")
+load("data/derived/season_names.Rda")
+```
+
+```{r}
+#| eval: false
 # Transform to SF object, so we can get centroids
-seasonsSF <- map(seasons_orig, ~.x %>%
+seasons_sf <- map(removed_nighttime, ~.x %>%
                    sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
@@
 
-save(seasonsSF, file = "data/derived/seasonsSF.Rda")
-load("data/derived/seasonsSF.Rda")
-
 ## Get centroids, so we can see who's "southern" for that season.
-centroids <- map(seasonsSF, ~.x %>%
+centroids <- map(seasons_sf, ~.x %>%
                    group_by(Nili_id) %>%
@@
 ## Get southern individuals for each season, so we can filter the data
-southernIndivs <- map(centroids, ~.x %>%
+southern_indivs <- map(centroids, ~.x %>%
                         filter(st_coordinates(.)[,2] < 3550000) %>%
@@
 ## Remove individuals not in the south
-before <- seasons
-seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
-after <- seasons
+before <- removed_nighttime
+removed_northern <- map2(.x = removed_nighttime, 
+                         .y = southern_indivs, ~.x 
+                         %>% filter(Nili_id %in% .y))
+after <- removed_northern
 
@@
 afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
-df <- data.frame(season = seasonNames_orig, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
+df <- data.frame(season = season_names, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
 df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
 df
+
+save(removed_northern, file = "data/derived/removed_northern.Rda")
+
+# tidy up
+rm(removed_nighttime)
+rm(season_names)
+rm(seasons_sf)
 rm(centroids)
-rm(seasonsSF)
+rm(southern_indivs)
+rm(before)
+rm(after)
+rm(removed_northern)
+rm(df)
+rm(afterIndivs)
+rm(afterRows)
+rm(beforeIndivs)
+rm(beforeRows)
+gc()
 ```
@@
 
--   `seasons`
-
-*Outputs:*
-
--   `seasons`
-
-```{r}
-#| eval: false
-# Include only individuals with enough points per day ------------------------
-# Investigate battery percentage vs points per day ------------------------
-# Is there some way we can relate battery charge percentage to number of points per day?
-test <- seasons[[5]]
-names(test)
-
-ppd <- map_dfr(seasons, ~.x %>%
+-   `data/derived/removed_northern.Rda`
+
+*Outputs:*
+
+-   `data/derived/removed_lowppd.Rda`
+
+```{r}
+load("data/derived/removed_northern.Rda")
+```
+
+```{r}
+#| eval: false
+# Calculate number of points per day for each individual
+ppd <- map_dfr(removed_northern, ~.x %>%
                  st_drop_geometry() %>%
@@
                            medBatt = mean(battery_charge_percent, na.rm = T)))
+
+# Sometimes, fewer points per day is due to low battery charge percent. Let's examine the relationship between battery charge and number of points per day.
+ppd %>%
+  filter(n < 100) %>% # only interested in lower numbers of points per day
+  ggplot(aes(x = minBatt, y = n))+
+  geom_point()+
+  facet_wrap(~seasonUnique)+
+  theme_classic()
+
+# What if we decide to exclude based on the combination of battery charge and points per day, with the minimum battery threshold at 50% and the points threshold at 10?
+
+# That is, if the minimum battery threshold was <= 50, a bird must have > 10 points on a given day for that day to be included.
+But if the minimum battery threshold was never less than 50%, it's okay to keep the ones with fewer points per day, since we know it was due to some reason other than battery charge.
 
@@
 
-# # cumulative histogram to look at the results of removing days based only on ppd
-# ppd %>%
-#   filter(n < 100) %>%
-#   ggplot(aes(x = n)) +
-#   stat_ecdf(aes(y = after_stat(y)*100))+
-#   ylab("Percentage of vulture-days included")+
-#   xlab("Points per day")+
-#   theme_classic()+
-#   facet_wrap(~seasonUnique)
-# 
-# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
-# ppd %>%
-#   filter(n < 100, minBatt > 50) %>%
-#   ggplot(aes(x = n)) +
-#   stat_ecdf(aes(y = after_stat(y)*100))+
-#   ylab("Percentage of vulture-days included")+
-#   xlab("Points per day")+
-#   theme_classic()+
-#   facet_wrap(~seasonUnique)
-# 
-# # same thing, but with batery set at 75
-# # cumulative histogram to look at results of removing days based on ppd after battery is taken into account (min pct < 50)
-# ppd %>%
-#   filter(n < 100, minBatt > 75) %>%
-#   ggplot(aes(x = n)) +
-#   stat_ecdf(aes(y = after_stat(y)*100))+
-#   ylab("Percentage of vulture-days included")+
-#   xlab("Points per day")+
-#   theme_classic()+
-#   facet_wrap(~seasonUnique)
-# 
-# # What if instead we decide we want a certain number of ppd--battery cuttoff may be sharper.
-Say, 30 ppd: is there a discernible battery cutoff that distinguishes "real" from "battery artifact" days?
-# ppd %>%
-#   filter(n < 30) %>%
-#   ggplot(aes(x = minBatt))+
-#   stat_ecdf(aes(y = after_stat(y)*100))+
-#   ylim(c(0, 100))+
-#   facet_wrap(~seasonUnique)+
-#   ylab("Percentage of vulture-days included")+
-#   xlab("Minimum battery percentage") # we do seem to see crooks here around 75%!
-# 
-# # Same thing but with 10 points per day? Predict much sharper cutoffs.
-# ppd %>%
-#   filter(n < 10) %>%
-#   ggplot(aes(x = minBatt))+
-#   stat_ecdf(aes(y = ..y..*100))+
-#   ylim(c(0, 100))+
-#   facet_wrap(~seasonUnique)+
-#   ylab("Percentage of vulture-days included")+
-#   xlab("Minimum battery percentage") # crooks around 50% battery minimum for 10ppd. This seems like a pretty good way to go!
-
-# Let's throw out anything with < 10ppd unless the battery has a min charge > 50%.
-
-# Orr suggests to just restrict to vulture*days with >30 points per day. Let's see what that would look like.
-thresh <- 30
-# ppd %>%
-#   mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
-#   ggplot(aes(x = n))+
-#   geom_histogram(aes(fill = Status))+
-#   facet_wrap(~seasonUnique, scales ="free")+
-#   geom_vline(aes(xintercept = thresh), col = "red")+
-#   scale_fill_manual(values = c("darkgray", "red"))+
-#   theme_classic()+
-#   ggtitle(paste("n removed =", thresh, sep = " "))+
-#   ylab("# vulture*days")+
-#   xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
-# 
-# ppd <- ppd %>%
-#   ungroup() %>%
-#   mutate(removed = ifelse(n < thresh, T, F))
-# 
-# # what proportion of vulture*days are removed overall?
-# (propRemoved <- ppd %>%
-#     group_by(seasonUnique, season) %>%
-#     summarize(nTot = n(),
-#               nRemoved = sum(removed),
-#               propRemoved = nRemoved/nTot))
-# propRemoved %>%
-#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
-#   geom_point(size = 10)+
-#   theme_classic()+
-#   scale_color_manual(values = c("blue", "red", "darkorange"))
-# 
-# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
-# 
-# # what about by individual over each season?
-# ppd %>%
-#   group_by(seasonUnique, season, Nili_id) %>%
-#   summarize(propRemoved = sum(removed)/n()) %>%
-#   ggplot(aes(x = seasonUnique, y = propRemoved))+
-#   geom_boxplot(aes(group = seasonUnique, fill = season))+
-#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
-#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
-#   theme_classic()+ # tells the same story
-#   ylab("Proportion of days removed")+
-#   xlab("")
-
-# Based on the battery charge analysis above, going to remove individuals that have 10 or fewer points per day, but only if the minimum battery charge is less than 50%.
-
 ## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
-seasons <- map(seasons, ~.x %>%
+removed_lowppd <- map(removed_northern, ~.x %>%
                  group_by(Nili_id, dateOnly) %>%
-                 filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>%
+                 filter(n() >= 10 | (n() < 10 & 
+                                       min(battery_charge_percent, 
+                                           na.rm = T) > 50)) %>%
                  ungroup())
+
+save(removed_lowppd, file = "data/derived/removed_lowppd.Rda")
+
+# tidying up
+rm(removed_northern)
+rm(ppd)
+rm(excl_batt)
+rm(excl_pts)
+rm(removed_lowppd)
+gc()
 ```
@@
 
-*Inputs:*
-
--   `seasons`
-
-*Outputs:*
-
--   `seasons`
-
-```{r}
-#| eval: false
+Now that we've removed days with too few points per day, we are left with only "valid" movement days for each individual.
+Now we want to exclude *individuals* (rather than individual-days) if the individual has too few valid days of movement data to accurately calculate movement metrics.
+
+Somewhat arbitrarily, we are removing individuals that are tracked for **fewer than 30 days in a given season.**
+
+*Inputs:*
+
+-   `data/derived/removed_lowppd.Rda`
+-   `data/derived/season_names.Rda`
+
+*Outputs:*
+
+-   `data/derived/removed_too_few_days.Rda`
+-   `data/derived/season_durations.Rda`
+
+```{r}
+load("data/derived/removed_lowppd.Rda")
+load("data/derived/season_names.Rda")
+```
+
+```{r}
+#| eval: false
+
+# How long are the seasons?
+season_durations <- map_dbl(removed_lowppd, ~max(.x$dateOnly)-min(.x$dateOnly))
+names(season_durations) <- season_names
+season_durations
+
 # Include only individuals with enough days tracked -----------------------
-## Must be tracked for at least 30 days per season
-durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))
-
-# walk2(seasons, durs, ~.x %>%
-#         st_drop_geometry() %>%
-#         group_by(Nili_id) %>%
-#         summarize(nDaysTracked = length(unique(dateOnly)),
-#                   propDaysTracked = nDaysTracked/.y) %>%
-#         pull(propDaysTracked) %>%
-#         hist())
-
-indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
+indivsToKeep <- map(removed_lowppd, ~.x %>%
                        st_drop_geometry() %>%
                        group_by(Nili_id) %>%
-                       summarize(nDaysTracked = length(unique(dateOnly)),
-                                 propDaysTracked = nDaysTracked/.y) %>%
+                       summarize(nDaysTracked = 
+                                   length(unique(dateOnly))) %>%
                        filter(nDaysTracked >= 30) %>%
@@
 ## remove those individuals not tracked for enough days
-before <- seasons
-after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))
-
-beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
-afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
-seasons <- after
+removed_too_few_days <- map2(.x = removed_lowppd, 
+                             .y = indivsToKeep, ~.x %>% 
+                filter(Nili_id %in% .y))
+
+beforeIndivs <- map_dbl(removed_lowppd, ~length(unique(.x$Nili_id)))
+afterIndivs <- map_dbl(removed_too_few_days, ~length(unique(.x$Nili_id)))
+
+save(removed_too_few_days, file = "data/derived/removed_too_few_days.Rda")
+save(season_durations, file = "data/derived/season_durations.Rda")
+
+# tidy up
+rm(removed_lowppd)
+rm(season_names)
+rm(season_durations)
+rm(indivsToKeep)
+rm(removed_too_few_days)
+rm(beforeIndivs)
+rm(afterIndivs)
+gc()
 ```
@@
 
--   `seasons`
-
-*Outputs:*
-
--   `data/derived/elevs_z10_seasons.Rda`
-
--   `data/derived/groundElev_z10.Rda`
-
--   `seasons`
-
-```{r}
-#| eval: false
+-   `data/derived/removed_too_few_days`
+
+*Outputs:*
+
+-   `data/derived/elevs_z10.Rda`
+
+-   `data/derived/with_altitudes.Rda`
+
+```{r}
+load("data/derived/removed_too_few_days.Rda")
+```
+
+```{r}
 # Get elevation rasters ---------------------------------------------------
-# Remove the first element of `seasons` because it doesn't have any data
-seasons <- seasons[-1] # removing summer 2020
-# # XXXXXXX start here
-#seasons <- map(seasons, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
-#elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x , z = 10))
-#save(elevs_z10_seasons, file = "data/derived/elevs_z10_seasons.Rda")
-load("data/derived/elevs_z10_seasons.Rda")
-
-# groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.
-# save(groundElev_z10, file = "data/derived/groundElev_z10.Rda")
-load("data/derived/groundElev_z10.Rda")
+sf <- map(removed_too_few_days, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
+elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
+save(elevs_z10, file = "data/derived/elevs_z10.Rda")
+```
+
+```{r}
+load("data/derived/elevs_z10.Rda")
+```
+
+```{r}
+#| eval: false
+groundElev_z10 <- map2(elevs_z10, sf, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.
 
 ## use elevation rasters to calculate height above ground level
-before <- seasons
-seasons <- map2(.x = seasons, .y = groundElev_z10, 
+with_altitudes <- map2(.x = removed_too_few_days, .y = groundElev_z10, 
                 ~.x %>% mutate(groundElev = .y) %>%
@@
                                                          TRUE ~ height_above_ground)))
-after <- seasons
 
 # There should be no change in the number of rows
-all(map2_lgl(before, after, ~nrow(.x) == nrow(.y)))
-```
-
-## Restrict to mode10 individuals
-
-XXX what is happening here
-
-*Inputs:*
-
--   `data/derived/toKeep_fixrate.Rda`
-
--   `seasons`
-
-*Outputs:*
-
--   `data/derived/seasons.Rda`
-
--   `data/derived/seasons_mode10.Rda`
-
-```{r}
-#| eval: false
-# Now restrict to the mode10 individuals ----------------------------------
-load("data/derived/toKeep_fixrate.Rda")
-toKeep_fixrate <- toKeep_fixrate[-1]
-seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
-
-save(seasons, file = "data/derived/seasons.Rda")
-save(seasons_mode10, file = "data/derived/seasons_mode10.Rda")
-load("data/derived/seasons.Rda")
-load("data/derived/seasons_mode10.Rda")
+all(map2_lgl(removed_too_few_days, with_altitudes, ~nrow(.x) == nrow(.y)))
+
+save(with_altitudes, file = "data/derived/with_altitudes.Rda")
+
+# tidy up
+rm(removed_too_few_days)
+rm(sf)
+rm(elevs_z10)
+rm(groundElev_z10)
+rm(with_altitudes)
 ```
@@
 
-Downsampling data to 10-minute intervals exclusively
-
-*Inputs:*
-
--   `seasons`
-
--   `seasons_mode10`
-
-*Outputs:*
-
--   `data/derived/seasons_10min.Rda`
-
--   `data/derived/seasons_mode10_10min.Rda`
-
-XXX this is confusing. Need to rename after concepts, not minutes.
-
-```{r}
-#| eval: false
-# Downsample the data and save the downsampled data
-subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
+We already removed individuals that had too low a fix rate.
+We are left with individuals that had a fix rate of 10 minutes or more frequent.
+But occasionally, there were periods of high-frequency sampling.
+This might mess up our movement calculations, so let's downsample the data so that any points that were sampled at a frequency much higher than 1 fix/10 minutes will be reduced to the same frequency as the others.
+
+*Inputs:*
+
+-   `data/derived/with_altitudes.Rda`
+
+-   `data/derived/season_names.Rda`
+
+*Outputs:*
+
+-   `data/derived/downsampled_10min.Rda`
+
+```{r}
+load("data/derived/with_altitudes.Rda")
+load("data/derived/season_names.Rda")
+```
+
+```{r}
+#| eval: false
+
+# Function to downsample the data
+subsample <- function(df, idCol = "Nili_id", 
+                      timestampCol = "timestamp", mins = 10){
   sub <- df %>%
@@
 # create and save downsampled datasets for later use
-seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
-seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
-
-map_dbl(seasons, nrow)
-map_dbl(seasons_10min, nrow)
-map_dbl(seasons_mode10, nrow)
-
-seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
-seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly))))
-
-save(seasons_10min, file = "data/derived/seasons_10min.Rda")
-save(seasons_mode10_10min, file = "data/derived/seasons_mode10_10min.Rda")
-```
+downsampled_10min <- map(with_altitudes, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
+
+# Add information on # of days tracked for each individual
+downsampled_10min <- map(downsampled_10min, ~.x %>% group_by(Nili_id) %>% 
+                       mutate(daysTracked = length(unique(dateOnly))))
+
+save(downsampled_10min, file = "data/derived/downsampled_10min.Rda")
+
+map_dbl(with_altitudes, nrow) %>% setNames(season_names)
+map_dbl(downsampled_10min, nrow) %>% setNames(season_names)
+
+# tidying
+rm(with_altitudes)
+rm(downsampled_10min)
+gc()
+```

### 2024-01-13 00.0_dataPrep.qmd
fixed file paths to point to the new data/dataPrep/ folder

@@
 
--   `data/derived/inpa.feather`
+-   `data/dataPrep/inpa.feather`
 
@@
   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
-write_feather(inpa, "data/derived/inpa.feather")
+write_feather(inpa, "data/dataPrep/inpa.feather")
 
@@
 
--   `data/derived/ornitela.feather`
+-   `data/``dataPrep``/ornitela.feather`
 
@@
 ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-write_feather(ornitela, "data/derived/ornitela.feather")
+write_feather(ornitela, "data/dataPrep/ornitela.feather")
 
@@
 
--   `data/derived/inpa.feather`
-
--   `data/derived/ornitela.feather`
-
-*Outputs:*
-
--   `data/derived/joined.feather`
+-   `data/dataPrep/inpa.feather`
+
+-   `data/dataPrep/ornitela.feather`
+
+*Outputs:*
+
+-   `data/dataPrep/joined.feather`
 
 ```{r load-inpa-ornitela}
-inpa <- read_feather("data/derived/inpa.feather")
-ornitela <- read_feather("data/derived/ornitela.feather")
+inpa <- read_feather("data/dataPrep/inpa.feather")
+ornitela <- read_feather("data/dataPrep/ornitela.feather")
 ```
@@
 length(unique(joined$Nili_id))
-write_feather(joined, "data/derived/joined.feather")
+write_feather(joined, "data/dataPrep/joined.feather")
 
@@
 
--   `data/derived/joined.feather`
-
-*Outputs:*
-
--   `data/derived/removed_periods.feather`
+-   `data/``dataPrep``/joined.feather`
+
+*Outputs:*
+
+-   `data/``dataPrep``/removed_periods.feather`
 
 ```{r load-hospital}
-joined <- read_feather("data/derived/joined.feather")
+joined <- read_feather("data/dataPrep/joined.feather")
 periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
@@
 
-write_feather(removed_periods, "data/derived/removed_periods.feather")
+write_feather(removed_periods, "data/dataPrep/removed_periods.feather")
 
@@
 
--   `data/derived/removed_periods.feather`
+-   `data/``dataPrep``/removed_periods.feather`
 
@@
 
--   `data/derived/roosts0.Rda`
-
--   `data/derived/removed_captures.feather`
-
-```{r}
-removed_periods <- read_feather("data/derived/removed_periods.feather")
+-   `data/``dataPrep``/roosts0.Rda`
+
+-   `data/``dataPrep``/removed_captures.feather`
+
+```{r}
+removed_periods <- read_feather("data/dataPrep/removed_periods.feather")
 ```
@@
 roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-save(roosts0, file = "data/derived/roosts0.Rda")
+save(roosts0, file = "data/dataPrep/roosts0.Rda")
 ```
@@
 capture_sites <- read.csv("data/raw/capture_sites.csv")
-load("data/derived/roosts0.Rda")
+load("data/dataPrep/roosts0.Rda")
 AllCarmelDates <- read.csv("data/raw/all_captures_carmel_2010-2021.csv")
@@
 
-write_feather(removed_captures, "data/derived/removed_captures.feather")
+write_feather(removed_captures, "data/dataPrep/removed_captures.feather")
 
@@
 
--   `data/derived/removed_captures.feather`
-
-*Outputs:*
-
--   `data/derived/with_age_sex.feather`
+-   `data/dataPrep/removed_captures.feather`
+
+*Outputs:*
+
+-   `data/dataPrep/with_age_sex.feather`
 
 ```{r attach-age-sex}
-removed_captures <- read_feather("data/derived/removed_captures.feather")
+removed_captures <- read_feather("data/dataPrep/removed_captures.feather")
 age_sex <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
@@
 
-write_feather(with_age_sex, "data/derived/with_age_sex.feather")
+write_feather(with_age_sex, "data/dataPrep/with_age_sex.feather")
 
@@
 
--   `data/derived/with_age_sex.feather`
-
-*Outputs:*
-
--   `data/derived/cleaned.feather`
+-   `data/dataPrep/with_age_sex.feather`
+
+*Outputs:*
+
+-   `data/dataPrep/cleaned.feather`
 
@@
 mask <- sf::st_read("data/raw/CutOffRegion.kml")
-with_age_sex <- read_feather("data/derived/with_age_sex.feather")
+with_age_sex <- read_feather("data/dataPrep/with_age_sex.feather")
 ```
@@
 
-save(cleaned, file = "data/derived/cleaned.Rda")
+save(cleaned, file = "data/dataPrep/cleaned.Rda")
 
@@
 
--   `data/derived/cleaned.Rda`
-
-*Outputs:*
-
--   `data/derived/seasons_list.Rda`
-
--   `data/derived/season_names.Rda`
-
-```{r}
-load("data/derived/cleaned.Rda")
+-   `data/dataPrep/cleaned.Rda`
+
+*Outputs:*
+
+-   `data/``dataPrep``/seasons_list.Rda`
+
+-   `data/``dataPrep``/season_names.Rda`
+
+```{r}
+load("data/dataPrep/cleaned.Rda")
 ```
@@
 season_names <- map_chr(seasons_list, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
-save(seasons_list, file = "data/derived/seasons_list.Rda")
-save(season_names, file = "data/derived/season_names.Rda")
+save(seasons_list, file = "data/dataPrep/seasons_list.Rda")
+save(season_names, file = "data/dataPrep/season_names.Rda")
 
@@
 
--   `data/derived/seasons_list.Rda`
-
-*Outputs:*
-
--   `data/derived/removed_lfr.Rda`
-
-```{r}
-load("data/derived/seasons_list.Rda")
+-   `data/dataPrep/seasons_list.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_lfr.Rda`
+
+```{r}
+load("data/dataPrep/seasons_list.Rda")
 ```
@@
 save(removed_lfr, 
-     file = "data/derived/removed_lfr.Rda")
+     file = "data/dataPrep/removed_lfr.Rda")
 
@@
 
--   `data/derived/removed_lfr`
-
-*Outputs:*
-
--   `data/derived/roosts.Rda`
-
-```{r}
-load("data/derived/removed_lfr.Rda")
+-   `data/dataPrep/removed_lfr`
+
+*Outputs:*
+
+-   `data/dataPrep/roosts.Rda`
+
+```{r}
+load("data/dataPrep/removed_lfr.Rda")
 ```
@@
                    remove = F))
-save(roosts, file = "data/derived/roosts.Rda")
+save(roosts, file = "data/dataPrep/roosts.Rda")
 
@@
 
--   `data/derived/removed_lfr.Rda`
--   `data/derived/season_names.Rda`
-
-*Outputs:*
-
--   `data/derived/removed_nightime.Rda`
-
-```{r}
-load("data/derived/removed_lfr.Rda")
-load("data/derived/season_names.Rda")
+-   `data/dataPrep/removed_lfr.Rda`
+-   `data/dataPrep/season_names.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_nightime.Rda`
+
+```{r}
+load("data/dataPrep/removed_lfr.Rda")
+load("data/dataPrep/season_names.Rda")
 ```
@@
 
-save(removed_nighttime, file = "data/derived/removed_nighttime.Rda")
+save(removed_nighttime, file = "data/dataPrep/removed_nighttime.Rda")
 
@@
 
--   `data/derived/removed_nighttime.Rda`
--   `data/derived/season_names.Rda`
-
-*Outputs:*
-
--   `data/derived/removed_northern.Rda`
-
-```{r}
-load("data/derived/removed_nighttime.Rda")
-load("data/derived/season_names.Rda")
+-   `data/dataPrep/removed_nighttime.Rda`
+-   `data/dataPrep/season_names.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_northern.Rda`
+
+```{r}
+load("data/dataPrep/removed_nighttime.Rda")
+load("data/dataPrep/season_names.Rda")
 ```
@@
 
-save(removed_northern, file = "data/derived/removed_northern.Rda")
+save(removed_northern, file = "data/dataPrep/removed_northern.Rda")
 
@@
 
--   `data/derived/removed_northern.Rda`
-
-*Outputs:*
-
--   `data/derived/removed_lowppd.Rda`
-
-```{r}
-load("data/derived/removed_northern.Rda")
+-   `data/dataPrep/removed_northern.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_lowppd.Rda`
+
+```{r}
+load("data/dataPrep/removed_northern.Rda")
 ```
@@
 
-save(removed_lowppd, file = "data/derived/removed_lowppd.Rda")
+save(removed_lowppd, file = "data/dataPrep/removed_lowppd.Rda")
 
@@
 
--   `data/derived/removed_lowppd.Rda`
--   `data/derived/season_names.Rda`
-
-*Outputs:*
-
--   `data/derived/removed_too_few_days.Rda`
--   `data/derived/season_durations.Rda`
-
-```{r}
-load("data/derived/removed_lowppd.Rda")
-load("data/derived/season_names.Rda")
+-   `data/dataPrep/removed_lowppd.Rda`
+-   `data/dataPrep/season_names.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_too_few_days.Rda`
+-   `data/dataPrep/season_durations.Rda`
+
+```{r}
+load("data/dataPrep/removed_lowppd.Rda")
+load("data/dataPrep/season_names.Rda")
 ```
@@
 
-save(removed_too_few_days, file = "data/derived/removed_too_few_days.Rda")
-save(season_durations, file = "data/derived/season_durations.Rda")
+save(removed_too_few_days, file = "data/dataPrep/removed_too_few_days.Rda")
+save(season_durations, file = "data/dataPrep/season_durations.Rda")
 
@@
 
--   `data/derived/removed_too_few_days`
-
-*Outputs:*
-
--   `data/derived/elevs_z10.Rda`
-
--   `data/derived/with_altitudes.Rda`
-
-```{r}
-load("data/derived/removed_too_few_days.Rda")
+-   `data/dataPrep/removed_too_few_days`
+
+*Outputs:*
+
+-   `data/dataPrep/elevs_z10.Rda`
+
+-   `data/dataPrep/with_altitudes.Rda`
+
+```{r}
+load("data/dataPrep/removed_too_few_days.Rda")
 ```
@@
 elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
-save(elevs_z10, file = "data/derived/elevs_z10.Rda")
-```
-
-```{r}
-load("data/derived/elevs_z10.Rda")
+save(elevs_z10, file = "data/dataPrep/elevs_z10.Rda")
+```
+
+```{r}
+load("data/dataPrep/elevs_z10.Rda")
 ```
@@
 
-save(with_altitudes, file = "data/derived/with_altitudes.Rda")
+save(with_altitudes, file = "data/dataPrep/with_altitudes.Rda")
 
@@
 
--   `data/derived/with_altitudes.Rda`
-
--   `data/derived/season_names.Rda`
-
-*Outputs:*
-
--   `data/derived/downsampled_10min.Rda`
-
-```{r}
-load("data/derived/with_altitudes.Rda")
-load("data/derived/season_names.Rda")
+-   `data/dataPrep/with_altitudes.Rda`
+
+-   `data/dataPrep/season_names.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/downsampled_10min.Rda`
+
+```{r}
+load("data/dataPrep/with_altitudes.Rda")
+load("data/dataPrep/season_names.Rda")
 ```
@@
 
-save(downsampled_10min, file = "data/derived/downsampled_10min.Rda")
+save(downsampled_10min, file = "data/dataPrep/downsampled_10min.Rda")
 

### 2024-01-17 00.0_dataPrep.qmd
More explanation, remove lfr individuals later on in the pipeline

@@
 
-## Remove individuals with low fix rate
+## Identify individuals with low fix rate
 
@@
 To do this, we will find the mode (most common) fix frequency for each individual on each day.
-Then, we will exclude individuals who never have observations at 10-minute intervals.
+Then, we will identify individuals who never have observations at 10-minute intervals.
 We will keep any that theoretically have a 10-minute fix rate (i.e.
@@
 
+Because tags can be put on and off and the fix rate can change, I'm calculating the fix rate on a per-season basis (which is why I didn't put this step earlier in the pipeline–it had to come after the season splitting).
+
+I suppose we could leave in the low fix rate individuals, especially for co-roosting.
+\[Actually, maybe we should leave them in for co-roosting, but not the others...
+which makes me think maybe we should leave them in for roost and social, just not movement, because they can contribute to the background information.
+Aaaarrrrrrggghhhhh....
+XXX\]
+
 *Inputs:*
@@
 
--   `data/dataPrep/removed_lfr.Rda`
+-   `data/dataPrep/low_fix_rate_indivs.Rda`
 
@@
 load("data/dataPrep/seasons_list.Rda")
+load("data/dataPrep/season_names.Rda")
 ```
@@
 
-# Exclude individuals that never have a daily mode of 10 minutes
-toRemove_lfr <- map(with_modes, ~.x %>%
+# Identify individuals that never have a daily mode of 10 minutes
+low_fix_rate_indivs <- map(with_modes, ~.x %>%
                         sf::st_drop_geometry() %>%
@@
                         unique())
-
-removed_lfr <- map2(seasons_list, toRemove_lfr, ~.x %>% 
-                                   filter(!(Nili_id %in% .y)))
-
-save(removed_lfr, 
-     file = "data/dataPrep/removed_lfr.Rda")
+names(low_fix_rate_indivs) <- season_names
+
+save(low_fix_rate_indivs, 
+     file = "data/dataPrep/low_fix_rate_indivs.Rda")
+# Don't need to save with_modes because we've already used this step to identify low fix rate individuals.
 
@@
 rm(seasons_list)
-rm(removed_lfr)
+rm(season_names)
+rm(low_fix_rate_indivs)
 rm(with_modes)
-rm(toRemove_lfr)
 gc()
@@
 But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric.
-So when we do the social analysis, we will use `removed_lfr` and not anything further along.
+So when we do the social analysis, we will use `seasons_list` and not anything further along.
 
@@
 
-We are going to need to get nightly roost locations for all of the remaining vultures.
+We are going to need to get nightly roost locations for all of the vultures.
 We have to do this now because getting roost locations requires using points that occur at night, which we will eventually want to remove in order to do the movement analysis.
@@
 
--   `data/dataPrep/removed_lfr`
+-   `data/dataPrep/seasons_list.Rda`
 
@@
 ```{r}
-load("data/dataPrep/removed_lfr.Rda")
+load("data/dataPrep/seasons_list.Rda")
 ```
@@
 # Get roosts for each season ----------------------------------------------
-roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts <- purrr::map(seasons_list, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
 roosts <- roosts %>%
@@
 # Tidy up
+rm(seasons_list)
+rm(roosts)
+gc()
+```
+
+# Further processing for movement analysis
+
+## Remove low fix rate individuals
+
+*Inputs:*
+
+-   `data/dataPrep/seasons_list.Rda`
+
+-   `data/dataPrep/low_fix_rate_indivs.Rda`
+
+```{r}
+load("data/dataPrep/seasons_list.Rda")
+load("data/dataPrep/low_fix_rate_indivs.Rda")
+```
+
+```{r}
+removed_lfr <- map2(seasons_list, low_fix_rate_indivs,
+                    ~.x %>% filter(!(Nili_id %in% .y)))
+save(removed_lfr, file = "data/dataPrep/removed_lfr.Rda")
+
+# tidy up
+rm(seasons_list)
+rm(low_fix_rate_indivs)
 rm(removed_lfr)
-rm(roosts)
-gc()
-```
-
-# Further processing for movement data
+gc()
+```
 
@@
 
-## Remove individuals with too few points per day
+## Remove vulture-days with too few points per day
 
@@
 
-## Remove individuals with too few days tracked
+## Remove individuals with too few high-ppd days tracked
 

### 2024-01-17 00.0_dataPrep.qmd
committed to this order for now

@@
 
-I suppose we could leave in the low fix rate individuals, especially for co-roosting.
-\[Actually, maybe we should leave them in for co-roosting, but not the others...
-which makes me think maybe we should leave them in for roost and social, just not movement, because they can contribute to the background information.
-Aaaarrrrrrggghhhhh....
-XXX\]
-
 *Inputs:*
@@
 ```{r}
+#| eval: false
 # Get elevation rasters ---------------------------------------------------
@@
 save(elevs_z10, file = "data/dataPrep/elevs_z10.Rda")
+save(sf, file = "data/dataPrep/sf.Rda")
 ```
@@
 load("data/dataPrep/elevs_z10.Rda")
+load("data/dataPrep/sf.Rda")
 ```
@@
 rm(with_altitudes)
+gc()
 ```

### 2024-01-17 scripts/01.0_calcMovement.R
updated with new-named files and re-ran, storing created intermediates in their own folder

@@ -31,64 +31,67 @@ distviz <- function(data, varname, seasoncol){
 }
 
 ## Load data ---------------------------------------------------------------
-base::load("data/derived/seasons_10min.Rda") # data rarefied to 10 minute intervals. Going to use that for everything.
-base::load("data/orphan/datasetAssignments.Rda")
-base::load("data/derived/roosts_seasons.Rda") # don't need a separate roost dataset for the rarefied data, because roosts are only once per day, and it makes sense to use the most detailed dataset possible.
-roosts_seasons <- map(roosts_seasons, ~.x %>% group_by(Nili_id) %>% 
+base::load("data/dataPrep/downsampled_10min.Rda") # data rarefied to 10 minute intervals. Going to use that for everything.
+base::load("data/dataPrep/season_names.Rda")
+base::load("data/orphan/datasetAssignments.Rda") # XXX START HERE--THIS IS 9 LONG AND NEEDS TO BE SHORTER
+datasetAssignments <- datasetAssignments[-1]
+base::load("data/dataPrep/roosts.Rda")
+roosts <- map(roosts, ~.x %>% group_by(Nili_id) %>% 
                         mutate(daysTracked = length(unique(roost_date))) %>% ungroup())
 
-roosts_seasons <- roosts_seasons[-1] # remove summer 2020
-datasetAssignments <- datasetAssignments[-1] # remove summer 2020
-test <- map2(roosts_seasons, datasetAssignments, ~left_join(.x, .y, by = "Nili_id"))
-seasonNames <- map_chr(seasons_10min, ~as.character(.x$seasonUnique[1]))
+test <- map2(roosts, datasetAssignments, ~left_join(.x, .y, by = "Nili_id"))
 roostPolygons <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
 
-durs <- map_dbl(seasons_10min, ~{
+durs <- map_dbl(downsampled_10min, ~{
   dates <- lubridate::ymd(.x$dateOnly)
   dur <- difftime(max(dates), min(dates), units = "days")
 })
 
-
 # Get number of days tracked for each individual --------------------------
-daysTracked_seasons <- map2(seasons_10min, durs, ~.x %>%
+daysTracked_seasons <- map2(downsampled_10min, durs, ~.x %>%
                               st_drop_geometry() %>%
                               group_by(Nili_id) %>%
[…]

### 2024-01-17 scripts/01.5_KDERarefaction.R
deprecated

@@ -1,139 +1,0 @@
-# Side quest: rarefaction plots for the KDE's
-library(tidyverse)
-library(adehabitatHR)
-library(sf)
-library(sp)
-library(raster)
-
-load("data/derived/hrList_indivs.Rda")
-load("data/derived/seasons.Rda")
-s <- map_chr(seasons, ~.x$seasonUnique[1])
-
-# For starters, let's see what href values we get when we run home range KUDs naively
-getH <- function(df){
-  sp <- SpatialPoints(df[,c("x", "y")])
-  k <- kernelUD(sp, h = "href", grid = 100, extent = 1)
-  h <- k@h$h
-}
-
-hs <- map_dbl(hrList_indivs[[3]], getH)
-hist(hs) # looks like our mean is a little over 5000. That gives me an idea of what ballpark to be in.
-summary(hs)
-
-# Some sample plots to visualize data
-## test one individual
-df <- hrList_indivs[[2]][[25]]
-sp <- SpatialPoints(df[,c("x", "y")])
-k <- kernelUD(sp, h = "href", grid = 100, extent = 1)
-j <- kernelUD(sp, h = 10000, grid = 100, extent = 1)
-
-image(k)
-poly <- getverticeshr(k, percent = 95) 
-plot(poly, add = TRUE)
-plot(sp, add = TRUE, cex = 0.01) # with image, this looks like a really bad representation, but with the 95% kernel it's a lot better.
-
-# okay, this is always going to be somewhat arbitrary. But maybe let's just go with 4000 for now?
-load("data/orphan/seasons_10min.Rda")
-load("data/orphan/seasons_20min.Rda")
-load("data/orphan/seasons_30min.Rda")
-load("data/orphan/seasons_60min.Rda")
-load("data/orphan/seasons_120min.Rda")
-
-# Function to calculate home range size
-calcHR <- function(df, pct, h = 4000, idCol = "Nili_id"){
-  id <- df[[idCol]][1]
-  sp <- SpatialPoints(df[,c("x", "y")])
-  k <- kernelUD(sp, h = h, grid = 100, extent = 1)
-  hr <- kernel.area(k, percent = pct)
-  n <- nrow(df)
-  return(data.frame(id = id, n = n, pct = pct, area = hr))
-}
-
-getHRs <- function(list){
-  cat("Preparing data\n")
-  indivs <- map(list, ~.x %>%
-                  dplyr::select(Nili_id) %>%
[…]

### 2024-01-17 scripts/03.0_PCA.R
re-ran just the PCA with the updated movement data

@@ -15,15 +15,11 @@ roostingColor = "brown1";
 flightColor = "skyblue3";
 
 ## Load data ------------------------------------------------------------
-load("data/derived/movementBehavior.Rda")
-load("data/derived/movementBehaviorScaled.Rda")
-load("data/derived/networkMetrics.Rda")
-seasonNames <- map_chr(movementBehavior, ~.x$seasonUnique[1])
-all(networkMetrics$season %in% seasonNames) # check that season names match up--good.
-load("data/derived/seasons_10min.Rda")
-
-# The summer season is mid-May through mid-September, but in 2020 we only have data starting September 4th. So I'm going to exclude summer 2020, since I don't think it's well represented.
-movementBehavior <- movementBehavior[seasonNames != "2020_summer"]
+load("data/calcMovement/movementBehavior.Rda")
+load("data/calcMovement/movementBehaviorScaled.Rda")
+### load("data/calcSocial/networkMetrics.Rda") #XXX
+load("data/dataPrep/season_names.Rda")
+load("data/derived/seasons_10min.Rda") # XXX
 
 # Create a dataset with all seasons combined
 allMovementBehavior <- purrr::list_rbind(movementBehavior)
@@ -117,7 +113,7 @@ mvmt <- allMovementBehavior %>%
   dplyr::select(Nili_id, birth_year, sex, seasonUnique, coreArea, homeRange, coreAreaFidelity, mnDailyMaxAlt, mnDailyMedAlt, propSwitch, shannon, meanDFD, meanDMD, meanDDT, mnTort) %>%
   left_join(mvmtPCVals, by = c("Nili_id", "seasonUnique"))
 
-# COMBINE movement and social data ---------------------------------------------
+# COMBINE movement and social data # XXX waiting for social ---------------------------------------------
 linked <- networkMetrics %>%
   left_join(mvmt, by = c("Nili_id", "season" = "seasonUnique")) %>%
   mutate(year = factor(as.numeric(str_extract(season, "[0-9]{4}"))),

### 2024-01-17 scripts/01.0_calcMovement.R
removed vars at the end

@@ -380,3 +380,39 @@ movementBehaviorScaled <- map(movementBehavior, ~.x %>%
 ## Save raw movement data ------------------------------------------------------
 save(movementBehavior, file = "data/calcMovement/movementBehavior.Rda")
 save(movementBehaviorScaled, file = "data/calcMovement/movementBehaviorScaled.Rda")
+
+# tidy up
+rm(ageSex)
+rm(areas)
+rm(cc)
+rm(coreAreas_indivs)
+rm(dailyAltitudes)
+rm(dailyAltitudesSumm)
+rm(dailyMovementList_seasons)
+rm(datasetAssignments)
+rm(daysTracked_seasons)
+rm(dfd)
+rm(dfddf)
+rm(dfdSumm)
+rm(downsampled_10min)
+rm(downsampled_10min_sf)
+rm(dts)
+rm(homeRanges_indivs)
+rm(hrList_indivs)
+rm(hrList_indivs_SP)
+rm(ikuda_df)
+rm(indivs)
+rm(indsKUDAreas)
+rm(kuds_indivs)
+rm(mbdf)
+rm(mm)
+rm(mnMvmt)
+rm(movementBehavior)
+rm(movementBehaviorScaled)
+rm(roostIDs)
+rm(roostPolygons)
+rm(roosts)
+rm(roostSequences)
+rm(roostSwitches)
+rm(shannon)
+rm(test)

### 2024-01-19 00.0_dataPrep.qmd
moved the low fix rate removal earlier in the process

@@
 
--   `data/``dataPrep``/ornitela.feather`
+-   `data/dataPrep/ornitela.feather`
 
@@
 
--   `data/``dataPrep``/joined.feather`
-
-*Outputs:*
-
--   `data/``dataPrep``/removed_periods.feather`
+-   `data/dataPrep/joined.feather`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_periods.feather`
 
@@
 
--   `data/``dataPrep``/removed_periods.feather`
+-   `data/dataPrep/removed_periods.feather`
 
@@
 
--   `data/``dataPrep``/roosts0.Rda`
-
--   `data/``dataPrep``/removed_captures.feather`
+-   `data/dataPrep/roosts0.Rda`
+
+-   `data/dataPrep/removed_captures.feather`
 
@@
 
--   `data/``dataPrep``/seasons_list.Rda`
-
--   `data/``dataPrep``/season_names.Rda`
+-   `data/dataPrep/seasons_list.Rda`
+
+-   `data/dataPrep/season_names.Rda`
 
@@
 
+## Remove low fix rate individuals
+
+*Inputs:*
+
+-   `data/dataPrep/seasons_list.Rda`
+
+-   `data/dataPrep/low_fix_rate_indivs.Rda`
+
+```{r}
+load("data/dataPrep/seasons_list.Rda") 
+load("data/dataPrep/low_fix_rate_indivs.Rda")
+```
+
+```{r}
+removed_lfr <- map2(seasons_list, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
+save(removed_lfr, file = "data/dataPrep/removed_lfr.Rda")  
+# tidy up 
+rm(seasons_list) 
+rm(low_fix_rate_indivs) 
+rm(removed_lfr) 
+gc()
+```
+
 ### Data for social analysis
@@
 ```{r}
-load("data/dataPrep/seasons_list.Rda")
+load("data/dataPrep/removed_lfr.Rda")
 ```
@@
 # Get roosts for each season ----------------------------------------------
-roosts <- purrr::map(seasons_list, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
 roosts <- roosts %>%
@@
 # Tidy up
-rm(seasons_list)
+rm(removed_lfr)
 rm(roosts)
@@
 # Further processing for movement analysis
-
-## Remove low fix rate individuals
-
-*Inputs:*
-
--   `data/dataPrep/seasons_list.Rda`
-
--   `data/dataPrep/low_fix_rate_indivs.Rda`
-
-```{r}
-load("data/dataPrep/seasons_list.Rda")
-load("data/dataPrep/low_fix_rate_indivs.Rda")
-```
-
-```{r}
-removed_lfr <- map2(seasons_list, low_fix_rate_indivs,
-                    ~.x %>% filter(!(Nili_id %in% .y)))
-save(removed_lfr, file = "data/dataPrep/removed_lfr.Rda")
-
-# tidy up
-rm(seasons_list)
-rm(low_fix_rate_indivs)
-rm(removed_lfr)
-gc()
-```
 

### 2024-01-25 testinterpolation.R
a hopefully-working interpolation pipeline for one season of data

@@ -1,0 +1,130 @@
+load("data/dataPrep/removed_lfr.Rda")
+library(tidyverse)
+library(sf)
+
+interpFun <- function(df, season){
+  rowsToAdd <- data.frame()
+  whichGap <- which(df$gap)
+  n <- length(whichGap)
+  for(i in whichGap){
+    cat(paste0(season, ": Interpolating gap ", which(whichGap == i), " of ", n,
+               "\n"))
+    Nili_id <- df$Nili_id[i]
+    start <- lubridate::ymd_hms(df$timestamp[i])
+    end <- lubridate::ymd_hms(df$timestamp[i+1])
+    timediff <- as.numeric(difftime(end, start, units = "mins"))
+    nPeriods <- ceiling(timediff/10)
+    nTimes <- nPeriods - 1
+    times <- seq.POSIXt(from = start, to = end, length.out = 2+nTimes)
+    times <- times[-c(1, length(times))] # remove the anchor times because those are already in the dataset
+    newrows <- data.frame(Nili_id = Nili_id,
+                          timestamp = times)
+    rowsToAdd <- bind_rows(rowsToAdd, newrows)
+  }
+  return(rowsToAdd)
+}
+
+# JUST ONE SEASON OF DATA
+season <- removed_lfr[[2]]
+pp <- season %>%
+  select(Nili_id, dateOnly, timestamp, location_long, location_lat) %>%
+  sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = FALSE) %>%
+  arrange(Nili_id, dateOnly, timestamp) %>%
+  group_by(Nili_id, dateOnly) %>%
+  mutate(timeToNext = as.numeric(difftime(lead(timestamp), timestamp, units = "mins")),
+         distToNext = sf::st_distance(geometry, geometry[row_number()+1],
+                                      by_element = T),
+         distToNext = as.numeric(distToNext))
+
+gp <- pp %>%
+  mutate(gap = case_when(timeToNext > 10 &
+                           timeToNext <= 60 &
+                           !is.na(timeToNext) &
+                           !is.na(lead(timeToNext))&
+                           distToNext <= 100
+                         ~ TRUE,
+                         TRUE ~ FALSE))
+whichGap <- which(gp$gap)
+
+toInterpolate <- interpFun(gp, season = "test")
+
+out <- gp %>% 
[…]

### 2024-01-25 scripts/test.R
obsolete

@@ -1,25 +1,0 @@
-# Frequentist models parallel to Bayesian models
-# 1. degree ~ PC1
-# 2. degree ~ situ + PC1
-# 3. degree ~ situ + PC1 + PC2
-# 4. degree ~ situ + PC1 + PC2 + age_group
-
-library(tidyverse)
-library(lme4)
-load("data/derived/forModeling.Rda")
-
-# 1.
-mod1 <- lm(degree_scl ~ PC1, data = forModeling)
-summary(mod1)
-
-# 2.
-mod2 <- lm(degree_scl ~ situ + PC1, data = forModeling)
-summary(mod2)
-
-# 3.
-mod3 <- lm(degree_scl ~ situ + PC1 + PC2, data = forModeling)
-summary(mod3)
-
-# 4.
-mod4 <- lm(degree_scl ~ situ + PC1 + PC2 + age_group, data = forModeling)
-summary(mod4)

### 2024-01-25 testinterpolation.R
full example of interpolation with one season of data, showing that interpolation (at least at the level proposed here) does NOT change the number of individuals eventually retained.

@@ -123,8 +123,58 @@ gg <- ff %>%
 length(tokeep)
 length(unique(ff$Nili_id)) # retained 28 out of 33 individuals
 
+df <- data.frame(rows = c(nrow(bb), nrow(cc), nrow(dd), nrow(ee), nrow(ff), nrow(gg)),
+                 indivs = c(length(unique(bb$Nili_id)), length(unique(cc$Nili_id)), length(unique(dd$Nili_id)), length(unique(ee$Nili_id)), length(unique(ff$Nili_id)), length(unique(gg$Nili_id))),
+                 step = c("1_after removing lfr indivs", "2_after interpolating", "3_after removing nighttime points", "4_after removing southern indivs", "5_after removing days with too few points", "6_after removing indivs not tracked for enough days"),
+                 group = "interpolated")
+
+df %>%
+  ggplot(aes(x = step, y = indivs))+
+  geom_point()+
+  geom_line(aes(group = group))+
+  scale_x_discrete(labels = label_wrap_gen(width = 20))+
+  theme_classic()
+
+  
 ####################
 # Now compare this result to how many individuals were retained in the real dataset
+load("data/dataPrep/removed_lfr.Rda")
+BB <- removed_lfr[[2]]
+load("data/dataPrep/removed_nighttime.Rda")
+DD <- removed_nighttime[[2]]
+rm(removed_nighttime)
+load("data/dataPrep/removed_northern.Rda")
+EE <- removed_northern[[2]]
+rm(removed_northern)
+load("data/dataPrep/removed_lowppd.Rda")
+FF <- removed_lowppd[[2]]
+rm(removed_lowppd)
 load("data/dataPrep/removed_too_few_days.Rda")
-forComparison <- removed_too_few_days[[2]]
+GG <- removed_too_few_days[[2]]
+rm(removed_too_few_days)
+
+DF <- data.frame(rows = c(nrow(BB), NA, nrow(DD), nrow(EE), nrow(FF), nrow(GG)),
+                 indivs = c(length(unique(BB$Nili_id)), NA, length(unique(DD$Nili_id)), length(unique(EE$Nili_id)), length(unique(FF$Nili_id)), length(unique(GG$Nili_id))),
+                 step = c("1_after removing lfr indivs", "2_after interpolating", "3_after removing nighttime points", "4_after removing southern indivs", "5_after removing days with too few points", "6_after removing indivs not tracked for enough days"),
[…]

### 2024-01-25 00.0_dataPrep.qmd
updated to new who's who, and added a new season of data

@@
 ```{r download-inpa}
-#| eval: false
 inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
@@
          year = as.numeric(lubridate::year(timestamp))) %>%
-  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
+  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
 write_feather(inpa, "data/dataPrep/inpa.feather")
 
-length(unique(inpa$trackId)) # 77
+length(unique(inpa$trackId))
 
@@
 ```{r download-ornitela}
-#| eval: false
 # Ornitela: download data from movebank (just a subset of the times for now)
 minDate <- "2020-09-01 00:00"
-maxDate <- "2023-05-15 11:59"
+maxDate <- "2023-09-15 11:59"
 ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
@@
 
-length(unique(ornitela$trackId)) # 127
+length(unique(ornitela$trackId))
 
@@
 # Add the Nili_id
-ww <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+ww <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
   dplyr::select(Nili_id, Movebank_id) %>%
@@
 
--   `data/raw/whoswho_vultures_20230315_new.xlsx`
+-   `data/raw/whoswho_vultures_20230920_new.xlsx`
 
@@
 joined <- read_feather("data/dataPrep/joined.feather")
-periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")
+periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove")
 ```
@@
 removal_annotated <- joined %>%
-  left_join(toRemove, by = c("Nili_id", "dateOnly")) %>%
+  left_join(periods_to_remove, by = c("Nili_id", "dateOnly")) %>%
   mutate(status = replace_na(status, "valid"))
@@
 save(roosts0, file = "data/dataPrep/roosts0.Rda")
+
+# tidy up
+rm(roosts0)
 ```
@@
 ```{r attach-age-sex}
+#| message: false
 removed_captures <- read_feather("data/dataPrep/removed_captures.feather")
-age_sex <- read_excel("data/raw/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+age_sex <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
   dplyr::select(Nili_id, birth_year, sex) %>%
@@
   left_join(age_sex, by = "Nili_id")
-nrow(with_age_sex) == nrow(removed_captures)
-length(unique(with_age_sex$Nili_id)) # 196
+nrow(with_age_sex) == nrow(removed_captures) # should be TRUE
+length(unique(with_age_sex$Nili_id)) #201
 
@@
          month = lubridate::month(dateOnly),
-         year = lubridate::year(dateOnly))
+         year = lubridate::year(dateOnly)) # XXX consider removing the downsample argument. Also this function is insanely slow and needs to be fixed, good grief.
 
@@
            dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
-  filter(dateOnly != "2023-05-15") # remove May 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
+  filter(dateOnly != "2023-09-15") # remove September 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
 
@@
   # explicitly set the factor levels in temporal order
-  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding")),
+  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding", "2023_summer")),
          season = factor(season, levels = c("breeding", "summer", "fall")))
@@
   filter(seasonUnique != "2020_summer")
-length(unique(with_seasons$seasonUnique))
+length(unique(with_seasons$seasonUnique)) # should be 9
 
@@
   group_split(.keep = T)
-length(seasons_list) # should be 8
+length(seasons_list) # should be 9
 # extract the season names in case we need a separate vector at some point.
 season_names <- map_chr(seasons_list, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
+length(season_names) # should also be 9
 save(seasons_list, file = "data/dataPrep/seasons_list.Rda")
@@
 -   `data/dataPrep/seasons_list.Rda`
+-   `data/dataPrep/season_names.Rda`
 
@@
 #| eval: false
-# Optionally remove individuals with a low fix rate -----------------------
 # function to calculate mode (most common) fix rate
@@
 
--   `data/dataPrep/seasons_list.Rda`
+-   `data/dataPrep/removed_lfr.Rda`
 
@@
     dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
-    dplyr::select(-c(sunrise, sunset)) %>%
     dplyr::filter(daylight == "day")
@@
                  summarize(n = n(),
-                           minBatt = min(battery_charge_percent, na.rm = T),
-                           meanBatt = mean(battery_charge_percent, na.rm = T),
-                           medBatt = mean(battery_charge_percent, na.rm = T)))
+                           minBatt = min(battery_charge_percent, na.rm = T)))
 

### 2024-02-03 scripts/00.0_dataPrep_copy.qmd
script for Cassidy--reduced version of the dataPrep qmd to be suitable for 1 year of data

--- 2024-01-25 00.0_dataPrep.qmd
@@
 library(elevatr)
+library(here) # for file paths
 ```
@@
 
-```{r}
-base::load("movebankCredentials/pw.Rda")
-MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
-rm(pw)
-```
+\[code omitted\]
 
@@
 
-Data collected by Ohad Hatzofe and team with GPS loggers deployed for their study.
-
-Inputs:
-
--   `MB.LoginObject`
-
-Outputs:
-
--   `data/dataPrep/inpa.feather`
-
-```{r download-inpa}
-inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
-inpa <- methods::as(inpa, "data.frame")
-inpa <- inpa %>%
-  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
-         year = as.numeric(lubridate::year(timestamp))) %>%
-  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
-write_feather(inpa, "data/dataPrep/inpa.feather")
-
-length(unique(inpa$trackId))
-
-# tidy up
-rm(inpa)
-gc()
-```
+\[code omitted\]
 
@@
 
-Inputs:
-
--   `MB.LoginObject`
-
-Outputs:
-
--   `data/dataPrep/ornitela.feather`
-
-```{r download-ornitela}
-# Ornitela: download data from movebank (just a subset of the times for now)
-minDate <- "2020-09-01 00:00"
-maxDate <- "2023-09-15 11:59"
-ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-write_feather(ornitela, "data/dataPrep/ornitela.feather")
-
-length(unique(ornitela$trackId))
-
-# tidy up
-rm(ornitela)
-rm(minDate)
-rm(maxDate)
-gc()
-```
+\[code omitted\]
 
@@
 
-*Inputs:*
-
--   `data/dataPrep/inpa.feather`
-
--   `data/dataPrep/ornitela.feather`
-
-*Outputs:*
-
--   `data/dataPrep/joined.feather`
-
-```{r load-inpa-ornitela}
-inpa <- read_feather("data/dataPrep/inpa.feather")
-ornitela <- read_feather("data/dataPrep/ornitela.feather")
-```
-
-```{r join-inpa-ornitela}
-#| eval: false
-
-# Will there be a problem joining these?
-names(ornitela)[!(names(ornitela) %in% names(inpa))] # all names in "ornitela" are found in inpa
-names(inpa)[!(names(inpa) %in% names(ornitela))] # many are not found in "ornitela" that are found in inpa
-
-# Add the dataset names so we can keep track of where the data comes from
-inpa_tojoin <- inpa[,names(ornitela)] %>%
-  mutate(dataset = "inpa")
-ornitela <- ornitela %>%
-  mutate(dataset = "ornitela")
-
-# Join the two datasets
-joined0 <- bind_rows(inpa_tojoin, ornitela)
-dim(joined0)
-
-rm(inpa)
-rm(inpa_tojoin)
-rm(ornitela)
-
-## fix trackId
-joined0 <- joined0 %>%
-  mutate(trackId = as.character(trackId),
-         trackId = case_when(trackId == "E03" ~ "E03w",
-                             TRUE ~ trackId))
-
-# number of unique individuals
-length(unique(joined0$trackId))
-
-# Add the Nili_id
-ww <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
-  dplyr::select(Nili_id, Movebank_id) %>%
-  distinct()
-
-# Are there any other individuals who have two different associated trackId's?
-ww %>% filter(Movebank_id %in% joined0$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
-ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
-
-all(joined0$trackId %in% ww$Movebank_id) # false.
-
-# Okay, let's investigate the ones that aren't included
-joined0 %>%
-  filter(!(trackId %in% ww$Movebank_id)) %>%
-  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
-
-joined0$trackId <- str_replace_all(tolower(joined0$trackId), "\\.", "")
-ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
-
-sort(unique(ww$mb_id))
-unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
-joined0$trackId[joined0$trackId == "y01t60w"] <- "t60white"
-joined0$trackId[joined0$trackId == "e86white"] <- "e86"
-unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
-# Investigate the remaining ones:
-joined0 %>% filter(!(trackId %in% ww$mb_id)) %>% 
-  dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>%
-  group_by(trackId, local_identifier, ring_id, dataset) %>% 
-  summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
-# I guess for now let's just make their trackId's into Nili_id's
-
-joined <- left_join(joined0, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
-                  by = c("trackId" = "mb_id"))
-nrow(joined) == nrow(joined0) # good, same number of rows. (check that the join worked)
-length(unique(joined$Nili_id))
-# For any that didn't get Nili_ids assigned, make it the trackId
-joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expected.
-joined <- joined %>%
-  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
-                             TRUE ~ Nili_id))
-length(unique(joined$Nili_id))
-write_feather(joined, "data/dataPrep/joined.feather")
-
-# tidy up
-rm(joined0)
-rm(joined)
-gc()
+\[code omitted\]
+
+## (for Cassidy only): Filter to just 1 year of data (2022)
+
+```{r}
+# joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
+# year22 <- joined %>%
+#   filter(lubridate::year(dateOnly) == "2022")
+write_csv(year22, file = here::here("data", "dataPrep", "year22.csv"))
 ```
@@
 
--   `data/dataPrep/joined.feather`
+-   `data/dataPrep/year22.csv`
 
@@
 ```{r load-hospital}
-joined <- read_feather("data/dataPrep/joined.feather")
+year22 <- read_csv("data/dataPrep/year22.csv")
 periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove")
@@
 # Join to the original data
-removal_annotated <- joined %>%
+removal_annotated <- year22 %>%
   left_join(periods_to_remove, by = c("Nili_id", "dateOnly")) %>%
   mutate(status = replace_na(status, "valid"))
-nrow(removal_annotated) == nrow(joined) #T: good, same number of rows.
+nrow(removal_annotated) == nrow(year22) #T: good, same number of rows.
 
@@
 length(unique(removal_annotated$Nili_id))
-nrow(joined) == nrow(removal_annotated) # true
+nrow(year22) == nrow(removal_annotated) # true
 nrow(removed_periods) != nrow(removal_annotated) # true
@@
 rm(periods_to_remove)
-rm(joined)
+rm(year22)
 rm(removal_annotated)

### 2024-02-03 scripts/03.0_PCA.R
various small changes?

@@ -19,7 +19,6 @@ load("data/calcMovement/movementBehavior.Rda")
 load("data/calcMovement/movementBehaviorScaled.Rda")
 ### load("data/calcSocial/networkMetrics.Rda") #XXX
 load("data/dataPrep/season_names.Rda")
-load("data/derived/seasons_10min.Rda") # XXX
 
 # Create a dataset with all seasons combined
 allMovementBehavior <- purrr::list_rbind(movementBehavior)
@@ -100,6 +99,7 @@ ggsave(biplot, filename = "fig/biplot.png", width = 6, height = 4)
 contrib <- round(pca_all$rotation[,1:3]*100, 2) %>%
   as.data.frame()
 contrib
+save(contrib, file = "data/PCA/contrib.Rda")
 
 fviz_screeplot(pca_all, addLabels = T)+
   ggtitle("Overall PCA: Sep 2020-Nov 2022")
@@ -113,7 +113,7 @@ mvmt <- allMovementBehavior %>%
   dplyr::select(Nili_id, birth_year, sex, seasonUnique, coreArea, homeRange, coreAreaFidelity, mnDailyMaxAlt, mnDailyMedAlt, propSwitch, shannon, meanDFD, meanDMD, meanDDT, mnTort) %>%
   left_join(mvmtPCVals, by = c("Nili_id", "seasonUnique"))
 
-# COMBINE movement and social data # XXX waiting for social ---------------------------------------------
+# COMBINE movement and social data # XXX waiting for social, start here ---------------------------------------------
 linked <- networkMetrics %>%
   left_join(mvmt, by = c("Nili_id", "season" = "seasonUnique")) %>%
   mutate(year = factor(as.numeric(str_extract(season, "[0-9]{4}"))),

### 2024-02-03 scripts/01.1_categorizeMovement.qmd

---
title: "More straightforward movement variable definitions"
execute: 
  echo: false
format: 
  html:
    embed-resources: true
    code-fold: true
    code-summary: "Show code"
editor: visual
---

```{r setup}
#| include: false
library(tidyverse)
library(sf)
library(factoextra)
library(ade4)
load(here::here("data", "calcMovement", "movementBehavior.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
load(here::here("data", "derived", "cc.Rda"))

# retrieve age information, since it wasn't carried through in the movement analysis (oops)
load(here::here("data", "dataPrep", "downsampled_10min.Rda"))
ages <- map(downsampled_10min, ~.x %>%
              dplyr::select(Nili_id, seasonUnique, season, year, age, age_group) %>%
              distinct()) %>%
  purrr::list_rbind()
rm(downsampled_10min) # XXX need to do this in a faster way--attach age at an earlier stage probably

# Create a dataset with all seasons combined
allMovementBehavior <- purrr::list_rbind(movementBehavior)
movementBehaviorScaled <- allMovementBehavior %>%
  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex, dataset, 
                   homeRange, coreArea, coreAreaFidelity), # not scaling the space use params, because I'm not putting those into a PCA.
                function(x){as.numeric(as.vector(scale(x)))}))
```

## Categorize the movement behaviors, and reduce the number of movement behaviors that we're including

```{r}
#| include: false
demo <- movementBehaviorScaled %>%
  dplyr::select(Nili_id, seasonUnique, daysTracked, propDaysTracked, dataset, birth_year, sex)
space_use <- movementBehaviorScaled %>%
  dplyr::select(Nili_id, seasonUnique, coreArea, homeRange, coreAreaFidelity)
movement <- movementBehaviorScaled %>%
  dplyr::select(Nili_id, seasonUnique, mnDailyMaxAlt, mnDailyMedAlt, meanDFD, maxDFD, meanDMD, meanDDT, mnTort)
roost_behavior <- movementBehaviorScaled %>%
  dplyr::select(Nili_id, seasonUnique, propSwitch, shannon, uniqueRoosts)
```

I first thought to do separate PCA's of each of these categories. But that doesn't make sense, especially for something like space use, where there are really only two metrics (and then one that's the combination of each of them)

Instead, I want to make these more biologically motivated. Let's do the following:

1\) home range only (but make it using AKDE method so it's actually defensible)

2\) PCA of movement behaviors only

3\) PCA of roost behaviors

## Movement PCA

```{r}
pca_movement <- prcomp(movement[,-c(1:2)])
fviz_pca_biplot(pca_movement, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_movement, addLabels = T) # that's pretty strongly loaded on the first principal component, so I think I can justify just using that as an overall measure of movement. To double check, let's see the contributions:
contrib_movement <- round(pca_movement$rotation[,1:3]*-100, 2) %>%
  as.data.frame() 
contrib_movement # Okay, PC1 does seem like it can be a general measure of movement.
movement <- movement %>%
  mutate(movement_pc1 = pca_movement$x[,1]*-1)
```

### Interpretation

```{r}
#| warning: false
#| message: false

movement %>%
  pivot_longer(cols = c("mnDailyMaxAlt", "mnDailyMedAlt", "meanDFD", "maxDFD", "meanDMD", "meanDDT", "mnTort"),
               names_to = "measure", values_to = "value") %>%
  mutate(measure = fct_recode(measure,
                              "Daily max. altitude (seasonal avg)" =
                                "mnDailyMaxAlt",
                              "Daily median altitude (seasonal avg)" = 
                                "mnDailyMedAlt",
                              "Daily flight time (seasonal avg)" = 
                                "meanDFD",
                              "Daily flight time (seasonal max)" = 
                                "maxDFD",
                              "Daily max displacement (seasonal avg)" = 
                                "meanDMD",
                              "Daily distance traveled (seasonal avg)" = 
                                "meanDDT",
                              "Tortuosity (seasonal avg)" = 
                                "mnTort")) %>%
  ggplot(aes(x = movement_pc1, y = value))+
  geom_point()+geom_smooth(method = "lm")+theme_minimal()+
  facet_wrap(~measure, labeller = labeller(measure = label_wrap_gen(30)))+
  ylab("Value (scaled)")+
  theme(legend.position = "none")
```

Individuals with higher movement PC1 values fly higher and farther, spend more time flying, and have straighter paths than individuals with lower movement PC1 values. Let's call this measure just "movement".

## Roost behavior PCA

```{r}
pca_roost <- prcomp(roost_behavior[,-c(1:2)])
fviz_pca_biplot(pca_roost, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_roost, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
contrib_roost <- round(pca_roost$rotation[,1:3]*-100, 2) %>%
  as.data.frame()
contrib_roost
roost_behavior <- roost_behavior %>%
  mutate(roost_pc1 = pca_roost$x[,1]*-1)
```

### Interpretation

```{r}
#| warning: false
#| message: false

roost_behavior %>%
  rename("Shannon roost diversity" = "shannon", "Unique roosts" = "uniqueRoosts", "Prop. nights roost-switching" = "propSwitch") %>%
  pivot_longer(cols = c("Shannon roost diversity", "Unique roosts", "Prop. nights roost-switching"),
               names_to = "measure", values_to = "value") %>%
  ggplot(aes(x = roost_pc1, y = value))+
  geom_point()+geom_smooth(method = "lm")+theme_minimal()+
  facet_wrap(~measure)+
  ylab("Value (scaled)")+
  theme(legend.position = "none")
```

Individuals that are higher on the roost PC1 have more unique roosts, a higher roost diversity, and switch roosts more often. Let's call this a measure of "**roost site diversification**" or something like that.

## Variable distributions

```{r}
#| message: false
#| warning: false
#| include: false
# Create a reduced dataset that only includes our measures of interest
red <- demo %>%
  left_join(space_use %>% dplyr::select(Nili_id, seasonUnique, homeRange)) %>%
  left_join(movement %>% dplyr::select(Nili_id, seasonUnique, movement_pc1)) %>%
  left_join(roost_behavior %>% dplyr::select(Nili_id, seasonUnique, roost_pc1)) %>%
  rename("space_use" = "homeRange",
         "movement" = "movement_pc1",
         "roost_div" = "roost_pc1")
nrow(red) == nrow(demo) # checking that the number of rows hasn't changed
red_long <- red %>%
  pivot_longer(cols = c("space_use", "movement", "roost_div"),
               names_to = "var",
               values_to = "value")
```

Let's examine these predictor variables

```{r}
#| message: false
#| warning: false
colors <- c("firebrick2", "blue", "gold")
red_long %>%
  ggplot(aes(x = value, col = var, fill = var))+
  geom_histogram()+ theme_minimal()+
  facet_wrap(~var, scales = "free")+
  scale_color_manual(values = colors)+
  scale_fill_manual(values = colors)+
  theme(legend.position = "none")+
  ylab("Frequency")+xlab("")
```

We can see here that we have another problem. Since `movement` and `roost_div` are composite variables derived from PCAs, they are pretty close to normally distributed. Meanwhile, our space use variable is very right-skewed, because it's a measure of area.

Let's log-tranform the space use variable to make the distributions more comparable and therefore make the modeling easier.

```{r}
#| include: false
red <- red %>%
  mutate(space_use_log = log(space_use))
red_long <- red %>%
  pivot_longer(cols = c("space_use", 
                        "space_use_log", 
                        "movement", 
                        "roost_div"),
               names_to = "var",
               values_to = "value")
```

Re-examine the distributions of the variables after log-transforming:

```{r}
#| message: false
#| warning: false
red_long %>%
  filter(var != "space_use") %>%
  ggplot(aes(x = value, col = var, fill = var))+
  geom_histogram()+ theme_minimal()+
  facet_wrap(~var, scales = "free")+
  scale_fill_manual(values = colors)+
  scale_color_manual(values = colors)+
  theme(legend.position = "none")+
  ylab("Frequency")+xlab("")
```

Okay that's still not perfect, but it's a lot better.

*Note also that this outcome may change after we calculate the home ranges by better means (AKDE).*

## Correlations

These are still three measures of movement, so I expect that they will be at least somewhat correlated. Let's take a look!

**Movement** and **roost diversification** are definitely correlated, but the correlation isn't super high: `r round(cor(red$movement, red$roost_div), 2)`.

```{r}
#| message: false
#| warning: false
red %>%
  ggplot(aes(x = movement, y = roost_div))+
  geom_point(color = "purple")+ geom_smooth(method = "lm", col = "purple")+ theme_minimal()
```

**Roost diversification** and **log-transformed home range** have a stronger correlation of `r round(cor(red$roost_div, red$space_use_log), 2)`.

```{r}
#| message: false
#| warning: false
red %>%
  ggplot(aes(x = roost_div, y = space_use_log))+
  geom_point(color = "chartreuse3")+ geom_smooth(method = "lm", col = "chartreuse3")+ theme_minimal()
```

Finally, **log-transformed home range** and **movement** have a correlation of `r round(cor(red$space_use_log, red$movement), 2)`.

```{r}
#| message: false
#| warning: false
red %>%
  ggplot(aes(x = space_use_log, y = movement))+
  geom_point(color = "orange2")+ geom_smooth(method = "lm", col = "orange2")+ theme_minimal()
```

There is enough correlation here that we're going to have to be careful, but I don't think there's enough that we have to not proceed. These predictors seem more biologically defensible because they are created a priori instead of derived from a single PCA.

## Examine effects of age and sex

```{r}
#| include: false
# Add age information
joined_ages <- red %>%
  left_join(ages)
nrow(joined_ages) == nrow(red) # FALSE because individuals can have multiple ages during a season. We can't have that!
joined_ages %>% group_by(Nili_id, seasonUnique) %>% filter(n() > 1)  %>% pull(seasonUnique) %>% unique() # indeed, it's just during the breeding seasons that this happens.

# For now, let's go with the older age, but I'm going to double-check with Marta about this. # XXX
ages_older <- ages %>%
  group_by(Nili_id, seasonUnique) %>%
  arrange(desc(age), .by_group = T) %>%
  slice(1) %>%
  ungroup()

red <- red %>%
  left_join(ages_older)
```

Let's see if the relationships between these variables and age are consistent with what Marta found (they should be!)

```{r}
#| warning: false
#| message: false
red %>%
  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
               names_to = "var", values_to = "value") %>%
  ggplot(aes(x = age, y = value, group = seasonUnique, col = var))+
  geom_point(alpha = 0.5) + geom_smooth(se = F)+
  theme_minimal()+
  xlab("Age")+
  facet_wrap(~var, scales = "free")+
  scale_color_manual(values = colors)+
  theme(legend.position = "none")
```

We sort of recover the same patterns as Marta, but they're not perfectly clear-cut.

Just in case we had any doubts about vultures' lack of sexual dimorphism...

```{r}
#| warning: false
#| message: false
red %>%
  filter(!is.na(sex), sex != "NA") %>%
  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
               names_to = "var", values_to = "value") %>%
  ggplot(aes(x = sex, y = value, col = var))+
  geom_jitter(height = 0, width = 0.2, alpha = 0.7)+
  geom_violin(fill = NA)+
  theme_minimal()+
  xlab("Age")+
  facet_wrap(~var, scales = "free")+
  scale_color_manual(values = colors)+
  theme(legend.position = "none")+ylab("")

red %>%
  filter(!is.na(sex), sex != "NA") %>%
  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
               names_to = "var", values_to = "value") %>%
  ggplot(aes(x = sex, y = value, col = var))+
  geom_jitter(height = 0, width = 0.2, alpha = 0.7)+
  geom_violin(fill = NA)+
  theme_minimal()+
  xlab("Age")+
  facet_grid(rows = vars(var), cols = vars(season), scales = "free")+
  scale_color_manual(values = colors)+
  theme(legend.position = "none")+ylab("")
```

Interestingly, if anything, the breeding season is when they show the *least* difference by sex! Weird.

Okay, yeah, so about those seasons:

```{r}
red %>%
  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
               names_to = "var", values_to = "value") %>%
  filter(var != "space_use") %>%
  ggplot(aes(x = value, group = seasonUnique, col = season))+
  geom_density()+ theme_minimal()+
  scale_color_manual(name = "Season", 
                     values = c(cc$breedingColor, 
                                cc$summerColor, 
                                cc$fallColor))+
  facet_wrap(~var, scales = "free")+
  ylab("Frequency")+xlab("")

red %>%
  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
               names_to = "var", values_to = "value") %>%
  filter(var != "space_use") %>%
  ggplot(aes(x = season, y = value))+
  geom_violin(aes(fill = season, col = season))+
  facet_wrap(~var, scales = "free")+
  theme_minimal()+
  scale_color_manual(name = "Season", 
                     values = c(cc$breedingColor, 
                                cc$summerColor, 
                                cc$fallColor))+
  scale_fill_manual(name = "Season", 
                     values = c(cc$breedingColor, 
                                cc$summerColor, 
                                cc$fallColor))+
  theme(legend.position = "none")+
  ylab("Value")+
  xlab("Season")
```

There's not a huge amount of differentiation by season, but there's a little bit. As expected, we do see the most movement in the summer and the least during the breeding season.

### 2024-02-03 scripts/01.0_calcMovement.R
running through the ctmm akde tutorial

@@ -13,6 +13,7 @@ library(elevatr) # for getting elevation information
 library(raster) # for getting elevation information
 library(parallel) # for running long distance calculations in parallel
 library(future) # for parallel computing
+library(ctmm)
 load("data/derived/cc.Rda")
 
 distviz <- function(data, varname, seasoncol){
@@ -31,37 +32,90 @@ distviz <- function(data, varname, seasoncol){
 }
 
 ## Load data ---------------------------------------------------------------
-base::load("data/derived/seasons_10min.Rda") # data rarefied to 10 minute intervals. Going to use that for everything.
+base::load("data/dataPrep/downsampled_10min.Rda") # data rarefied to 10 minute intervals. Going to use that for everything.
 base::load("data/orphan/datasetAssignments.Rda")
-base::load("data/derived/roosts_seasons.Rda") # don't need a separate roost dataset for the rarefied data, because roosts are only once per day, and it makes sense to use the most detailed dataset possible.
-roosts_seasons <- map(roosts_seasons, ~.x %>% group_by(Nili_id) %>% 
+base::load("data/dataPrep/season_names.Rda")
+base::load("data/dataPrep/roosts.Rda") # don't need a separate roost dataset for the rarefied data, because roosts are only once per day, and it makes sense to use the most detailed dataset possible.
+roosts <- map(roosts, ~.x %>% group_by(Nili_id) %>% 
                         mutate(daysTracked = length(unique(roost_date))) %>% ungroup())
 
-roosts_seasons <- roosts_seasons[-1] # remove summer 2020
-datasetAssignments <- datasetAssignments[-1] # remove summer 2020
-test <- map2(roosts_seasons, datasetAssignments, ~left_join(.x, .y, by = "Nili_id"))
-seasonNames <- map_chr(seasons_10min, ~as.character(.x$seasonUnique[1]))
+test <- map2(roosts, datasetAssignments, ~left_join(.x, .y, by = "Nili_id"))
 roostPolygons <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
 
-durs <- map_dbl(seasons_10min, ~{
+durs <- map_dbl(downsampled_10min, ~{
   dates <- lubridate::ymd(.x$dateOnly)
   dur <- difftime(max(dates), min(dates), units = "days")
 })
 
 
 # Get number of days tracked for each individual --------------------------
[…]

### 2024-02-06 scripts/00.0_dataPrep.qmd
updated file path to use here::here
re-did subsampling for movement data, and added subsampling for social data too

--- 2024-01-25 00.0_dataPrep.qmd
@@
 ```{r}
-base::load("movebankCredentials/pw.Rda")
+base::load("movebankCredentials", "pw.Rda")
 MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
@@
 ```{r}
-load("data/dataPrep/cleaned.Rda")
+load(here::here("data", "dataPrep", "cleaned.Rda"))
 ```
@@
 
+```{r}
+load(here::here("data", "dataPrep", "removed_lfr.Rda"))
+```
+
+```{r}
+#| eval: false
+
+# Function to downsample the data
+subsample <- function(df, idCol = "Nili_id", 
+                      timestampCol = "timestamp", mins = 10){
+  sub <- df %>%
+    mutate(d = lubridate::date(.data[[timestampCol]])) %>% 
+    group_by(.data[[idCol]]) %>%
+    arrange(timestampCol, .by_group = T) %>%
+    mutate(tc = cut(.data[[timestampCol]], 
+                    breaks = paste(as.character(mins), 
+                                   "min", sep = " "))) %>%
+    ungroup() %>%
+    group_by(.data[[idCol]], d, tc) %>%
+    slice(1) %>%
+    ungroup() %>%
+    dplyr::select(-c("d", "tc"))
+  return(sub)
+}
+
+downsampled_10min_forSocial <- map(removed_lfr, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
+
+# Add information on # of days tracked for each individual
+downsampled_10min_forSocial <- map(downsampled_10min_forSocial, ~.x %>% group_by(Nili_id) %>% 
+                       mutate(daysTracked = length(unique(dateOnly))))
+save(downsampled_10min_forSocial, file = here::here("data", "dataPrep", "downsampled_10min_forSocial.Rda"))
+
+map_dbl(removed_lfr, nrow) %>% setNames(season_names)
+map_dbl(downsampled_10min_forSocial, nrow) %>% setNames(season_names)
+
+# tidying
+rm(removed_lfr)
+rm(downsampled_10min_forSocial)
+gc()
+```
+
 ### Data for social analysis
@@
 ```{r}
-load("data/dataPrep/with_altitudes.Rda")
-load("data/dataPrep/season_names.Rda")
-```
-
-```{r}
-#| eval: false
-
-# Function to downsample the data
-subsample <- function(df, idCol = "Nili_id", 
-                      timestampCol = "timestamp", mins = 10){
-  sub <- df %>%
-    arrange(idCol, timestampCol) %>%
-    group_by(.data[[idCol]]) %>%
-    mutate(tc = cut(.data[[timestampCol]], 
-                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
-    group_by(.data[[idCol]], 
-             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
-    slice(1) %>%
-    ungroup() %>%
-    dplyr::select(-c("d", "tc"))
-  return(sub)
-}
+load(here::here("data", "dataPrep", "with_altitudes.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
+```
+
+```{r}
+#| eval: false
+
+# defined the "subsample" function previously above
 
@@
 
-save(downsampled_10min, file = "data/dataPrep/downsampled_10min.Rda")
+save(downsampled_10min, file = here::here("data", "dataPrep", "downsampled_10min.Rda"))
 

### 2024-02-06 scripts/akde_test.R
full akde test workflow

@@ -1,0 +1,112 @@
+# Testing out ctmm
+
+library(ctmm)
+library(tidyverse)
+library(furrr)
+library(purrr)
+
+## Load data ---------------------------------------------------------------
+base::load("data/dataPrep/downsampled_10min.Rda") # data rarefied to 10 minute intervals. Going to use that for everything.
+
+# ctmm_season_test <- downsampled_10min[[2]] %>%
+#   dplyr::select("ID" = Nili_id, timestamp, "longitude" = location_long, "latitude" = location_lat)
+# ctmm_season_test_animals <- ctmm_season_test %>%
+#   group_by(ID) %>%
+#   group_split()
+# save(ctmm_season_test_animals, file = "data/ctmm_season_test_animals.Rda")
+load("data/ctmm_season_test_animals.Rda")
+base::load("data/dataPrep/season_names.Rda")
+
+# Testing out AKDE --------------------------------------------------------
+animals <- map(downsampled_10min, ~unique(.x$Nili_id))
+seasons_split <- map(downsampled_10min, ~.x %>%
+                       dplyr::select("ID" = Nili_id,
+                                     timestamp,
+                                     "longitude" = location_long,
+                                     "latitude" = location_lat) %>%
+                       group_by(ID) %>%
+                       group_split())
+telems_list <- map(seasons_split, ~map(.x, as.telemetry))
+# check out a few examples
+dt.plot(telems_list[[1]][[3]])
+dt.plot(telems_list[[1]][[4]])
+dt.plot(telems_list[[1]][[2]])# the minimum time interval shouldn't be this small; I'm not sure what's going on. Need to go back and figure out why this didn't get downsampled properly. # XXX WHY ARE THERE SHORT SAMPLING INTERVALS?
+
+# Check the range-residency assumption
+variograms_list <- map(telems_list, ~map(.x, variogram, .progress = T))
+walk(variograms_list[[1]], plot) # just get a brief glimpse of the different individuals. They generally look okay; there's a fair bit of periodicity that I will have to look into (i.e. does it affect the ability of the package to handle the data?) # XXX DOES PERIODICITY MATTER?
[…]

### 2024-02-06 scripts/01.1_categorizeMovement.qmd
Updates before vulture meeting 2/5/24

@@
 movement <- movementBehaviorScaled %>%
-  dplyr::select(Nili_id, seasonUnique, mnDailyMaxAlt, mnDailyMedAlt, meanDFD, maxDFD, meanDMD, meanDDT, mnTort)
+  dplyr::select(Nili_id, seasonUnique, mnDailyMnAlt, meanDFD, meanDMD, meanDDT, mnTort)
 roost_behavior <- movementBehaviorScaled %>%
@@
 
-I first thought to do separate PCA's of each of these categories.
-But that doesn't make sense, especially for something like space use, where there are really only two metrics (and then one that's the combination of each of them)
-
-Instead, I want to make these more biologically motivated. Let's do the following:
-
-1\) home range only (but make it using AKDE method so it's actually defensible)
-
-2\) PCA of movement behaviors only
-
-3\) PCA of roost behaviors
+The PCA movement variables are a little hard to interpret, and I want the predictor variables to directly describe aspects of individual behavior/movement/space use.\
+\
+New proposal is to make a PCA of the strictly movement-related variables, a PC of the strictly roosting-related variables, and then use AKDE home ranges as the third variable measuring space use.
+Here is what this would look like:
 
@@
 movement %>%
-  pivot_longer(cols = c("mnDailyMaxAlt", "mnDailyMedAlt", "meanDFD", "maxDFD", "meanDMD", "meanDDT", "mnTort"),
+  pivot_longer(cols = c("mnDailyMnAlt", "meanDFD", "meanDMD", "meanDDT", "mnTort"),
                names_to = "measure", values_to = "value") %>%
   mutate(measure = fct_recode(measure,
-                              "Daily max. altitude (seasonal avg)" =
-                                "mnDailyMaxAlt",
-                              "Daily median altitude (seasonal avg)" = 
-                                "mnDailyMedAlt",
+                              "Daily mean altitude (seasonal avg)" = 
+                                "mnDailyMnAlt",
                               "Daily flight time (seasonal avg)" = 
                                 "meanDFD",
-                              "Daily flight time (seasonal max)" = 
-                                "maxDFD",
                               "Daily max displacement (seasonal avg)" = 
@@
 
-*Note also that this outcome may change after we calculate the home ranges by better means (AKDE).*
+*Note also that this outcome may change quantitatively after we calculate the home ranges by better means (AKDE), but I don't expect it to be qualitatively different.
+AKDE calculations are currently in progress.*
 
@@
 
-**Movement** and **roost diversification** are definitely correlated, but the correlation isn't super high: `r round(cor(red$movement, red$roost_div), 2)`.
+**Movement** and **roost diversification** have a correlation of `r round(cor(red$movement, red$roost_div), 2)`.
 
@@
 There is enough correlation here that we're going to have to be careful, but I don't think there's enough that we have to not proceed.
-These predictors seem more biologically defensible because they are created a priori instead of derived from a single PCA.
+These predictors seem more biologically defensible because they are created *a priori* instead of derived from a single PCA.
 
@@
   theme(legend.position = "none")+ylab("")
-
-red %>%
-  filter(!is.na(sex), sex != "NA") %>%
-  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
-               names_to = "var", values_to = "value") %>%
-  ggplot(aes(x = sex, y = value, col = var))+
-  geom_jitter(height = 0, width = 0.2, alpha = 0.7)+
-  geom_violin(fill = NA)+
-  theme_minimal()+
-  xlab("Age")+
-  facet_grid(rows = vars(var), cols = vars(season), scales = "free")+
-  scale_color_manual(values = colors)+
-  theme(legend.position = "none")+ylab("")
+# 
+# red %>%
+#   filter(!is.na(sex), sex != "NA") %>%
+#   pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+#                names_to = "var", values_to = "value") %>%
+#   ggplot(aes(x = sex, y = value, col = var))+
+#   geom_jitter(height = 0, width = 0.2, alpha = 0.7)+
+#   geom_violin(fill = NA)+
+#   theme_minimal()+
+#   xlab("Age")+
+#   facet_grid(rows = vars(var), cols = vars(season), scales = "free")+
+#   scale_color_manual(values = colors)+
+#   theme(legend.position = "none")+ylab("")
 ```

### 2024-02-06 scripts/02.0_calcSocial.R
Re-running the edge calculations on the downsampled social data, with new season included

@@ -4,63 +4,80 @@ library(vultureUtils)
 library(igraph)
 library(tidyverse)
 library(sf)
-source("scripts/evenness.R")
 
 ## Data ---------------------------------------------------------------
-load("data/derived/seasons_forSoc.Rda")
-load("data/derived/seasons_forSoc_mode10.Rda")
-seasons_forSoc <- map(seasons_forSoc, ~st_as_sf(.x, coords = c("location_long", "location_lat"), crs = "WGS84", remove = F))
-seasons_forSoc_mode10 <- map(seasons_forSoc_mode10, ~st_as_sf(.x, coords = c("location_long", "location_lat"), crs = "WGS84", remove = F))
+load("data/dataPrep/downsampled_10min_forSocial.Rda") # this is the last dataset we produced in the dataPrep script before moving on to the further cleaning for movement, so this is the one we're going to use for the social interactions.
+sfdata <- map(downsampled_10min_forSocial, ~st_as_sf(.x, coords = c("location_long", "location_lat"), crs = "WGS84", remove = F))
 roostPolygons <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
-load("data/derived/roosts_seasons.Rda")
-load("data/derived/roosts_seasons_mode10.Rda")
-load("data/orphan/datasetAssignments.Rda")
-
-seasonNames <- map_chr(seasons_forSoc, ~as.character(.x$seasonUnique[1]))
-seasons_forSoc <- seasons_forSoc[-which(seasonNames == "2020_summer")]
-seasons_forSoc_mode10 <- seasons_forSoc_mode10[-which(seasonNames == "2020_summer")]
-datasetAssignments <- datasetAssignments[-which(seasonNames == "2020_summer")]
-roosts_seasons <- roosts_seasons[-which(seasonNames == "2020_summer")]
-roosts_seasons_mode10 <- roosts_seasons_mode10[-which(seasonNames == "2020_summer")]
-seasonNames <- seasonNames[-which(seasonNames == "2020_summer")]
+rm(downsampled_10min_forSocial)
+gc()
+load("data/dataPrep/season_names.Rda")
+load("data/dataPrep/roosts.Rda") # XXX load
 
 # Investigate the data ----------------------------------------------------
 # How many individuals do we have with vs. without excluding those with a lower sampling rate?
[…]

### 2024-02-06 scripts/02.0_calcSocial.R
switched name of roostingSRI to be consistent with flight and feeding

@@ -63,7 +63,7 @@ rm(feedingEdges)
 rm(feedingSRI)
 gc()
 
-roosting <- map(roosts, ~{
+roostingSRI <- map(roosts, ~{
   vultureUtils::getRoostEdges(.x, mode = "polygon", 
                               roostPolygons = roostPolygons, 
                               return = "sri", 
@@ -73,8 +73,8 @@ roosting <- map(roosts, ~{
                               dateCol = "roost_date")
   }, .progress = T)
 
-save(roosting, file = "data/calcSocial/roosting.Rda")
-rm(roosting)
+save(roostingSRI, file = "data/calcSocial/roostingSRI.Rda")
+rm(roostingSRI)
 
 
 # XXX START HERE

### 2024-02-06 scripts/00.0_dataPrep.qmd
added graph to examine the number of rows left at the four stages of data cleaning (before and after downsampling

@@
 ```
+
+```{r}
+#| eval: false
+# Examine the number of rows that get removed by the downsampling at different stages
+load(here::here("data", "dataPrep", "removed_lfr.Rda"))
+load(here::here("data", "dataPrep", "downsampled_10min_forSocial.Rda"))
+load(here::here("data", "dataPrep", "with_altitudes.Rda"))
+load(here::here("data", "dataPrep", "downsampled_10min_forSocial.Rda"))
+
+df <- data.frame(seasonUnique = factor(season_names, levels =season_names), removed_lfr = map_dbl(removed_lfr, nrow), downsampled_10min_forSocial = map_dbl(downsampled_10min_forSocial, nrow), with_altitudes = map_dbl(with_altitudes, nrow), downsampled_10min = map_dbl(downsampled_10min, nrow)) %>% pivot_longer(-seasonUnique, names_to = "stage", values_to = "rows") %>% mutate(stage = factor(stage, levels = unique(stage)))
+
+df %>% ggplot(aes(x = stage, y = rows, col = seasonUnique, group = seasonUnique))+geom_line()+theme_minimal()
+```
+
+This looks crazy, but it's actually correct.
+Note that the stages aren't actually linear– the `removed_lfr` –\> `downsampled_10min_forSocial` pathway is a separate offshoot.
+`downsampled_10min_forSocial` doesn't then get passed into the further cleaning for movement data; rather, it goes straight to the social analysis while `removed_lfr` passes along to the further cleaning.
+In 2023 summer, there were a lot of high-resolution fixes, so this matters a whole lot.
+I didn't notice the effects of the downsampling much before because the other seasons just weren't too big, but this new season is an order of magnitude larger, which makes the downsampling essential.

### 2024-02-06 scripts/00.0_dataPrep.qmd
investigating the results of downsampling

@@
 I didn't notice the effects of the downsampling much before because the other seasons just weren't too big, but this new season is an order of magnitude larger, which makes the downsampling essential.
+
+```{r}
+#| eval: false
+
+# Let's make sure the downsampling actually worked
+testseason <- downsampled_10min[[1]]
+with_ints <- testseason %>%
+  group_by(Nili_id, dateOnly) %>%
+  arrange(timestamp, .by_group = T) %>%
+  mutate(int = difftime(lead(timestamp), timestamp, units = "mins"))
+with_ints %>% 
+  select(Nili_id, dateOnly, timestamp, int) %>%
+  ggplot(aes(x = as.numeric(int), col = Nili_id))+
+  geom_density()+
+  theme_minimal()+
+  theme(legend.position = "none") # Good, none of them are less than 0.
+
+# Do we have any less than 10 minutes?
+with_ints %>% filter(as.numeric(int) < 10) # yes, there are a lot that are less than 10 minutes
+
+# What about less than 9 minutes?
+with_ints %>% filter(as.numeric(int) < 9) # okay, so something in my code is not working to filter out the really short intervals. What's happening here?
+
+# Oh...
+the way I did it actually guarantees this to happen.
+I cut the times into groups and then took the first one for each group.
+So if the first one for the previous group happened to be at the end of that group and then the first one of the next group happened to be at the beginning of that group, we can end up with arbitrarily small time intervals.
+
+# Now, I kind of don't care, since the main point here was just to get rid of the points. But I wonder if there's a different way to do this.
+```

### 2024-02-06 scripts/00.0_dataPrep.qmd
literature precedent for thinning data, and trying atlastools

@@
 editor: visual
+bibliography: references.bib
 ---
@@
 ```
+
+I just read up on subsampling vs.
+thinning in [@gupte2022].
+It seems like I've reinvented the wheel here.
+What I'm doing is subsampling (choosing one point from each time interval), as opposed to aggregation (summarizing all the points within a given time interval, e.g.
+taking the mean).
+They both have their faults; for example, aggregation adds gps error (which can be measured, but still) while subsampling risks selecting outlier points just by chance but does preserve the location error structure without further effort.
+They note that you need to do the rest of the cleaning first before doing the subsampling, which is exactly what I've done, so yay!
+
+There is a function in `atlastools` for thinning the data, either by aggregation or subsampling. I'm going to test it and see if it's faster.
+
+I tried out the atlastools function and it was quite buggy and not necessarily faster, so I'm going to stick with my own function.
+Can still cite this paper for a general guideline of how to do it though.
+
+```{r}
+# library(atlastools)
+# library(tictoc)
+# testseason <- removed_lfr[[1]] %>% mutate(x = "location_long", y = "location_lat") %>% select(x, y, timestamp, Nili_id, dateOnly) %>% filter(!is.na(timestamp), !is.na(x), !is.na(y), !is.na(Nili_id), !is.na(Nili_id)) %>% mutate(time = as.numeric(timestamp))
+# tic()
+# atl_thinned <- atl_thin_data(data = testseason,
+#               interval = 600,
+#               id_columns = c("Nili_id", "dateOnly"),
+#               method = "subsample")
+# toc()
+```
+
+Okay, so in sum, I am *not* worried about there being a few really brief time intervals remaining in my data.
+For the AKDE's, I'll want to use a 10-minute `dt` value to avoid applying the shortest interval over the entire dataset.
+But for proceeding, it's okay to have these results of the downsampling stand.

### 2024-02-07 scripts/04.0_mixedModels.R
re-working the data ingest portion of the modeling script

@@ -1,5 +1,4 @@
 library(car) # for p-values in mixed models?
-# In lieu of tidyverse
 library(tidyverse)
 library(sf)
 library(factoextra)
@@ -19,68 +18,58 @@ library(broom.mixed)
 library(jtools) # similar to sjplot, for forest plots
 library(emmeans) # estimated marginal means/trends
 
-load("data/derived/linked.Rda")
-
 # Set ggplot theme to classic
 theme_set(theme_classic())
 
+# Load network metrics and movement variables
+load("data/calcSocial/networkMetrics.Rda")
+networkMetrics <- networkMetrics %>% rename("seasonUnique" = "season")
+load("data/new_movement_vars.Rda")
+load("data/dataPrep/season_names.Rda")
 load("data/derived/cc.Rda")
 
-# Here's an interesting reference article for some modeling stuff: https://biol607.github.io/lab/12_gzlm.html
+# Join the two datasets (there will be 3 rows per individual per season, for the 3 different situations)
+linked <- new_movement_vars %>%
+  left_join(networkMetrics, by = c("Nili_id", "seasonUnique"))
+nrow(linked) == 3*nrow(new_movement_vars)
+
+# Housekeeping
 linked <- linked %>%
-  mutate(season = factor(season, levels = c("breeding", "summer", "fall")))
+  mutate(season = stringr::str_extract(seasonUnique, "[a-z]+"),
+         year = as.numeric(stringr::str_extract(seasonUnique, "[0-9]+")),
+         seasonUnique = factor(seasonUnique, levels = season_names),
+         sex = factor(sex),
+         season = factor(season, levels = c("breeding", "summer", "fall")))
 
-# Remove individuals that don't have movement information
-colSums(is.na(linked)) # lots of NA's
-linked <- linked %>%
-  filter(!is.na(PC1)) # note that because I joined the sex/age information along with the movement information (instead of along with the social information), if I ever want to *not* filter out those individuals, I'll have to go back and add sex/age info for the additional birds. Don't need this for now, so I'm skipping it. 
-colSums(is.na(linked)) # okay now we have no more NA's except in sex. Yay!
+# Check for NA's (we shouldn't have any because of the direction of the join)
+colSums(is.na(linked)) # awesome
 
-# Mixed models -------------------------------------------------------------------
-# Response variables:
-# Relative degree
-# Relative strength
-# Evenness
-
-# Predictors:
[…]

### 2024-02-07 scripts/02.0_calcSocial.R
re-calculated normalized degree and strength appropriately

@@ -13,6 +13,8 @@ datasetAssignments <- map(downsampled_10min_forSocial, ~.x %>% select(Nili_id, d
                                                        TRUE ~ dataset)) %>%
                             ungroup() %>%
                             distinct())
+save(datasetAssignments, file = "data/calcSocial/datasetAssignments.Rda")
+load("data/calcSocial/datasetAssignments.Rda")
 rm(downsampled_10min_forSocial)
 gc()
 roostPolygons <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
@@ -140,11 +142,12 @@ networkMetrics <- map2_dfr(flightGraphs, season_names, ~{
 # Add info about number of indivs and relative measures
 networkMetrics <- networkMetrics %>%
   group_by(season, type) %>%
-  mutate(n = length(unique(Nili_id))) %>% # CAREFUL HERE! Was getting the wrong degree bc conflating number of rows with number of individuals.
+  mutate(n = length(unique(Nili_id)), # CAREFUL HERE! Was getting the wrong degree bc conflating number of rows with number of individuals.
+         totalStrength = sum(strength)) %>% 
   ungroup() %>%
-  mutate(degreeRelative = degree/n,
-         strengthRelative = strength/n) %>%
-  dplyr::select(season, type, n, Nili_id, degree, degreeRelative, strength, strengthRelative)
+  mutate(normDegree = degree/(n-1),
+         normStrength = strength/totalStrength) %>%
+  dplyr::select(season, type, n, Nili_id, degree, normDegree, strength, normStrength)
 
 # Save network metrics ----------------------------------------------------
 # Assign datasets
@@ -165,3 +168,7 @@ table(zer$season, zer$dataset) # there are more consistently some zeroes in the 
 zer %>% group_split(season) %>% map(., ~.x %>% pull(Nili_id) %>% unique() %>% sort()) # okay we definitely have some repeats, but they aren't blatantly all the same.
 
 # So I guess I conclude from this that the zeroes are real and should be treated as real... 
+
+# Briefly check out the distributions of the normalized degree and strength values
+hist(networkMetrics$normDegree)
+hist(networkMetrics$normStrength)

### 2024-02-07 scripts/01.0_calcMovement.R
removed extraneous movement measures

@@ -37,7 +37,6 @@ base::load("data/dataPrep/roosts.Rda")
 roosts <- map(roosts, ~.x %>% group_by(Nili_id) %>% 
                         mutate(daysTracked = length(unique(roost_date))) %>% ungroup())
 
-test <- map2(roosts, datasetAssignments, ~left_join(.x, .y, by = "Nili_id"))
 roostPolygons <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
 
 durs <- map_dbl(downsampled_10min, ~{
@@ -193,20 +192,12 @@ dfd <- map(downsampled_10min_sf, ~.x %>%
 
 dfdSumm <- map(dfd, ~.x %>%
                  group_by(Nili_id) %>%
-                 summarize(meanDFD = mean(totalFlightDuration, na.rm = T),
-                           minDFD = min(totalFlightDuration, na.rm = T),
-                           maxDFD = max(totalFlightDuration, na.rm = T),
-                           medDFD = median(totalFlightDuration, na.rm = T)))
+                 summarize(meanDFD = mean(totalFlightDuration, na.rm = T)))
 
 names(dfdSumm) <- season_names
 dfddf <- purrr::list_rbind(dfdSumm, names_to = "season") %>%
-  mutate(meanDFD = meanDFD/60/60,
-         minDFD = minDFD/60/60,
-         maxDFD = maxDFD/60/60,
-         medDFD = medDFD/60/60)
+  mutate(meanDFD = meanDFD/60/60)
 distviz(dfddf, "meanDFD", "season")
-distviz(dfddf, "minDFD", "season")
-distviz(dfddf, "maxDFD", "season")
 
 ## Time spent flying in summer/fall/breeding
 dfddf <- dfddf %>%
@@ -219,22 +210,6 @@ dfddf %>%
   geom_density(linewidth = 1.5, alpha = 0.7)+
   theme_classic()+
   scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor))
-
-# what about min?
-dfddf %>%
-  ggplot(aes(x = minDFD, col = szn, group = factor(season)))+
-  geom_density(linewidth = 1.5, alpha = 0.7)+
-  theme_classic()+
-  scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor)) # fewer zero days in summer! Makes sense.
-
-# what about max?
-dfddf %>%
[…]

### 2024-02-07 scripts/01.1_categorizeMovement.qmd
re-ran movement vars and saved the output

@@
 movementBehaviorScaled <- allMovementBehavior %>%
-  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex, dataset, 
-                   homeRange, coreArea, coreAreaFidelity), # not scaling the space use params, because I'm not putting those into a PCA.
+  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex, homeRange, coreArea, coreAreaFidelity), # not scaling the space use params, because I'm not putting those into a PCA.
                 function(x){as.numeric(as.vector(scale(x)))}))
@@
 demo <- movementBehaviorScaled %>%
-  dplyr::select(Nili_id, seasonUnique, daysTracked, propDaysTracked, dataset, birth_year, sex)
+  dplyr::select(Nili_id, seasonUnique, daysTracked, propDaysTracked, birth_year, sex)
 space_use <- movementBehaviorScaled %>%
@@
 There's not a huge amount of differentiation by season, but there's a little bit. As expected, we do see the most movement in the summer and the least during the breeding season.
+
+```{r}
+head(red)
+new_movement_vars <- red
+save(new_movement_vars, file = here::here("data", "new_movement_vars.Rda"))
+```

### 2024-02-07 scripts/04.0_mixedModels.R
base model for degree

@@ -69,7 +69,22 @@ linked %>%
 networkMetrics %>% filter(degree == 0 | strength == 0) # lots of rows
 linked %>% filter(degree == 0 | strength == 0) # just one individual in one season!
 
-# It is possible this will pose a problem for modeling. I wonder if it's better to remove the zero or do some kind of other transformation?
+# # XXX It is possible this will pose a problem for modeling. I wonder if it's better to remove the zero or do some kind of other transformation?
+# I think I can just remove it for now
+linked <- linked %>%
+  filter(normDegree > 0, normStrength > 0)
+
+# Modeling ----------------------------------------------------------------
+degree_base <- lmer(normDegree ~ situ + movement + roost_div + space_use_log + age_group + season + (1|seasonUnique)+(1|Nili_id), data = linked)
+check_model(degree_base) # the good news is that the variance inflation factor is quite low, so these predictors aren't prohibitively highly correlated (yay!!!!). The bad news is that the posterior predictive check and the residuals vs. fitted values plots look... bad. 
+plot(DHARMa::simulateResiduals(degree_base), pch=".")
+
+
+
+
+
+
+
 
 ## Degree ------------------------------------------------------------------
 # 2023-06-05 going back to gaussian for now

### 2024-02-07 scripts/04.0_mixedModels.R
degree model and associated plots for lab meeting 2/8/24

@@ -53,17 +53,44 @@ table(linked$situ)
 
 
 # Examine response variable distributions ---------------------------------
-linked %>%
+normDegree_years_seasons_flight <- linked %>%
+  filter(type == "flight") %>%
+  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
   ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-  geom_density()+
-  scale_color_manual(name = "Season", values = c(cc$breedingColor, cc$summerColor, cc$fallColor))+
-  theme_minimal() # what is up with the summers??? Why do they have so much higher degree in summer? This much difference seems unexpected.
+  geom_density(linewidth = 1.5)+
+  theme_classic()+
+  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
+  ylab("Frequency")+
+  xlab("Degree (normalized)")+
+  facet_wrap(~season)+
+  theme(legend.position = "none", text = element_text(size = 20))
+ggsave(normDegree_years_seasons_flight, filename = "fig/normDegree_years_seasons_flight.png", width = 9, height = 7)
 
-linked %>%
-  ggplot(aes(x = normStrength, col = season, group = seasonUnique))+
-  geom_density()+
-  scale_color_manual(name = "Season", values = c(cc$breedingColor, cc$summerColor, cc$fallColor))+
-  theme_minimal() # XXX what is up with the summers??? Why do they have so much higher degree in summer? This much difference seems unexpected.
+normDegree_years_seasons_feeding <- linked %>%
+  filter(type == "feeding") %>%
+  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
+  ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
+  geom_density(linewidth = 1.5)+
+  theme_classic()+
+  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
+  ylab("Frequency")+
+  xlab("Degree (normalized)")+
+  facet_wrap(~season)+
+  theme(legend.position = "none", text = element_text(size = 20))
+ggsave(normDegree_years_seasons_feeding, filename = "fig/normDegree_years_seasons_feeding.png", width = 9, height = 7)
+
+normDegree_years_seasons_roosting <- linked %>%
+  filter(type == "roosting") %>%
[…]

### 2024-02-07 scripts/01.1_categorizeMovement.qmd
made plots prettier

@@
   ylab("Value (scaled)")+
-  theme(legend.position = "none")
+  theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
 ```
@@
 #| warning: false
+cor <- round(cor(red$movement, red$roost_div), 2)
 red %>%
   ggplot(aes(x = movement, y = roost_div))+
-  geom_point(color = "purple")+ geom_smooth(method = "lm", col = "purple")+ theme_minimal()
+  geom_point(color = "purple")+ geom_smooth(method = "lm", col = "purple")+ theme_minimal()+
+  ylab("Roost diversification")+
+  xlab("Movement")+
+  ggplot2::annotate(geom = "text", x = 4.5, y = -2, label = paste0("Correlation = ", cor))
 ```
@@
 #| warning: false
+cor <- round(cor(red$roost_div, red$space_use_log), 2)
 red %>%
   ggplot(aes(x = roost_div, y = space_use_log))+
-  geom_point(color = "chartreuse3")+ geom_smooth(method = "lm", col = "chartreuse3")+ theme_minimal()
+  geom_point(color = "chartreuse3")+ geom_smooth(method = "lm", col = "chartreuse3")+ theme_minimal()+
+  xlab("Roost diversification")+ ylab("Space use (log)")+
+  ggplot2::annotate(geom = "text", x = 4, y = 11, label = paste0("Correlation = ", cor))
 ```
@@
 #| warning: false
+cor <- round(cor(red$space_use_log, red$movement), 2)
 red %>%
   ggplot(aes(x = space_use_log, y = movement))+
-  geom_point(color = "orange2")+ geom_smooth(method = "lm", col = "orange2")+ theme_minimal()
+  geom_point(color = "orange2")+ geom_smooth(method = "lm", col = "orange2")+ theme_minimal()+
+  xlab("Space use (log)") + ylab("Movement")+
+  ggplot2::annotate(geom = "text", x = 14, y = -4.5, label = paste0("Correlation = ", cor))
 ```
@@
 Let's see if the relationships between these variables and age are consistent with what Marta found (they should be!)
+
+```{r}
+#| warning: false
+#| message: false
+## TEMPORARY: FOR PRESENTATION
+load(here::here("data", "derived", "cc.Rda"))
+red %>%
+  mutate(season = stringr::str_extract(seasonUnique, "[a-z]+")) %>%
+  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+               names_to = "var", values_to = "value") %>%
+  ggplot(aes(x = age, y = value, group = seasonUnique, col = season))+
+  geom_point(alpha = 0.5) + geom_smooth(se = F)+
+  theme_minimal()+
+  xlab("Age")+
+  facet_wrap(~var, scales = "free")+
+  scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor))+
+  theme()
+```
 

### 2024-02-07 scripts/02.0_calcSocial.R
A whole bunch of pretty network plots!

@@ -172,3 +172,31 @@ zer %>% group_split(season) %>% map(., ~.x %>% pull(Nili_id) %>% unique() %>% so
 # Briefly check out the distributions of the normalized degree and strength values
 hist(networkMetrics$normDegree)
 hist(networkMetrics$normStrength)
+
+# Plotting ----------------------------------------------------------------
+library(ggraph)
+library(tidygraph)
+grph_deg <- function(graph, season_name, type){
+  g <- graph %>%
+    as_tbl_graph() %>%
+    mutate(normDegree = centrality_degree()/n()) %>%
+    ggraph(layout = "fr")+
+    geom_edge_link(alpha = 0.2)+
+    geom_node_point(aes(size = normDegree, col = normDegree))+
+    scale_color_viridis()+
+    theme(legend.position = "none")+
+    ggtitle(season_name)
+  return(g)
+  cat("Done with ", title, "\n")
+}
+
+flightPlots <- map2(.x = flightGraphs, .y = season_names, ~grph_deg(.x, .y, type = "flight"))
+feedingPlots <- map2(.x = feedingGraphs, .y = season_names, ~grph_deg(.x, .y, type = "feeding"))
+roostingPlots <- map2(.x = roostingGraphs, .y = season_names, ~grph_deg(.x, .y, type = "roosting"))
+
+fl <- patchwork::wrap_plots(flightPlots, nrow = 3, ncol = 3)
+fe <- patchwork::wrap_plots(feedingPlots, nrow = 3, ncol = 3)
+ro <- patchwork::wrap_plots(roostingPlots, nrow = 3, ncol = 3)
+ggsave(fl, filename = "fig/networkGraphs/degree/flight.png", width = 9, height = 7)
+ggsave(fe, filename = "fig/networkGraphs/degree/feeding.png", width = 9, height = 7)
+ggsave(ro, filename = "fig/networkGraphs/degree/roosting.png", width = 9, height = 7)

### 2024-02-07 scripts/05.1_viz_plotModels.R
Removed response variable distributions, since we have those elsewhere now

@@ -14,25 +14,6 @@ load("data/derived/mods.Rda")
 load("data/derived/cc.Rda")
 load("data/derived/scaled_d.Rda") # for un-scaling the degree model
 
-# Response variable distributions -----------------------------------------
-forModeling %>%
-  pivot_longer(cols = c("degree", "strength", "evenness"), names_to = "measure", values_to = "value") %>%
-  ggplot(aes(x = value, col = season, group = interaction(season, year)))+
-  geom_density()+
-  facet_wrap(~measure, scales = "free")+
-  theme_classic()+
-  scale_color_manual(name = "Season", values = c(cc[["breedingColor"]], cc[["fallColor"]], cc[["summerColor"]]))+
-  ylab("")+xlab("")+theme(text = element_text(size = 16))
-
-forModeling %>%
-  pivot_longer(cols = c("degree_scl", "strength_scl", "evenness_scl"), names_to = "measure", values_to = "value") %>%
-  ggplot(aes(x = value, col = season, group = interaction(season, year)))+
-  geom_density()+
-  facet_grid(~measure, scales = "free")+
-  theme_classic()+
-  scale_color_manual(name = "Season", values = c(cc[["breedingColor"]], cc[["fallColor"]], cc[["summerColor"]]))+
-  ylab("")+xlab("")+theme(text = element_text(size = 16))
-
 # Degree ------------------------------------------------------------------
 d <- mods[["d"]]
 scale_d <- attributes(scaled_d)$`scaled:scale`

### 2024-02-08 scripts/04.0_mixedModels.R
strength mixed model

@@ -17,6 +17,7 @@ library(sjPlot)
 library(broom.mixed)
 library(jtools) # similar to sjplot, for forest plots
 library(emmeans) # estimated marginal means/trends
+library(ggeffects)
 
 # Set ggplot theme to classic
 theme_set(theme_classic())
@@ -102,6 +103,8 @@ linked <- linked %>%
   filter(normDegree > 0, normStrength > 0)
 
 # Modeling ----------------------------------------------------------------
+
+# Degree ------------------------------------------------------------------
 degree_base <- glmmTMB(normDegree ~ situ + movement + roost_div + space_use_log + age_group + season + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian())
 check_model(degree_base) # the good news is that the variance inflation factor is quite low, so these predictors aren't prohibitively highly correlated (yay!!!!). The bad news is that the posterior predictive check and the residuals vs. fitted values plots look... bad. 
 check_predictions(degree_base) # yeah this model is not a good fit for the data. No wonder, given the really squinched up values in the summers. I would have thought that including season and situation as predictors would fix this, but apparently not.
@@ -223,6 +226,74 @@ roost_div_situ_forest <- roost_div_situ_emt %>%
   xlab("Situation")+
   coord_flip()
 ggsave(roost_div_situ_forest, file = "fig/roost_div_situ_forest.png", width = 10, height = 6)
+
+# Strength ----------------------------------------------------------------
+strength_base <- glmmTMB(normStrength ~ situ + movement + roost_div + space_use_log + age_group + season + (1|seasonUnique)+(1|Nili_id), data = linked, family = beta_family())
+check_model(strength_base) # this actually looks pretty good, shockingly! And I only had to exclude one individual for having a 0 for strength.
+simulationOutput <- DHARMa::simulateResiduals(strength_base)
+plot(simulationOutput, pch=".") # this is kinda bad but still better than the degree model!
+
+# Let's see if we want any interactions
+strength_full <- glmmTMB(normStrength ~ situ*movement + situ*roost_div + situ*space_use_log + season*movement + season*roost_div + season*space_use_log + age_group + (1|seasonUnique)+(1|Nili_id), data = linked, family = beta_family())
[…]

### 2024-02-08 scripts/05.2_viz_plotPCExtremes.R
deleted some obsolete scripts

@@ -1,350 +1,0 @@
-# Script for creating maps showing individuals' trajectories on either end of the principal component spectrum.
-library(tidyverse)
-library(ggspatial)
-library(sf)
-library(grid)
-load("data/derived/kuds_indivs.Rda")
-load("data/derived/hrList_indivs.Rda")
-
-theme_set(theme_classic(base_size = 9))
-
-load("data/derived/linked.Rda")
-load("data/derived/seasons_10min.Rda")
-seasonNames <- map_chr(seasons_10min, ~as.character(.x$seasonUnique[1]))
-names(seasons_10min) <- seasonNames
-load("data/derived/cc.Rda")
-
-daysTracked <- map(seasons_10min, ~.x %>% st_drop_geometry() %>% dplyr::select(Nili_id, seasonUnique, daysTracked) %>% distinct()) %>% purrr::list_rbind()
-
-# Visualize PC extremes ---------------------------------------------------
-# Only want to include individuals that participated in all three
-forplots <- linked %>%
-  group_by(Nili_id, season, year) %>%
-  filter(length(unique(type)) == 3) %>%
-  ungroup() %>%
-  filter(!is.na(PC1)) %>%
-  mutate(seasonUnique = paste(year, season, sep = "_")) %>%
-  left_join(daysTracked, by = c("seasonUnique", "Nili_id"))
-
-# Take a look at the biplot again, this time facetted by season
-forplots %>%
-  ggplot(aes(x = PC1, y = PC2, col = daysTracked))+
-  geom_point(size = 2)+
-  geom_vline(aes(xintercept = 0))+
-  geom_hline(aes(yintercept = 0))+
-  theme_bw()+
-  scale_color_viridis_c()+
-  facet_wrap(~seasonUnique)
-
-# colored by days tracked
-forplots %>%
-  ggplot(aes(x = PC1, y = PC2, col = daysTracked))+
-  geom_point(size = 2)+
-  geom_vline(aes(xintercept = 0))+
-  geom_hline(aes(yintercept = 0))+
-  theme_bw()+
-  scale_color_viridis_c()
-
-# colored by year*season
-forplots %>%
-  ggplot(aes(x = PC1, y = PC2, col = seasonUnique))+
-  geom_point(size = 2)+
-  geom_vline(aes(xintercept = 0))+
-  geom_hline(aes(yintercept = 0))+
-  theme_bw()+
[…]

### 2024-02-08 scripts/ggeffects_ggpredict_plotTesting_2023-08-04.R

@@ -1,27 +1,0 @@
-library(tidyverse)
-library(sjPlot)
-library(ggeffects)
-load("d.Rda")
-
-# Using plot_model
-# Conditional (predict) plot with plot_model
-plot_model(d, type = "pred", terms = c("PC2", "situ"), show.data = T)+
-  theme_classic()
-
-# Marginal effects plot with plot_model
-plot_model(d, type = "eff", terms = c("PC2", "situ"), show.data = T)+
-  theme_classic()
-
-# Using ggeffects and then ggplot
-pred <- as.data.frame(ggpredict(d, terms = c("PC2", "situ")))
-eff <- as.data.frame(ggeffect(d, terms = c("PC2", "situ")))
-ggplot(pred, aes(x, predicted)) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
-              alpha = 0.2, linewidth = 0.6, show.legend = F)+
-  geom_line(aes(col = group), linewidth = 1.5)
-
-ggplot(eff, aes(x, predicted)) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
-              alpha = 0.2, linewidth = 0.6, show.legend = F)+
-  geom_line(aes(col = group), linewidth = 1.5)+
-  geom_point(data = forModeling, aes(x = PC2, y = degree_scl, col = situ))

### 2024-02-08 scripts/02.1_visualizeNetworks.qmd
moved network visualization to its own script, and created degree/strength vizes

---
title: "Network visualizations"
format: html
editor: visual
---

## Visualizing social networks

```{r}
#| include: false
library(tidyverse)
library(ggraph)
library(tidygraph)
library(igraph)
library(here)
```

```{r}
#| include: false
load(here("data", "calcSocial", "feedingGraphs.Rda"))
load(here("data", "calcSocial", "flightGraphs.Rda"))
load(here("data", "calcSocial", "roostingGraphs.Rda"))
load(here("data", "calcSocial", "networkMetrics.Rda"))
load(here("data", "mixedModels", "linked.Rda"))

# Have to join `linked` back to `networkMetrics` so we maintain the network values for those that don't ultimately get included in movement
info <- networkMetrics %>%
  left_join(linked %>%
              select(Nili_id, "season" = seasonUnique, daysTracked, propDaysTracked, birth_year, sex, space_use, movement, roost_div, space_use_log, age, age_group, type)) %>%
  mutate(focal = case_when(is.na(movement) & is.na(space_use) & is.na(roost_div) ~ FALSE,
                           TRUE ~ TRUE))
nrow(info) == nrow(networkMetrics) # TRUE
nrow(linked) < nrow(info) # TRUE

infos_flight <- info %>%
  filter(type == "flight") %>%
  group_by(season) %>%
  group_split()
infos_feeding <- info %>%
  filter(type == "feeding") %>%
  group_by(season) %>%
  group_split()
infos_roosting <- info %>%
  filter(type == "roosting") %>%
  group_by(season) %>%
  group_split()

flightGraphs <- map2(flightGraphs, infos_flight, ~{
  .x %>% 
    as_tbl_graph() %>%
    activate(nodes) %>%
    left_join(.y, by = c("name" = "Nili_id"))
})
feedingGraphs <- map2(feedingGraphs, infos_feeding, ~{
  .x %>% 
    as_tbl_graph() %>%
    activate(nodes) %>%
    left_join(.y, by = c("name" = "Nili_id"))
})
roostingGraphs <- map2(roostingGraphs, infos_roosting, ~{
  .x %>% 
    as_tbl_graph() %>%
    activate(nodes) %>%
    left_join(.y, by = c("name" = "Nili_id"))
})
```

```{r}
# Network visualizations ----------------------------------------------------------------
grph_deg <- function(graph, season_name, type){
  g <- graph %>%
    ggraph(layout = "fr")+
    geom_edge_link(alpha = 0.2)+
    geom_node_point(aes(size = normDegree, col = normDegree))+
    scale_color_viridis()+
    theme(legend.position = "none")+
    ggtitle(season_name)
  return(g)
}

grph_str <- function(graph, season_name, type){
  g <- graph %>%
    ggraph(layout = "fr")+
    geom_edge_link(alpha = 0.2)+
    geom_node_point(aes(size = normStrength, col = normStrength))+
    scale_color_viridis()+
    theme(legend.position = "none")+
    ggtitle(season_name)
  return(g)
}

flightPlots_degree <- map2(.x = flightGraphs, .y = season_names, ~grph_deg(.x, .y, type = "flight"))
feedingPlots_degree <- map2(.x = feedingGraphs, .y = season_names, ~grph_deg(.x, .y, type = "feeding"))
roostingPlots_degree <- map2(.x = roostingGraphs, .y = season_names, ~grph_deg(.x, .y, type = "roosting"))

flightPlots_strength <- map2(.x = flightGraphs, .y = season_names, ~grph_str(.x, .y, type = "flight"))
feedingPlots_strength <- map2(.x = feedingGraphs, .y = season_names, ~grph_str(.x, .y, type = "feeding"))
roostingPlots_strength <- map2(.x = roostingGraphs, .y = season_names, ~grph_str(.x, .y, type = "roosting"))

fl_d <- patchwork::wrap_plots(flightPlots_degree, nrow = 3, ncol = 3)
fe_d <- patchwork::wrap_plots(feedingPlots_degree, nrow = 3, ncol = 3)
ro_d <- patchwork::wrap_plots(roostingPlots_degree, nrow = 3, ncol = 3)
ggsave(fl_d, filename = here("fig", "networkGraphs", "degree", "flight.png"), width = 9, height = 7)
ggsave(fe_d, filename = here("fig", "networkGraphs", "degree", "feeding.png"), width = 9, height = 7)
ggsave(ro_d, filename = here("fig", "networkGraphs", "degree", "roosting.png"), width = 9, height = 7)

fl_s <- patchwork::wrap_plots(flightPlots_strength, nrow = 3, ncol = 3)
fe_s <- patchwork::wrap_plots(feedingPlots_strength, nrow = 3, ncol = 3)
ro_s <- patchwork::wrap_plots(roostingPlots_strength, nrow = 3, ncol = 3)
ggsave(fl_s, filename = here("fig", "networkGraphs", "strength", "flight.png"), width = 9, height = 7)
ggsave(fe_s, filename = here("fig", "networkGraphs", "strength", "feeding.png"), width = 9, height = 7)
ggsave(ro_s, filename = here("fig", "networkGraphs", "strength", "roosting.png"), width = 9, height = 7)
```

### 2024-02-08 scripts/02.0_calcSocial.R

@@ -92,6 +92,9 @@ load("data/calcSocial/roostingSRI.Rda")
 flightGraphs <- map(flightSRI, ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
 feedingGraphs <- map(feedingSRI,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
 roostingGraphs <- map(roostingSRI,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
+save(flightGraphs, file = "data/calcSocial/flightGraphs.Rda")
+save(feedingGraphs, file = "data/calcSocial/feedingGraphs.Rda")
+save(roostingGraphs, file = "data/calcSocial/roostingGraphs.Rda")
 
 # Describing the networks -------------------------------------------------
 # Network-level measures
@@ -173,30 +176,4 @@ zer %>% group_split(season) %>% map(., ~.x %>% pull(Nili_id) %>% unique() %>% so
 hist(networkMetrics$normDegree)
 hist(networkMetrics$normStrength)
 
-# Plotting ----------------------------------------------------------------
-library(ggraph)
-library(tidygraph)
-grph_deg <- function(graph, season_name, type){
-  g <- graph %>%
-    as_tbl_graph() %>%
-    mutate(normDegree = centrality_degree()/n()) %>%
-    ggraph(layout = "fr")+
-    geom_edge_link(alpha = 0.2)+
-    geom_node_point(aes(size = normDegree, col = normDegree))+
-    scale_color_viridis()+
-    theme(legend.position = "none")+
-    ggtitle(season_name)
-  return(g)
-  cat("Done with ", title, "\n")
-}
 
-flightPlots <- map2(.x = flightGraphs, .y = season_names, ~grph_deg(.x, .y, type = "flight"))
-feedingPlots <- map2(.x = feedingGraphs, .y = season_names, ~grph_deg(.x, .y, type = "feeding"))
-roostingPlots <- map2(.x = roostingGraphs, .y = season_names, ~grph_deg(.x, .y, type = "roosting"))
-
-fl <- patchwork::wrap_plots(flightPlots, nrow = 3, ncol = 3)
-fe <- patchwork::wrap_plots(feedingPlots, nrow = 3, ncol = 3)
-ro <- patchwork::wrap_plots(roostingPlots, nrow = 3, ncol = 3)
[…]

### 2024-02-14 scripts/01.1_categorizeMovement.qmd
updated with new space use PC

@@
 movementBehaviorScaled <- allMovementBehavior %>%
-  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex, homeRange, coreArea, coreAreaFidelity), # not scaling the space use params, because I'm not putting those into a PCA.
+  mutate(coreArea_log = log(coreArea),
+         homeRange_log = log(homeRange)) %>%
+  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex), 
                 function(x){as.numeric(as.vector(scale(x)))}))
@@
   dplyr::select(Nili_id, seasonUnique, daysTracked, propDaysTracked, birth_year, sex)
+
 space_use <- movementBehaviorScaled %>%
-  dplyr::select(Nili_id, seasonUnique, coreArea, homeRange, coreAreaFidelity)
+  dplyr::select(Nili_id, seasonUnique, coreArea_log, homeRange_log, coreAreaFidelity)
+
 movement <- movementBehaviorScaled %>%
-  dplyr::select(Nili_id, seasonUnique, mnDailyMnAlt, meanDFD, meanDMD, meanDDT, mnTort)
+  dplyr::select(Nili_id, seasonUnique, mnDailyMnAlt, meanDFD, meanDMD, meanDDT, mnTort) %>%
+  filter(!is.na(mnDailyMnAlt))
+
 roost_behavior <- movementBehaviorScaled %>%
-  dplyr::select(Nili_id, seasonUnique, propSwitch, shannon, uniqueRoosts)
+  dplyr::select(Nili_id, seasonUnique, propSwitch, shannon, uniqueRoosts) %>%
+  filter(!is.na(shannon))
 ```
@@
 
+## Space use PCA
+
+```{r}
+pca_space <- prcomp(space_use[,-c(1:3)]) # only including home range and coreAreaFidelity (because it would be redundant to include all three)
+fviz_pca_biplot(pca_space, axes = c(1,2), 
+                geom = c("point"), 
+                alpha.ind = 0.3,
+                repel = T)+
+  theme_classic()
+fviz_screeplot(pca_space, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
+contrib_space <- round(pca_roost$rotation[,1:3]*-100, 2) %>%
+  as.data.frame()
+contrib_space
+space_use <- space_use %>%
+  mutate(space_pc1 = pca_space$x[,1]*-1)
+```
+
 ## Variable distributions
@@
 red <- demo %>%
-  left_join(space_use %>% dplyr::select(Nili_id, seasonUnique, homeRange)) %>%
+  left_join(space_use %>% dplyr::select(Nili_id, seasonUnique, space_pc1)) %>%
   left_join(movement %>% dplyr::select(Nili_id, seasonUnique, movement_pc1)) %>%
   left_join(roost_behavior %>% dplyr::select(Nili_id, seasonUnique, roost_pc1)) %>%
-  rename("space_use" = "homeRange",
+  rename("space_use" = "space_pc1",
          "movement" = "movement_pc1",
-         "roost_div" = "roost_pc1")
-nrow(red) == nrow(demo) # checking that the number of rows hasn't changed
+         "roost_div" = "roost_pc1") %>%
+  filter(!is.na(movement))
+nrow(red) == nrow(movement) # checking that the number of rows hasn't changed
 red_long <- red %>%
@@
 
-Let's examine these predictor variables
+Let's examine these predictor variables.
 
@@
 
-We can see here that we have another problem.
-Since `movement` and `roost_div` are composite variables derived from PCAs, they are pretty close to normally distributed.
-Meanwhile, our space use variable is very right-skewed, because it's a measure of area.
-
-Let's log-tranform the space use variable to make the distributions more comparable and therefore make the modeling easier.
-
-```{r}
-#| include: false
-red <- red %>%
-  mutate(space_use_log = log(space_use))
-red_long <- red %>%
-  pivot_longer(cols = c("space_use", 
-                        "space_use_log", 
-                        "movement", 
-                        "roost_div"),
-               names_to = "var",
-               values_to = "value")
-```
-
-Re-examine the distributions of the variables after log-transforming:
-
-```{r}
-#| message: false
-#| warning: false
-red_long %>%
-  filter(var != "space_use") %>%
-  ggplot(aes(x = value, col = var, fill = var))+
-  geom_histogram()+ theme_minimal()+
-  facet_wrap(~var, scales = "free")+
-  scale_fill_manual(values = colors)+
-  scale_color_manual(values = colors)+
-  theme(legend.position = "none")+
-  ylab("Frequency")+xlab("")
-```
-
-Okay that's still not perfect, but it's a lot better.
-
-*Note also that this outcome may change quantitatively after we calculate the home ranges by better means (AKDE), but I don't expect it to be qualitatively different.
-AKDE calculations are currently in progress.*
+All variables are fairly close to normally distributed, since they come from PCAs and are scaled.
 
@@
 
-**Roost diversification** and **log-transformed home range** have a stronger correlation of `r round(cor(red$roost_div, red$space_use_log), 2)`.
+**Roost diversification** and **log-transformed home range** have a stronger correlation of `r round(cor(red$roost_div, red$space_use), 2)`.
 
@@
 #| warning: false
-cor <- round(cor(red$roost_div, red$space_use_log), 2)
+cor <- round(cor(red$roost_div, red$space_use), 2)
 red %>%
-  ggplot(aes(x = roost_div, y = space_use_log))+
+  ggplot(aes(x = roost_div, y = space_use))+
   geom_point(color = "chartreuse3")+ geom_smooth(method = "lm", col = "chartreuse3")+ theme_minimal()+
   xlab("Roost diversification")+ ylab("Space use (log)")+
-  ggplot2::annotate(geom = "text", x = 4, y = 11, label = paste0("Correlation = ", cor))
+  ggplot2::annotate(geom = "text", x = 3, y = -2, label = paste0("Correlation = ", cor))
 ```
@@
 #| warning: false
-cor <- round(cor(red$space_use_log, red$movement), 2)
+cor <- round(cor(red$space_use, red$movement), 2)
 red %>%
-  ggplot(aes(x = space_use_log, y = movement))+
+  ggplot(aes(x = space_use, y = movement))+
   geom_point(color = "orange2")+ geom_smooth(method = "lm", col = "orange2")+ theme_minimal()+
   xlab("Space use (log)") + ylab("Movement")+
-  ggplot2::annotate(geom = "text", x = 14, y = -4.5, label = paste0("Correlation = ", cor))
+  ggplot2::annotate(geom = "text", x = 2.5, y = -3, label = paste0("Correlation = ", cor))
 ```
@@
   mutate(season = stringr::str_extract(seasonUnique, "[a-z]+")) %>%
-  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+  pivot_longer(cols = c("movement", "roost_div", "space_use"),
                names_to = "var", values_to = "value") %>%
@@
 red %>%
-  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+  pivot_longer(cols = c("movement", "roost_div", "space_use"),
                names_to = "var", values_to = "value") %>%
@@
   filter(!is.na(sex), sex != "NA") %>%
-  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+  pivot_longer(cols = c("movement", "roost_div", "space_use"),
                names_to = "var", values_to = "value") %>%
@@
 #   filter(!is.na(sex), sex != "NA") %>%
-#   pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+#   pivot_longer(cols = c("movement", "roost_div", "space_use"),
 #                names_to = "var", values_to = "value") %>%
@@
 red %>%
-  pivot_longer(cols = c("movement", "roost_div", "space_use_log"),
+  pivot_longer(cols = c("movement", "roost_div", "space_use"),
                names_to = "var", values_to = "value") %>%
-  filter(var != "space_use") %>%
   ggplot(aes(x = season, y = value))+

### 2024-02-14 scripts/00.0_dataPrep.R
deleted data prep r script; now using the qmd only

@@ -1,664 +1,0 @@
-# Prepare data for movement analysis
-# Setup -------------------------------------------------------------------
-## Load packages -----------------------------------------------------------
-library(vultureUtils)
-library(sf)
-library(tidyverse)
-library(move)
-library(feather)
-library(readxl)
-library(elevatr)
-
-## Download data from movebank
-base::load("movebankCredentials/pw.Rda")
-MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
-rm(pw)
-
-# Hatzofe INPA data
-# inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
-# inpadf <- methods::as(inpa, "data.frame")
-# inpadf_touse <- inpadf %>%
-#   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
-#          year = as.numeric(lubridate::year(timestamp))) %>%
-#   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
-# write_feather(inpadf_touse, "data/derived/inpadf_touse.feather")
-inpadf_touse <- read_feather("data/derived/inpadf_touse.feather")
-# Ornitela: download data from movebank (just a subset of the times for now)
-minDate <- "2020-09-01 00:00"
-maxDate <- "2023-05-15 11:59"
-# dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-# write_feather(dat, "data/derived/dat.feather")
-dat <- read_feather("data/derived/dat.feather")
-# number of unique individuals
-length(unique(dat$trackId)) # 127
-length(unique(inpadf_touse$trackId)) # 77
-# 
-# # Will there be a problem joining these?
-# names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
-# names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.
-# 
-# # Add the dataset names so we can keep track of where the data comes from
-# inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
[…]

### 2024-02-14 scripts/00.0_dataPrep.qmd
fixed file names to use here::here()

@@
 library(elevatr)
+library(here)
 ```
@@
 ```{r}
-base::load("movebankCredentials", "pw.Rda")
+base::load(here::here("movebankCredentials", "pw.Rda"))
 MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
@@
   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
-write_feather(inpa, "data/dataPrep/inpa.feather")
+write_feather(inpa, here::here("data", "dataPrep", "inpa.feather"))
 
@@
 ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
-write_feather(ornitela, "data/dataPrep/ornitela.feather")
+write_feather(ornitela, here::here("data", "dataPrep", "ornitela.feather"))
 
@@
 ```{r load-inpa-ornitela}
-inpa <- read_feather("data/dataPrep/inpa.feather")
-ornitela <- read_feather("data/dataPrep/ornitela.feather")
+inpa <- read_feather(here::here("data", "dataPrep", "inpa.feather"))
+ornitela <- read_feather(here::here("data", "dataPrep", "ornitela.feather"))
 ```
@@
 # Add the Nili_id
-ww <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+ww <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
   dplyr::select(Nili_id, Movebank_id) %>%
@@
 length(unique(joined$Nili_id))
-write_feather(joined, "data/dataPrep/joined.feather")
+write_feather(joined, here::here("data", "dataPrep", "joined.feather"))
 
@@
 ```{r load-hospital}
-joined <- read_feather("data/dataPrep/joined.feather")
-periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove")
+joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
+periods_to_remove <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
 ```
@@
 
-write_feather(removed_periods, "data/dataPrep/removed_periods.feather")
+write_feather(removed_periods, here::here("data", "dataPrep", "removed_periods.feather"))
 
@@
 ```{r}
-removed_periods <- read_feather("data/dataPrep/removed_periods.feather")
+removed_periods <- read_feather(here::here("data", "dataPrep", "removed_periods.feather"))
 ```
@@
 roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-save(roosts0, file = "data/dataPrep/roosts0.Rda")
+save(roosts0, file = here::here("data", "dataPrep", "roosts0.Rda"))
 
@@
 # Load information about the capture sites
-capture_sites <- read.csv("data/raw/capture_sites.csv")
-load("data/dataPrep/roosts0.Rda")
-AllCarmelDates <- read.csv("data/raw/all_captures_carmel_2010-2021.csv")
+capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
+load(here::here("data", "dataPrep", "roosts0.Rda"))
+AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))
 ```
@@
 
-write_feather(removed_captures, "data/dataPrep/removed_captures.feather")
+write_feather(removed_captures, here::here("data", "dataPrep", "removed_captures.feather"))
 
@@
 #| message: false
-removed_captures <- read_feather("data/dataPrep/removed_captures.feather")
-age_sex <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
+removed_captures <- read_feather(here::here("data", "dataPrep", "removed_captures.feather"))
+age_sex <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
   dplyr::select(Nili_id, birth_year, sex) %>%
@@
 
-write_feather(with_age_sex, "data/dataPrep/with_age_sex.feather")
+write_feather(with_age_sex, here::here("data", "dataPrep", "with_age_sex.feather"))
 
@@
 ```{r}
-mask <- sf::st_read("data/raw/CutOffRegion.kml")
-with_age_sex <- read_feather("data/dataPrep/with_age_sex.feather")
+mask <- sf::st_read(here::here("data", "raw", "CutOffRegion.kml"))
+with_age_sex <- read_feather(here::here("data", "dataPrep", "with_age_sex.feather"))
 ```
@@
 
-save(cleaned, file = "data/dataPrep/cleaned.Rda")
+save(cleaned, file = here::here("data", "dataPrep", "cleaned.Rda"))
 
@@
 length(season_names) # should also be 9
-save(seasons_list, file = "data/dataPrep/seasons_list.Rda")
-save(season_names, file = "data/dataPrep/season_names.Rda")
+save(seasons_list, file = here::here("data", "dataPrep", "seasons_list.Rda"))
+save(season_names, file = here::here("data", "dataPrep", "season_names.Rda"))
 
@@
 ```{r}
-load("data/dataPrep/seasons_list.Rda")
-load("data/dataPrep/season_names.Rda")
+load(here::here("data", "dataPrep", "seasons_list.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
 ```
@@
 save(low_fix_rate_indivs, 
-     file = "data/dataPrep/low_fix_rate_indivs.Rda")
+     file = here::here("data", "dataPrep", "low_fix_rate_indivs.Rda"))
 # Don't need to save with_modes because we've already used this step to identify low fix rate individuals.
@@
 ```{r}
-load("data/dataPrep/seasons_list.Rda") 
-load("data/dataPrep/low_fix_rate_indivs.Rda")
+load(here::here("data", "dataPrep", "seasons_list.Rda")) 
+load(here::here("data", "dataPrep", "low_fix_rate_indivs.Rda"))
 ```
@@
 removed_lfr <- map2(seasons_list, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
-save(removed_lfr, file = "data/dataPrep/removed_lfr.Rda")  
+save(removed_lfr, file = here::here("data", "dataPrep", "removed_lfr.Rda"))  
 # tidy up 
@@
 ```{r}
-load("data/dataPrep/removed_lfr.Rda")
+load(here::here("data", "dataPrep", "removed_lfr.Rda"))
 ```
@@
                    remove = F))
-save(roosts, file = "data/dataPrep/roosts.Rda")
+save(roosts, file = here::here("data", "dataPrep", "roosts.Rda"))
 
@@
 ```{r}
-load("data/dataPrep/removed_lfr.Rda")
-load("data/dataPrep/season_names.Rda")
+load(here::here("data", "dataPrep", "removed_lfr.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
 ```
@@
 
-save(removed_nighttime, file = "data/dataPrep/removed_nighttime.Rda")
+save(removed_nighttime, file = here::here("data", "dataPrep", "removed_nighttime.Rda"))
 
@@
 ```{r}
-load("data/dataPrep/removed_nighttime.Rda")
-load("data/dataPrep/season_names.Rda")
+load(here::here("data", "dataPrep", "removed_nighttime.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
 ```
@@
 
-save(removed_northern, file = "data/dataPrep/removed_northern.Rda")
+save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))
 
@@
 ```{r}
-load("data/dataPrep/removed_northern.Rda")
+load(here::here("data", "dataPrep", "removed_northern.Rda"))
 ```
@@
 
-save(removed_lowppd, file = "data/dataPrep/removed_lowppd.Rda")
+save(removed_lowppd, file = here::here("data", "dataPrep", "removed_lowppd.Rda"))
 
@@
 ```{r}
-load("data/dataPrep/removed_lowppd.Rda")
-load("data/dataPrep/season_names.Rda")
+load(here::here("data", "dataPrep", "removed_lowppd.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
 ```
@@
 
-save(removed_too_few_days, file = "data/dataPrep/removed_too_few_days.Rda")
-save(season_durations, file = "data/dataPrep/season_durations.Rda")
+save(removed_too_few_days, file = here::here("data", "dataPrep", "removed_too_few_days.Rda"))
+save(season_durations, file = here::here("data", "dataPrep", "season_durations.Rda"))
 
@@
 ```{r}
-load("data/dataPrep/removed_too_few_days.Rda")
+load(here::here("data", "dataPrep", "removed_too_few_days.Rda"))
 ```
@@
 elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
-save(elevs_z10, file = "data/dataPrep/elevs_z10.Rda")
-save(sf, file = "data/dataPrep/sf.Rda")
-```
-
-```{r}
-load("data/dataPrep/elevs_z10.Rda")
-load("data/dataPrep/sf.Rda")
+save(elevs_z10, file = here::here("data", "dataPrep", "elevs_z10.Rda"))
+save(sf, file = here::here("data", "dataPrep", "sf.Rda"))
+```
+
+```{r}
+load(here::here("data", "dataPrep", "elevs_z10.Rda"))
+load(here::here("data", "dataPrep", "sf.Rda"))
 ```
@@
 
-save(with_altitudes, file = "data/dataPrep/with_altitudes.Rda")
+save(with_altitudes, file = here::here("data", "dataPrep", "with_altitudes.Rda"))
 

### 2024-02-14 scripts/00.0_dataPrep_copy.qmd
fixed file names to use here::here()

@@
 #   filter(lubridate::year(dateOnly) == "2022")
-write_csv(year22, file = here::here("data", "dataPrep", "year22.csv"))
+# write_csv(year22, file = here::here("data", "dataPrep", "year22.csv"))
 ```
@@
 ```{r load-hospital}
-year22 <- read_csv("data/dataPrep/year22.csv")
-periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove")
+year22 <- read_csv(here("data", "dataPrep", "year22.csv"))
+periods_to_remove <- read_excel(here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
 ```
@@
 length(unique(removal_annotated$Nili_id))
-nrow(year22) == nrow(removal_annotated) # true
-nrow(removed_periods) != nrow(removal_annotated) # true
 

### 2024-02-14 scripts/01.1_akde.R
updated to seasonUnique to avoid causing problems for joins

@@ -104,7 +104,7 @@ for(i in 1:length(stats_w_95)){
            "dof_bandwidth" = "bandwidth") %>%
     mutate(n_abs_area = map_dbl(telems_list[[i]], nrow),
            eff_ss_prop = round(n_eff_area/n_abs_area, 2),
-           season = season_names[i])
+           seasonUnique = season_names[i])
   stats_w_95[[i]] <- stats_weighted
 }
 stats_w_95_df <- purrr::list_rbind(stats_w_95) %>% mutate(level = 0.95)
@@ -125,7 +125,7 @@ for(i in 1:length(stats_w_50)){
            "dof_bandwidth" = "bandwidth") %>%
     mutate(n_abs_area = map_dbl(telems_list[[i]], nrow),
            eff_ss_prop = round(n_eff_area/n_abs_area, 2),
-           season = season_names[i])
+           seasonUnique = season_names[i])
   stats_w_50[[i]] <- stats_weighted
 }
 stats_w_50_df <- purrr::list_rbind(stats_w_50) %>% mutate(level = 0.5)

### 2024-02-20 scripts/repeatability.R
removed daily graphs and added interval ones

@@ -319,62 +319,135 @@ for(i in 1:length(ndays)){
     }
     return(out)
   })
+  rm(fl)
+  rm(fe)
+  rm(ro)
 }
-metrics_flight <- map(metrics_flight, ~purrr::list_rbind(.x))
-metrics_feeding <- map(metrics_feeding, ~purrr::list_rbind(.x))
-metrics_roosting <- map(metrics_roosting, ~purrr::list_rbind(.x))
+metrics_flight_df <- map(metrics_flight, ~purrr::list_rbind(.x))
+metrics_feeding_df <- map(metrics_feeding, ~purrr::list_rbind(.x))
+metrics_roosting_df <- map(metrics_roosting, ~purrr::list_rbind(.x))
 
-metrics <- bind_rows(purrr::list_rbind(metrics_flight), 
-                     purrr::list_rbind(metrics_feeding), 
-                     purrr::list_rbind(metrics_roosting))
+metrics <- bind_rows(purrr::list_rbind(metrics_flight_df), 
+                     purrr::list_rbind(metrics_feeding_df), 
+                     purrr::list_rbind(metrics_roosting_df))
 
-### XXX START HERE
-# Daily graph plots -------------------------------------------------------
-tbl_graphs_flight <- map(flightDays_graphs, ~as_tbl_graph(.x) %>%
-                           activate(nodes) %>%
-                           mutate(degree = igraph::degree(.))) 
-tbl_graphs_feeding <- map(feedingDays_graphs, ~as_tbl_graph(.x) %>%
-                            activate(nodes) %>%
-                            mutate(degree = igraph::degree(.))) 
-tbl_graphs_roosting <- map(roostDays_graphs, ~as_tbl_graph(.x) %>%
-                             activate(nodes) %>%
-                             mutate(degree = igraph::degree(.))) 
+# Create tbl graphs and join metrics
+graphs_flight_tbl <- vector(mode = "list", length = length(ndays))
+graphs_feeding_tbl <- vector(mode = "list", length = length(ndays))
+graphs_roosting_tbl <- vector(mode = "list", length = length(ndays))
+for(i in 1:length(ndays)){
+  fl <- graphs_flight[[i]]
+  fe <- graphs_feeding[[i]]
+  ro <- graphs_roosting[[i]]
+  
[…]

### 2024-02-20 scripts/00.0_dataPrep.qmd
removed northern indivs from social as well as movement. Began to parallelize.

@@
 library(here)
+library(furrr)
+future::plan(future::multisession, workers = 10) # this will last for the whole R session.
 ```
@@
 
+## Restrict to southern population
+
+Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, <https://epsg.io/32636)>
+
+*Inputs:*
+
+-   `data/dataPrep/removed_lfr.Rda`
+-   `data/dataPrep/season_names.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_northern.Rda`
+
+```{r}
+load(here::here("data", "dataPrep", "removed_lfr.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
+```
+
+```{r}
+#| eval: false
+# Transform to SF object, so we can get centroids
+seasons_sf <- map(removed_lfr, ~.x %>%
+                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
+                   sf::st_set_crs("WGS84") %>%
+                   sf::st_transform(32636))
+
+## Get centroids, so we can see who's "southern" for that season.
+centroids <- map(seasons_sf, ~.x %>%
+                   group_by(Nili_id) %>%
+                   summarize(geometry = st_union(geometry)) %>%
+                   st_centroid())
+
+## Examine a histogram of centroid latitudes 
+walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
+
+## Get southern individuals for each season, so we can filter the data
+southern_indivs <- map(centroids, ~.x %>%
+                        filter(st_coordinates(.)[,2] < 3550000) %>%
+                        pull(Nili_id))
+
+## Remove individuals not in the south
+removed_northern <- map2(.x = removed_lfr, 
+                         .y = southern_indivs, ~.x 
+                         %>% filter(Nili_id %in% .y))
+save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))
+
+# tidy up
+rm(removed_lfr)
+rm(season_names)
+rm(seasons_sf)
+rm(centroids)
+rm(southern_indivs)
+rm(removed_northern)
+rm(df)
+gc()
+```
+
 ### Data for social analysis
@@
 
--   `data/dataPrep/removed_lfr.Rda`
+-   `data/dataPrep/removed_northern.Rda`
 
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_lfr.Rda"))
+load(here::here("data", "dataPrep", "removed_northern.Rda"))
 ```
@@
 # Get roosts for each season ----------------------------------------------
-roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
+roosts <- furrr::future_map(removed_northern, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"), .progress = T)
 roosts <- roosts %>%
-  map(., ~st_as_sf(.x, crs = "WGS84", 
+  furrr::future_map(., ~st_as_sf(.x, crs = "WGS84", 
                    coords = c("location_long", "location_lat"), 
-                   remove = F))
+                   remove = F), .progress = T)
 save(roosts, file = here::here("data", "dataPrep", "roosts.Rda"))
@@
 # Tidy up
-rm(removed_lfr)
+rm(removed_northern)
 rm(roosts)
@@
 
--   `data/dataPrep/removed_lfr.Rda`
+-   `data/dataPrep/removed_northern.Rda`
 -   `data/dataPrep/season_names.Rda`
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_lfr.Rda"))
+load(here::here("data", "dataPrep", "removed_northern.Rda"))
 load(here::here("data", "dataPrep", "season_names.Rda"))
@@
 # Remove nighttime points -------------------------------------------------
-removed_nighttime <- map(removed_lfr, ~{
+removed_nighttime <- map(removed_northern, ~{
   times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
@@
 rm(removed_nighttime)
-rm(removed_lfr)
+rm(removed_northern)
 rm(season_names)
-gc()
-```
-
-## Restrict to southern population
-
-Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, <https://epsg.io/32636)>
-
-*Inputs:*
-
--   `data/dataPrep/removed_nighttime.Rda`
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/removed_northern.Rda`
-
-```{r}
-load(here::here("data", "dataPrep", "removed_nighttime.Rda"))
-load(here::here("data", "dataPrep", "season_names.Rda"))
-```
-
-```{r}
-#| eval: false
-# Transform to SF object, so we can get centroids
-seasons_sf <- map(removed_nighttime, ~.x %>%
-                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
-                   sf::st_set_crs("WGS84") %>%
-                   sf::st_transform(32636))
-
-## Get centroids, so we can see who's "southern" for that season.
-centroids <- map(seasons_sf, ~.x %>%
-                   group_by(Nili_id) %>%
-                   summarize(geometry = st_union(geometry)) %>%
-                   st_centroid())
-
-## Examine a histogram of centroid latitudes 
-walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
-
-## Get southern individuals for each season, so we can filter the data
-southern_indivs <- map(centroids, ~.x %>%
-                        filter(st_coordinates(.)[,2] < 3550000) %>%
-                        pull(Nili_id))
-
-## Remove individuals not in the south
-before <- removed_nighttime
-removed_northern <- map2(.x = removed_nighttime, 
-                         .y = southern_indivs, ~.x 
-                         %>% filter(Nili_id %in% .y))
-after <- removed_northern
-
-beforeRows <- map_dbl(before, nrow)
-afterRows <- map_dbl(after, nrow)
-beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
-afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
-df <- data.frame(season = season_names, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
-df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
-df
-
-save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))
-
-# tidy up
-rm(removed_nighttime)
-rm(season_names)
-rm(seasons_sf)
-rm(centroids)
-rm(southern_indivs)
-rm(before)
-rm(after)
-rm(removed_northern)
-rm(df)
-rm(afterIndivs)
-rm(afterRows)
-rm(beforeIndivs)
-rm(beforeRows)
 gc()

### 2024-02-21 scripts/repeatability.R
created time plots

@@ -9,6 +9,7 @@ library(zoo)
 
 # Repeatability across seasons --------------------------------------------
 load("data/mixedModels/linked.Rda")
+future::plan(future::multisession(), workers = 10)
 
 ## Just using the traits we've already calculated and plotting across seasons, and coloring by the average value for each individual across all of its seasons, let's see if we get gradients
 
@@ -87,7 +88,7 @@ annotated <- map(breaks, ~testseason %>%
 
 # XXX I'll have to find a way to include the last few dates in these (i.e. add one more element to "breaks" that's [last one + 1 interval unit]), but for now it's fine to leave them as NA.
 
-split_sfData <- purrr::map(annotated, ~{
+split_sfData <- map(annotated, ~{
   cat("grouping\n")
   lst <- .x %>%
     group_by(int) %>%
@@ -138,8 +139,9 @@ for(split in 1:length(ndays)){
   datalist <- split_sfData[[split]]
   roostlist <- split_sfData_roosts[[split]]
   cat("working on flight\n")
-  future::plan(future::multisession, workers = 16)
+  future::plan(future::multisession, workers = 5)
   fl <- suppressWarnings(furrr::future_map(datalist, ~{
+    library(sf)
     vultureUtils::getFlightEdges(.x, roostPolygons = roostPolygons,
                                  distThreshold = 1000, idCol = "Nili_id",
                                  return = "sri")
@@ -242,11 +244,11 @@ for(i in 1:length(ndays)){
 graphs_flight <- vector(mode = "list", length = length(ndays))
 graphs_feeding <- vector(mode = "list", length = length(ndays))
 graphs_roosting <- vector(mode = "list", length = length(ndays))
+future::plan(future::multisession, workers = 5)
 for(i in 1:length(ndays)){
   flight <- flout[[i]]
   feeding <- feout[[i]]
   roosting <- roout[[i]]
-  future::plan(future::multisession, workers = 16)
   flightgraphs <- furrr::future_map(flight, ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T),
                                     .progress = T)
   feedinggraphs <- furrr::future_map(feeding, ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T),
@@ -330,6 +332,7 @@ metrics_roosting_df <- map(metrics_roosting, ~purrr::list_rbind(.x))
[…]

### 2024-02-21 scripts/00.0_dataPrep.qmd
restricted to southern individuals for both movement and social

@@
 
+## Restrict to southern population
+
+Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, <https://epsg.io/32636)>
+
+*Inputs:*
+
+-   `data/dataPrep/seasons_list.Rda`
+-   `data/dataPrep/season_names.Rda`
+
+*Outputs:*
+
+-   `data/dataPrep/removed_northern.Rda`
+
+```{r}
+load(here::here("data", "dataPrep", "seasons_list.Rda"))
+load(here::here("data", "dataPrep", "season_names.Rda"))
+```
+
+```{r}
+#| eval: false
+# Transform to SF object, so we can get centroids
+seasons_sf <- map(seasons_list, ~.x %>%
+                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
+                   sf::st_set_crs("WGS84") %>%
+                   sf::st_transform(32636))
+
+## Get centroids, so we can see who's "southern" for that season.
+centroids <- map(seasons_sf, ~.x %>%
+                   group_by(Nili_id) %>%
+                   summarize(geometry = st_union(geometry)) %>%
+                   st_centroid())
+
+## Examine a histogram of centroid latitudes 
+#walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
+
+## Get southern individuals for each season, so we can filter the data
+southern_indivs <- map(centroids, ~.x %>%
+                        filter(st_coordinates(.)[,2] < 3550000) %>%
+                        pull(Nili_id))
+
+## Remove individuals not in the south
+removed_northern <- map2(.x = seasons_list, 
+                         .y = southern_indivs, ~.x 
+                         %>% filter(Nili_id %in% .y))
+save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))
+
+# tidy up
+rm(seasons_list)
+rm(season_names)
+rm(seasons_sf)
+rm(centroids)
+rm(southern_indivs)
+rm(removed_northern)
+rm(df)
+gc()
+```
+
 ## Identify individuals with low fix rate
@@
 
--   `data/dataPrep/seasons_list.Rda`
+-   `data/dataPrep/removed_northern.Rda`
 -   `data/dataPrep/season_names.Rda`
@@
 ```{r}
-load(here::here("data", "dataPrep", "seasons_list.Rda"))
+load(here::here("data", "dataPrep", "removed_northern.Rda"))
 load(here::here("data", "dataPrep", "season_names.Rda"))
@@
 # Calculate daily modes and add them to the dataset (using the function defined above)
-with_modes <- map(seasons_list, ~.x %>%
+with_modes <- map(removed_northern, ~.x %>%
                  group_by(dateOnly, Nili_id) %>%
@@
 # Tidy up
-rm(seasons_list)
+rm(removed_northern)
 rm(season_names)
@@
 
--   `data/dataPrep/seasons_list.Rda`
+-   `data/dataPrep/removed_northern.Rda`
 
@@
 ```{r}
-load(here::here("data", "dataPrep", "seasons_list.Rda")) 
+load(here::here("data", "dataPrep", "removed_northern.Rda")) 
 load(here::here("data", "dataPrep", "low_fix_rate_indivs.Rda"))
@@
 ```{r}
-removed_lfr <- map2(seasons_list, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
+removed_lfr <- map2(removed_northern, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
 save(removed_lfr, file = here::here("data", "dataPrep", "removed_lfr.Rda"))  
 # tidy up 
-rm(seasons_list) 
+rm(removed_northern) 
 rm(low_fix_rate_indivs) 
@@
 ```
+
+## Remove high-frequency fixes
 
@@
 
-## Restrict to southern population
-
-Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, <https://epsg.io/32636)>
-
-*Inputs:*
-
--   `data/dataPrep/removed_lfr.Rda`
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/removed_northern.Rda`
-
-```{r}
-load(here::here("data", "dataPrep", "removed_lfr.Rda"))
-load(here::here("data", "dataPrep", "season_names.Rda"))
-```
-
-```{r}
-#| eval: false
-# Transform to SF object, so we can get centroids
-seasons_sf <- map(removed_lfr, ~.x %>%
-                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
-                   sf::st_set_crs("WGS84") %>%
-                   sf::st_transform(32636))
-
-## Get centroids, so we can see who's "southern" for that season.
-centroids <- map(seasons_sf, ~.x %>%
-                   group_by(Nili_id) %>%
-                   summarize(geometry = st_union(geometry)) %>%
-                   st_centroid())
-
-## Examine a histogram of centroid latitudes 
-walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.
-
-## Get southern individuals for each season, so we can filter the data
-southern_indivs <- map(centroids, ~.x %>%
-                        filter(st_coordinates(.)[,2] < 3550000) %>%
-                        pull(Nili_id))
-
-## Remove individuals not in the south
-removed_northern <- map2(.x = removed_lfr, 
-                         .y = southern_indivs, ~.x 
-                         %>% filter(Nili_id %in% .y))
-save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))
-
-# tidy up
-rm(removed_lfr)
-rm(season_names)
-rm(seasons_sf)
-rm(centroids)
-rm(southern_indivs)
-rm(removed_northern)
-rm(df)
-gc()
-```
-
 ### Data for social analysis
@@
 But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric.
-So when we do the social analysis, we will use `seasons_list` and not anything further along.
+So when we do the social analysis, we will use `removed_lfr` and not anything further along.
 
@@
 
--   `data/dataPrep/removed_northern.Rda`
+-   `data/dataPrep/downsampled_10min_forSocial.Rda`
 
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_northern.Rda"))
+load(here::here("data", "dataPrep", "downsampled_10min_forSocial.Rda"))
 ```
@@
 # Get roosts for each season ----------------------------------------------
-roosts <- furrr::future_map(removed_northern, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"), .progress = T)
+roosts <- map(downsampled_10min_forSocial, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"), .progress = T)
 roosts <- roosts %>%
-  furrr::future_map(., ~st_as_sf(.x, crs = "WGS84", 
+  map(., ~st_as_sf(.x, crs = "WGS84", 
                    coords = c("location_long", "location_lat"), 
@@
 # Tidy up
-rm(removed_northern)
+rm(downsampled_10min_forSocial)
 rm(roosts)
@@
 
--   `data/dataPrep/removed_northern.Rda`
+-   `data/dataPrep/removed_lfr.Rda`
 -   `data/dataPrep/season_names.Rda`
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_northern.Rda"))
+load(here::here("data", "dataPrep", "removed_lfr.Rda"))
 load(here::here("data", "dataPrep", "season_names.Rda"))
@@
 # Remove nighttime points -------------------------------------------------
-removed_nighttime <- map(removed_northern, ~{
+removed_nighttime <- map(removed_lfr, ~{
   times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
@@
     dplyr::filter(daylight == "day")
-})
+}, .progress = T)
 
@@
 rm(removed_nighttime)
-rm(removed_northern)
+rm(removed_lfr)
 rm(season_names)
@@
 
--   `data/dataPrep/removed_northern.Rda`
+-   `data/dataPrep/removed_nighttime.Rda`
 
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_northern.Rda"))
+load(here::here("data", "dataPrep", "removed_nighttime.Rda"))
 ```
@@
 # Calculate number of points per day for each individual
-ppd <- map_dfr(removed_northern, ~.x %>%
+ppd <- map_dfr(removed_nighttime, ~.x %>%
                  st_drop_geometry() %>%
@@
                  summarize(n = n(),
-                           minBatt = min(battery_charge_percent, na.rm = T)))
+                           minBatt = min(battery_charge_percent, na.rm = T)), .progress = T)
 
@@
 ## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50%
-removed_lowppd <- map(removed_northern, ~.x %>%
+removed_lowppd <- map(removed_nighttime, ~.x %>%
                  group_by(Nili_id, dateOnly) %>%
@@
                        filter(nDaysTracked >= 30) %>%
-                       pull(Nili_id))
+                       pull(Nili_id), .progress = T)
 
@@
                 filter(Nili_id %in% .y))
-
-beforeIndivs <- map_dbl(removed_lowppd, ~length(unique(.x$Nili_id)))
-afterIndivs <- map_dbl(removed_too_few_days, ~length(unique(.x$Nili_id)))
 
@@
 rm(removed_too_few_days)
-rm(beforeIndivs)
-rm(afterIndivs)
 gc()
@@
 sf <- map(removed_too_few_days, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
-elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
+elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10), .progress = T)
 save(elevs_z10, file = here::here("data", "dataPrep", "elevs_z10.Rda"))
@@
 #| eval: false
-groundElev_z10 <- map2(elevs_z10, sf, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84.
+groundElev_z10 <- map2(elevs_z10, sf, ~raster::extract(x = .x, y = .y), .progress = T) # have to convert the crs because the raster from elevatr is in WGS84.
 

### 2024-02-21 scripts/02.0_calcSocial.R
running edges in future

@@ -4,11 +4,13 @@ library(vultureUtils)
 library(igraph)
 library(tidyverse)
 library(sf)
+library(future)
+future::plan(future::multisession(), workers = 10)
 
 ## Data ---------------------------------------------------------------
 load("data/dataPrep/downsampled_10min_forSocial.Rda") # this is the last dataset we produced in the dataPrep script before moving on to the further cleaning for movement, so this is the one we're going to use for the social interactions.
 sfdata <- map(downsampled_10min_forSocial, ~st_as_sf(.x, coords = c("location_long", "location_lat"), crs = "WGS84", remove = F))
-datasetAssignments <- map(downsampled_10min_forSocial, ~.x %>% select(Nili_id, dataset, seasonUnique) %>% distinct() %>% group_by(Nili_id, seasonUnique) %>%
+datasetAssignments <- map(downsampled_10min_forSocial, ~.x %>% dplyr::select(Nili_id, dataset, seasonUnique) %>% distinct() %>% group_by(Nili_id, seasonUnique) %>%
                             mutate(dataset = case_when(n() > 1 ~ "both",
                                                        TRUE ~ dataset)) %>%
                             ungroup() %>%
@@ -22,15 +24,13 @@ load("data/dataPrep/season_names.Rda")
 load("data/dataPrep/roosts.Rda") # XXX load
 load("data/derived/cc.Rda")
 
-
-# Investigate the data ----------------------------------------------------
-# How many individuals do we have with vs. without excluding those with a lower sampling rate?
-all <- map_dbl(sfdata, ~length(unique(.x$Nili_id)))
-nindivs <- map(sfdata, ~.x %>% sf::st_drop_geometry() %>% select(Nili_id, dataset) %>% distinct() %>% group_by(dataset) %>% summarize(n = length(unique(Nili_id)))) %>% map2(.x = ., .y = season_names, ~.x %>% mutate(seasonUnique = .y)) %>% purrr::list_rbind() 
-nindivs # so generally more ornitela than inpa individuals, but by the later seasons we do have double digits of inpa individuals.
-
 # Social Networks ---------------------------------------------------------
 flight <- vector(mode = "list", length = length(sfdata))
+future::plan(future::multisession(), workers = 10)
+flight <- furrr::future_map(sfdata, ~vultureUtils::getFlightEdges(.x, roostPolygons = roostPolygons,
+                                                                  distThreshold = 1000, idCol = "Nili_id",
[…]

### 2024-02-21 scripts/temp_multilayer_test.R
tested out hourly/daily

@@ -4,26 +4,18 @@ library(lubridate)
 library(tnet)
 library(sna)
 
-load("data/derived/flightSeasons_mode10_edges.Rda")
-fs <- flightSeasons_mode10_edges
-testseason <- fs[[5]] # just doing this for one arbitrary season
+load("data/calcSocial/flightEdges.Rda")
+testseason <- flightEdges[[5]] # just doing this for one arbitrary season
 
-# Make a binary edgelist
+# Make a binary edgelist, grouping by day
 bin <- testseason %>%
   mutate(day = lubridate::date(minTimestamp)) %>%
-  mutate(hour = floor_date(minTimestamp, unit = "hours")) %>%
-  select(ID1, ID2, day, hour) %>%
+  #mutate(hour = floor_date(minTimestamp, unit = "hours")) %>%
+  select(ID1, ID2, day#, 
+         #hour
+         ) %>%
   distinct() %>%
   mutate(weight = 1)
-
-# Plot associations
-bin %>%
-  group_by(hour) %>% 
-  summarize(n = n()) %>%
-  ggplot(aes(x = hour, y = n))+
-  geom_point()+
-  geom_path(alpha = 0.5)+
-  geom_smooth(method = "lm") # generally more in-flight interactions over the course of the season, which makes sense because it's warming up.
 
 opp_edges <- bin
 names(opp_edges)[1:2] <- c("ID2", "ID1")
@@ -32,7 +24,119 @@ all_bin <- bind_rows(bin, opp_edges) # attach
 
 # Add layer numbers (days) to make the next step easier
 all_bin <- all_bin %>%
+  arrange(day, ID1, ID2) %>%
+  mutate(layer = as.integer(as.factor(day))) %>%
   select(-day) %>%
+  rename("layer1" = "layer") %>%
+  mutate(layer2 = layer1) %>%
+  relocate(layer1, .after = "ID1") %>%
+  relocate(layer2, .after = "ID2")
+
+self_edges <- expand.grid(as.character(unique(all_bin$ID1)), 1:(max(all_bin$layer1)-1)) # create all possible combinations of indiv ID and layer
+colnames(self_edges) <- c("ID1", "layer1")
+
+self_edges <- self_edges %>%
+  mutate(ID2 = ID1, layer2 = layer1 + 1, weight = 1)
+self_edges2 <- self_edges %>%
+  mutate(layer1 = layer1 + 1, layer2 = layer2-1)
[…]

### 2024-02-21 scripts/muxLib DF.R
included some information points to debug multilayer function

@@ -1124,11 +1124,12 @@ GetMultilayerReducibility <- function(SupraAdjacencyMatrix,Layers,Nodes,Method,T
       
       print(paste("DEBUG:", i, SingleLayerEntropy[i]))
     }
-    
+    cat("finished singlelayerentropy\n")
     ###########################
     #JSD
     ###########################
     JSD <- Matrix(0, Layers, Layers)
+    cat("initialized jsd\n")
     
     if(Type=="Ordinal"){
       #this line is required, because a 0 means no distance..
@@ -1139,21 +1140,26 @@ GetMultilayerReducibility <- function(SupraAdjacencyMatrix,Layers,Nodes,Method,T
         JSD[i,j] <- GetJensenShannonDivergence(NodesTensor[[i]], NodesTensor[[j]], 
                                                SingleLayerEntropy[i],SingleLayerEntropy[j])
         #JSD[j,i] <- JSD[i,j]
-      }  
+      }
+      cat("got jsd for ordinal\n")
     }
     if(Type=="Categorical"){
+      cat("Calculating JSDs\n")
       for(i in 1:(Layers-1)){
         for(j in (i+1):Layers){
           JSD[i,j] <- GetJensenShannonDivergence(NodesTensor[[i]], NodesTensor[[j]], 
                                                  SingleLayerEntropy[i],SingleLayerEntropy[j])
           JSD[j,i] <- JSD[i,j]
         }
+        cat("Finished with layer", i, "out of", layers, "\n")
       }
+      cat("got jsd for categorical\n")
     }
     JSD <- sqrt(JSD)
-    
+
     #Quality function
     hc <- hclust(as.dist(JSD), method=Method)
+    cat("clustered \n")
     
     ## list of merging operations
     #see http://stackoverflow.com/questions/18215184/how-to-print-the-order-of-hierarchical-clustering-in-r
@@ -1165,9 +1171,11 @@ GetMultilayerReducibility <- function(SupraAdjacencyMatrix,Layers,Nodes,Method,T
     for(i in 2:Layers){
       AggregateMatrix <- AggregateMatrix + NodesTensor[[i]]
     }
+    cat("got aggregate matrix\n")
     
     #aggregate entropy
     AggregateEntropy <- GetRenyiEntropyFromAdjacencyMatrix(AggregateMatrix,1)
+    cat("got aggregate entropy\n")
     
     print(paste("DEBUG: Aggregate entropy", AggregateEntropy))
[…]

### 2024-02-21 scripts/02.1_visualizeNetworks.qmd
updates to graphs

@@
               select(Nili_id, "season" = seasonUnique, daysTracked, propDaysTracked, birth_year, sex, space_use, movement, roost_div, space_use_log, age, age_group, type)) %>%
-  mutate(focal = case_when(is.na(movement) & is.na(space_use) & is.na(roost_div) ~ FALSE,
-                           TRUE ~ TRUE))
+  mutate(focal = case_when(is.na(movement) & 
+                             is.na(space_use) & 
+                             is.na(roost_div) ~ FALSE,
+                           TRUE ~ TRUE),
+         label = case_when(!focal ~ Nili_id,
+                           TRUE ~ NA_character_))
 nrow(info) == nrow(networkMetrics) # TRUE
@@
 ```{r}
+#| message: false
 # Network visualizations ----------------------------------------------------------------
@@
     geom_edge_link(alpha = 0.2)+
-    geom_node_point(aes(size = normDegree, col = normDegree))+
+    geom_node_point(aes(size = normDegree, col = normDegree, shape = dataset))+
+    geom_node_text(aes(label = label,
+                       hjust = -0.2, vjust = 0.1))+
+    scale_shape_manual(name = "Focal", values = c(17, 16))+
     scale_color_viridis()+
@@
     geom_edge_link(alpha = 0.2)+
-    geom_node_point(aes(size = normStrength, col = normStrength))+
+    geom_node_point(aes(size = normStrength, 
+                        col = normStrength, 
+                        shape = dataset))+
+    geom_node_text(aes(label = label,
+                       hjust = -0.2, vjust = 0.1))+
+    scale_shape_manual(name = "Focal", values = c(17, 16))+
     scale_color_viridis()+
@@
 
+grph_test <- function(graph, season_name, type){
+  g <- graph %>%
+    ggraph(layout = "fr")+
+    geom_edge_link(alpha = 0.2)+
+    geom_node_point(aes(size = normStrength, 
+                        col = normStrength, 
+                        shape = focal))+
+    geom_node_text(aes(label = label,
+                       hjust = -0.2, vjust = 0.1))+
+    scale_shape_manual(name = "Focal", values = c(17, 16))+
+    scale_color_viridis()+
+    theme(legend.position = "none")+
+    ggtitle(season_name)
+  return(g)
+}
+
+grph_test(flightGraphs[[2]], "2021_breeding", type = "flight")
+```
+
+```{r}
 flightPlots_degree <- map2(.x = flightGraphs, .y = season_names, ~grph_deg(.x, .y, type = "flight"))

### 2024-02-21 scripts/01.0_akde.R
renamed scripts so they're in the right order

@@ -27,6 +27,7 @@ seasons_split <- map(downsampled_10min, ~.x %>%
                        group_split())
 telems_list <- map(seasons_split, ~map(.x, as.telemetry))
 save(telems_list, file = "data/telems_list.Rda")
+load("data/telems_list.Rda")
 # check out a few examples
 dt.plot(telems_list[[1]][[3]])
 dt.plot(telems_list[[1]][[4]])
@@ -43,17 +44,19 @@ walk(variograms_list[[1]], plot) # just get a brief glimpse of the different ind
 # AAA in the ctmm google group, it seems like people check for range-residency and dispersal movements for every animal. I don't think that is practical here, and I could go down basically an infinite rabbit hole trying to do that, so since most of the animals are range resident I am going to assume they all are, generally, and hope that variation in home range sizes basically gets washed out.
 
 # Selecting the best-fit movement model through model selection
-guesses_list <- map(telems_list, ~map(.x, ~ctmm.guess(.x, interactive = FALSE), .progress = T))
+future::plan(future::multisession(), workers = 20)
+guesses_list <- furrr::future_map(telems_list, ~map(.x, ~ctmm.guess(.x, interactive = FALSE), .progress = T))
 save(guesses_list, file = "data/akde/guesses_list.Rda")
 load("data/akde/guesses_list.Rda")
 
 # Now we do the fits, which I suspect will take a long time. I'm going to do these in a for loop so they will save along the way.
-# for(i in 1:length(telems_list)){
-#   cat("processing season", i, "\n")
-#   fits <- map2(.x = telems_list[[i]], .y = guesses_list[[i]], ~ctmm.select(.x, .y, method = "pHREML"), .progress = TRUE)
-#   filename <- paste0("data/akde/pHREML_fits_", season_names[i], ".Rda")
-#   save(fits, file = filename)
-# }
+for(i in 2:length(telems_list)){ # XXX STARTING FROM 2
+  cat("processing season", i, "\n")
+  fits <- furrr::future_map2(.x = telems_list[[i]], .y = guesses_list[[i]], 
+                             ~ctmm.select(.x, .y, method = "pHREML"), .progress = TRUE)
+  filename <- paste0("data/akde/pHREML_fits_", season_names[i], ".Rda")
+  save(fits, file = filename)
+}
 # Load each of the fits objects in. Because I stupidly saved them all as "fits", need to rename them after loading them in.
[…]

### 2024-02-21 scripts/01.1_calcMovement.R
Re-ran with northern indivs removed

@@ -179,7 +179,7 @@ dfddf %>%
 ### Daily Distance Traveled/Mean Displacement -------------------------------
 
 # set up parallel backend with 4 sessions (change this to match your machine)
-future::plan(future::multicore, workers = 4)
+future::plan(future::multicore, workers = 10)
 
 # function to calculate displacements within a group
 calc_displacements <- function(group) {
@@ -217,7 +217,7 @@ calc_metrics <- function(data){
 }
 
 # apply the calc_metrics() function to each season in the list using purrr::map()
-dailyMovementList_seasons <- purrr::map(downsampled_10min_sf, calc_metrics, .progress = T)
+dailyMovementList_seasons <- map(downsampled_10min_sf, calc_metrics, .progress = T)
 save(dailyMovementList_seasons, file = "data/calcMovement/dailyMovementList_seasons.Rda")
 load("data/calcMovement/dailyMovementList_seasons.Rda") 
 

### 2024-02-21 scripts/03.0_PCA.R
removed old PCA, now obsolete

@@ -1,148 +1,0 @@
-# Analysis script for movement/social position
-
-# Setup -------------------------------------------------------------------
-## Load packages -----------------------------------------------------------
-library(vultureUtils)
-library(sf)
-library(tidyverse)
-library(igraph)
-library(mapview) # for quick maps
-library(factoextra)
-library(ade4)
-
-feedingColor = "gold3";
-roostingColor = "brown1";
-flightColor = "skyblue3";
-
-## Load data ------------------------------------------------------------
-load("data/calcMovement/movementBehavior.Rda")
-load("data/calcMovement/movementBehaviorScaled.Rda")
-### load("data/calcSocial/networkMetrics.Rda") #XXX
-load("data/dataPrep/season_names.Rda")
-
-# Create a dataset with all seasons combined
-allMovementBehavior <- purrr::list_rbind(movementBehavior)
-allMovementBehavior$sex[allMovementBehavior$sex == "NA"] <- NA
-colSums(is.na(allMovementBehavior))
-# Scale relative to all the seasons, not just one season (movementBehaviorScaled has vars scaled per season)
-allMovementBehavior_scaledAll <- allMovementBehavior %>%
-  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex, dataset), function(x){as.numeric(as.vector(scale(x)))}))
-allMovementBehavior_scaledAll <- allMovementBehavior_scaledAll[complete.cases(allMovementBehavior_scaledAll),] # remove a couple NA values
-colSums(is.na(allMovementBehavior_scaledAll))
-
-# Multivariate index of movement behavior (PCA) ------------------------------
-# Decided to create one single PCA for the full time range, and then apply those loadings to individual seasons as needed. 
-
-# Variables available:
-glimpse(movementBehavior)
-# HR
-# coreArea [keep]
-# homeRange [keep]
-# coreAreaFidelity [keep]
-
-# Altitude
-# mnDailyMaxAlt [keep]
-# mnDailyMnAlt [remove -- redundant with mnDailyMedAlt]
-# mnDailyPropOver1km [remove--redundant with mnDailyMaxAlt]
-# mnDailyMedAlt [keep, but need justification for why we're keeping means for other traits but median here]
-
-# Roosts
[…]

### 2024-02-22 scripts/02.1_visualizeNetworks.qmd
updated model effect plots and var distribution plots

--- 2024-02-08 scripts/02.1_visualizeNetworks.qmd
@@
   left_join(linked %>%
-              select(Nili_id, "season" = seasonUnique, daysTracked, propDaysTracked, birth_year, sex, space_use, movement, roost_div, space_use_log, age, age_group, type)) %>%
-  mutate(focal = case_when(is.na(movement) & is.na(space_use) & is.na(roost_div) ~ FALSE,
-                           TRUE ~ TRUE))
+              select(Nili_id, "season" = seasonUnique, daysTracked, propDaysTracked, birth_year, sex, space_use, movement, roost_div, age, age_group, type)) %>%
+  mutate(focal = case_when(is.na(movement) & 
+                             is.na(space_use) & 
+                             is.na(roost_div) ~ FALSE,
+                           TRUE ~ TRUE),
+         label = case_when(!focal ~ Nili_id,
+                           TRUE ~ NA_character_)) %>%
+  mutate(season = factor(season, levels = season_names))
 nrow(info) == nrow(networkMetrics) # TRUE
@@
 ```{r}
+#| message: false
 # Network visualizations ----------------------------------------------------------------
+# This process only works for the color scaling because we know our minimum degree value is 0 and our maximum degree value is 1. For strength, it's going to be more complicated.
 grph_deg <- function(graph, season_name, type){
+  min <- graph %>% activate(nodes) %>% pull(normDegree) %>% min()
+  max <- graph %>% activate(nodes) %>% pull(normDegree) %>% max()
    g <- graph %>%
@@
     geom_edge_link(alpha = 0.2)+
-    geom_node_point(aes(size = normDegree, col = normDegree))+
+    geom_node_point(aes(size = normDegree, col = normDegree, shape = dataset))+
+    # geom_node_text(aes(label = label,
+    #                    hjust = -0.2, vjust = 0.1))+
+    scale_shape_manual(name = "Dataset", values = c(17, 16, 18))+
+    scale_color_viridis(begin = min, end = max)+
+    theme(legend.position = "none")+
+    ggtitle(season_name)
+  return(g)
+}
+
+minstr <- min(info$normStrength)
+maxstr <- max(info$normStrength) # XXX CHECK THIS
+grph_str <- function(graph, season_name, type){
+  min <- graph %>% activate(nodes) %>% pull(normStrength) %>% min()
+  max <- graph %>% activate(nodes) %>% pull(normStrength) %>% max()
+  g <- graph %>%
+    ggraph(layout = "fr")+
+    geom_edge_link(alpha = 0.2)+
+    geom_node_point(aes(size = normStrength, 
+                        col = normStrength*(1/maxstr), 
+                        shape = dataset))+
+    # geom_node_text(aes(label = label,
+    #                    hjust = -0.2, vjust = 0.1))+
+    scale_shape_manual(name = "Dataset", values = c(17, 16, 18))+
     scale_color_viridis()+
@@
 
-grph_str <- function(graph, season_name, type){
-  g <- graph %>%
-    ggraph(layout = "fr")+
-    geom_edge_link(alpha = 0.2)+
-    geom_node_point(aes(size = normStrength, col = normStrength))+
-    scale_color_viridis()+
-    theme(legend.position = "none")+
-    ggtitle(season_name)
-  return(g)
-}
+grph_str(flightGraphs[[1]], "2020_fall", type = "flight")
+grph_deg(flightGraphs[[2]], "2021_breeding", type = "flight")
+grph_deg(flightGraphs[[3]], "2021_summer", type = "flight")
+grph_deg(roostingGraphs[[6]], "2022_summer", type = "roosting")
+```
 
+```{r}
 flightPlots_degree <- map2(.x = flightGraphs, .y = season_names, ~grph_deg(.x, .y, type = "flight"))
@@
 ro_d <- patchwork::wrap_plots(roostingPlots_degree, nrow = 3, ncol = 3)
-ggsave(fl_d, filename = here("fig", "networkGraphs", "degree", "flight.png"), width = 9, height = 7)
-ggsave(fe_d, filename = here("fig", "networkGraphs", "degree", "feeding.png"), width = 9, height = 7)
-ggsave(ro_d, filename = here("fig", "networkGraphs", "degree", "roosting.png"), width = 9, height = 7)
+ggsave(fl_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "flight.png"), width = 9, height = 7)
+ggsave(fe_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "feeding.png"), width = 9, height = 7)
+ggsave(ro_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "roosting.png"), width = 9, height = 7)
 
@@
 ro_s <- patchwork::wrap_plots(roostingPlots_strength, nrow = 3, ncol = 3)
-ggsave(fl_s, filename = here("fig", "networkGraphs", "strength", "flight.png"), width = 9, height = 7)
-ggsave(fe_s, filename = here("fig", "networkGraphs", "strength", "feeding.png"), width = 9, height = 7)
-ggsave(ro_s, filename = here("fig", "networkGraphs", "strength", "roosting.png"), width = 9, height = 7)
+ggsave(fl_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "flight.png"), width = 9, height = 7)
+ggsave(fe_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "feeding.png"), width = 9, height = 7)
+ggsave(ro_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "roosting.png"), width = 9, height = 7)
 ```

### 2024-02-22 scripts/04.0_mixedModels.R

@@ -68,6 +68,7 @@ normDegree_years_seasons_flight <- linked %>%
   xlab("Degree (normalized)")+
   facet_wrap(~season)+
   theme(legend.position = "none", text = element_text(size = 20))
+normDegree_years_seasons_flight
 ggsave(normDegree_years_seasons_flight, filename = "fig/normDegree_years_seasons_flight.png", width = 9, height = 7)
 
 normDegree_years_seasons_feeding <- linked %>%
@@ -133,7 +134,7 @@ degree_3 <- glmmTMB(normDegree ~ situ*movement + situ*space_use + season*movemen
 summary(degree_3) # after removing season*space_use, we still don't have a significant interaction between movement and season
 
 degree_4 <- glmmTMB(normDegree ~ situ*movement + situ*space_use + season + roost_div  + age_group + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian())
-summary(degree_4) # made sure to add season bask as a main effect. 
+summary(degree_4) # made sure to add season back as a main effect. 
 check_collinearity(degree_4) # as a bonus, now none of our VIFs are above 3. Yay!
 
 degree_mod <- degree_4
@@ -195,7 +196,6 @@ plot_d_eff_situ_movement <- ggplot(d_eff_situ_movement, aes(x, predicted))+
   theme(text = element_text(size = 16))
 plot_d_eff_situ_movement
 ggsave(plot_d_eff_situ_movement, file = "fig/plot_d_eff_situ_movement.png", width = 9, height = 6)
-
 
 ## situ:movement (forest plot) ---------------------------------------------
 d_emt_situ_movement <- emmeans::emtrends(degree_mod, "situ", var = "movement") %>%

### 2024-02-22 scripts/01.2_categorizeMovement.qmd
Showed new space use PC variable against individual vars

--- 2024-02-14 scripts/01.1_categorizeMovement.qmd
@@
 
+```{r}
+#| warning: false
+#| message: false
+
+space_use %>%
+  rename("Home range (log-transformed)" = "homeRange_log", "CA/HR ratio" = "coreAreaFidelity") %>%
+  pivot_longer(cols = c("Home range (log-transformed)", "CA/HR ratio"),
+               names_to = "measure", values_to = "value") %>%
+  ggplot(aes(x = space_pc1, y = value))+
+  geom_point()+geom_smooth(method = "lm")+theme_minimal()+
+  facet_wrap(~measure)+
+  ylab("Value (scaled)")+
+  theme(legend.position = "none",
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
+```
+
 ## Variable distributions
@@
   geom_point(color = "chartreuse3")+ geom_smooth(method = "lm", col = "chartreuse3")+ theme_minimal()+
-  xlab("Roost diversification")+ ylab("Space use (log)")+
+  xlab("Roost diversification")+ ylab("Space use")+
   ggplot2::annotate(geom = "text", x = 3, y = -2, label = paste0("Correlation = ", cor))
@@
 
-Finally, **log-transformed home range** and **movement** have a correlation of `r round(cor(red$space_use_log, red$movement), 2)`.
+Finally, **log-transformed home range** and **movement** have a correlation of `r round(cor(red$space_use, red$movement), 2)`.
 
@@
   geom_point(color = "orange2")+ geom_smooth(method = "lm", col = "orange2")+ theme_minimal()+
-  xlab("Space use (log)") + ylab("Movement")+
+  xlab("Space use") + ylab("Movement")+
   ggplot2::annotate(geom = "text", x = 2.5, y = -3, label = paste0("Correlation = ", cor))

### 2024-02-22 scripts/repeatability.R
re-did time plots, and made graph of timescales

@@ -266,17 +266,18 @@ save(graphs_flight, file = "data/graphs_flight.Rda")
 save(graphs_feeding, file = "data/graphs_feeding.Rda")
 save(graphs_roosting, file = "data/graphs_roosting.Rda")
 
-# Get metrics for each graph
-metrics_flight <- vector(mode = "list", length = length(ndays))
-metrics_feeding <- vector(mode = "list", length = length(ndays))
-metrics_roosting <- vector(mode = "list", length = length(ndays))
+
+# Individual-level metrics ------------------------------------------------
+metrics_flight_indiv <- vector(mode = "list", length = length(ndays))
+metrics_feeding_indiv <- vector(mode = "list", length = length(ndays))
+metrics_roosting_indiv <- vector(mode = "list", length = length(ndays))
 
 for(i in 1:length(ndays)){
   fl <- graphs_flight[[i]]
   fe <- graphs_feeding[[i]]
   ro <- graphs_roosting[[i]]
   
-  metrics_flight[[i]] <- map2(fl, brks[[i]], ~{
+  metrics_flight_indiv[[i]] <- map2(fl, brks[[i]], ~{
     if(length(.x) > 0){
       out <- data.frame(degree = igraph::degree(.x),
                         strength = igraph::strength(.x),
@@ -291,7 +292,7 @@ for(i in 1:length(ndays)){
     }
     return(out)
   })
-  metrics_feeding[[i]] <- map2(fe, brks[[i]], ~{
+  metrics_feeding_indiv[[i]] <- map2(fe, brks[[i]], ~{
     if(length(.x) > 0){
       out <- data.frame(degree = igraph::degree(.x),
                         strength = igraph::strength(.x),
@@ -306,7 +307,7 @@ for(i in 1:length(ndays)){
     }
     return(out)
   })
-  metrics_roosting[[i]] <- map2(ro, brks_roosts[[i]], ~{
+  metrics_roosting_indiv[[i]] <- map2(ro, brks_roosts[[i]], ~{
     if(length(.x) > 0){
       out <- data.frame(degree = igraph::degree(.x),
                         strength = igraph::strength(.x),
@@ -325,14 +326,14 @@ for(i in 1:length(ndays)){
   rm(fe)
   rm(ro)
 }
-metrics_flight_df <- map(metrics_flight, ~purrr::list_rbind(.x))
[…]

### 2024-02-22 scripts/02.1_visualizeNetworks.qmd
saved seasonal networks with transparent backgrounds

@@
 load(here("data", "mixedModels", "linked.Rda"))
+load(here("data/dataPrep/season_names.Rda"))
 
@@
 ro_d <- patchwork::wrap_plots(roostingPlots_degree, nrow = 3, ncol = 3)
-ggsave(fl_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "flight.png"), width = 9, height = 7)
-ggsave(fe_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "feeding.png"), width = 9, height = 7)
-ggsave(ro_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "roosting.png"), width = 9, height = 7)
+ggsave(fl_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "flight.png"), width = 9, height = 7, bg = "transparent")
+ggsave(fe_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "feeding.png"), width = 9, height = 7, bg = "transparent")
+ggsave(ro_d, filename = here("fig", "networkGraphs", "seasonal", "degree", "roosting.png"), width = 9, height = 7, bg = "transparent")
 
@@
 ro_s <- patchwork::wrap_plots(roostingPlots_strength, nrow = 3, ncol = 3)
-ggsave(fl_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "flight.png"), width = 9, height = 7)
-ggsave(fe_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "feeding.png"), width = 9, height = 7)
-ggsave(ro_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "roosting.png"), width = 9, height = 7)
+ggsave(fl_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "flight.png"), width = 9, height = 7, bg = "transparent")
+ggsave(fe_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "feeding.png"), width = 9, height = 7, bg = "transparent")
+ggsave(ro_s, filename = here("fig", "networkGraphs", "seasonal", "strength", "roosting.png"), width = 9, height = 7, bg = "transparent")
 ```

### 2024-02-22 scripts/04.0_mixedModels.R
moved model effect plots to their own folder

@@ -157,7 +157,7 @@ plot_d_eff_situ_space <- ggplot(d_eff_situ_space, aes(x, predicted))+
   ggtitle("")+theme_classic()+
   theme(text = element_text(size = 16))
 plot_d_eff_situ_space
-ggsave(plot_d_eff_situ_space, file = "fig/plot_d_eff_situ_space.png", width = 9, height = 6)
+ggsave(plot_d_eff_situ_space, file = "fig/mmPlots/plot_d_eff_situ_space.png", width = 9, height = 6)
 
 ## situ:space_use (forest plot) --------------------------------------------
 d_emt_situ_space <- emmeans::emtrends(degree_mod, "situ", var = "space_use") %>%
@@ -178,7 +178,7 @@ plot_d_emt_situ_space <- d_emt_situ_space %>%
   ylab("Space use effect")+
   xlab("Situation")+
   coord_flip()
-ggsave(plot_d_emt_situ_space, file = "fig/plot_d_emt_situ_space.png", width = 10, height = 6)
+ggsave(plot_d_emt_situ_space, file = "fig/mmPlots/plot_d_emt_situ_space.png", width = 10, height = 6)
 
 
 ## situ:movement (effect plot) ---------------------------------------------
@@ -195,7 +195,7 @@ plot_d_eff_situ_movement <- ggplot(d_eff_situ_movement, aes(x, predicted))+
   ggtitle("")+theme_classic()+
   theme(text = element_text(size = 16))
 plot_d_eff_situ_movement
-ggsave(plot_d_eff_situ_movement, file = "fig/plot_d_eff_situ_movement.png", width = 9, height = 6)
+ggsave(plot_d_eff_situ_movement, file = "fig/mmPlots/plot_d_eff_situ_movement.png", width = 9, height = 6)
 
 ## situ:movement (forest plot) ---------------------------------------------
 d_emt_situ_movement <- emmeans::emtrends(degree_mod, "situ", var = "movement") %>%
@@ -216,7 +216,7 @@ plot_d_emt_situ_movement <- d_emt_situ_movement %>%
   ylab("Movement effect")+
   xlab("Situation")+
   coord_flip()
-ggsave(plot_d_emt_situ_movement, file = "fig/plot_d_emt_situ_movement.png", width = 10, height = 6)
+ggsave(plot_d_emt_situ_movement, file = "fig/mmPlots/plot_d_emt_situ_movement.png", width = 10, height = 6)
 
[…]

### 2024-02-28 scripts/04.0_mixedModels.R
commented out some portions to make this easier to source in the manuscript

@@ -57,45 +57,45 @@ save(linked, file = "data/mixedModels/linked.Rda")
 
 
 # Examine response variable distributions ---------------------------------
-normDegree_years_seasons_flight <- linked %>%
-  filter(type == "flight") %>%
-  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
-  ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-  geom_density(linewidth = 1.5)+
-  theme_classic()+
-  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
-  ylab("Frequency")+
-  xlab("Degree (normalized)")+
-  facet_wrap(~season)+
-  theme(legend.position = "none", text = element_text(size = 20))
-normDegree_years_seasons_flight
-ggsave(normDegree_years_seasons_flight, filename = "fig/normDegree_years_seasons_flight.png", width = 9, height = 7)
-
-normDegree_years_seasons_feeding <- linked %>%
-  filter(type == "feeding") %>%
-  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
-  ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-  geom_density(linewidth = 1.5)+
-  theme_classic()+
-  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
-  ylab("Frequency")+
-  xlab("Degree (normalized)")+
-  facet_wrap(~season)+
-  theme(legend.position = "none", text = element_text(size = 20))
-ggsave(normDegree_years_seasons_feeding, filename = "fig/normDegree_years_seasons_feeding.png", width = 9, height = 7)
-
-normDegree_years_seasons_roosting <- linked %>%
-  filter(type == "roosting") %>%
-  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
-  ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-  geom_density(linewidth = 1.5)+
-  theme_classic()+
-  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
-  ylab("Frequency")+
-  xlab("Degree (normalized)")+
-  facet_wrap(~season)+
-  theme(legend.position = "none", text = element_text(size = 20))
[…]

### 2024-02-28 temporal.qmd
more work with reducibility and temporal analysis in advance of meeting with Sandra

---
title: "Examining vulture social networks across time scales"
author: "Kaija Gahm"
date: 2024-02-28
format: 
  html:
    self-contained: true
execute:
  message: false
  warning: false
editor: visual
---

```{r}
#| include: false
library(tidyverse)
library(here)
theme_set(theme_classic())
```

```{r}
#| include: false
# Graphs of metrics -------------------------------------------------------
load(here::here("data/metrics_indiv.Rda"))
load(here::here("data/derived/cc.Rda"))
glimpse(metrics_indiv)
metrics_indiv <- metrics_indiv %>%
  group_by(ndays, int) %>%
  mutate(normDegree = degree/(n-1),
         normStrength = strength/sum(strength, na.rm = T)) # XXX WHAT IS WRONG HERE?
```

```{r}
#| echo: false
metrics_indiv %>%
  ggplot(aes(x = ndays, y = normDegree, col = type))+
  ylab("Degree (normalized)")+
  xlab("Time window (days)")+
  geom_jitter(alpha = 0.1, width = 0.6)+
  theme(axis.title = element_text(size = 14))+
  facet_wrap(~type) # There are a few individuals that don't roost with the others and therefore have a low degree, even at the high time windows. I thought those were Carmel birds, but I've now removed the northern indivs and we still see this pattern.

metrics_indiv %>%
  ggplot(aes(x = ndays, y = normStrength, col = type))+
  geom_jitter(alpha = 0.1, width = 0.6)+
  xlab("Time window (days)")+
  ylab("Strength (normalized)")+
  theme(axis.title = element_text(size = 14))+
  facet_wrap(~type) # I do not understand why this goes DOWN. Am I normalizing it incorrectly?

# How do network-level measures change over the time points?
metrics_net <- metrics_indiv %>%
  group_by(ndays, int, type, n) %>%
  summarize(mndeg = mean(degree),
            mnstr = mean(strength),
            mnnormdeg = mean(normDegree),
            mnnormstr = mean(normStrength)) %>%
  mutate(int = lubridate::ymd(int))
```

```{r}
#| echo: false
metrics_net %>%
  ggplot(aes(x = int, y = mnnormdeg, col = factor(ndays)))+
  ylab("Mean degree (normalized)")+
  xlab("Date")+
  geom_point()+
  geom_line()+
  facet_wrap(~type, ncol = 1, scales = "free")+
  theme(axis.title = element_text(size = 14))+
  labs(color = "Time window\n(days)")+
  geom_hline(aes(yintercept = 0.5), col = "black", linetype = 2)
```

```{r}
#| echo: false
#| fig-height: 10
metrics_net %>%
  ggplot(aes(x = mnnormdeg, col = type))+
  xlab("Mean degree (normalized)")+
  geom_density(linewidth = 1)+
  facet_wrap(~interaction(type, ndays), scales = "free", ncol = 3, nrow = 7)+
  scale_color_manual(name = "Situation",
                     values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14))+
  geom_vline(aes(xintercept = 0.5), linetype = 2)
```

```{r}
#| echo: false
metrics_net %>%
  group_by(ndays, type) %>%
  summarize(mn_mnnormdeg = mean(mnnormdeg)) %>%
  ggplot(aes(x = ndays, y = mn_mnnormdeg, col = type))+
  geom_point(size = 5)+
  geom_line(linewidth = 2)+
  ylab("Mean degree (normalized)")+
  xlab("Time window (days)")+
  scale_color_manual(name = "Situation",
                     values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  geom_hline(aes(yintercept = 0.5), linetype = 2)+
  theme(text = element_text(size = 14))
```

### 2024-02-28 scripts/reduc.R

@@ -1,0 +1,158 @@
+# Load libraries
+library(tidyverse)
+library(lubridate)
+library(tnet)
+library(sna)
+
+# Data ingest (already sliced into appropriate chunks)
+load("data/calcSocial/feedingEdges.Rda")
+feedingEdges <- purrr::list_rbind(feedingEdges) %>% mutate(date = lubridate::date(minTimestamp))
+load("data/calcSocial/flightEdges.Rda")
+flightEdges <- purrr::list_rbind(flightEdges) %>% mutate(date = lubridate::date(minTimestamp))
+load("data/calcSocial/roostingEdges.Rda")
+roostingEdges <- purrr::list_rbind(roostingEdges)
+source("scripts/muxLib DF.R")
+
+cutfun <- function(vec, days){
+  min <- min(vec)
+  max <- max(vec)
+  ngroups <- ceiling(as.numeric(max-min)/days)
+  brks <- seq(from = min, by = days, length.out = ngroups + 1)
+  cut <- cut(vec, breaks = brks, include.lowest = T, right = F)
+}
+
+# Feed into code to do the rest
+doReduce <- function(data, groupingCol){
+  # Make a binary edgelist, grouping by the group variable
+  bin <- data[,c("ID1", "ID2", groupingCol)] %>%
+    distinct() %>%
+    mutate(weight = 1)
+  
+  opp_edges <- bin
+  names(opp_edges)[1:2] <- c("ID2", "ID1")
+  
+  all_bin <- bind_rows(bin, opp_edges) # attach 
+  
+  # Add layer numbers (days) to make the next step easier
+  all_bin <- all_bin %>%
+    arrange(.data[[groupingCol]], ID1, ID2) %>%
+    mutate(layer = as.numeric(.data[[groupingCol]])) %>%
+    select(-all_of(groupingCol)) %>%
+    rename("layer1" = "layer") %>%
+    mutate(layer2 = layer1) %>%
+    relocate(layer1, .after = "ID1") %>%
+    relocate(layer2, .after = "ID2")
+  
+  self_edges <- expand.grid(as.character(unique(all_bin$ID1)), 1:(max(all_bin$layer1)-1)) # create all possible combinations of indiv ID and layer
+  colnames(self_edges) <- c("ID1", "layer1")
+  
+  self_edges <- self_edges %>%
+    mutate(ID2 = ID1, layer2 = layer1 + 1, weight = 1)
[…]

### 2024-02-28 manuscript.qmd
added some figures after sourcing modeling code

@@
 library(tidyverse)
+library(here)
 ```
@@
 But flight is a continuous process driven by many decisions resulting from individual and social information [@sassi2023], and feeding events are concentrated in time on a given day.
-Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every [NUMBER] days and persist for hours to a few days (Orr Spiegel, pers.
+Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every \[NUMBER\] days and persist for hours to a few days (Orr Spiegel, pers.
 comm.).
@@
 
-To answer these questions, we drew on data from a free-flying population of Eurasian griffon vultures (Gyps fulvus) fitted with GPS tags.
-\[tag specs\].
-Over the period 2020-09-01 to 2023-05-15, the dataset contained \[number\] of tagged vultures.
+To answer these questions, we analyzed data from a free-flying population of Eurasian griffon vultures (Gyps fulvus) fitted with GPS tags.
+
+\[tag specs\]
+
+Over the period \[date\] to \[date\], the dataset contained \[number\] of tagged vultures.
 Tags recorded GPS fixes, as well as speed and heading readings, every 10 minutes when battery conditions allowed.
@@
 We subsetted the dataset to exclude periods of time that vultures spent in capture cages, in the wildlife hospital following poisonings or injuries, or in captivity prior to wild release (in the case of captive-born birds).
-We cleaned the remaining data to remove invalid GPS fixes \[more details\].
+We cleaned the data to remove invalid GPS fixes.
 Then we split data from each into three seasons.
@@
 For each season, we calculated movement measures only for individuals tracked for at least 30 days (after removing low-point days as described above).
-We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of [level–more details about resolution?] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of \[level–more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
 Finally, since some tags were occasionally set to a much higher fix rate (e.g.
@@
 2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
-Using the `adehabitatHR` package [@calenge2023], we calculated home range as the 95% kernel utilization distribution (KUD) of a vulture's GPS fixes in a season; core area as the 50% KUD, and core area use intensity as the ratio of the 50% to the 95% KUD.
+We calculated home range as the 95% AKDE of a vulture's GPS fixes in a season @
 
@@
 
-To gain a complete picture of individual vultures' spatial phenotype, we summarized the above-described movement measures (centered and scaled) in a principal component analysis.
-The first two principal components together explained 65% of the variance in spatial phenotype (Table X).
+To summarize individuals' spatial phenotype, we created three composite indices: **movement**, **space use**, and **roost diversification**.
+Each index was derived from a principal component analysis of several spatial phenotype variables, all of which were scaled and centered over the entire date range of the study before putting them into the PCA.
 
-PC1 explained 47.6% of the variance in spatial phenotype.
-All movement variables except tortuosity loaded positively on PC1, with the strongest contributors being mean daily maximum displacement, roost diversity, and proportion of nights roost-switching (Figure X).
-Overall, we can understand PC1 as an aggregate measure of individual movement.
-Vultures with high PC1 values cover more area, use more different roosts, travel farther, spend more time flying, and have straighter trajectories.
-
-PC2 explained 17.4% of the variance in spatial phenotype.
-The space use variables loaded positively on this principal component, while flight variables (both altitude measures, daily distance traveled, and daily flight time) contributed strongly negatively (Figure X).
-We can understand PC2 as a rough measure of exploratory behavior.
-Vultures with high PC2 values use a larger spatial area, while those with low PC2 values fly higher and spend more time flying (in a more concentrated area).
+**Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity.
+**Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE.
+**Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night).
+The composite measures accounted for %, %, and % of the variation in their respective PCAs.
 
@@
 ```{r}
-#| echo: false
-#| output: true
-#| label: fig-nvultures
-#| fig-cap: "Number of vultures in different social situations over the study period.
-Lines for the 'feeding' situation have been jittered upward because they overlap completely with the 'flight' situation.
-Dashed lines and open circles represent the total number of vultures tracked in the population.
-Solid lines and filled circles represent the number of focal vultures."
-
-load("data/derived/forModeling.Rda")
-load("data/derived/cc.Rda")
-load("data/derived/seasonNames_orig.Rda")
-seasonNames <- seasonNames_orig[-1]
-focals <- forModeling %>%
-  filter(!is.na(PC1)) %>%
-  group_by(season, year, type) %>%
-  summarize(total = n[1],
-            focal = length(unique(Nili_id)), .groups = "drop") %>%
-  arrange(year, season, type) %>%
-  mutate(seasonUnique = paste(year, season, sep = "_")) %>%
-  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding"))) %>%
-  mutate(seasonU = str_replace(seasonUnique, "breeding", "B")) %>%
-  mutate(seasonU = str_replace(seasonU, "fall", "F")) %>%
-  mutate(seasonU = str_replace(seasonU, "summer", "S")) %>%
-  mutate(seasonU = str_remove(seasonU, "_")) %>%
-  mutate(seasonU = str_remove(seasonU, "(?<=^)20")) %>%
-  mutate(seasonU = factor(seasonU, levels = c("20F", "21B", "21S", "21F", "22B", "22S", "22F", "23B")))  %>%
-  pivot_longer(cols = c(total, focal), names_to = "Vultures", values_to = "n") %>%
-  rename("Situation" = "type") %>%
-  ungroup()
-
-focals %>%
-  mutate(n = case_when(Situation == "feeding" ~ n + 1, # just adding a tiny bit of jitter so as to more easily see the different lines 
-                       TRUE ~ n)) %>%
-  ggplot(aes(x = seasonU, y = n, col = Situation))+
-  geom_line(aes(group = interaction(Situation, Vultures), linetype = Vultures), linewidth = 1)+
-  geom_point(size = 4, aes(pch = Vultures))+
-  scale_shape_manual(values = c(19, 1))+
-  theme_classic()+
-  ylab("# vultures")+
-  xlab("Season")+
-  scale_color_manual(values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
-  theme(text = element_text(size = 16))
-
-wide <- focals %>%
-  pivot_wider(id_cols = c("season", "year", "Situation", "seasonUnique", "seasonU"), names_from = "Vultures", values_from = "n")
+#| include: false
+source(here("scripts/04.0_mixedModels.R"))
 ```
-
-Networks varied in the total and focal number of individuals, from a minimum of `r min(wide$total)` vultures (`r wide$focal[wide$total == min(wide$total)][1]` focal) in the `r wide$year[wide$total == min(wide$total)][1]` `r wide$season[wide$total == min(wide$total)][1]` season to a maximum of `r max(wide$total)` vultures (`r wide$focal[wide$total == max(wide$total)]` focal) in the `r wide$year[wide$total == max(wide$total)]` `r wide$season[wide$total == max(wide$total)]` season in the co-roosting network (@fig-nvultures) .
-The number of vultures tagged generally increased over time as more individuals were fitted with tags over the course of the study period.
 
@@
 #| echo: false
-#| output: true
-#| label: fig-density
-#| fig-cap: "Density of the proximity-based social networks created for each season and social situation."
-
-load("data/derived/flightSeasons_mode10.Rda")
-load("data/derived/feedingSeasons_mode10.Rda")
-load("data/derived/roostSeasons.Rda")
-
-flightSeasons_mode10_g <- map(flightSeasons_mode10, ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
-feedingSeasons_mode10_g <- map(feedingSeasons_mode10,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
-roostSeasons_g <- map(roostSeasons,  ~vultureUtils::makeGraph(mode = "sri", data = .x, weighted = T))
-
-# Describing the networks -------------------------------------------------
-# Network-level measures
-fldens <- map(flightSeasons_mode10_g, igraph::edge_density) %>% unlist()
-fedens <- map(feedingSeasons_mode10_g, igraph::edge_density) %>% unlist()
-rodens <- map(roostSeasons_g, igraph::edge_density) %>% unlist()
-dens <- data.frame(season = factor(seasonNames, levels = seasonNames), flight = fldens, feeding = fedens, roosting = rodens) %>%
-  pivot_longer(cols = c("flight", "feeding", "roosting"), names_to = "Situation", values_to = "Density") %>%
-  mutate(seasonU = str_replace(season, "breeding", "B")) %>%
-  mutate(seasonU = str_replace(seasonU, "fall", "F")) %>%
-  mutate(seasonU = str_replace(seasonU, "summer", "S")) %>%
-  mutate(seasonU = str_remove(seasonU, "_")) %>%
-  mutate(seasonU = str_remove(seasonU, "(?<=^)20")) %>%
-  mutate(seasonU = factor(seasonU, levels = c("20F", "21B", "21S", "21F", "22B", "22S", "22F", "23B")))
-
-dens %>%
-  ggplot(aes(x = seasonU, y = Density, col = Situation))+
-  geom_line(aes(group = Situation), linewidth = 1)+
-  geom_point(size = 4)+
-  theme_classic()+
-  ylab("Network density")+
-  xlab("Season")+
-  scale_color_manual(values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
-  theme(text = element_text(size = 16))
+#| label: fig-d-space
+#| fig-cap: "Relationship between normalized degree and space use"
+plot_d_eff_space
 ```
 
-Seasonal social networks varied in density from `r min(dens$Density)` (`r stringr::str_extract(dens$season[dens$Density == min(dens$Density)], "[a-z]+")` `r str_extract(dens$season[dens$Density == min(dens$Density)], "[0-9]+")` ) to `r max(dens$Density)` (`r stringr::str_extract(dens$season[dens$Density == max(dens$Density)], "[a-z]+")` `r str_extract(dens$season[dens$Density == max(dens$Density)], "[0-9]+")`) (@fig-density).
-Co-roosting networks were generally denser than co-feeding or co-flight networks.
+```{r}
+#| echo: false
+#| label: fig-d-roost
+#| fig-cap: "Relationship between normalized degree and roost diversification"
+plot_d_eff_roost
+```
 
-Two-way interactions between situation and both movement (PC1) and exploration (PC2) were retained in the degree model.
-No interaction effects including season were retained, and the main effect of season was not significant (Table XX).
+```{r}
+#| echo: false
+#| label: fig-d-space-situ
+#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
+plot_d_eff_situ_space
+plot_d_emt_situ_space
+```
+
+```{r}
+#| echo: false
+#| label: fig-s-movement
+#| fig-cap: "Relationship between normalized strength and movement"
+plot_s_eff_movement
+```
+
+```{r}
+#| echo: false
+#| label: fig-d-movement-situ
+#| fig-cap: "Relationship between normalized degree and movement, with separate trend lines for each social situation"
+plot_d_eff_situ_movement
+plot_d_emt_situ_movement
+```
+
+```{r}
+#| echo: false
+#| label: fig-s-roost
+#| fig-cap: "Relationship between normalized strength and roost diversification"
+plot_s_eff_roost
+```
+
+```{r}
+#| echo: false
+#| label: fig-s-space
+#| fig-cap: "Relationship between normalized strength and space use"
+plot_s_eff_situ_space
+plot_s_emt_situ_space
+```
 
 # Discussion
-
-\[text here\]

### 2024-02-28 scripts/muxLib DF.R

@@ -1151,7 +1151,7 @@ GetMultilayerReducibility <- function(SupraAdjacencyMatrix,Layers,Nodes,Method,T
                                                  SingleLayerEntropy[i],SingleLayerEntropy[j])
           JSD[j,i] <- JSD[i,j]
         }
-        cat("Finished with layer", i, "out of", layers, "\n")
+        cat("Finished with layer", i, "out of", length(1:Layers-1), "\n")
       }
       cat("got jsd for categorical\n")
     }

### 2024-02-28 scripts/temp_multilayer_test.R

@@ -4,16 +4,18 @@ library(lubridate)
 library(tnet)
 library(sna)
 
-load("data/calcSocial/flightEdges.Rda")
-testseason <- flightEdges[[5]] # just doing this for one arbitrary season
+load("data/calcSocial/feedingEdges.Rda")
+testseason <- feedingEdges[[9]] # just doing this for one arbitrary season
+testseason <- testseason %>%
+  mutate(date = lubridate::date(minTimestamp))
+dates <- seq(from = min(testseason$date), to = min(testseason$date) + 4, by = 1)
+slice <- testseason %>% dplyr::filter(date %in% dates)
 
 # Make a binary edgelist, grouping by day
-bin <- testseason %>%
-  mutate(day = lubridate::date(minTimestamp)) %>%
-  #mutate(hour = floor_date(minTimestamp, unit = "hours")) %>%
-  select(ID1, ID2, day#, 
-         #hour
-         ) %>%
+bin <- slice %>%
+  mutate(day = lubridate::date(minTimestamp)) %>% # for feeding/flight
+  #mutate(day = date) %>% # for roosing
+  select(ID1, ID2, day) %>%
   distinct() %>%
   mutate(weight = 1)
 
@@ -103,123 +105,15 @@ gq <- inters_reduce$gQualityFunction
 
 df <- data.frame("Amount of aggregation" = 1:length(gq),
                  "Difference in relative entropy" = gq)
-df %>%
+feedingentropy <- df %>%
   ggplot(aes(x = Amount.of.aggregation, y = Difference.in.relative.entropy))+
   geom_point()+
   geom_line()+
-  theme_classic()
+  theme_classic()+
+  ylab("Difference in relative entropy")+
+  xlab("Amount of aggregation")+
+  theme(axis.title = element_text(size = 14))
+ggsave(feedingentropy, file = "fig/entropy/feedingentropy.png", width = 7, height = 5)
+
 
 # It sort of seems like this will always have this pattern, at least for days. Or maybe that just indicates that the vultures change on a smaller time scale than days.
-
-
-# Hourly ------------------------------------------------------------------
-# Let me try with an hourly timescale, but only for one week, so we don't go insane with the number of layers.
-min <- min(lubridate::date(testseason$minTimestamp))
-minplusweek <- min + 7
-max <- max(lubridate::date(testseason$maxTimestamp))
-
-testweek <- testseason %>%
[…]

### 2024-02-28 scripts/02.0_calcSocial.R
saved edges and SRI for roosting for easier use in temporal analysis

@@ -72,18 +72,23 @@ rm(feedingEdges)
 rm(feedingSRI)
 gc()
 
-roostingSRI <- map(roosts, ~{
+roosting <- map(roosts, ~{
   vultureUtils::getRoostEdges(.x, mode = "polygon", 
                               roostPolygons = roostPolygons, 
-                              return = "sri", 
+                              return = "both", 
                               latCol = "location_lat", 
                               longCol = "location_long", 
                               idCol = "Nili_id", 
                               dateCol = "roost_date")
   }, .progress = T)
 
+roostingEdges <- map(roosting, "edges")
+roostingSRI <- map(roosting, "sri")
+save(roosting, file = "data/calcSocial/roosting.Rda")
+save(roostingEdges, file = "data/calcSocial/roostingEdges.Rda")
 save(roostingSRI, file = "data/calcSocial/roostingSRI.Rda")
 rm(roostingSRI)
+rm(roostingEdges)
 
 load("data/calcSocial/flightSRI.Rda")
 load("data/calcSocial/feedingSRI.Rda")

### 2024-02-28 scripts/01.0_akde.R
plotting some home ranges for SoCAB presentation

@@ -140,6 +140,3 @@ plothr <- function(season, animal, level = 0.95){
   season_name <- season_names[season]
   title(paste(stringr::str_to_title(animals[[season]][animal]), season_name))
 }
-
-plothr(3, 3)
-plothr(3, 3, level = 0.5)

### 2024-02-28 scripts/06.0_postModeling_viz.R
centroid visualization

@@ -1,0 +1,129 @@
+library(tidyverse)
+library(sf)
+library(mapview)
+load("data/dataPrep/downsampled_10min_forSocial.Rda")
+roostpolys <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
+load("data/dataPrep/season_names.Rda")
+load("data/dataPrep/roosts.Rda")
+load("data/mixedModels/linked.Rda")
+
+# Edge effects ------------------------------------------------------------
+
+# Central roosts ----------------------------------------------------------
+roosts <- map2(roosts, season_names, ~.x %>% mutate(seasonUnique = .y))
+roosts <- purrr::list_rbind(roosts) %>%
+  select(Nili_id, seasonUnique, date, roost_date, location_lat, location_long) %>%
+  sf::st_as_sf(coords=c("location_long", "location_lat"), remove = F, crs = "WGS84")
+
+linked_tojoin_deg <- linked %>%
+  select(Nili_id, seasonUnique, space_use, movement, roost_div, season, year, type, normDegree, n) %>%
+  pivot_wider(id_cols = c("Nili_id", "seasonUnique", "space_use", "movement", "roost_div", "season", "year"), names_from = "type", values_from = "normDegree", names_prefix = "deg_")
+linked_tojoin_str <- linked %>%
+  select(Nili_id, seasonUnique, space_use, movement, roost_div, season, year, type, normStrength, n) %>%
+  pivot_wider(id_cols = c("Nili_id", "seasonUnique", "space_use", "movement", "roost_div", "season", "year"), names_from = "type", values_from = "normStrength", names_prefix = "str_")
+linked_tojoin <- left_join(linked_tojoin_deg, linked_tojoin_str)
+linked_tojoin %>% group_by(Nili_id, seasonUnique) %>% filter(n()>1) # should be 0 rows, good.
+
+roosts_joined <- roosts %>%
+  left_join(linked_tojoin, by = c("Nili_id", "seasonUnique")) %>%
+  mutate(seasonUnique = factor(seasonUnique, levels = season_names))
+
+roosts_joined %>%
+  ggplot(aes(x = movement, y = location_long, col = season))+
+  geom_point(alpha = 0.2)+
+  theme_classic()+
+  scale_color_manual(values = c(cc$breedingColor, cc$summerColor, cc$fallColor))+
+  facet_wrap(~seasonUnique, scales = "free")+
+  theme(legend.position = "none")+
[…]

### 2024-02-28 scripts/05.4_viz_vultureNatHist.R
replaced with new season names

@@ -1,5 +1,5 @@
 # Vulture natural history visualizations, general
-# 2023-09-13
+# 2024-02-24
 
 library(vultureUtils)
 library(tidyverse)
@@ -9,8 +9,7 @@ library(ggspatial)
 library(grid)
 library(ggmap)
 library(Polychrome)
-load("data/derived/seasonNames_orig.Rda")
-seasonNames <- seasonNames_orig[-1]
+load("data/dataPrep/season_names.Rda")
 
 # Roosts map --------------------------------------------------------------
 r <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")

### 2024-02-29 docs/manuscript.qmd
moved manuscript and other docs to separate folder

--- 2024-02-28 manuscript.qmd
@@
 
+#### Social network centrality is known to be important in animals.
+
 The social structure of animal groups has consequences for disease transmission, information sharing, and population dynamics.
@@
 
+#### We observe that some individuals are more central than others in a social network. What causes this variation?
+
 An individual's social centrality may be determined by many factors.
 Differences in social position have been attributed to individual phenotypic characteristics, such as age, sex, or personality \[REFS\].
+
+##### Individual traits
+
+-   Multiple spatial behaviors → Social centrality @albery2021
+
+-   Exploration → Weak associations with more conspecifics, more movement between flocks. @aplin2013
+
+-   Slow-exploring → More stable relationships @aplin2013
+
+-   Personality → Social structure @aplin2013
+
+-   Stress → Weaker parental bonds, in captivity @boogert2014
+
+-   Stress → Foraging with more others, less choosy, more central, in captivity @boogert2014
+
+-   Stress → Foraging with more others, less choosy, more central, in the wild @brandl2019
+
+-   More exploratory → higher degree, in finches @moyers2018
+
+-   Quick explorers → More interactions @moyers2018
+
+-   Slower explorers → less central ("Yet, slower explorers overall did not travel shorter distances than faster explorers, indicating that a less central social network position was not merely a consequence of lower activity.") @snijders2014
+
+-   Females → higher strength, more preferred partners (only during lactation) @wey2013
+
+-   Females → more stable partnerships @wey2013
+
+##### Individuals' spatial phenotypes are important determinants of their social centrality
+
 Of course, how an animal uses space is also an important driver of how many conspecifics it interacts with, and how frequently.
@@
 In this light, the "spatial phenotype" may be considered in addition to more conventional phenotypic measures to explain why some individuals are more central than others in their social networks.
+
+-   Definition of "spatial phenotype" @webber2023
+
+-   Definition of "spatial behaviour" @albery2021
+
+-   Note that explicit spatial context is also important, and we didn't consider it. (We should have! How do we justify not doing that?)
+
+    -   An individual's observed spatial phenotype results from a combination of its internal tendency toward movement and the external environmental conditions through which it moves [(Nathan et al.
+2008)](https://www.zotero.org/google-docs/?IyWvRB)
+
+##### Social interactions occur when individuals encounter each other, in various behavioral situations.
+
+-   Co-occurrence in space and time is necessary for social interaction to occur. Therefore, more movement → more social interaction, in general.
+
+-   However, movement is also an expression of an individual's underlying tendency to move, as well as potentially an emergent consequence of things like dominance rank or social environment.
+
+-   Social centrality arises from a combination of the number of interactions (strength) and unique individuals encountered.
+An individual can achieve high social centrality by being more geographically central in the population, versus by being more versatile.
+
+-   No one has investigated the link between spatial phenotype and social centrality.
+We know that individuals vary in both spatial phenotype and social centrality, but we don't know how the one influences the other.
+
+    -   Furthermore, social networks are not monolithic.
+Because the biological mechanisms behind social network structure are different in different social situations, it stands to reason that spatial phenotype should be related to social centrality in different ways in the different social contexts.
+
+        -   For example, different components of spatial phenotype might be expected to relate differently to networks in different social situations.
+
+        -   Also, we might just see different-direction effects.
+For example, individuals that move a lot might interact more frequently in flight but interact less frequently at feeding stations (either because they spend less time feeding or because they feed at more peripheral locations).
+Therefore, it is important to consider different behavioral situations separately when teasing apart the relationship between the social and spatial phenotypes.
 
@@
 
+##### We set out to uncover whether spatial phenotypes could explain observed variation in social centrality \[should we add centroid location as a factor?\]
+
+-   Hypotheses/predictions from slides
+
+##### Research Questions
+
+Q1: What is the relationship between spatial phenotype and social interactions?
+
+-   H1A: Individuals that move more will have more, and stronger, encounters with other unique individuals.
+
+    -   P1A: Vultures that move more will encounter more others (higher degree centrality in the social network) and have more interactions (higher strength in the social network) than vultures who move less.
+
+-   H1B: Individuals that are more exploratory will have more but weaker encounters with other unique individuals.
+
+    -   P1B: Exploratory vultures will encounter more unique individuals (higher degree centrality), but will have fewer interactions (lower strength) compared to less exploratory vultures.
+
+Q2: How is the relationship between spatial phenotype and social interactions impacted by social situation and season?
+
+-   H2A: Social situations in which animals move more will have a greater impact on social interactions
+
+    -   P2A: Social centrality in co-flight networks will be more strongly associated with spatial phenotype than in co-roosting and co-feeding behavioral situations.
+
+-   H2B: Seasons in which animals move more will have a stronger impact on social interactions
+
+    -   P2B: The relationship between movement and social position will be stronger in the non-breeding seasons (summer and fall) than during the breeding season.
+
+Q3: How does the time scale of movement predict social interactions?
+
+-   H3A: The relationship between movement and social centrality will be strongest within the time scales at which change occurs in the social network.
+
+    -   P3A: Co-flight networks will change hourly to daily, co-feeding networks will change every few days, and co-roosting networks will change over longer periods of time (weeks to months).
+
+    -   P3B: Social position in each of the social situations will be best predicted by individuals' movement patterns aggregated over the corresponding time scale.
+
+### Study system
+
 Eurasian griffon vultures (*Gyps fulvus*) provide a study system well suited to ascertaining the relationship between individuals' spatial phenotypes and their social centrality.
@@
 
--   Unusual GPS tracking coverage of the population–high proportion of individuals tracked, so we can get a population-level sense of movement behavior and social interaction patterns
+-   Unusual GPS tracking coverage of the population--high proportion of individuals tracked, so we can get a population-level sense of movement behavior and social interaction patterns
 
@@
     -   Communal roosts: many individuals together at once, variability in size/density of roosts.
-Used for information exchange \[REFS\] and also important for disease transmission \[or not–cf Elvira Nili\].
+Used for information exchange \[REFS\] and also important for disease transmission \[or not--cf Elvira Nili\].
 Breeding/nesting also occurs at/near roost sites, on cliffs.
@@
 
-# Research Questions
-
-Q1: What is the relationship between spatial phenotype and social interactions?
-
--   H1A: Individuals that move more will have more, and stronger, encounters with other unique individuals.
-
-    -   P1A: Vultures that move more will encounter more others (higher degree centrality in the social network) and have more interactions (higher strength in the social network) than vultures who move less.
-
--   H1B: Individuals that are more exploratory will have more but weaker encounters with other unique individuals.
-
-    -   P1B: Exploratory vultures will encounter more unique individuals (higher degree centrality), but will have fewer interactions (lower strength) compared to less exploratory vultures.
-
-Q2: How is the relationship between spatial phenotype and social interactions impacted by social situation and season?
-
--   H2A: Social situations in which animals move more will have a greater impact on social interactions
-
-    -   P2A: Social centrality in co-flight networks will be more strongly associated with spatial phenotype than in co-roosting and co-feeding behavioral situations.
-
--   H2B: Seasons in which animals move more will have a stronger impact on social interactions
-
-    -   P2B: The relationship between movement and social position will be stronger in the non-breeding seasons (summer and fall) than during the breeding season.
-
-Q3: How does the time scale of movement predict social interactions?
-
--   H3A: The relationship between movement and social centrality will be strongest within the time scales at which change occurs in the social network.
-
-    -   P3A: Co-flight networks will change hourly to daily, co-feeding networks will change every few days, and co-roosting networks will change over longer periods of time (weeks to months).
-
-    -   P3B: Social position in each of the social situations will be best predicted by individuals' movement patterns aggregated over the corresponding time scale.
-
-# Methods
-
 ## Study system
@@
 
-## Tagging and aging
+# Methods
+
+## Field methods
 
@@
 
-We subsetted the dataset to exclude periods of time that vultures spent in capture cages, in the wildlife hospital following poisonings or injuries, or in captivity prior to wild release (in the case of captive-born birds).
-We cleaned the data to remove invalid GPS fixes.
+-   Download data from movebank (ornitela, inpa). Give general dates.
+
+-   (Fix IDs)\--don't need to report this in the manuscript
+
+-   Remove invalid periods in the hospital and capture cages (1 sentence)
+
+    -   We subsetted the dataset to exclude periods of time that vultures spent in capture cages, in the wildlife hospital following poisonings or injuries, or in captivity prior to wild release (in the case of captive-born birds).
+
+-   Data cleaning
+
+    -   Basic data quality filters
+
+        -   We cleaned the data to remove invalid GPS fixes.
+
+    -   Remove spiky speeds
+
+-   Double-masking process
+
+-   Split into seasons
+
+    -   Then we split data from each into three seasons.
+The fall, or "pre-breeding," season includes dates from 15 September through 14 December; the breeding season lasts from 15 December through 14 May of the following year; and the summer lasts from 15 May through 14 September.
+These season boundaries were chosen based on biological knowledge of the vultures' breeding timelines (see more details in the Study System section).
+Therefore, the analysis contained a total of 8 seasons, from the fall of 2020 through the 2023 breeding season (Table X).
+
+-   Further cleaning for movement data
+
 Then we split data from each into three seasons.
@@
 For each season, we calculated movement measures only for individuals tracked for at least 30 days (after removing low-point days as described above).
-We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of \[level–more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of \[level--more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
 Finally, since some tags were occasionally set to a much higher fix rate (e.g.
@@
 2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
-We calculated home range as the 95% AKDE of a vulture's GPS fixes in a season @
+We calculated home range as the 95% AKDE of a vulture's GPS fixes in a season \@
 

### 2024-03-06 scripts/dataCleaning.qmd

---
title: "dataCleaning"
format: html
editor: visual
---

```{r}
#| include: false
library(tidyverse)
#devtools::install_github("kaijagahm/vultureUtils")
library(vultureUtils)
library(here)
library(readxl)
```

```{r}
base::load(here::here("movebankCredentials/pw.Rda"))
MB.LoginObject <- move::movebankLogin(username = "kaijagahm", 
                                      password = pw)
rm(pw)
```

## Download data

```{r}
# Set min and max date
minDate <- "2023-12-01 00:00"
maxDate <- "2023-12-07 00:00"

mydata <- vultureUtils::downloadVultures(loginObject = MB.LoginObject,
                                         removeDup = T, 
                                         dfConvert = T, 
                                         quiet = T, 
                                         dateTimeStartUTC = minDate,
                                         dateTimeEndUTC = maxDate)
```

## Join `Nili_id`s

```{r}
ww <- read_excel(here("data/raw/whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")
ww_tojoin <- ww %>% select(Nili_id, Movebank_id) %>% distinct() # pull out just the names columns, nothing else, and remove any duplicates

# join by movebank ID
mydata <- left_join(mydata, ww_tojoin, 
                        by = c("local_identifier" = "Movebank_id"))
```

## Remove invalid periods (hospital etc.)

```{r}
periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove") %>%
  select(Nili_id, remove_start, remove_end) %>%
  mutate(across(contains("remove"), lubridate::ymd)) %>%
  filter(!is.na(remove_end)) %>%
  # The following steps generate a sequence of days between the start and end date, so we can then join them to the original data
  group_by(Nili_id) %>%
  mutate(dateOnly = map2(remove_start, remove_end, seq, by = "1 day")) %>%
  unnest(cols = c(dateOnly)) %>%
  select(Nili_id, dateOnly) %>%
  mutate(remove = T)

removal_annotated <- mydata %>%
  mutate(dateOnly = lubridate::date(timestamp)) %>%
  left_join(periods_to_remove, by = c("Nili_id", "dateOnly"))

removed_periods <- removal_annotated %>%
  filter(is.na(remove))
```

## GPS jamming filter

```{r}
#XXX Ryan's working on this
removeJammed <- vultureUtils::[functionname](removed_periods)
```

## Data cleaning

```{r}
cleaned <- cleanData(dataset = removed_periods,
                     precise = F, 
                     longCol = "location_long",
                     latCol = "location_lat",
                     idCol = "Nili_id",
                     gpsJam = T) # need to remove all gps-jamming-related things in this function
```

### 2024-03-06 scripts/dataCleaning.qmd
tried out the gps jamming and cleaning functions, and started working on a capture cage function

@@
 library(readxl)
+library(mapview)
 ```
@@
 
-## Download data
+## *Download data*
 
@@
 
-## Join `Nili_id`s
+## *Join `Nili_id`s*
 
 ```{r}
-ww <- read_excel(here("data/raw/whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")
+ww <- read_excel(here("data/raw/whoswho_vultures_20230920_new.xlsx"), 
+                 sheet = "all gps tags")
 ww_tojoin <- ww %>% select(Nili_id, Movebank_id) %>% distinct() # pull out just the names columns, nothing else, and remove any duplicates
@@
 
-## Remove invalid periods (hospital etc.)
+## *Remove invalid periods (hospital etc.)*
 
 ```{r}
-periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove") %>%
+periods_to_remove <- read_excel(here("data/raw/whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove") %>%
   select(Nili_id, remove_start, remove_end) %>%
@@
 removed_periods <- removal_annotated %>%
-  filter(is.na(remove))
+  filter(is.na(remove)) %>%
+  sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = F)
 ```
 
-## GPS jamming filter
+## *GPS jamming filter*
 
 ```{r}
-#XXX Ryan's working on this
-removeJammed <- vultureUtils::[functionname](removed_periods)
+jamPolygons <- sf::st_read(here("data/raw/GPS_jamming_3.kml"))
+
+# Visualize the jam polygons with the points on top
+mapview(jamPolygons, col.regions = "red") + 
+  mapview(removed_periods, col.regions = "blue", color = "blue", cex = 3, alpha = 0.5)
+
+# Remove points that fall into the jam polygons
+removed_jammed <- gpsJamFilter(dataset = removed_periods, mask = jamPolygons,
+                               longCol = "location_long", 
+                               latCol = "location_lat", idCol = "Nili_id")
+
+# and now visualize the after
+mapview(jamPolygons, col.regions = "red") +
+  mapview(removed_jammed, col.regions = "skyblue", color = "skyblue", cex = 3, alpha = 0.5)
 ```
 
-## Data cleaning
+## *Data cleaning*
 
 ```{r}
-cleaned <- cleanData(dataset = removed_periods,
+cleaned <- cleanData(dataset = removed_jammed,
                      precise = F, 
@@
                      idCol = "Nili_id",
-                     gpsJam = T) # need to remove all gps-jamming-related things in this function
+                     report = F)
+
+cleaned_precise <- cleanData(dataset = removed_jammed,
+                             precise = T,
+                             longCol = "location_long",
+                             latCol = "location_lat",
+                             idCol = "Nili_id",
+                             report = F)
+
+dim(removed_jammed)
+dim(cleaned)
+dim(cleaned_precise)
 ```
+
+## *Remove time spent in capture cages*
+
+```{r}
+# Get roosts (for the purposes of capture cage assignment)
+roosts <- vultureUtils::get_roosts_df(cleaned, 
+                                     id = "Nili_id")
+
+# Load information about capture sites
+capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
+load(here::here("data", "dataPrep", "roosts0.Rda"))
+AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))
+
+removeCaptures <- function(data, roosts, captureSites, distance = 500){
+  # Identify the period of time during which the capture sites are open (when we need to do this exclusion)
+  start.day <- 01
+  start.month <- 08
+  end.day <- 30
+  end.month <- 11
+
+  # Get roosts that fall in the capture period
+  sub.roosts <- roosts %>%
+    mutate(start_date = lubridate::dmy(paste(start.day, start.month,
+                                      lubridate::year(date), sep = "-")),
+           end_date = lubridate::dmy(paste(end.day, end.month, 
+                                           lubridate::year(date), sep = "-"))) %>%
+    filter(date >= start_date & date <= end_date)
+
+  if(nrow(sub.roosts) > 0){
+    # Calculate the roost distance to each of the capture cages. If it is less than 500m, keep that line.
+  crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
+                 nrow = nrow(sub.roosts), ncol = 2) # roost locs as simple lat/long coords
+  distanceMat <- matrix(ncol = nrow(captureSites),
+                        nrow = nrow(crds))
+  colnames(distanceMat) <-  unique(captureSites$name)
+
+  for(i in 1:nrow(crds)){
+    distanceMat[i,] <- round(geosphere::distm(crds[i,], 
+                                              captureSites[,c(3,2)]), 2)
+    }
+  }
+
+  closestCaptureSite <- colnames(distanceMat)[apply(distanceMat, 1, which.min)] # ID of closest capture site
+  closestCaptureDist <- apply(distanceMat, 1, min) # distance from closest capture site
+
+  sub.roosts <- cbind(sub.roosts, closestCaptureSite, closestCaptureDist)
+  sub.roosts$captured <- ifelse(sub.roosts$closestCaptureDist <= distance, T, F)
+  sub.captured.dates <- sub.roosts %>% 
+    filter(captured) %>%
+    select(Nili_id, date, closestCaptureSite, 
+           closestCaptureDist, captured)
+  # XXX START HERE
+
+}
+
+## For Carmel--different protocol.
+sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
+sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")
+
+AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")
+
+AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
+AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
+AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))
+
+AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)
+
+sub.captured.carmel <- sub.captured.carmel %>%
+  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
+         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
+  filter(Captured == "yes") %>%
+  dplyr::select(-c(known_capture))
+
+names(sub.captured.no.carmel)
+names(sub.captured.carmel)
+
+sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)
+
+# We also need to exclude the day after the bird was captured
+sub.captured.dates.1 <- sub.captured.dates
+sub.captured.dates.1$date <- sub.captured.dates.1$date+1
+
+sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
+sub.captured.dates <- sub.captured.dates %>%
+  dplyr::distinct(Nili_id, date, .keep_all = T)
+
+# It all looks ok, so we can subset the dataset to exclude the capture dates
+removed_captures <- removed_periods %>%
+  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
+nrow(removed_periods) == nrow(removed_captures) # should be TRUE. NOW we can filter.
+removed_captures <- removed_captures %>%
+  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured
+
+# How many rows and individuals did we remove?
+before <- nrow(removed_periods)
+after <- nrow(removed_captures)
+(propChange <- (before-after)/before) # approx. 1% of data removed.
+length(unique(removed_periods$Nili_id)) == length(unique(removed_captures$Nili_id)) # no change in number of individuals! As expected.
+```

### 2024-03-13 scripts/reduc.R
moved temporal scripts to temporal project

@@ -1,220 +1,0 @@
-# Load libraries
-library(tidyverse)
-library(lubridate)
-library(tnet)
-library(sna)
-
-# Data ingest (already sliced into appropriate chunks)
-load("data/calcSocial/feedingEdges.Rda")
-feedingEdges <- purrr::list_rbind(feedingEdges) %>% mutate(date = lubridate::date(minTimestamp))
-load("data/calcSocial/flightEdges.Rda")
-flightEdges <- purrr::list_rbind(flightEdges) %>% mutate(date = lubridate::date(minTimestamp))
-load("data/calcSocial/roostingEdges.Rda")
-roostingEdges <- purrr::list_rbind(roostingEdges)
-source("scripts/muxLib DF.R")
-
-cutfun <- function(vec, days){
-  min <- min(vec)
-  max <- max(vec)
-  ngroups <- ceiling(as.numeric(max-min)/days)
-  brks <- seq(from = min, by = days, length.out = ngroups + 1)
-  cut <- cut(vec, breaks = brks, include.lowest = T, right = F)
-}
-
-# Feed into code to do the rest
-doReduce <- function(data, groupingCol){
-  # Make a binary edgelist, grouping by the group variable
-  bin <- data[,c("ID1", "ID2", groupingCol)] %>%
-    distinct() %>%
-    mutate(weight = 1)
-  
-  opp_edges <- bin
-  names(opp_edges)[1:2] <- c("ID2", "ID1")
-  
-  all_bin <- bind_rows(bin, opp_edges) # attach 
-  
-  # Add layer numbers (days) to make the next step easier
-  all_bin <- all_bin %>%
-    arrange(.data[[groupingCol]], ID1, ID2) %>%
-    mutate(layer = as.numeric(.data[[groupingCol]])) %>%
-    select(-all_of(groupingCol)) %>%
-    rename("layer1" = "layer") %>%
-    mutate(layer2 = layer1) %>%
-    relocate(layer1, .after = "ID1") %>%
-    relocate(layer2, .after = "ID2")
-  
-  self_edges <- expand.grid(as.character(unique(all_bin$ID1)), 1:(max(all_bin$layer1)-1)) # create all possible combinations of indiv ID and layer
-  colnames(self_edges) <- c("ID1", "layer1")
-  
-  self_edges <- self_edges %>%
-    mutate(ID2 = ID1, layer2 = layer1 + 1, weight = 1)
[…]

### 2024-03-13 scripts/repeatability.R

@@ -1,705 +1,0 @@
-# Exploring repeatability
-library(tidyverse)
-library(vultureUtils)
-library(igraph)
-library(sf)
-library(tidygraph)
-library(ggraph)
-library(zoo)
-
-# Repeatability across seasons --------------------------------------------
-load("data/mixedModels/linked.Rda")
-future::plan(future::multisession(), workers = 10)
-
-## Just using the traits we've already calculated and plotting across seasons, and coloring by the average value for each individual across all of its seasons, let's see if we get gradients
-
-plotRepeat <- function(df, var, situation = "flight", ylab){
-  legend_title <- paste0("Mean\n", var, "\nvalue")
-  plt <- df %>%
-    filter(type == situation) %>%
-    group_by(Nili_id) %>%
-    mutate(avg = mean(.data[[var]], na.rm = T)) %>%
-    ggplot(aes(x = seasonUnique, y = get(var), group = Nili_id, color = avg))+
-    geom_point()+
-    geom_line(alpha = 0.7)+
-    theme_classic()+
-    scale_color_viridis_c(name = legend_title)+
-    xlab("Season")+
-    ylab(ylab)+
-    theme(axis.title = element_text(size = 16))
-  return(plt)
-}
-
-# PREDICTORS
-## note that for these, the situation doesn't matter (defaults to flight) because the predictors are just repeated over and over for each season
-plotRepeat(linked, "movement", ylab = "Movement")
-plotRepeat(linked, "space_use", ylab = "Space use (log-transformed)")
-plotRepeat(linked, "roost_div", ylab = "Roost diversification")
-
-# SOCIAL METRICS
-plotRepeat(linked, "normDegree", situation = "flight", ylab = "Normalized degree (flight)")
-plotRepeat(linked, "normDegree", situation = "feeding", ylab = "Normalized degree (feeding)")
-plotRepeat(linked, "normDegree", situation = "roosting", ylab = "Normalized degree (roosting)")
-plotRepeat(linked, "normStrength", situation = "flight", ylab = "Normalized strength (flight)")
-plotRepeat(linked, "normStrength", situation = "feeding", ylab = "Normalized strength (feeding)")
-plotRepeat(linked, "normStrength", situation = "roosting", ylab = "Normalized strength (roosting)")
-
-# Just by looking at these colors, I would say that the measures (both predictors and responses) are somewhat repeatable, but not strongly.
-
[…]

### 2024-03-13 scripts/temp_multilayer_test.R

@@ -1,119 +1,0 @@
-# Setup for temporal multilayer network analysis
-library(tidyverse)
-library(lubridate)
-library(tnet)
-library(sna)
-
-load("data/calcSocial/feedingEdges.Rda")
-testseason <- feedingEdges[[9]] # just doing this for one arbitrary season
-testseason <- testseason %>%
-  mutate(date = lubridate::date(minTimestamp))
-dates <- seq(from = min(testseason$date), to = min(testseason$date) + 4, by = 1)
-slice <- testseason %>% dplyr::filter(date %in% dates)
-
-# Make a binary edgelist, grouping by day
-bin <- slice %>%
-  mutate(day = lubridate::date(minTimestamp)) %>% # for feeding/flight
-  #mutate(day = date) %>% # for roosing
-  select(ID1, ID2, day) %>%
-  distinct() %>%
-  mutate(weight = 1)
-
-opp_edges <- bin
-names(opp_edges)[1:2] <- c("ID2", "ID1")
-
-all_bin <- bind_rows(bin, opp_edges) # attach 
-
-# Add layer numbers (days) to make the next step easier
-all_bin <- all_bin %>%
-  arrange(day, ID1, ID2) %>%
-  mutate(layer = as.integer(as.factor(day))) %>%
-  select(-day) %>%
-  rename("layer1" = "layer") %>%
-  mutate(layer2 = layer1) %>%
-  relocate(layer1, .after = "ID1") %>%
-  relocate(layer2, .after = "ID2")
-
-self_edges <- expand.grid(as.character(unique(all_bin$ID1)), 1:(max(all_bin$layer1)-1)) # create all possible combinations of indiv ID and layer
-colnames(self_edges) <- c("ID1", "layer1")
-
-self_edges <- self_edges %>%
-  mutate(ID2 = ID1, layer2 = layer1 + 1, weight = 1)
-self_edges2 <- self_edges %>%
-  mutate(layer1 = layer1 + 1, layer2 = layer2-1)
-self_edges_all <- bind_rows(self_edges, self_edges2)
-
-# Combine with symmetrical edgelist
-all_edges <- bind_rows(all_bin, self_edges)
-
-# If we're staying in R, muxViz needs each edgelist to number the vultures from 1 to the number of vultures
-
-all_edges <- all_edges %>%
-  mutate(ID1 = as.integer(factor(ID1)),
-         ID2 = as.integer(factor(ID2)))
-
-# Save
[…]

### 2024-03-13 scripts/00.0_dataPrep.qmd
test gps jamming data for ryan

@@
 
+# Also get some to test GPS jamming (for Ryan 2/28/24)
+mindate_jamming <- "2023-10-01"
+maxdate_jamming <- "2024-02-01"
+test_jamming_ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = mindate_jamming, dateTimeEndUTC = maxdate_jamming)
+write_csv(test_jamming_ornitela, file = here("data/dataPrep/test_jamming_ornitela.csv"))
+
 # tidy up
@@
 rm(maxDate)
+rm(mindate_jamming)
+rm(maxdate_jamming)
+rm(test_jamming_ornitela)
 gc()

### 2024-03-13 scripts/01.0_akde.R
got overall season centroids for use in network graphs

@@ -92,6 +92,9 @@ for(i in 1:length(fits_list)){
 }
 tictoc::toc()
 save(uds_w, file = "data/akde/uds_w.Rda")
+uds_w_sf <- map(uds_w, ~map(.x, ~as.sf(.x) %>% sf::st_transform("WGS84"), 
+                            .progress = T))
+save(uds_w_sf, file = "data/akde/uds_w_sf.Rda")
 
 stats_w_95 <- vector(mode = "list", length = length(uds_w))
 for(i in 1:length(stats_w_95)){
@@ -140,3 +143,68 @@ plothr <- function(season, animal, level = 0.95){
   season_name <- season_names[season]
   title(paste(stringr::str_to_title(animals[[season]][animal]), season_name))
 }
+
+# PKDE --------------------------------------------------------------------
+test_telems <- telems_list[[1]]
+test_uds <- uds_w[[1]]
+test_pkde <- pkde(data = test_telems, UD = test_uds, kernel = "individual", weights = TRUE, ref = "Gaussian")
+pkde(data = DATA, UD = uds, kernel = "individual", weights = FALSE, ref = "Gaussian")
+#DATA needs to be a list of telemetry objects (with the same length as uds) and not just one telemetry object.
+# This isn't working....
+
+# What about just some ellipses for basic visualization?
+test_season <- downsampled_10min[[5]]
+test_season_minimal <- test_season %>% ungroup() %>% sample_n(ceiling(nrow(test_season)/10)) # make it 1/10 the size
+
+test_season %>%
+  filter(Nili_id %in% unique(test_season$Nili_id)[1:10]) %>%
+  ggplot(aes(x = location_long, y = location_lat, col = Nili_id))+
+  geom_point(alpha = 0.7, size = 0.2)+
+  theme_void()+
+  theme(legend.position = "none")+
+  stat_ellipse()
+
+# What about the actual sf objects?
+load("data/akde/uds_w_sf.Rda")
+uds_w_sf_seasons <- map(uds_w_sf, ~do.call(rbind, .x))
+load("data/dataPrep/season_names.Rda")
+uds_w_sf_seasons <- map2(uds_w_sf_seasons, season_names, ~.x %>% mutate(seasonUnique = .y))
+sfs <- do.call(rbind, uds_w_sf_seasons)
+sfs <- sfs %>%
[…]

### 2024-03-13 scripts/02.0_calcSocial.R
removed note

@@ -21,7 +21,7 @@ rm(downsampled_10min_forSocial)
 gc()
 roostPolygons <- sf::st_read("data/raw/roosts50_kde95_cutOffRegion.kml")
 load("data/dataPrep/season_names.Rda")
-load("data/dataPrep/roosts.Rda") # XXX load
+load("data/dataPrep/roosts.Rda")
 load("data/derived/cc.Rda")
 
 # Social Networks ---------------------------------------------------------

### 2024-03-13 scripts/04.0_mixedModels.R
added centroids to models

@@ -29,10 +29,23 @@ networkMetrics <- networkMetrics %>% rename("seasonUnique" = "season")
 load("data/new_movement_vars.Rda")
 load("data/dataPrep/season_names.Rda")
 load("data/derived/cc.Rda")
+load("data/akde/sfs_est_centroids.Rda")
+centrs <- sfs_est_centroids %>%
+  select(seasonUnique, Nili_id, dist_szn_centr) %>%
+  group_by(seasonUnique) %>%
+  mutate(mn = mean(dist_szn_centr),
+         sd = sd(dist_szn_centr)) %>%
+  mutate(centr = (dist_szn_centr-mn)/sd,
+         centr = -1*centr) %>%
+  sf::st_drop_geometry() %>%
+  select(seasonUnique, Nili_id, centr) %>%
+  ungroup() %>%
+  distinct()
 
 # Join the two datasets (there will be 3 rows per individual per season, for the 3 different situations)
 linked <- new_movement_vars %>%
-  left_join(networkMetrics, by = c("Nili_id", "seasonUnique"))
+  left_join(networkMetrics, by = c("Nili_id", "seasonUnique")) %>%
+  left_join(centrs)
 nrow(linked) == 3*nrow(new_movement_vars)
 
 # Housekeeping
@@ -109,8 +122,8 @@ linked <- linked %>%
 # Modeling ----------------------------------------------------------------
 
 # Degree ------------------------------------------------------------------
-# degree_base <- glmmTMB(normDegree ~ situ + movement + roost_div + space_use + age_group + season + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian())
-# check_model(degree_base) # the good news is that the variance inflation factor is quite low, so these predictors aren't prohibitively highly correlated (yay!!!!). The bad news is that the posterior predictive check and the residuals vs. fitted values plots look... bad. 
+# degree_base <- glmmTMB(normDegree ~ situ + movement + roost_div + space_use + age_group + season + centr + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian())
+# check_model(degree_base) # the good news is that the variance inflation factor is quite low, so these predictors aren't prohibitively highly correlated (yay!!!!). The bad news is that the posterior predictive check and the residuals vs. fitted values plots look... bad.
[…]

### 2024-03-25 scripts/dataCleaning.qmd
prepared dataCleaning.qmd for new vultureUtils functions

@@
 capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
-load(here::here("data", "dataPrep", "roosts0.Rda"))
 AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))
 
-removeCaptures <- function(data, roosts, captureSites, distance = 500){
+# Full code as it currently exists in the data cleaning pipeline:
 # Identify the period of time during which the capture sites are open (when we need to do this exclusion)
@@
 end.month <- 11
+distance <- 500 # distance, in meters, to calculate from the cage
 
-  # Get roosts that fall in the capture period
-  sub.roosts <- roosts %>%
-    mutate(start_date = lubridate::dmy(paste(start.day, start.month,
-                                      lubridate::year(date), sep = "-")),
-           end_date = lubridate::dmy(paste(end.day, end.month, 
-                                           lubridate::year(date), sep = "-"))) %>%
+# Subset the roost dataset with the start and end dates for the capture period
+sub.roosts <- roosts0 %>%
+  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
+                              format="%d-%m-%Y"),
+         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
+                            format="%d-%m-%Y")) %>%
   filter(date >= start_date & date <= end_date)
 
-  if(nrow(sub.roosts) > 0){
-    # Calculate the roost distance to each of the capture cages. If it is less than 500m, keep that line.
+unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.
+
+# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
 crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
-                 nrow = nrow(sub.roosts), ncol = 2) # roost locs as simple lat/long coords
-  distanceMat <- matrix(ncol = nrow(captureSites),
-                        nrow = nrow(crds))
-  colnames(distanceMat) <-  unique(captureSites$name)
+               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates
 
-  for(i in 1:nrow(crds)){
-    distanceMat[i,] <- round(geosphere::distm(crds[i,], 
-                                              captureSites[,c(3,2)]), 2)
-    }
+DistanceMat <- matrix(ncol = nrow(capture_sites), nrow = nrow(crds))
+colnames(DistanceMat) <- unique(capture_sites$name)
+
+for(i in 1:nrow(crds)){ # for each roost point...
+  DistanceMat[i,] <- round(geosphere::distm(crds[i,], capture_sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
+  #print(i)
 }
 
-  closestCaptureSite <- colnames(distanceMat)[apply(distanceMat, 1, which.min)] # ID of closest capture site
-  closestCaptureDist <- apply(distanceMat, 1, min) # distance from closest capture site
+ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
+ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site
 
-  sub.roosts <- cbind(sub.roosts, closestCaptureSite, closestCaptureDist)
-  sub.roosts$captured <- ifelse(sub.roosts$closestCaptureDist <= distance, T, F)
-  sub.captured.dates <- sub.roosts %>% 
-    filter(captured) %>%
-    select(Nili_id, date, closestCaptureSite, 
-           closestCaptureDist, captured)
-  # XXX START HERE
+sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist)
+sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")
 
-}
+sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage
+
+sub.captured.dates <- subset(sub.captured,
+                             select = c("Nili_id",
+                                        "date",
+                                        "ClosestCaptureSite",
+                                        "ClosestCaptureDist",
+                                        "Captured"))
 

### 2024-03-25 scripts/testinterpolation.R
moved to old

@@ -1,180 +1,0 @@
-load("data/dataPrep/removed_lfr.Rda")
-library(tidyverse)
-library(sf)
-
-interpFun <- function(df, season){
-  rowsToAdd <- data.frame()
-  whichGap <- which(df$gap)
-  n <- length(whichGap)
-  for(i in whichGap){
-    cat(paste0(season, ": Interpolating gap ", which(whichGap == i), " of ", n,
-               "\n"))
-    Nili_id <- df$Nili_id[i]
-    start <- lubridate::ymd_hms(df$timestamp[i])
-    end <- lubridate::ymd_hms(df$timestamp[i+1])
-    timediff <- as.numeric(difftime(end, start, units = "mins"))
-    nPeriods <- ceiling(timediff/10)
-    nTimes <- nPeriods - 1
-    times <- seq.POSIXt(from = start, to = end, length.out = 2+nTimes)
-    times <- times[-c(1, length(times))] # remove the anchor times because those are already in the dataset
-    newrows <- data.frame(Nili_id = Nili_id,
-                          timestamp = times)
-    rowsToAdd <- bind_rows(rowsToAdd, newrows)
-  }
-  return(rowsToAdd)
-}
-
-# JUST ONE SEASON OF DATA
-season <- removed_lfr[[2]]
-pp <- season %>%
-  select(Nili_id, dateOnly, timestamp, location_long, location_lat) %>%
-  sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = FALSE) %>%
-  arrange(Nili_id, dateOnly, timestamp) %>%
-  group_by(Nili_id, dateOnly) %>%
-  mutate(timeToNext = as.numeric(difftime(lead(timestamp), timestamp, units = "mins")),
-         distToNext = sf::st_distance(geometry, geometry[row_number()+1],
-                                      by_element = T),
-         distToNext = as.numeric(distToNext))
-
-gp <- pp %>%
-  mutate(gap = case_when(timeToNext > 10 &
-                           timeToNext <= 60 &
-                           !is.na(timeToNext) &
-                           !is.na(lead(timeToNext))&
-                           distToNext <= 100
-                         ~ TRUE,
-                         TRUE ~ FALSE))
-whichGap <- which(gp$gap)
-
-toInterpolate <- interpFun(gp, season = "test")
-
-out <- gp %>% 
-  bind_rows(toInterpolate %>% mutate(interpolated = TRUE)) %>%
[…]

### 2024-03-25 scripts/04.0_mixedModels.R
updated the whole analysis with the roost network calculated by 500m distance instead of shared polygons

@@ -83,7 +83,7 @@ save(linked, file = "data/mixedModels/linked.Rda")
 #   theme(legend.position = "none", text = element_text(size = 20))
 # normDegree_years_seasons_flight
 # ggsave(normDegree_years_seasons_flight, filename = "fig/normDegree_years_seasons_flight.png", width = 9, height = 7)
-# 
+
 # normDegree_years_seasons_feeding <- linked %>%
 #   filter(type == "feeding") %>%
 #   mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
@@ -96,19 +96,19 @@ save(linked, file = "data/mixedModels/linked.Rda")
 #   facet_wrap(~season)+
 #   theme(legend.position = "none", text = element_text(size = 20))
 # ggsave(normDegree_years_seasons_feeding, filename = "fig/normDegree_years_seasons_feeding.png", width = 9, height = 7)
-# 
-# normDegree_years_seasons_roosting <- linked %>%
-#   filter(type == "roosting") %>%
-#   mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
-#   ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-#   geom_density(linewidth = 1.5)+
-#   theme_classic()+
-#   scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
-#   ylab("Frequency")+
-#   xlab("Degree (normalized)")+
-#   facet_wrap(~season)+
-#   theme(legend.position = "none", text = element_text(size = 20))
-# ggsave(normDegree_years_seasons_roosting, filename = "fig/normDegree_years_seasons_roosting.png", width = 9, height = 7)
+
+normDegree_years_seasons_roosting <- linked %>%
+  filter(type == "roosting") %>%
+  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
+  ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
+  geom_density(linewidth = 1.5)+
+  theme_classic()+
+  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
+  ylab("Frequency")+
+  xlab("Degree (normalized)")+
+  facet_wrap(~season)+
+  theme(legend.position = "none", text = element_text(size = 20))
+ggsave(normDegree_years_seasons_roosting, filename = "fig/normDegree_years_seasons_roosting.png", width = 9, height = 7)
 
[…]

### 2024-03-25 scripts/05.0_viz_descriptive.R

@@ -92,3 +92,4 @@ focals %>%
   xlab("Season")+
   scale_color_manual(values = c(cc[["feedingColor"]], cc[["flightColor"]], cc[["roostingColor"]]))+
   theme(text = element_text(size = 16))
+

### 2024-03-27 docs/manuscript.qmd
edits to methods 3/27

@@
 
-##### Individuals' spatial phenotypes are important determinants of their social centrality
+#### Individuals' spatial phenotypes are important determinants of their social centrality
 
@@
 
-##### Social interactions occur when individuals encounter each other, in various behavioral situations.
+#### Social interactions occur when individuals encounter each other, in various behavioral situations.
 
@@
 
-##### We set out to uncover whether spatial phenotypes could explain observed variation in social centrality \[should we add centroid location as a factor?\]
+#### We set out to uncover whether spatial phenotypes could explain observed variation in social centrality.
 
@@
 
-Q3: How does the time scale of movement predict social interactions?
-
--   H3A: The relationship between movement and social centrality will be strongest within the time scales at which change occurs in the social network.
-
-    -   P3A: Co-flight networks will change hourly to daily, co-feeding networks will change every few days, and co-roosting networks will change over longer periods of time (weeks to months).
-
-    -   P3B: Social position in each of the social situations will be best predicted by individuals' movement patterns aggregated over the corresponding time scale.
+# Methods {data-link="methods"}
 
@@
 
-Eurasian griffon vultures (*Gyps fulvus*) provide a study system well suited to ascertaining the relationship between individuals' spatial phenotypes and their social centrality.
+This work was conducted on a population of GPS-tagged Eurasian griffon vultures in southern Israel, a study system well suited to studying the relationship between movement and sociality.
+Griffon vultures are large obligate scavengers that use soaring flight to locate food resources (carcasses) on the landscape.
+They may travel long distances without expending very much energy.
+Griffon vultures are locally endangered in Israel, where they have declined significantly since \[DATE\], and they face significant anthropogenic threats, most commonly through pesticide and NSAID poisonings, power line collisions, and shootings.
+The population in southern Israel has an unusually high GPS tracking coverage, with a high proportion of individuals (approximately 80-90%) fitted with GOS tags.
+The tags send a position reading ever ten minutes if battery condition is good, or less frequently otherwise.
+\[paper\] says that we only need \[percentage\] of individuals in order to get a good sense of the population-level social structure, and our coverage far exceeds that.
+We are able to get a very good sense of individuals' movement behavior and social interaction patterns.
 
--   obligate scavengers, large soaring fliers, lots of movement
+Griffon vultures interact in several social situations, each with a different biological relevance.
+In particular, griffons rely on social information to locate food resources.
+Interactions in flight (i.e.
+when individuals fly within each other's perceptual distance) provide an opportunity for local enhancement of carcass resources, or following to a carcass site.
+Therefore, they are important for information transfer [@harel2017], finding locations of thermal uplift [@sassi2023], and feeding sites [@spiegel2013; @cortés-avizanda2014].
+Vultures also obtain social information about food resources at communal roosting sites.
+Roost sites can vary in size and density, with many vultures roosting together at once \[REF\].
+These roosts also serve as breeding colonies during the breeding season (approximately December through May).
+Finally, vultures interact frequently at or around carcasses while eating.
+The free-living griffon vultures in the southern Israel population are provisioned at several feeding sites, where livestock carcasses are deposited irregularly to supplement the availability of naturally-occurring carcasses on the landscape.
+Vultures may co-feed, associate in close proximity near a carcass, or even fight over the food resource.
 
--   Unusual GPS tracking coverage of the population--high proportion of individuals tracked, so we can get a population-level sense of movement behavior and social interaction patterns
-
--   Three clear social situations in which they interact, each with a different significance/relevance.
-
-    -   Communal roosts: many individuals together at once, variability in size/density of roosts.
-Used for information exchange \[REFS\] and also important for disease transmission \[or not--cf Elvira Nili\].
-Breeding/nesting also occurs at/near roost sites, on cliffs.
-
-    -   Flight: interactions in flight are important for information transfer [@harel2017], finding locations of thermal uplift [@sassi2023] and feeding sites [@spiegel2013; @cortés-avizanda2014].
-
-    -   Feeding: Communal feeding at carcasses.
-Also sometimes standing around the carcass site.
-Sometimes associate for a while before beginning to feed, e.g.
-waiting for critical mass to be reached (Orr Spiegel, pers.
-comm.) Potentially also important for disease transmission \[but not, cf.
-ref Elvira Nili\].
-
-While some griffon vultures make long-range forays \[REFS\], there is a mostly resident population in a relatively small area in southern Israel \[REF\].
-The population is small and restricted enough that individuals' home ranges are highly overlapping, with most individuals using a large fraction of available space in the region.
-This makes habitat features likely to affect most individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement [@harel2016].
-GPS fixes are taken once every 10 minutes, frequently enough to measure flight movements with good accuracy, and most individuals are tracked for several months to several years.
-Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population [@sharma2023].\
-\
-\[General introduction to the importance of time scale in social network analysis\]
-
--   [@blonder2012; @cantor2012; @hobson2013; @rasmussen2013; @holme2015; @farine2018; @fisher2021; @pasquaretta2021]
-
-Social situations occur over different time scales, and this is an important consideration when studying the causes and consequences of their structure.
-In defining co-roosting, co-flight, and co-feeding interactions based on proximity in each of these behavioral situations, we have created social networks in which edges already conceptually represent interactions over different amounts of time.
-Because individuals have one roosting location per night, edge weights in the co-roosting network are measured in units of days, while pairs of individuals can interact many times in flight or at a carcass on a given day.\
-\
-The biological processes driving the execution of these behaviors and the formation of these networks are likewise happening over different time scales.
-An individual must decide once per day which roost to go to [@harel2016].
-But flight is a continuous process driven by many decisions resulting from individual and social information [@sassi2023], and feeding events are concentrated in time on a given day.
-Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every \[NUMBER\] days and persist for hours to a few days (Orr Spiegel, pers.
-comm.).
-Individuals may return to the same carcass on successive days.
-While at a carcass, individuals may fight and interact many times.
-
-## Study system
-
-To answer these questions, we analyzed data from a free-flying population of Eurasian griffon vultures (Gyps fulvus) fitted with GPS tags.
-
-\[tag specs\]
-
-Over the period \[date\] to \[date\], the dataset contained \[number\] of tagged vultures.
-Tags recorded GPS fixes, as well as speed and heading readings, every 10 minutes when battery conditions allowed.
-The GPS data are stored on Movebank [@kays2022] and were loaded into R using the `move` package\[ @kranstauber2023\].
-
-# Methods
+The southern Israel griffon vulture population is mostly resident, though some individuals (especially juveniles) occasionally make long-range forays \[REF\].
+The population is small and restricted enough that individuals' home ranges are highly overlapping, with most individuals using a large fraction of available habitat in the region.
+This makes habitat features likely to affect many individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement [@harel2016].
+GPS fixes are taken frequently enough to monitor flight movements with good accuracy, and most individuals are tracked for several months to several years.
+Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population [@sharma2023].
 
@@
 
-## Data cleaning
+\[tag specs\]
 
--   Download data from movebank (ornitela, inpa). Give general dates.
+## Data preparation
 
--   (Fix IDs)\--don't need to report this in the manuscript
+For this study, we began with data between September 1, 2020 and September 15, 2023 from the \[ornitela\] and \[inpa\] datasets, which we downloaded from Movebank [@kays2022].
+For each vulture, we removed any periods of time spent in the capture cage or in the wildlife hospital (e.g.
+while recovering from an illness, injury, or poisoning), or in captivity prior to wild release (in the case of captive-born birds).
+We cleaned the data by filtering out GPS fixes deemed unreliable (e.g.
+too few satellites for triangulation, nonsensical heading readings, impossible jumps from previous fixes).
+Altogether, we removed \<1% of fixes this way (xxx LITERATE PROGRAMMING--PUT ACTUAL NUMBER WHEN DATA CLEANING IS FINAL).
+The data were divided into seasons based on biological knowledge of vultures' breeding timelines: a breeding season (December 15 through May 14 of the following year); a summer or post-breeding season (May 15th through September 14th), and a transient or pre-breeding season (September 15th through December 14th).
+For simplicity, we refer to each breeding season by the calendar year in which it ends (e.g.
+the 2021 breeding season lasts from December 15, 2020 through May 14, 2021).
+We analyzed data from the 2020 pre-breeding/transient season through the 2023 post-breeding/summer season, for a total of nine seasons across four calendar years (three of each type of season).
 
--   Remove invalid periods in the hospital and capture cages (1 sentence)
+For each season, we included only vultures that registered at least 30% of their GPS fixes within the general area of Israel (to exclude, for instance, any birds that were hardly detected in the area before leaving for a long-range foray).
+The vulture population in Israel is geographically separated, with most individuals in the south but a small and relatively isolated population at the Hai Bar Carmel Nature Reserve and in the Golan Heights.
+While vultures occasionally move between the northern and southern population, such movements are rare, and northern vultures behave differently than southern vultures due to more consistent feeding and a more restricted range in the north.
+Therefore, for each season, we restricted our analysis to individuals whose spatial centroid (including all GPS points after data cleaning) lay south of UTM Northing 3550000 (Zone 36N), equivalent to approximately 32.086238 North latitude.
+This threshold was determined by visual examination of a histogram of centroid latitudes, which was bimodal and disjoint.
 
-    -   We subsetted the dataset to exclude periods of time that vultures spent in capture cages, in the wildlife hospital following poisonings or injuries, or in captivity prior to wild release (in the case of captive-born birds).
+#### Determining roost locations
 
--   Data cleaning
+We used GPS fixes to determine vultures' nightly roost locations.
+The solar-powered GPS tags are set to record very few fixes overnight, and cell reception (needed for data transmission) is often very poor on the cliffs where roosting sites are located.
+Therefore, we combined information from individuals' last locations of the evening and first locations of the morning to assign a roost site for each individual on each night.
+Full details about the data cleaning and roost assignment procedure can be found in the supplementary material.
 
-    -   Basic data quality filters
-
-        -   We cleaned the data to remove invalid GPS fixes.
-
-    -   Remove spiky speeds
-
--   Double-masking process
-
--   Split into seasons
-
-    -   Then we split data from each into three seasons.
-The fall, or "pre-breeding," season includes dates from 15 September through 14 December; the breeding season lasts from 15 December through 14 May of the following year; and the summer lasts from 15 May through 14 September.
-These season boundaries were chosen based on biological knowledge of the vultures' breeding timelines (see more details in the Study System section).
-Therefore, the analysis contained a total of 8 seasons, from the fall of 2020 through the 2023 breeding season (Table X).
-
--   Further cleaning for movement data
-
-Then we split data from each into three seasons.
-The fall, or "pre-breeding," season includes dates from 15 September through 14 December; the breeding season lasts from 15 December through 14 May of the following year; and the summer lasts from 15 May through 14 September.
-These season boundaries were chosen based on biological knowledge of the vultures' breeding timelines (see more details in the Study System section).
-Therefore, the analysis contained a total of 8 seasons, from the fall of 2020 through the 2023 breeding season (Table X).
-All of these data was reserved for use in proximity-based social network analysis.
-Since the GPS tags give few to no fixes overnight, and since reception is often very poor on the cliffs where vultures roost, we combined information from individuals' last locations of the evening and first locations of the morning to assign a roost site for each individual on each night \[more details\].
-Full details about the data cleaning and roost assignment procedure can be found in the supplementary material.
+### Measuring spatial phenotype
 

### 2024-03-27 docs/manuscript.qmd
complete draft of data cleaning description

@@
 For this study, we began with data between September 1, 2020 and September 15, 2023 from the \[ornitela\] and \[inpa\] datasets, which we downloaded from Movebank [@kays2022].
-For each vulture, we removed any periods of time spent in the capture cage or in the wildlife hospital (e.g.
+For each vulture, we removed any periods of time spent in the wildlife hospital (e.g.
 while recovering from an illness, injury, or poisoning), or in captivity prior to wild release (in the case of captive-born birds).
 We cleaned the data by filtering out GPS fixes deemed unreliable (e.g.
-too few satellites for triangulation, nonsensical heading readings, impossible jumps from previous fixes).
-Altogether, we removed \<1% of fixes this way (xxx LITERATE PROGRAMMING--PUT ACTUAL NUMBER WHEN DATA CLEANING IS FINAL).
-The data were divided into seasons based on biological knowledge of vultures' breeding timelines: a breeding season (December 15 through May 14 of the following year); a summer or post-breeding season (May 15th through September 14th), and a transient or pre-breeding season (September 15th through December 14th).
+too few satellites for triangulation, nonsensical heading readings, impossible jumps from previous fixes), and we removed any time that vultures spent in capture cages during the tagging season in the fall.
+Altogether, approximately \[PERCENT\] of fixes were removed as compared to the raw data.
+Next, we removed any vultures that spent fewer than 30 days, in total, within a focal region encompassing the general vicinity of Israel (to exclude, for instance, birds that were briefly detected in the area before dispersing far away).
+For vultures detected in the area for at least 30 days during the study period, we subsequently removed any fixes for those individuals that occurred outside of the same focal region, in order to restrict our analyses to movements that occurred near the main population locus and ensure that most social interactions would be with other tagged individuals.
+(We accounted for missing time periods, including removed data, by calculating association indices in the social networks).
+
+We divided the remaining data into seasons based on biological knowledge of vultures' breeding timelines: a breeding season (December 15 through May 14 of the following year); a summer or post-breeding season (May 15th through September 14th), and a transient or pre-breeding season (September 15th through December 14th).
 For simplicity, we refer to each breeding season by the calendar year in which it ends (e.g.
@@
 
-For each season, we included only vultures that registered at least 30% of their GPS fixes within the general area of Israel (to exclude, for instance, any birds that were hardly detected in the area before leaving for a long-range foray).
 The vulture population in Israel is geographically separated, with most individuals in the south but a small and relatively isolated population at the Hai Bar Carmel Nature Reserve and in the Golan Heights.
@@
 This threshold was determined by visual examination of a histogram of centroid latitudes, which was bimodal and disjoint.
+Finally, to account for older GPS tags that were set to lower fix frequencies, we excluded any individuals from each season that never had their tags set to a 10-minute frequency during that season.
+In recent months, some tags were temporarily set to record fixes at much higher frequencies, so those individuals were downsampled to a 10-minute frequency in order to make data analysis more tractable.
+
+After this cleaning, the data used for social network analysis contained \[NUMBER\] of unique vultures in total.
+Table \[NUMBER\] shows the number of individuals included in the social networks for each season.
 
@@
 
-### Measuring spatial phenotype
+### Further cleaning for movement analysis
 
@@
 Fixes were ideally collected every 10 minutes, but sometimes they were less frequent due to bad cellular reception in the cliff areas where vultures often sit or due to low battery charge.
-For each individual, only days with 10 or more GPS fixes per day were retained, unless the battery charge never dipped below 50% during that day.
-In those cases, the low number of points was attributed to the vultures spending most of their time sitting on a cliff without much movement, rather than to missed points due to low battery.
+For each individual, only days with 10 or more GPS fixes per day were retained, unless the battery charge remained above 50% for the entire day.
+In those cases, the low number of points was attributed to the vultures spending most of their time sitting on a cliff (with limited reception) without much movement, rather than to actual missing data due to low battery.
 Since these sedentary days represent an important component of vulture movement behavior, we retained this information (Orr Spiegel, pers.
 obs.).
+
+After removing days with too few fixes per day, we removed any vultures that did not have at least 30 days of valid movement data remaining in a given season, to ensure that movement estimates and averages were based on a high enough sample size of days to capture the range of the vulture's behavior.
+Since this was done on a per-season basis, individuals that were removed from one season's analysis due to not having enough valid days could still have movement measures calculated for a different season.
+Finally, we downsampled the movement data to a 10-minute interval to remove high-frequency fixes.
 

### 2024-03-28 docs/manuscript.qmd
some updates to methods

@@
 bibliography: references.bib
-csl: the-american-naturalist.csl
 ---
@@
 
--   Multiple spatial behaviors → Social centrality @albery2021
+-   Multiple spatial behaviors → Social centrality [@albery2021]
 
--   Exploration → Weak associations with more conspecifics, more movement between flocks. @aplin2013
+-   Exploration → Weak associations with more conspecifics, more movement between flocks. [@aplin2013]
 
--   Slow-exploring → More stable relationships @aplin2013
+-   Slow-exploring → More stable relationships [@aplin2013]
 
--   Personality → Social structure @aplin2013
+-   Personality → Social structure [@aplin2013]
 
--   Stress → Weaker parental bonds, in captivity @boogert2014
+-   Stress → Weaker parental bonds, in captivity [@boogert2014]
 
--   Stress → Foraging with more others, less choosy, more central, in captivity @boogert2014
+-   Stress → Foraging with more others, less choosy, more central, in captivity [@boogert2014]
 
--   Stress → Foraging with more others, less choosy, more central, in the wild @brandl2019
+-   Stress → Foraging with more others, less choosy, more central, in the wild [@brandl2019]
 
--   More exploratory → higher degree, in finches @moyers2018
+-   More exploratory → higher degree, in finches [@moyers2018]
 
--   Quick explorers → More interactions @moyers2018
+-   Quick explorers → More interactions [@moyers2018]
 
--   Slower explorers → less central ("Yet, slower explorers overall did not travel shorter distances than faster explorers, indicating that a less central social network position was not merely a consequence of lower activity.") @snijders2014
+-   Slower explorers → less central ("Yet, slower explorers overall did not travel shorter distances than faster explorers, indicating that a less central social network position was not merely a consequence of lower activity.") [@snijders2014]
 
--   Females → higher strength, more preferred partners (only during lactation) @wey2013
+-   Females → higher strength, more preferred partners (only during lactation) [@wey2013]
 
--   Females → more stable partnerships @wey2013
+-   Females → more stable partnerships [@wey2013]
 
@@
 
--   Definition of "spatial phenotype" @webber2023
+-   Definition of "spatial phenotype" [@webber2023]
 
--   Definition of "spatial behaviour" @albery2021
+-   Definition of "spatial behaviour" [@albery2021]
 
@@
 Vultures also obtain social information about food resources at communal roosting sites.
-Roost sites can vary in size and density, with many vultures roosting together at once \[REF\].
+Roost sites can vary in size and density, with many vultures roosting together at once.
 These roosts also serve as breeding colonies during the breeding season (approximately December through May).
@@
 
+```{r}
+#| include: false
+load(here("data/dataPrep/dataPrep_report_stats.Rda"))
+dp <- dataPrep_report_stats
+cl <- which(dp$dataset == "data_cleaned")
+pctremovedcleaned <- round(100*(dp$nrow[[cl-1]] - dp$nrow[[cl]])/(dp$nrow[[cl-1]]), 2)
+```
+
 For this study, we began with data between September 1, 2020 and September 15, 2023 from the \[ornitela\] and \[inpa\] datasets, which we downloaded from Movebank [@kays2022].
@@
 too few satellites for triangulation, nonsensical heading readings, impossible jumps from previous fixes), and we removed any time that vultures spent in capture cages during the tagging season in the fall.
-Altogether, approximately \[PERCENT\] of fixes were removed as compared to the raw data.
-Next, we removed any vultures that spent fewer than 30 days, in total, within a focal region encompassing the general vicinity of Israel (to exclude, for instance, birds that were briefly detected in the area before dispersing far away).
+Altogether, approximately `r pctremovedcleaned`% of fixes were removed as compared to the raw data.
+Next, we removed any vultures that spent fewer than 30 days, in total, within a focal region encompassing the general vicinity of Israel (in order to exclude, for instance, birds that were briefly detected in the area before dispersing far away).
 For vultures detected in the area for at least 30 days during the study period, we subsequently removed any fixes for those individuals that occurred outside of the same focal region, in order to restrict our analyses to movements that occurred near the main population locus and ensure that most social interactions would be with other tagged individuals.
@@
 
-After this cleaning, the data used for social network analysis contained \[NUMBER\] of unique vultures in total.
-Table \[NUMBER\] shows the number of individuals included in the social networks for each season.
+```{r}
+#| include: false
+soc <- which(dp$dataset == "downsampled_10min_forSocial")
+vulturessoc <- unique(unlist(dp$vultures[[soc]]))
+nvulturessoc <- length(vulturessoc)
+nvulturessoc_szns <- dp$nvultures[[soc]]
+```
+
+After this cleaning, the data used for social network analysis contained `r nvulturessoc` unique vultures in total (range: `r min(nvulturessoc_szns)`-`r max(nvulturessoc_szns)`; mean = `r round(mean(nvulturessoc_szns), 2)` vultures per season, sd = `r round(sd(nvulturessoc_szns), 2)` vultures per season).
 
@@
 
-### Further cleaning for movement analysis
+#### Further cleaning for movement analysis
 
 To characterize individual movement behavior and spatial phenotype, we cleaned the data further.
-We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package suncalc [@thieurmel2022].
+We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package {suncalc} [@thieurmel2022].
 We restricted movement calculations to a set of focal individuals with activity regions centered in the Negev region of Israel, the site of most of our tagging effort.
@@
 
+```{r}
+#| include: false
+mov <- which(dp$dataset == "downsampled_10min")
+vulturesmov <- unique(unlist(dp$vultures[[mov]]))
+nvulturesmov <- length(vulturesmov)
+nvulturesmov_szns <- dp$nvultures[[mov]]
+```
+
 After removing days with too few fixes per day, we removed any vultures that did not have at least 30 days of valid movement data remaining in a given season, to ensure that movement estimates and averages were based on a high enough sample size of days to capture the range of the vulture's behavior.
@@
 Finally, we downsampled the movement data to a 10-minute interval to remove high-frequency fixes.
+Ultimately, we calculated movement measures for a total of `r nvulturesmov` **focal individuals** in total (range: `r min(nvulturesmov_szns)`-`r max(nvulturesmov_szns)`; mean = `r round(mean(nvulturesmov_szns), 2)` vultures per season, sd = `r round(sd(nvulturesmov_szns), 2)` vultures per season).
 
-\[Geography\]
-
-For each season, we calculated movement measures only for individuals tracked for at least 30 days (after removing low-point days as described above).
-We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of \[level--more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
-Finally, since some tags were occasionally set to a much higher fix rate (e.g.
-every 1 minute) for short periods of time, we downsampled the data to a 10 minute fix rate to make sure that all individuals had comparable fix rates.
-
-## Movement calculations
+## Measuring spatial phenotype
 
@@
 2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
-We calculated home range as the 95% AKDE of a vulture's GPS fixes in a season \@
+We used several measures to capture an individual's space use, roosting behavior, and flight movements.
+
+### Space use
+
+We calculated seasonal home ranges and core areas for each focal vulture, using autocorrelated kernel density estimates (AKDEs) via the {ctmm} R package [@fleming2023].
+We defined the home range as the 95% AKDE and the core area as the 50% AKDE.
+To measure differences in how individuals used their home ranges, we calculated the ratio of the seasonal core area size to the seasonal home range size.
+We defined a general "space use" variable as the first principal component of a PCA containing the home range size and the core area/home range ratio.
+The "space use" principal component explained \[PERCENT\] of the variation in the data, with individuals low on the principal component having \[CHARACTERISTICS\] and individuals high on the principal component having \[CHARACTERISTICS\].
+
+### Roost diversification
 
@@
 Then, for each vulture in each season, we calculated the number of unique roosts used, the proportion of nights switching roosts (versus returning to the same roost as the previous night), and the diversity of roosts used (Shannon diversity index).
+
+### Movement
+
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the {elevatr} package [@hollister2022], at a zoom level of \[level--more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
+Finally, since some tags were occasionally set to a much higher fix rate (e.g.
+every 1 minute) for short periods of time, we downsampled the data to a 10 minute fix rate to make sure that all individuals had comparable fix rates.
 
@@
 
-We used the R package `igraph` [@csardi2005] to calculate each individual's degree and strength for each social situation in each season.
+We used the R package {igraph} [@csardi2005] to calculate each individual's degree and strength for each social situation in each season.
 Degree is defined as the number of other individuals that a vulture is connected to (i.e.
@@
 
-We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the lme4 and lmerTest packages [@bates2015; @kuznetsova2017].
+We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the {lme4} and {lmerTest} packages [@bates2015; @kuznetsova2017].
 We fit separate models for each response variable (degree and strength).
@@
 
-For each response variable, we began by fitting a model without any interaction terms to check for multicollinearity between the predictors and check model assumptions, using the performance and DHARMa packages [@lüdecke2021; @hartig2022].
+For each response variable, we began by fitting a model without any interaction terms to check for multicollinearity between the predictors and check model assumptions, using the {performance} and {DHARMa} packages [@lüdecke2021; @hartig2022].
 Next, we fitted a model containing two three-way interaction terms (between season, situation, and each of the spatial phenotype variables), all of the included two-way interaction terms, age as a main effect, and the two random effects (individual ID; season by year).
@@
 
-For each model, we calculated marginal effects using the emmeans package [@lenth2023] and created visualizations using ggeffects and ggplot2 [@wickham2016; @lüdecke2018].
+For each model, we calculated marginal effects using the {emmeans} package [@lenth2023] and created visualizations using {ggeffects} and {ggplot2} [@wickham2016; @lüdecke2018].
 
@@
 ```{r}
-#| include: false
+#| eval: false
 source(here("scripts/04.0_mixedModels.R"))
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-d-space
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-d-roost
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-d-space-situ
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-s-movement
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-d-movement-situ
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-s-roost
@@
 ```{r}
-#| echo: false
+#| eval: false
 #| label: fig-s-space

### 2024-03-28 scripts/00.0_dataPrep.qmd
Beginning to simplify with the new vultureUtils functions

--- 2024-02-21 scripts/00.0_dataPrep.qmd
@@
 
-Inputs:
-
--   `MB.LoginObject`
-
-Outputs:
-
--   `data/dataPrep/inpa.feather`
-
 ```{r download-inpa}
@@
 write_feather(inpa, here::here("data", "dataPrep", "inpa.feather"))
-
-length(unique(inpa$trackId))
-
-# tidy up
-rm(inpa)
-gc()
 ```
@@
 
-Inputs:
-
--   `MB.LoginObject`
-
-Outputs:
-
--   `data/dataPrep/ornitela.feather`
+Data collected by the TAU-UCLA partnership under the NSF-BSF grant. Ornitela tags.
 
@@
 write_feather(ornitela, here::here("data", "dataPrep", "ornitela.feather"))
-
-length(unique(ornitela$trackId))
-
-# tidy up
-rm(ornitela)
-rm(minDate)
-rm(maxDate)
-gc()
 ```
@@
 ## Join INPA and Ornitela data
-
-*Inputs:*
-
--   `data/dataPrep/inpa.feather`
-
--   `data/dataPrep/ornitela.feather`
-
-*Outputs:*
-
--   `data/dataPrep/joined.feather`
 
@@
 # Will there be a problem joining these?
-names(ornitela)[!(names(ornitela) %in% names(inpa))] # all names in "ornitela" are found in inpa
-names(inpa)[!(names(inpa) %in% names(ornitela))] # many are not found in "ornitela" that are found in inpa
+length(names(ornitela)[!(names(ornitela) %in% names(inpa))]) # all names in "ornitela" are found in inpa
+length(names(inpa)[!(names(inpa) %in% names(ornitela))]) # many are not found in "ornitela" that are found in inpa
 
@@
 rm(ornitela)
+```
 
-## fix trackId
+```{r join-nili-ids}
+ww <- read_excel(here("data/raw/whoswho_vultures_20230920_new.xlsx"), 
+                 sheet = "all gps tags")
+ww_tojoin <- ww %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() # pull out just the names columns, nothing else, and remove any duplicates
+
+# Prepare for join: are there any individuals in the `local_identifier` column of `joined0` that don't appear in the `Movebank_id` column of `ww_tojoin`?
+problems <- joined0 %>% filter(!(local_identifier %in% ww_tojoin$Movebank_id)) %>% pull(local_identifier) %>% unique()
+problems #let's check these against the who's who and see if we can make some reasonable changes.
+
+## Fixes:
+# Typo in the Movebank_id column of the who's who:
+ww_tojoin <- ww_tojoin %>% mutate(Movebank_id = case_when(Movebank_id == "A65 Whiite" ~ "A65 White",
+                                                          .default = Movebank_id))
+# Fixes to joined0:
 joined0 <- joined0 %>%
-  mutate(trackId = as.character(trackId),
-         trackId = case_when(trackId == "E03" ~ "E03w",
-                             TRUE ~ trackId))
+  mutate(local_identifier = case_when(local_identifier == "E86 White" ~ "E86",
+                                      local_identifier == "E88 White" ~ "E88w",
+                                      .default = local_identifier))
 
-# number of unique individuals
-length(unique(joined0$trackId))
+## Look for any remaining problems:
+problems <- joined0 %>% filter(!(local_identifier %in% ww_tojoin$Movebank_id)) %>% pull(local_identifier) %>% unique()
+problems #going to fix both of these afterward.
+E66 isn't listed in the Who's who at all, so we'll just call it "E66" in the Nili_id.
+The other one, Y01>T60 W, I've manually determined is Nili_id "tammy".
 
-# Add the Nili_id
-ww <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
-  dplyr::select(Nili_id, Movebank_id) %>%
-  distinct()
+# join by movebank ID
+joined <- left_join(joined0, ww_tojoin, 
+                        by = c("local_identifier" = "Movebank_id"))
+joined <- joined %>%
+  mutate(Nili_id = case_when(is.na(Nili_id) & local_identifier == "E66 White" ~ "E66",
+                             is.na(Nili_id) & local_identifier == "Y01>T60 W" ~ "tammy",
+                             .default = Nili_id))
 
-# Are there any other individuals who have two different associated trackId's?
-ww %>% filter(Movebank_id %in% joined0$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
-ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's
+# Are there any remaining NA's for Nili_id?
+nas <- joined %>% filter(is.na(Nili_id)) %>% pull(local_identifier) %>% unique()
+length(nas) # yay, no more!
 
-all(joined0$trackId %in% ww$Movebank_id) # false.
-
-# Okay, let's investigate the ones that aren't included
-joined0 %>%
-  filter(!(trackId %in% ww$Movebank_id)) %>%
-  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces
-
-joined0$trackId <- str_replace_all(tolower(joined0$trackId), "\\.", "")
-ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")
-
-sort(unique(ww$mb_id))
-unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
-joined0$trackId[joined0$trackId == "y01t60w"] <- "t60white"
-joined0$trackId[joined0$trackId == "e86white"] <- "e86"
-unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
-# Investigate the remaining ones:
-joined0 %>% filter(!(trackId %in% ww$mb_id)) %>% 
-  dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>%
-  group_by(trackId, local_identifier, ring_id, dataset) %>% 
-  summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
-# I guess for now let's just make their trackId's into Nili_id's
-
-joined <- left_join(joined0, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
-                  by = c("trackId" = "mb_id"))
-nrow(joined) == nrow(joined0) # good, same number of rows. (check that the join worked)
-length(unique(joined$Nili_id))
-# For any that didn't get Nili_ids assigned, make it the trackId
-joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expected.
-joined <- joined %>%
-  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
-                             TRUE ~ Nili_id))
-length(unique(joined$Nili_id))
-write_feather(joined, here::here("data", "dataPrep", "joined.feather"))
-
-# tidy up
-rm(joined0)
-rm(joined)
-gc()
+write_feather(joined, here("data/dataPrep/joined.feather"))
 ```
@@
 
-*Inputs:*
-
--   `data/raw/whoswho_vultures_20230920_new.xlsx`
-
--   `data/dataPrep/joined.feather`
-
-*Outputs:*
-
--   `data/dataPrep/removed_periods.feather`
-
 ```{r load-hospital}
@@
 ```{r remove-periods}
-## annotate the data with periods to remove
-periods_to_remove <- periods_to_remove %>%
-  dplyr::select(Nili_id,
-                remove_start,
-                remove_end,
-                reason) %>%
-  mutate(across(c(remove_start, remove_end), .fns = function(x){
-    lubridate::ymd(x)})) %>%
-  dplyr::filter(!is.na(remove_end)) %>%
-  group_by(Nili_id, reason) %>%
-  # sequence of daily dates for each corresponding start, end elements
-  dplyr::mutate(dateOnly = map2(remove_start, remove_end, 
-                                seq, by = "1 day")) %>%
-  # unnest the list column
-  unnest(cols = c(dateOnly)) %>%
-  # remove any duplicate rows
-  distinct() %>%
-  dplyr::select(-c(remove_start, remove_end)) %>%
-  rename("status" = reason)
-
-# Join to the original data
-removal_annotated <- joined %>%
-  left_join(periods_to_remove, by = c("Nili_id", "dateOnly")) %>%
-  mutate(status = replace_na(status, "valid"))
-nrow(removal_annotated) == nrow(joined) #T: good, same number of rows.
-
-# Actually REMOVE the periods to remove...
-removed_periods <- removal_annotated %>%
-  filter(status == "valid")
-nrow(removed_periods)
-length(unique(removal_annotated$Nili_id))
-nrow(joined) == nrow(removal_annotated) # true
-nrow(removed_periods) != nrow(removal_annotated) # true
-
-write_feather(removed_periods, here::here("data", "dataPrep", "removed_periods.feather"))
-
-# tidy up
-rm(periods_to_remove)
-rm(joined)
-rm(removal_annotated)
-rm(removed_periods)
-gc()
+removed_periods <- vultureUtils::removeInvalidPeriods(dataset = joined, periodsToRemove = periods_to_remove)
+save(removed_periods, file = here("data/dataPrep/removed_periods.Rda"))
 ```
 
-### Capture cages
-
-This is Marta's code for identifying periods when the vultures were in the capture cages and removing days of data around those occasions.
-
-To identify the capture dates, first we need to classify the roosts (now using the `get_roosts()` function).
-Then, if the bird roosted within 500m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset.
-This has been validated.
-This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it.
-So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data.
-
-Note: I'm calling the roosts data frame generated here `roosts0.Rda` because I want to distinguish it from the more final roost location *list* that we will generate further along in the cleaning process.
-
-*Inputs:*
-
--   `data/dataPrep/removed_periods.feather`
-
--   `data/raw/capture_sites.csv`
-
--   `data/raw/all_captures_carmel_2010-2021.csv`
-
-*Outputs:*
-
--   `data/dataPrep/roosts0.Rda`
-
--   `data/dataPrep/removed_captures.feather`
+## Clean the data based on filters for invalid data points (see `vultureUtils` package code for details)
 
 ```{r}
-removed_periods <- read_feather(here::here("data", "dataPrep", "removed_periods.feather"))
+# warning: this takes a long time and uses a lot of memory!
+data_cleaned <- vultureUtils::cleanData(dataset = removed_periods,
+                     precise = F,
+                     longCol = "location_long",
+                     latCol = "location_lat",
+                     idCol = "Nili_id",
+                     report = F)
+rm(joined)
+gc()
+save(data_cleaned, file = here("data/dataPrep/data_cleaned.Rda"))
+
+(nrow(removed_periods)-nrow(data_cleaned))/nrow(removed_periods) # removed less than 1% of the points.
 ```
 
-```{r remove-capture-cage-getroosts}
-#| eval: false
-# Identify and remove capture dates using Marta's code --------------------
-roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
-save(roosts0, file = here::here("data", "dataPrep", "roosts0.Rda"))
+## Remove time spent in capture cages
+```{r}
+capture_sites <- read.csv(here("data/raw/capture_sites.csv"))
+carmel <- read.csv(here("data/raw/all_captures_carmel_2010-2021.csv"))
+removed_captures <- removeCaptures(data = data_cleaned, 
+                                       captureSites = capture_sites, 
+                                       AllCarmelDates = carmel, 
+                                       distance = 500, idCol = "Nili_id")
 
-# tidy up
-rm(roosts0)
-```
-
-```{r}
-# Load information about the capture sites
-capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
-load(here::here("data", "dataPrep", "roosts0.Rda"))
-AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))
-```
-
-```{r remove-capture-cage}
-# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
-start.day <- 01
-start.month <- 08
-end.day <-  30
-end.month <- 11
-distance <- 500 # distance, in meters, to calculate from the cage
-
-# Subset the roost dataset with the start and end dates for the capture period
-sub.roosts <- roosts0 %>%
-  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
-                              format="%d-%m-%Y"),
-         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
-                            format="%d-%m-%Y")) %>%
-  filter(date >= start_date & date <= end_date)
-
-unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.
-
-# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
-crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
-               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates
-
-DistanceMat <- matrix(ncol = nrow(capture_sites), nrow = nrow(crds))
-colnames(DistanceMat) <- unique(capture_sites$name)
-
-for(i in 1:nrow(crds)){ # for each roost point...
-  DistanceMat[i,] <- round(geosphere::distm(crds[i,], capture_sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
-  #print(i)
-}
-
-ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
-ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site
-
-sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist)
-sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")
-
-sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage
-
-sub.captured.dates <- subset(sub.captured,
-                             select = c("Nili_id",
-                                        "date",
-                                        "ClosestCaptureSite",
-                                        "ClosestCaptureDist",
-                                        "Captured"))
-
-## For Carmel--different protocol.
-sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
-sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")
-
-AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")
-
-AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
-AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
-AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))
-
-AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)
-
-sub.captured.carmel <- sub.captured.carmel %>%
-  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
-         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
-  filter(Captured == "yes") %>%
-  dplyr::select(-c(known_capture))
-
-names(sub.captured.no.carmel)
-names(sub.captured.carmel)
-
-sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)
-
-# We also need to exclude the day after the bird was captured
-sub.captured.dates.1 <- sub.captured.dates
-sub.captured.dates.1$date <- sub.captured.dates.1$date+1
-
-sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
-sub.captured.dates <- sub.captured.dates %>%
-  dplyr::distinct(Nili_id, date, .keep_all = T)
-
-# It all looks ok, so we can subset the dataset to exclude the capture dates
-removed_captures <- removed_periods %>%
-  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
-nrow(removed_periods) == nrow(removed_captures) # should be TRUE. NOW we can filter.
-removed_captures <- removed_captures %>%
-  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured
-
-# How many rows and individuals did we remove?
-before <- nrow(removed_periods)
-after <- nrow(removed_captures)
-(propChange <- (before-after)/before) # approx. 1% of data removed.
-length(unique(removed_periods$Nili_id)) == length(unique(removed_captures$Nili_id)) # no change in number of individuals! As expected.
-
-removed_captures <- removed_captures %>%
-  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
-
-write_feather(removed_captures, here::here("data", "dataPrep", "removed_captures.feather"))
-
-# tidy up
-rm(AllCarmelDates)
-rm(AllCarmelDates.1)
-rm(AllCarmelDates.2)
-rm(AllCarmelDates.3)
-rm(AllCarmelDates.all)
-rm(capture_sites)
-rm(crds)
-rm(DistanceMat)
-rm(removed)
-rm(removed_captures)
-rm(roosts0)
-rm(sub.captured)
-rm(sub.captured.carmel)
-rm(sub.captured.dates)
-rm(sub.captured.dates.1)
-rm(sub.captured.no.carmel)
-rm(sub.roosts)
-rm(removed_periods)
-rm(after)
-rm(before)
-rm(ClosestCaptureDist)
-rm(ClosestCaptureSite)
-rm(distance)
-rm(start.day)
-rm(start.month)
-rm(end.day)
-rm(end.month)
-rm(propChange)
-
-gc()
+(nrow(removed_captures)-nrow(data_cleaned))/nrow(data_cleaned) # did not remove any data. Should probably check on this...
+save(removed_captures, file = here("data/dataPrep/removed_captures.Rda"))
 ```
@@
 
-*Inputs:*
-
--   `data/raw/whoswho_vultures_20230920_new.xlsx`
-
--   `data/dataPrep/removed_captures.feather`
-
-*Outputs:*
-
--   `data/dataPrep/with_age_sex.feather`
-
 ```{r attach-age-sex}
 #| message: false
-removed_captures <- read_feather(here::here("data", "dataPrep", "removed_captures.feather"))
-age_sex <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
+load(here("data/dataPrep/removed_captures.Rda"))
+age_sex <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), 
+                      sheet = "all gps tags")[,1:35] %>%
   dplyr::select(Nili_id, birth_year, sex) %>%
   distinct()
-```
 
-```{r}
-#| eval: false
 with_age_sex <- removed_captures %>%
@@
 nrow(with_age_sex) == nrow(removed_captures) # should be TRUE
-length(unique(with_age_sex$Nili_id)) #201
-
-write_feather(with_age_sex, here::here("data", "dataPrep", "with_age_sex.feather"))
+length(unique(with_age_sex$Nili_id)) #200
 
@@
 rm(removed_captures)
-rm(with_age_sex)
 rm(age_sex)
@@
 
-## Clean data
+## Mask data
 
@@
 
-*Inputs:*
+```{r region-masking}
+mask <- sf::st_read(here("data/raw/CutOffRegion.kml"))
 
--   `data/raw/CutOffRegion.kml`
+# XXX TEST
+test <- sample_n(with_age_sex, 50000)
+reMask <- T # XXX INCORPORATE THIS INTO THE FUNCTION!!
+masked_test <- vultureUtils::inMaskFilter(dataset = test, mask = mask, inMaskThreshold = 0.33)
+# XXX END TEST
 
--   `data/dataPrep/with_age_sex.feather`
+# Region masking
+data_masked <- vultureUtils::inMaskFilter(dataset = with_age_sex, mask = mask, inMaskThreshold = 0.33, crs = "WGS84") # this takes a long time!!
+save(data_masked, file = here("data/dataPrep/data_masked.Rda")) # note that this saves a 3-element list. Need to get Ryan to fix this.#XXX
+load(here("data/dataPrep/data_masked.Rda"))
+length(data_masked$Nili_id)
+data_masked_2 <- data_masked$dataset
 
-*Outputs:*
-
--   `data/dataPrep/cleaned.feather`
-
-```{r}
-mask <- sf::st_read(here::here("data", "raw", "CutOffRegion.kml"))
-with_age_sex <- read_feather(here::here("data", "dataPrep", "with_age_sex.feather"))
-```
-
-```{r clean-data}
-#| eval: false
-# Clean the data
-## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
-cleaned <- vultureUtils::cleanData(dataset = with_age_sex, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) 
-length(unique(cleaned$Nili_id)) # 172 individuals
 # Fix time zone so dates make sense ---------------------------------------
 ## Overwrite the dateOnly column from the new times
-cleaned <- cleaned %>%
+data_masked_2 <- data_masked_2 %>%
   mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
@@
          month = lubridate::month(dateOnly),
-         year = lubridate::year(dateOnly)) # XXX consider removing the downsample argument. Also this function is insanely slow and needs to be fixed, good grief.
-
-# This produces a huge file because it's an SF file. We don't really need it to be sf at this stage. Let's remove that.
-cleaned <- sf::st_drop_geometry(cleaned)
-
-save(cleaned, file = here::here("data", "dataPrep", "cleaned.Rda"))
+         year = lubridate::year(dateOnly))
+save(data_masked_2, file = here("data/dataPrep/data_masked_2.Rda")) # just the df #XXX fix
 
@@
 rm(with_age_sex)
-rm(cleaned)
 gc()
@@
 
-*Inputs:*
+```{r split-seasons}
+load(here("data/dataPrep/data_masked_2.Rda")) # XXX fix
 
--   `data/dataPrep/cleaned.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/seasons_list.Rda`
-
--   `data/dataPrep/season_names.Rda`
-
-```{r}
-load(here::here("data", "dataPrep", "cleaned.Rda"))
-```
-
-```{r split-seasons}
-#| eval: false
 # Split into Marta's 3 seasons --------------------------------------------
-cleaned <- cleaned %>%
+split <- data_masked_2 %>%
   mutate(start_breeding = ifelse(month %in% c(1:5),
@@
 
-with_seasons <- cleaned %>%
+with_seasons <- split %>%
   mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
@@
          season = factor(season, levels = c("breeding", "summer", "fall")))
-
-# remove columns that were used in the season calculation
-with_seasons <- with_seasons %>%
-  dplyr::select(-c(month, start_breeding, start_summer, start_fall))
 
@@
 
-*Inputs:*
-
--   `data/dataPrep/seasons_list.Rda`
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/removed_northern.Rda`
-
 ```{r}
-load(here::here("data", "dataPrep", "seasons_list.Rda"))
-load(here::here("data", "dataPrep", "season_names.Rda"))
+load(here("data/dataPrep/seasons_list.Rda"))
+load(here("data/dataPrep/season_names.Rda"))
 ```
@@
                          %>% filter(Nili_id %in% .y))
-save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))
+save(removed_northern, file = here("data/dataPrep/removed_northern.Rda"))
 
@@
 Because tags can be put on and off and the fix rate can change, I'm calculating the fix rate on a per-season basis (which is why I didn't put this step earlier in the pipeline–it had to come after the season splitting).
-
-*Inputs:*
-
--   `data/dataPrep/removed_northern.Rda`
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/low_fix_rate_indivs.Rda`
 
@@
 ## Remove low fix rate individuals
-
-*Inputs:*
-
--   `data/dataPrep/removed_northern.Rda`
-
--   `data/dataPrep/low_fix_rate_indivs.Rda`
 
@@
 
-*Inputs:*
-
--   `data/dataPrep/downsampled_10min_forSocial.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/roosts.Rda`
-
 ```{r}
@@
 ## Remove nighttime points
-
-*Inputs:*
-
--   `data/dataPrep/removed_lfr.Rda`
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/removed_nightime.Rda`
 
@@
 ## Remove vulture-days with too few points per day
-
-*Inputs:*
-
--   `data/dataPrep/removed_nighttime.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/removed_lowppd.Rda`
 
@@
 
-*Inputs:*
-
--   `data/dataPrep/removed_lowppd.Rda`
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/removed_too_few_days.Rda`
--   `data/dataPrep/season_durations.Rda`
-
 ```{r}
@@
 ## Calculate flight altitude
-
-*Inputs:*
-
--   `data/dataPrep/removed_too_few_days`
-
-*Outputs:*
-
--   `data/dataPrep/elevs_z10.Rda`
-
--   `data/dataPrep/with_altitudes.Rda`
 
@@
 This might mess up our movement calculations, so let's downsample the data so that any points that were sampled at a frequency much higher than 1 fix/10 minutes will be reduced to the same frequency as the others.
-
-*Inputs:*
-
--   `data/dataPrep/with_altitudes.Rda`
-
--   `data/dataPrep/season_names.Rda`
-
-*Outputs:*
-
--   `data/dataPrep/downsampled_10min.Rda`
 

### 2024-03-29 scripts/00.0_dataPrep.qmd
fixed downsampling function and reduced 8 million rows to 5 million

@@
 library(furrr)
-future::plan(future::multisession, workers = 10) # this will last for the whole R session.
+future::plan(future::multisession, workers = 5) # this will last for the whole R session.
+
+# function to downsample data to an arbitrary number of minutes, assuming df contains only one individual and timestamp column is called "timestamp"
+down <- function(df, mins){
+  df <- dplyr::arrange(df, timestamp)
+  howmuch <- paste(mins, "minute", sep = " ")
+  rounded <- lubridate::round_date(df$timestamp, howmuch)
+  keep <- !duplicated(rounded)
+  return(df[keep,])
+}
 ```
@@
 
+## Downsample to 5-minute frequency
+This data is just WAY too huge to handle now that we have the higher-frequency fixes from the summer/fall of 2023. Going to downsample it a bit.
+
+```{r}
+# split `joined` by `Nili_id`
+lst <- joined %>% group_split(Nili_id)
+
+# apply the function to each of the ~200 individuals
+interval_5min <- map(lst, ~down(df = .x, mins = 5), .progress = T)
+interval_5min <- purrr::list_rbind(interval_5min)
+write_feather(interval_5min, here("data/dataPrep/interval_5min.feather"))
+nrow(joined)
+nrow(interval_5min) # okay, that removed a lot!
+```
+
 ## Remove invalid periods
@@
 ```{r load-hospital}
-joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
-periods_to_remove <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
+interval_5min <- read_feather(here::here("data/dataPrep/interval_5min.feather"))
+periods_to_remove <- read_excel(here::here("data/raw/whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
 ```
@@
 ```{r remove-periods}
-removed_periods <- vultureUtils::removeInvalidPeriods(dataset = joined, periodsToRemove = periods_to_remove)
-save(removed_periods, file = here("data/dataPrep/removed_periods.Rda"))
+removed_periods <- vultureUtils::removeInvalidPeriods(dataset = interval_5min, periodsToRemove = periods_to_remove)
+write_feather(removed_periods, here("data/dataPrep/removed_periods.feather"))
 ```
@@
 #| eval: false
-
-# Function to downsample the data
-subsample <- function(df, idCol = "Nili_id", 
-                      timestampCol = "timestamp", mins = 10){
-  sub <- df %>%
-    mutate(d = lubridate::date(.data[[timestampCol]])) %>% 
-    group_by(.data[[idCol]]) %>%
-    arrange(timestampCol, .by_group = T) %>%
-    mutate(tc = cut(.data[[timestampCol]], 
-                    breaks = paste(as.character(mins), 
-                                   "min", sep = " "))) %>%
-    ungroup() %>%
-    group_by(.data[[idCol]], d, tc) %>%
-    slice(1) %>%
-    ungroup() %>%
-    dplyr::select(-c("d", "tc"))
-  return(sub)
-}
-
 downsampled_10min_forSocial <- map(removed_lfr, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)

### 2024-03-29 scripts/04.0_mixedModels.R
updated plots after most recent round of data cleaning revision

@@ -24,12 +24,12 @@ library(gtsummary)
 theme_set(theme_classic())
 
 # Load network metrics and movement variables
-load("data/calcSocial/networkMetrics.Rda")
+load(here("data/calcSocial/networkMetrics.Rda"))
 networkMetrics <- networkMetrics %>% rename("seasonUnique" = "season")
-load("data/new_movement_vars.Rda")
-load("data/dataPrep/season_names.Rda")
-load("data/derived/cc.Rda")
-load("data/akde/sfs_est_centroids.Rda")
+load(here("data/new_movement_vars.Rda"))
+load(here("data/dataPrep/season_names.Rda"))
+load(here("data/derived/cc.Rda"))
+load(here("data/akde/sfs_est_centroids.Rda"))
 centrs <- sfs_est_centroids %>%
   select(seasonUnique, Nili_id, dist_szn_centr) %>%
   group_by(seasonUnique) %>%
@@ -66,7 +66,7 @@ linked <- linked %>%
                           type == "roosting" ~ "Ro"))
 table(linked$situ)
 
-save(linked, file = "data/mixedModels/linked.Rda")
+save(linked, file = here("data/mixedModels/linked.Rda"))
 
 
 # Examine response variable distributions ---------------------------------
@@ -97,18 +97,18 @@ save(linked, file = "data/mixedModels/linked.Rda")
 #   theme(legend.position = "none", text = element_text(size = 20))
 # ggsave(normDegree_years_seasons_feeding, filename = "fig/normDegree_years_seasons_feeding.png", width = 9, height = 7)
 
-normDegree_years_seasons_roosting <- linked %>%
-  filter(type == "roosting") %>%
-  mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
-  ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-  geom_density(linewidth = 1.5)+
-  theme_classic()+
-  scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
-  ylab("Frequency")+
-  xlab("Degree (normalized)")+
-  facet_wrap(~season)+
-  theme(legend.position = "none", text = element_text(size = 20))
-ggsave(normDegree_years_seasons_roosting, filename = "fig/normDegree_years_seasons_roosting.png", width = 9, height = 7)
+# normDegree_years_seasons_roosting <- linked %>%
+#   filter(type == "roosting") %>%
[…]

### 2024-03-29 scripts/00.1_wrapAround.qmd
brought over wrap-around code

---
title: "Wrap Around"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
```

## Load data

```{r}
load
```

```{r}
sms <- seq(from = 1, to = 25, by = 1)

sms_sims_conveyor <- vector(mode = "list", length = length(sms))
for(i in 1:length(sms)){
  shift <- sms[i]
  cat("shiftmax = ", shift, "\n")
  sims_conveyor <- map(sims_xy, ~{
    realizations <- vector(mode = "list", length = n)
    for(j in 1:n){
      cat(".")
      realizations[[j]] <- rotate_data_table(dataset = .x, shiftMax = shift,
                                             idCol = "indiv", dateCol = "date",
                                             timeCol = "time")
    }
    cat("\n")
    return(realizations)
  })
  sms_sims_conveyor[[i]] <- sims_conveyor
}
save(sms_sims_conveyor, file = "data/simulations/sms_sims_conveyor.Rda")
load("data/simulations/sms_sims_conveyor.Rda")
```

## Create wrapped data

### 2024-03-29 scripts/00.0_functions.R

@@ -1,0 +1,53 @@
+# FUNCTION DEFINITIONS
+# To be used in the rest of the workflow for this project.
+library(dplyr)
+library(lubridate)
+library(data.table)
+library(spatsoc)
+library(purrr)
+#devtools::install_github("kaijagahm/vultureUtils")
+library(vultureUtils)
+
+# This function definition comes from the Vulture-Conveyor-Belt project. I've removed all the other functions since this is the only one we need.
+
+# 4. rotate_data_table ------------------------------------------------------------
+# Note: this no longer actually takes a data_table, just keeping the name for consistency with previous.
+# Unlike the previous function, this one requires that you separate the dates and times into separate columns beforehand.
+rotate_data_table <- function(dataset, shiftMax, idCol = "indiv", dateCol = "date", timeCol = "time"){
+  indivList <- dataset %>%
+    group_by(.data[[idCol]]) %>%
+    group_split(.keep = T)
+  joined <- vector(mode = "list", length = length(indivList))
+  for(indiv in 1:length(indivList)){
+    x <- indivList[[indiv]]
+    shift <- sample(-(shiftMax):shiftMax, size = 1)
+    #cat(shift, "\n")
+    # get all unique days that show up
+    days <- sort(unique(x[[dateCol]]))
+    
+    # get min and max dates to shift around (the "poles" of the conveyor)
+    selfMinDate <- min(days, na.rm = T)
+    selfMaxDate <- max(days, na.rm = T)
+    
+    # create a total sequence of dates to select from
+    daysFilled <- seq(lubridate::ymd(selfMinDate), lubridate::ymd(selfMaxDate), by = "day")
+    # converting to numbers so we can use %%--which dates are the ones we started with?
+    vec <- which(daysFilled %in% days)
+    shiftedvec <- vec + shift # shift
+    new <- (shiftedvec - min(vec)) %% (max(vec)-min(vec)+1)+1 # new dates as numbers
+    shiftedDates <- daysFilled[new] # select those dates from the possibilities
+    
+    # Make a data frame to hold the old and new dates
+    daysDF <- bind_cols({{dateCol}} := days, 
+                        "newDate" = shiftedDates,
+                        shift = shift)
+    nw <- left_join(x, daysDF, by = dateCol)
+    
+    if(!is.null(timeCol)){
[…]

### 2024-04-01 scripts/00.1_wrapAround.qmd
wrapping around and calculating graph metrics

@@
 library(here)
+library(future)
+source(here("scripts/00.0_functions.R"))
+roostPolygons <- sf::st_read(here("data/raw/roosts50_kde95_cutOffRegion.kml"))
 ```
@@
 ```{r}
-load
+load(here("data/dataPrep/downsampled_10min_forSocial.Rda"))
 ```
 
+It's completely not computationally tractable to first make the entire dataset with 100 replicates and then calculate networks on it.
+Instead, we're going to have to iterate through, do each one, calculate the networks on each one, calculate the metrics, and save those.
+
 ```{r}
-sms <- seq(from = 1, to = 25, by = 1)
+# for each season (map2 data, season_names)
+  # make list for metrics for each rep
+    # for each rep"
+      # rotate
+      # calc sri on the rotated data
+      # make graph from the sri
+      # calculate graph metrics
+      # save graph metrics to spot in list
+      # delete rotated data, sri, graph, and metrics
+  # Bind the metrics with the season name .y
 
-sms_sims_conveyor <- vector(mode = "list", length = length(sms))
-for(i in 1:length(sms)){
-  shift <- sms[i]
-  cat("shiftmax = ", shift, "\n")
-  sims_conveyor <- map(sims_xy, ~{
-    realizations <- vector(mode = "list", length = n)
-    for(j in 1:n){
-      cat(".")
-      realizations[[j]] <- rotate_data_table(dataset = .x, shiftMax = shift,
-                                             idCol = "indiv", dateCol = "date",
-                                             timeCol = "time")
+test <- downsampled_10min_forSocial[[1]] %>% 
+  mutate(time = substr(as.character(timestamp), 12, 19))
+roosts <- vultureUtils::get_roosts_df(test, id = "Nili_id") %>%
+  sf::st_as_sf(., crs = "WGS84", coords = c("location_long", "location_lat"), remove = F)
+reps <- 100
+shiftMax <- 2
+fn <- function(test, iter){
+  # rotate data
+  message("rotating")
+  rotated <- rotate_data_table(dataset = test, shiftMax = shiftMax, 
+                                   idCol = "Nili_id", dateCol = "dateOnly",
+                                   timeCol = "time") %>% 
+    mutate(rep = iter,
+           timestamp = newdatetime) %>%
+    sf::st_as_sf(., coords = c("location_long", "location_lat"), remove = F, crs = "WGS84")
+
+  tojoin_roosts <- rotated %>% st_drop_geometry() %>% 
+    select(Nili_id, "roost_date" = dateOnly, newDate) %>% distinct()
+  roosts_temp <- roosts %>%
+    left_join(tojoin_roosts, by = c("Nili_id", "roost_date")) %>%
+    mutate(roost_date = newDate)
+
+  # calculate sri for rotated data
+  message("calculating sri")
+  fl <- vultureUtils::getFlightEdges(rotated, roostPolygons = roostPolygons,
+                        distThreshold = 1000, idCol = "Nili_id",
+                        return = "sri")
+  fe <- vultureUtils::getFeedingEdges(rotated, roostPolygons = roostPolygons,
+                        distThreshold = 50, idCol = "Nili_id",
+                        return = "sri")
+  ro <- vultureUtils::getRoostEdges(roosts_temp, mode = "distance",
+                                    distThreshold = 500, latCol = "location_lat",
+                                    longCol = "location_long", idCol = "Nili_id",
+                                    dateCol = "roost_date", return = "sri")
+
+  # make graph from sri
+  message("making graph")
+  g_fl <- makeGraph(mode = "sri", data = fl, weighted = T)
+  g_fe <- makeGraph(mode = "sri", data = fe, weighted = T)
+  g_ro <- makeGraph(mode = "sri", data = ro, weighted = T)
+
+  # calculate graph metrics
+  message("calculating metrics")
+  graph_metrics <- data.frame(degree = degree(g_fl),
+                   strength = strength(g_fl),
+                   Nili_id = names(degree(g_fl)),
+                   situ = "flight") %>%
+    bind_rows(data.frame(degree = degree(g_fe),
+                         strength = strength(g_fe),
+                         Nili_id = names(degree(g_fe)),
+                         situ = "feeding")) %>%
+    bind_rows(data.frame(degree = degree(g_ro),
+                         strength = strength(g_ro),
+                         Nili_id = names(degree(g_ro)),
+                         situ = "roosting"))
+  row.names(graph_metrics) <- NULL
+
+  # save graph metrics to spot in list
+  return(graph_metrics)
 }
-    cat("\n")
-    return(realizations)
-  })
-  sms_sims_conveyor[[i]] <- sims_conveyor
-}
-save(sms_sims_conveyor, file = "data/simulations/sms_sims_conveyor.Rda")
-load("data/simulations/sms_sims_conveyor.Rda")
+
+metrics <- map(1:reps, ~fn(test, iter = .x), .progress = T)
 ```
-
-## Create wrapped data

### 2024-04-01 scripts/00.1_wrapAround.qmd
added vertices to deal with the no-edges problem

@@
   message("making graph")
-  g_fl <- makeGraph(mode = "sri", data = fl, weighted = T)
-  g_fe <- makeGraph(mode = "sri", data = fe, weighted = T)
-  g_ro <- makeGraph(mode = "sri", data = ro, weighted = T)
+  verts <- unique(rotated$Nili_id)
+  g_fl <- makeGraph(mode = "sri", data = fl, weighted = T, vertices = verts)
+  g_fe <- makeGraph(mode = "sri", data = fe, weighted = T, vertices = verts)
+  g_ro <- makeGraph(mode = "sri", data = ro, weighted = T, vertices = verts)
 

### 2024-04-03 scripts/00.1_wrapAround.qmd
completed wrap-around randomizations for vultures

--- 2024-03-29 scripts/00.1_wrapAround.qmd
@@
 ```{r}
-load
+load(here("data/dataPrep/downsampled_10min_forSocial.Rda"))
+```
+
+It's completely not computationally tractable to first make the entire dataset with 100 replicates and then calculate networks on it.
+Instead, we're going to have to iterate through, do each one, calculate the networks on each one, calculate the metrics, and save those.
+
+```{r}
+# prepare the data
+downsampled_10min_forSocial <- map(downsampled_10min_forSocial, ~{
+  .x %>% mutate(time = substr(as.character(timestamp), 12, 19))
+})
+
+roosts <- map(downsampled_10min_forSocial, ~{
+  vultureUtils::get_roosts_df(.x, id = "Nili_id") %>%
+    sf::st_as_sf(., crs = "WGS84", coords = c("location_long", "location_lat"), remove = F)
+})
 ```
@@
 ```{r}
-sms <- seq(from = 1, to = 25, by = 1)
+fn <- function(dataset, rst, iter){
+  # rotate data
+  #message("rotating")
+  rotated <- rotate_data_table(dataset = dataset, shiftMax = shiftMax, 
+                               idCol = "Nili_id", dateCol = "dateOnly",
+                               timeCol = "time") %>% 
+    mutate(timestamp = newdatetime) %>%
+    sf::st_as_sf(., coords = c("location_long", "location_lat"), remove = F, crs = "WGS84")
+  verts <- unique(rotated$Nili_id)
 
-sms_sims_conveyor <- vector(mode = "list", length = length(sms))
-for(i in 1:length(sms)){
-  shift <- sms[i]
-  cat("shiftmax = ", shift, "\n")
-  sims_conveyor <- map(sims_xy, ~{
-    realizations <- vector(mode = "list", length = n)
-    for(j in 1:n){
-      cat(".")
-      realizations[[j]] <- rotate_data_table(dataset = .x, shiftMax = shift,
-                                             idCol = "indiv", dateCol = "date",
-                                             timeCol = "time")
+  tojoin_roosts <- rotated %>% st_drop_geometry() %>% 
+    select(Nili_id, "roost_date" = dateOnly, newDate) %>% distinct()
+  roosts_temp <- rst %>%
+    left_join(tojoin_roosts, by = c("Nili_id", "roost_date")) %>%
+    mutate(roost_date = newDate)
+  vertsr <- unique(roosts_temp$Nili_id)
+
+  # calculate sri for rotated data
+  #message("calculating sri")
+  fl <- vultureUtils::getFlightEdges(rotated, roostPolygons = roostPolygons,
+                                     distThreshold = 1000, idCol = "Nili_id",
+                                     return = "sri") %>% 
+    mutate(sri = replace_na(sri, 0)) %>%
+    filter(sri != 0)
+  fe <- vultureUtils::getFeedingEdges(rotated, roostPolygons = roostPolygons,
+                                      distThreshold = 50, idCol = "Nili_id",
+                                      return = "sri") %>%
+    mutate(sri = replace_na(sri, 0)) %>%
+    filter(sri != 0)
+  ro <- vultureUtils::getRoostEdges(roosts_temp, mode = "distance",
+                                    distThreshold = 100, latCol = "location_lat",
+                                    longCol = "location_long", idCol = "Nili_id",
+                                    dateCol = "roost_date", return = "sri") %>%
+    mutate(sri = replace_na(sri, 0)) %>%
+    filter(sri != 0)
+
+  # make graph from sri
+  #message("making graph")
+  g_fl <- igraph::graph_from_data_frame(d = fl %>% mutate(weight = sri), 
+                                        directed = FALSE, vertices = verts)
+  g_fe <- igraph::graph_from_data_frame(d = fe %>% mutate(weight = sri), 
+                                        directed = FALSE, vertices = verts)
+  g_ro <- igraph::graph_from_data_frame(d = ro %>% mutate(weight = sri), 
+                                        directed = FALSE, vertices = vertsr)
+
+  # calculate graph metrics
+  #message("calculating metrics")
+  graph_metrics <- data.frame(degree = degree(g_fl),
+                              strength = strength(g_fl),
+                              Nili_id = names(degree(g_fl)),
+                              situ = "flight") %>%
+    bind_rows(data.frame(degree = degree(g_fe),
+                         strength = strength(g_fe),
+                         Nili_id = names(degree(g_fe)),
+                         situ = "feeding")) %>%
+    bind_rows(data.frame(degree = degree(g_ro),
+                         strength = strength(g_ro),
+                         Nili_id = names(degree(g_ro)),
+                         situ = "roosting")) %>%
+    mutate(rep = iter)
+  row.names(graph_metrics) <- NULL
+
+  # save graph metrics to spot in list
+  return(graph_metrics)
 }
-    cat("\n")
-    return(realizations)
-  })
-  sms_sims_conveyor[[i]] <- sims_conveyor
-}
-save(sms_sims_conveyor, file = "data/simulations/sms_sims_conveyor.Rda")
-load("data/simulations/sms_sims_conveyor.Rda")
+
+reps <- 100
+shiftMax <- 2
 ```
 
-## Create wrapped data
+Now actually do the thing
+
+```{r}
+future::plan(future::multisession, workers = 15)
+metrics_seasons <- vector(mode = "list", length = length(downsampled_10min_forSocial))
+for(i in 1:length(downsampled_10min_forSocial)){
+  cat("Working on season ", i)
+  dat <- downsampled_10min_forSocial[[i]]
+  rst <- roosts[[i]]
+  metrics_seasons[[i]] <- suppressWarnings(furrr::future_map(1:reps, ~fn(dataset = dat, 
+                                                        rst = rst, iter = .x),
+                              .progress = T))
+}
+
+for(i in 1:length(metrics_seasons)){
+  metrics_seasons[[i]] <- purrr::list_rbind(map2(metrics_seasons[[i]], 1:reps, ~.x %>% mutate(rep = .y))) %>% mutate(seasonUnique = season_names[i])
+}
+metrics_wrapped <- purrr::list_rbind(metrics_seasons)
+metrics_wrapped <- metrics_wrapped %>%
+  group_by(seasonUnique, situ, rep) %>%
+  mutate(n = length(unique(Nili_id)))
+save(metrics_wrapped, file = here("data/metrics_wrapped.Rda"))
+```
+
+```{r}
+load(here("data/metrics_wrapped.Rda"))
+metrics_wrapped <- metrics_wrapped %>%
+  group_by(seasonUnique, situ, rep) %>%
+  mutate(normDegree = degree/(n-1),
+         normStrength = strength/sum(strength, na.rm = T))
+
+load(here("data/calcSocial/networkMetrics.Rda"))
+
+allMetrics <- networkMetrics %>% select(season, "situ" = type, Nili_id, normDegree, normStrength)%>% left_join(metrics_wrapped %>% select("season" = seasonUnique, situ, Nili_id, rep, "wrapped_normDegree" = normDegree, "wrapped_normStrength" = normStrength)) %>%
+  mutate(Nili_id = factor(Nili_id))
+
+# See an example of the deviations for 10 random vultures
+vultures <- sample(allMetrics$Nili_id, size = 15)
+allMetrics %>% filter(season == "2023_summer", Nili_id %in% vultures) %>%
+  ggplot(aes(x = Nili_id, y = wrapped_normDegree))+
+  geom_violin()+
+  geom_point(aes(x = Nili_id, y = normDegree), col = "firebrick2", size = 2)+
+  facet_wrap(~situ, scales = "free")+
+  theme(axis.text.x = element_blank(), 
+        axis.ticks.x = element_blank())
+```
+
+```{r}
+metrics_summary <- allMetrics %>%
+  group_by(season, situ, Nili_id) %>%
+  summarize(normDegree = normDegree[1],
+            normStrength = normStrength[1],
+            diff_deg = normDegree[1]-mean(wrapped_normDegree, na.rm = T),
+            diff_str = normStrength[1]-mean(wrapped_normStrength, na.rm = T),
+            sd_deg = sd(wrapped_normDegree, na.rm = T),
+            sd_str = sd(wrapped_normStrength, na.rm = T)) %>%
+  mutate(z_deg = diff_deg/sd_deg,
+         z_str = diff_str/sd_str)
+
+save(metrics_summary, file = here("data/metrics_summary.Rda"))
+```
+
+Compare z scores to observed degree and strength values
+
+```{r}
+metrics_summary %>%
+  filter(!is.infinite(z_deg)) %>%
+  ggplot(aes(x = normDegree, y = z_deg, col = situ))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  theme_minimal()+
+  facet_wrap(~season, scales = "free")
+
+metrics_summary %>%
+  filter(!is.infinite(z_str)) %>%
+  ggplot(aes(x = normDegree, y = z_str, col = situ))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  theme_minimal()+
+  facet_wrap(~season, scales = "free")
+```
+Interesting that for co-roosting, the relationship between degree and z-score doesn't look the same.
+Maybe just because the co-roosting distributions are so incredibly left-skewed that everything gets thrown off by the few points that have a lower normDegree?
+Anyway, what matters is that we are now getting normal z-scores for roosting, not just 0 and 1.
+
+Let's look at the z-score distributions to see if we'll have to change our modeling method or not.
+```{r}
+metrics_summary %>%
+  ggplot(aes(x = z_deg, group = interaction(season, situ), col = situ))+
+  geom_density() # these look decently normally distributed, so we should be okay modeling normally.
+
+metrics_summary %>%
+  ggplot(aes(x = z_str, group = interaction(season, situ), col = situ))+
+  geom_density() # Likewise, nice! Seems like we may be able to model with a gaussian for this one too, which is nice.
+
+# Those distributions are totally different from modeling the normalized degree in absolute terms:
+metrics_summary %>%
+  ggplot(aes(x = normDegree, group = interaction(season, situ), col = situ))+
+  geom_density()
+
+metrics_summary %>%
+  ggplot(aes(x = normStrength, group = interaction(season, situ), col = situ))+
+  geom_density()
+
+# they're measuring different things.
+Normalized degree directly measures how many others an individual interacts with, while degree z-score models how much an individual deviates from the normalized degree it would have achieved just by change.
+```

### 2024-04-03 docs/manuscript.qmd

@@
 
-For this study, we began with data between September 1, 2020 and September 15, 2023 from the \[ornitela\] and \[inpa\] datasets, which we downloaded from Movebank [@kays2022].
+For this study, we used data between September 1, 2020 and September 15, 2023 from the \[ornitela\] and \[inpa\] datasets, which we downloaded from Movebank [@kays2022].
 For each vulture, we removed any periods of time spent in the wildlife hospital (e.g.
@@
 
-For each focal individual, we characterized a pattern-based spatial phenotype (Webber et al.
-2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
-We used several measures to capture an individual's space use, roosting behavior, and flight movements.
+We characterized the pattern-based spatial phenotype [@webber2023a] of each focal individual by measuring its overall space use, use of roost sites, and movement patterns.
 
-### Space use
-
-We calculated seasonal home ranges and core areas for each focal vulture, using autocorrelated kernel density estimates (AKDEs) via the {ctmm} R package [@fleming2023].
+We calculated seasonal home ranges and core areas using autocorrelated kernel density estimates (AKDEs) via the {ctmm} R package [@fleming2023].
 We defined the home range as the 95% AKDE and the core area as the 50% AKDE.
 To measure differences in how individuals used their home ranges, we calculated the ratio of the seasonal core area size to the seasonal home range size.
-We defined a general "space use" variable as the first principal component of a PCA containing the home range size and the core area/home range ratio.
-The "space use" principal component explained \[PERCENT\] of the variation in the data, with individuals low on the principal component having \[CHARACTERISTICS\] and individuals high on the principal component having \[CHARACTERISTICS\].
-
-### Roost diversification
 
@@
 
-### Movement
-
-We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the {elevatr} package [@hollister2022], at a zoom level of \[level--more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
-Finally, since some tags were occasionally set to a much higher fix rate (e.g.
-every 1 minute) for short periods of time, we downsampled the data to a 10 minute fix rate to make sure that all individuals had comparable fix rates.
-
 We identified in-flight fixes (ground speed \>= 5 m/s) and calculated measures of daily flight behavior, including maximum and median daily flight altitude and daily flight time (the sum of time intervals between adjacent in-flight fixes).
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the {elevatr} package at a ground resolution of approximately 250 meters [@hollister2022] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
 Finally, we used all fixes to calculate daily distance traveled (the sum of all distances between successive fixes); daily maximum displacement (the straight-line distance between the beginning of the daily movement track and the farthest point reached on that day); and movement tortuosity (cumulative distance to the point of maximum displacement divided by the straight-line distance to the point of maximum displacement [@benhamou2004]).
-We took seasonal means of each of these individual daily measures to get a single value per individual vulture, per season.
+We took seasonal means of each of these individual daily measures to get a single value per individual, per season.
 
 ## Summarizing movement measures
+```{r}
+#| include: false
+load(here("data/calcMovement/pca_space.Rda"))
+propvar_space <- as.data.frame(summary(pca_space)$importance)$PC1[2]
 
-To summarize individuals' spatial phenotype, we created three composite indices: **movement**, **space use**, and **roost diversification**.
-Each index was derived from a principal component analysis of several spatial phenotype variables, all of which were scaled and centered over the entire date range of the study before putting them into the PCA.
+load(here("data/calcMovement/pca_roost.Rda"))
+propvar_roost <- as.data.frame(summary(pca_roost)$importance)$PC1[2]
+
+load(here("data/calcMovement/pca_movement.Rda"))
+propvar_movement <- as.data.frame(summary(pca_movement)$importance)$PC1[2]
+```
+
+We characterized individuals' spatial phenotypes using three composite indices: movement, space use, and roost diversification.
+Each index was derived from a principal component analysis of several spatial phenotype variables, each scaled and centered over the entire date range of the study.
 
 **Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity.
+It explained `r 100*round(propvar_movement,2)`% of the variation in the PCA.
+Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall.
 **Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE.
+It explained `r 100*round(propvar_space,2)`% of the variation in the PCA.
+Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e.
+the core area made up a larger proportion of the home range).
 **Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night).
-The composite measures accounted for %, %, and % of the variation in their respective PCAs.
+It explained `r 100*round(propvar_roost,2)`% of the variation in the PCA.
+
+While the griffon vulture population in southern Israel is fairly well mixed and vultures do not have discrete, exclusive territories, there is still variability in where individuals spend their time within the landscape.
+Spatial position has the potential to directly affect how many other individuals a vulture encounters and which feeding and roost sites it frequents.
+In addition, though a high proportion of the vultures in the area are GPS-tagged, more peripheral vultures might be more likely to interact with untagged birds living farther from the population center, potentially diluting their measured measured strength or degree.
+To account for the influence of geographic centrality on social centrality, we calculated the geographic centroid of all GPS points for each individual in each season.
+Then we compared these centroids to the overall centroid of all individuals in the population in that season to obtain an index of centrality (versus peripherality).
+We incorporated spatial centrality as a covariate into our linear models.
 
@@
 
-Using GPS fixes from the tagged vultures, we created proximity-based social networks in three different social situations, following Sharma [-@sharma2023]: co-flight, co-feeding (diurnal ground interactions) and co-roosting (nocturnal ground interactions).
-Networks were created separately for each of the eight seasons.
-For each social situation in each season, we created a weighted, undirected network.
+Using GPS fixes from the tagged vultures, we created weighted, undirected proximity-based social networks in three different social situations for each of the nine seasons, following Sharma [-@sharma2023]: co-flight, co-feeding (diurnal ground interactions) and co-roosting (nocturnal ground interactions).
 Because not all individuals were tagged for exactly the same period of time, and to account for missing data due to low tag battery or poor reception, we calculated the simple ratio index (SRI) to quantify the strength of the association between each pair of vultures in each social situation.
@@
 
-Co-feeding
+### Co-feeding
 
@@
 comm.).
-Co-roosting Roost sites were assigned using [methods].
-Unlike flight and diurnal ground GPS fixes, vultures were only assigned one roosting location per night.
-Two vultures that roosted within the boundaries of the same roost site polygon on a given night were considered to be co-roosting.
+
+### Co-roosting 
+Vultures were considered to be co-roosting if they roosted within 500m of each other on a given night.
+Unlike interactions in the co-feeding and co-flight networks, co-roosting interactions could occur at a maximum frequency of once per night.
+Therefore, edge weights in the co-roosting network represented the number of nights on which two individuals co-roosted divided by the number of nights when both individuals had a valid roost location (i.e.
+they theoretically could have co-roosted), regardless of proximity.
 
@@
 We fit separate models for each response variable (degree and strength).
-The predictors included were movement (PC1), exploration (PC2), situation (feeding, flight, roosting), age group (juvenile/subadult, adult), and season (breeding, summer, fall).
+The predictors included were movement (PC1), exploration (PC2), situation (feeding, flight, roosting), age group (juvenile/subadult, adult), season (breeding, summer, fall), and geographic centrality.
 To account for the different numbers of individuals in the network in each unique season, we included a random effect for the interaction of season and year, e.g.
@@
 ```{r}
-#| eval: false
+#| include: false
+#| output: false
 source(here("scripts/04.0_mixedModels.R"))
@@
 
+## Degree
+
 ```{r}
 #| eval: false
+#| output: false
 #| label: fig-d-space
 #| fig-cap: "Relationship between normalized degree and space use"
-plot_d_eff_space
+# NOT SHOW
+plot_d_eff_space # null--no overall relationship between space use and degree
 ```
@@
 ```{r}
-#| eval: false
-#| label: fig-d-roost
-#| fig-cap: "Relationship between normalized degree and roost diversification"
-plot_d_eff_roost
-```
-
-```{r}
-#| eval: false
 #| label: fig-d-space-situ
@@
 ```
+There was a significant relationship between how vultures used space and how many other individuals they encountered in flight.
+Vultures that consistently ranged widely encountered more others in flight, while those that concentrated their movement in smaller home ranges and smaller proportional core areas flew with fewer others.
+There was a slight, albeit non-statistically-significant, trend toward lower degree in the co-feeding and co-roosting networks for vultures with larger ranges.
 
@@
 #| eval: false
-#| label: fig-s-movement
-#| fig-cap: "Relationship between normalized strength and movement"
-plot_s_eff_movement
+#| output: false
+#| label: fig-d-roost
+#| fig-cap: "Relationship between normalized degree and roost diversification"
+# NOT SHOW
+plot_d_eff_roost # null--no overall relationship between roost diversification and degree
 ```
@@
 #| eval: false
+#| output: false
 #| label: fig-d-movement-situ
 #| fig-cap: "Relationship between normalized degree and movement, with separate trend lines for each social situation"
+# NOT SHOW
 plot_d_eff_situ_movement
-plot_d_emt_situ_movement
+plot_d_emt_situ_movement # going to show the plot by season instead of this one
 ```
+
+```{r}
+#| label: fig-d-movement-situ-season
+#| fig-cap: "Relationship between normalized degree and movement, for each season and social situation."
+plot_d_emt_situ_movement_season
+```
+
+Movement behavior affected degree in all three social situations.
+Vultures with a higher movement score co-roosted with fewer conspecifics only in the transient, pre-breeding season; otherwise movement did not affect their co-roosting degree.
+In the breeding and post-breeding seasons, vultures that moved more encountered more conspecifics in flight and while feeding, but this relationship did not hold in the pre-breeding season.
+
+## Strength
+
+```{r}
+#| eval: false
+#| output: false
+#| label: fig-s-movement
+#| fig-cap: "Relationship between normalized strength and movement"
+plot_s_eff_movement # null
+```
+
+There was no relationship between movement and strength, either overall or in any season or situation.
 
@@
 
+However, individuals with more diversified roosts had lower strength (i.e. they interacted less) overall, considering seasons and situations together.
+
 ```{r}
@@
 ```
+Vultures with greater space use had significantly more co-flight interactions (higher strength in the co-flight network).
+They also tended to have a higher co-roosting strength, though the trend was not statistically significant.
+There was no relationship between seasonal space use and co-feeding strength.
 

### 2024-04-03 scripts/00.0_dataPrep.qmd

@@
 ## Downsample to 5-minute frequency
+
 This data is just WAY too huge to handle now that we have the higher-frequency fixes from the summer/fall of 2023. Going to downsample it a bit.
@@
 removed_periods <- vultureUtils::removeInvalidPeriods(dataset = interval_5min, periodsToRemove = periods_to_remove)
-write_feather(removed_periods, here("data/dataPrep/removed_periods.feather"))
+save(removed_periods, file = here("data/dataPrep/removed_periods.Rda"))
 ```
@@
 # warning: this takes a long time and uses a lot of memory!
-data_cleaned <- vultureUtils::cleanData(dataset = removed_periods,
+cleaned <- vultureUtils::cleanData(dataset = removed_periods,
                      precise = F,
@@
                      report = F)
-rm(joined)
+rm(interval_5min)
 gc()
-save(data_cleaned, file = here("data/dataPrep/data_cleaned.Rda"))
+save(cleaned, file = here("data/dataPrep/cleaned.Rda"))
 
-(nrow(removed_periods)-nrow(data_cleaned))/nrow(removed_periods) # removed less than 1% of the points.
+(nrow(removed_periods)-nrow(cleaned))/nrow(removed_periods) # removed less than 1% of the points.
 ```
@@
 ## Remove time spent in capture cages
+
 ```{r}
@@
 carmel <- read.csv(here("data/raw/all_captures_carmel_2010-2021.csv"))
-removed_captures <- removeCaptures(data = data_cleaned, 
+removed_captures <- removeCaptures(data = cleaned, 
                                        captureSites = capture_sites, 
@@
 
-(nrow(removed_captures)-nrow(data_cleaned))/nrow(data_cleaned) # did not remove any data. Should probably check on this...
+(nrow(removed_captures)-nrow(cleaned))/nrow(cleaned) # did not remove any data. Should probably check on this...
 save(removed_captures, file = here("data/dataPrep/removed_captures.Rda"))
@@
 length(unique(with_age_sex$Nili_id)) #200
+
+mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
+         sex = case_when(sex == "f/m" ~ "u",
+                         TRUE ~ sex)) %>%
 
@@
 # Region masking
-data_masked <- vultureUtils::inMaskFilter(dataset = with_age_sex, mask = mask, inMaskThreshold = 0.33, crs = "WGS84") # this takes a long time!!
-save(data_masked, file = here("data/dataPrep/data_masked.Rda")) # note that this saves a 3-element list. Need to get Ryan to fix this.#XXX
-load(here("data/dataPrep/data_masked.Rda"))
-length(data_masked$Nili_id)
-data_masked_2 <- data_masked$dataset
+data_masked <- vultureUtils::inMaskFilter(dataset = with_age_sex, mask = mask, inMaskThreshold = 0.33, crs = "WGS84")$dataset # this takes a long time!!
 
@@
 ## Overwrite the dateOnly column from the new times
-data_masked_2 <- data_masked_2 %>%
+data_masked <- data_masked %>%
   mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
@@
          year = lubridate::year(dateOnly))
-save(data_masked_2, file = here("data/dataPrep/data_masked_2.Rda")) # just the df #XXX fix
+save(data_masked, file = here("data/dataPrep/data_masked.Rda")) # just the df #XXX fix
 
@@
 ```{r split-seasons}
-load(here("data/dataPrep/data_masked_2.Rda")) # XXX fix
+load(here("data/dataPrep/data_masked.Rda")) # XXX fix
+data_masked <- sf::st_drop_geometry(data_masked) # to make it faster to work on
 
 # Split into Marta's 3 seasons --------------------------------------------
-split <- data_masked_2 %>%
-  mutate(start_breeding = ifelse(month %in% c(1:5),
-                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
-                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
-         start_summer = ifelse(month == 12,
-                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
-                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
-         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
-         start_breeding = as.Date(start_breeding),
-         start_summer = as.Date(start_summer),
-         season = case_when(
-           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
-           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
-           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
-  filter(dateOnly != "2023-09-15") # remove September 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
+with_seasons <- data_masked %>% mutate(month = lubridate::month(timestampIsrael),
+                day = lubridate::day(timestampIsrael),
+                year = lubridate::year(timestampIsrael)) %>%
+  mutate(season = case_when(((month == 12 & day >= 15) | 
+                              (month %in% 1:4) | 
+                              (month == 5 & day < 15)) ~ "breeding",
+                            ((month == 5 & day >= 15) | 
+                               (month %in% 6:8) | 
+                               (month == 9 & day < 15)) ~ "summer",
+                            .default = "fall")) %>%
+  filter(dateOnly != "2023-09-15") %>% # remove September 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
+  mutate(seasonUnique = case_when(season == "breeding" & month == 12 ~
+                                    paste(as.character(year + 1), season, sep = "_"),
+                                  .default = paste(as.character(year), season, sep = "_"))) %>%
+  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding", "2023_summer")),
+         season = factor(season, levels = c("breeding", "summer", "fall")))
 
-with_seasons <- split %>%
-  mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
-         sex = case_when(sex == "f/m" ~ "u",
-                         TRUE ~ sex)) %>%
+# Add ages (based on season, which is why this goes here rather than above)
+with_seasons <- with_seasons %>%
+  mutate(age = year - birth_year) %>%
   mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
@@
                                age >= 5 ~ "adult",
-                               TRUE ~ NA)) %>%
-  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
-                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
-                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
-  # explicitly set the factor levels in temporal order
-  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding", "2023_summer")),
-         season = factor(season, levels = c("breeding", "summer", "fall")))
+                               .default = NA)) %>%
+  ungroup()
 
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_northern.Rda"))
-load(here::here("data", "dataPrep", "season_names.Rda"))
+load(here::here("data/dataPrep/removed_northern.Rda"))
+load(here::here("data/dataPrep/season_names.Rda"))
 ```
@@
 ```{r}
-removed_lfr <- map2(removed_northern, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
+removed_lfr <- map2(removed_northern, low_fix_rate_indivs,                     
+                    ~.x %>% filter(!(Nili_id %in% .y))) 
 save(removed_lfr, file = here::here("data", "dataPrep", "removed_lfr.Rda"))  
@@
 rm(low_fix_rate_indivs) 
-rm(removed_lfr) 
 gc()
@@
 
-## Remove high-frequency fixes
+## Remove high-frequency fixes (downsample for social)
 
@@
 ```{r}
-#| eval: false
-downsampled_10min_forSocial <- map(removed_lfr, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
+lst <- map(removed_lfr, ~.x %>% group_split(Nili_id))
+downsampled_10min_forSocial <- map(lst, ~{
+  purrr::list_rbind(map(.x, ~down(df = .x, mins = 10), .progress = T))
+}, .progress = T)
 
@@
                        mutate(daysTracked = length(unique(dateOnly))))
-save(downsampled_10min_forSocial, file = here::here("data", "dataPrep", "downsampled_10min_forSocial.Rda"))
-
-map_dbl(removed_lfr, nrow) %>% setNames(season_names)
-map_dbl(downsampled_10min_forSocial, nrow) %>% setNames(season_names)
+save(downsampled_10min_forSocial, file = here::here("data/dataPrep/downsampled_10min_forSocial.Rda"))
 
@@
 rm(removed_lfr)
-rm(downsampled_10min_forSocial)
 gc()
@@
 ```{r}
-load(here::here("data", "dataPrep", "downsampled_10min_forSocial.Rda"))
+load(here::here("data/dataPrep/downsampled_10min_forSocial.Rda"))
 ```
@@
 rm(downsampled_10min_forSocial)
-rm(roosts)
 gc()
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_lfr.Rda"))
-load(here::here("data", "dataPrep", "season_names.Rda"))
+load(here::here("data/dataPrep/removed_lfr.Rda"))
+load(here::here("data/dataPrep/season_names.Rda"))
 ```
@@
 
-save(removed_nighttime, file = here::here("data", "dataPrep", "removed_nighttime.Rda"))
+save(removed_nighttime, file = here::here("data/dataPrep/removed_nighttime.Rda"))
 
 # tidy up
-rm(removed_nighttime)
 rm(removed_lfr)
@@
 ```{r}
-load(here::here("data", "dataPrep", "removed_nighttime.Rda"))
+load(here::here("data/dataPrep/removed_nighttime.Rda"))
 ```
@@
 # Calculate number of points per day for each individual
+# XXX START HERE can't find battery charge percent
 ppd <- map_dfr(removed_nighttime, ~.x %>%
@@
                  summarize(n = n(),
-                           minBatt = min(battery_charge_percent, na.rm = T)), .progress = T)
+                           minBatt = min(battery_charge_percent, na.rm = T)) %>% 
+                 ungroup(), .progress = T)
 
@@
 # tidying up
+rm(removed_nighttime)
 rm(removed_northern)
@@
 rm(excl_pts)
-rm(removed_lowppd)
 gc()
@@
 ```{r}
-#| eval: false
-
-# defined the "subsample" function previously above
-
-# create and save downsampled datasets for later use
-downsampled_10min <- map(with_altitudes, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling.
+lst_2 <- map(with_altitudes, ~.x %>% group_split(Nili_id))
+downsampled_10min <- map(lst_2, ~{
+  purrr::list_rbind(map(.x, ~down(df = .x, mins = 10), .progress = T))
+}, .progress = T)
 
@@
                        mutate(daysTracked = length(unique(dateOnly))))
-
-save(downsampled_10min, file = here::here("data", "dataPrep", "downsampled_10min.Rda"))
-
-map_dbl(with_altitudes, nrow) %>% setNames(season_names)
-map_dbl(downsampled_10min, nrow) %>% setNames(season_names)
+save(downsampled_10min, file = here::here("data/dataPrep/downsampled_10min.Rda"))
 

### 2024-04-03 scripts/01.2_categorizeMovement.qmd

@@
   mutate(movement_pc1 = pca_movement$x[,1]*-1)
+save(pca_movement, file = here("data/calcMovement/pca_movement.Rda"))
 ```
@@
   mutate(roost_pc1 = pca_roost$x[,1]*-1)
+save(pca_roost, file = here("data/calcMovement/pca_roost.Rda"))
 ```
@@
 fviz_screeplot(pca_space, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
-contrib_space <- round(pca_roost$rotation[,1:3]*-100, 2) %>%
+contrib_space <- round(pca_space$rotation[,1:3]*-100, 2) %>%
   as.data.frame()
 contrib_space
+# Multiply by -1 to flip these around, because right now `pca_space` shows these both being negatively loaded.
 space_use <- space_use %>%
   mutate(space_pc1 = pca_space$x[,1]*-1)
+save(pca_space, file = here("data/calcMovement/pca_space.Rda"))
+
+# Plots of the extremes
+high <- space_use %>% filter(space_pc1 == max(space_pc1, na.rm = T))
+low <- space_use %>% filter(space_pc1 == min(space_pc1, na.rm = T))
+
 ```

### 2024-04-04 scripts/02.0_calcSocial.R
Sticking to for loops for this

@@ -26,11 +26,6 @@ load("data/derived/cc.Rda")
 
 # Social Networks ---------------------------------------------------------
 flight <- vector(mode = "list", length = length(sfdata))
-future::plan(future::multisession(), workers = 10)
-flight <- furrr::future_map(sfdata, ~vultureUtils::getFlightEdges(.x, roostPolygons = roostPolygons,
-                                                                  distThreshold = 1000, idCol = "Nili_id",
-                                                                  return = "both", getLocs = T), .progress = T)
-
 for(i in 1:length(sfdata)){
   cat("Working on iteration", i, "\n")
   dat <- sfdata[[i]]

### 2024-04-04 scripts/01.2_categorizeMovement.qmd
re-ran with new pc's

@@
 fviz_screeplot(pca_space, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
-contrib_space <- round(pca_space$rotation[,1:3]*-100, 2) %>%
+contrib_space <- round(pca_space$rotation[,1:2]*-100, 2) %>%
   as.data.frame()
@@
 save(pca_space, file = here("data/calcMovement/pca_space.Rda"))
-
-# Plots of the extremes
-high <- space_use %>% filter(space_pc1 == max(space_pc1, na.rm = T))
-low <- space_use %>% filter(space_pc1 == min(space_pc1, na.rm = T))
-
 ```

### 2024-04-05 docs/manuscript.qmd
updated with a brief description of the degree and strength z-scores

@@
 ## Summarizing movement measures
+
 ```{r}
@@
 ### Co-roosting
+
 Vultures were considered to be co-roosting if they roosted within 500m of each other on a given night.
@@
 Strength measures how much an individual interacts, and is calculated as the sum of the weights of all edges that the focal vulture is involved in.
+We normalized degree by dividing it by the number of individuals present in the network in the current season minus one.
+We normalized strength by dividing it by the sum of all strength values in the network in the current season \[REF\].
+
+## Comparison to a reference model
+
+Since social interactions can arise from both spatial proximity and social preference (i.e.
+"passive" versus "active" interactions [@smith-aguilar2016]).
+We can use spatially explicit data-stream randomization to compare individuals' observed interaction rates to those expected to arise from chance encounters by individuals sharing space without synchronizing their movements [@gahm; @spiegel2016].
+This allows us to decouple social and spatial drivers of social interactions.
+It also helps us account for how individuals' locations in space (e.g.
+the locations of their home ranges) affect how many social interactions they end up having.
+
+Using the "wrap-around" method outlined in [@gahm], we conducted 100 iterations of the wrap-around randomization for each season of data separately, using a maximum time-shift of 1 day either backward or forward.
+We calculated normalized degree and strength for each individual in each season in each of the 100 replicate randomizations.
+Then, we compared each individual's observed normalized degree and strength to the distribution of 100 normalized degree and strength values obtained from the randomization replicates and calculated a z-score for the observed value to measure the extent to which it deviated from random.
+These z-scores were used as the response variables in our mixed models.
 
@@
 ```
+
 There was a significant relationship between how vultures used space and how many other individuals they encountered in flight.
@@
 ```
+
 Vultures with greater space use had significantly more co-flight interactions (higher strength in the co-flight network).

### 2024-04-06 scripts/00.1_wrapAround.qmd
re-ran models and organized scripts

@@
 reps <- 100
-shiftMax <- 2
+shiftMax <- 1
 ```
@@
 # See an example of the deviations for 10 random vultures
+set.seed(3)
 vultures <- sample(allMetrics$Nili_id, size = 15)
@@
 ```
+
 Interesting that for co-roosting, the relationship between degree and z-score doesn't look the same.
@@
 Let's look at the z-score distributions to see if we'll have to change our modeling method or not.
+
 ```{r}

### 2024-04-06 scripts/01.1_calcMovement.R

@@ -164,18 +164,6 @@ dfddf <- purrr::list_rbind(dfdSumm, names_to = "season") %>%
   mutate(meanDFD = meanDFD/60/60)
 distviz(dfddf, "meanDFD", "season")
 
-## Time spent flying in summer/fall/breeding
-dfddf <- dfddf %>%
-  mutate(szn = as.factor(str_extract(season, "summer|breeding|fall")),
-         year = str_extract(season, "[0-9]{4}"))
-
-# Do the seasons differ in mean daily flight duration?
-dfddf %>%
-  ggplot(aes(x = meanDFD, col = szn, group = factor(season)))+
-  geom_density(linewidth = 1.5, alpha = 0.7)+
-  theme_classic()+
-  scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor))
-
 ### Daily Distance Traveled/Mean Displacement -------------------------------
 
 # set up parallel backend with 4 sessions (change this to match your machine)

### 2024-04-06 scripts/04.0_mixedModels.R

@@ -24,8 +24,8 @@ library(gtsummary)
 theme_set(theme_classic())
 
 # Load network metrics and movement variables
-load(here("data/calcSocial/networkMetrics.Rda"))
-networkMetrics <- networkMetrics %>% rename("seasonUnique" = "season")
+load(here("data/metrics_summary.Rda"))
+metrics_summary <- metrics_summary %>% rename("seasonUnique" = "season")
 load(here("data/new_movement_vars.Rda"))
 load(here("data/dataPrep/season_names.Rda"))
 load(here("data/derived/cc.Rda"))
@@ -44,7 +44,7 @@ centrs <- sfs_est_centroids %>%
 
 # Join the two datasets (there will be 3 rows per individual per season, for the 3 different situations)
 linked <- new_movement_vars %>%
-  left_join(networkMetrics, by = c("Nili_id", "seasonUnique")) %>%
+  left_join(metrics_summary, by = c("Nili_id", "seasonUnique")) %>%
   left_join(centrs)
 nrow(linked) == 3*nrow(new_movement_vars)
 
@@ -57,10 +57,11 @@ linked <- linked %>%
          season = factor(season, levels = c("breeding", "summer", "fall")))
 
 # Check for NA's (we shouldn't have any because of the direction of the join)
-colSums(is.na(linked)) # awesome
+colSums(is.na(linked)) # why do we have missing data for some of these? That's worrying.
 
 # Create a shorter version of "type" for easier interpretability
 linked <- linked %>%
+  rename("type" = situ) %>%
   mutate(situ = case_when(type == "flight" ~ "Fl", 
                           type == "feeding" ~ "Fe",
                           type == "roosting" ~ "Ro"))
@@ -70,49 +71,49 @@ save(linked, file = here("data/mixedModels/linked.Rda"))
 
 
 # Examine response variable distributions ---------------------------------
-# normDegree_years_seasons_flight <- linked %>%
-#   filter(type == "flight") %>%
-#   mutate(season = factor(season, levels = c("fall", "breeding", "summer"))) %>%
-#   ggplot(aes(x = normDegree, col = season, group = seasonUnique))+
-#   geom_density(linewidth = 1.5)+
-#   theme_classic()+
-#   scale_color_manual(name = "Season", values = c(cc$fallColor, cc$breedingColor, cc$summerColor))+
-#   ylab("Frequency")+
-#   xlab("Degree (normalized)")+
-#   facet_wrap(~season)+
[…]

### 2024-04-06 R/functions.R
initial targets commit

@@ -1,0 +1,1035 @@
+# Themes ------------------------------------------------------------------
+theme_quals <- function(){ 
+  font <- "Calibri"   #assign font family up front
+  
+  theme_classic() %+replace%    #replace elements we want to change
+    theme(text = element_text(size = 18, family = "Calibri"))
+}
+
+theme_abs_2023 <- function(){
+  theme_classic() %+replace%
+    theme(panel.background = element_rect(fill = "#FFFCF6"),
+          plot.background = element_rect(fill = "#FFFCF6"),
+          text = element_text(color = "#7A695A", size = 18),
+          axis.text = element_text(color = "#7A695A"),
+          axis.ticks = element_line(color = "#7A695A"),
+          axis.line = element_line(color = "#7A695A"))
+}
+
+# Data prep ---------------------------------------------------------------
+get_loginObject <- function(pw){
+  load(pw)
+  loginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
+  rm(pw)
+  return(loginObject)
+}
+get_inpa <- function(loginObject){
+  inpa <- move::getMovebankData(study = 6071688, 
+                                login = loginObject, 
+                                removeDuplicatedTimestamps = TRUE,
+                                timestamp_start = "2020090100000",
+                                timestamp_end = "2023091500000")
+  inpa <- methods::as(inpa, "data.frame")
+  inpa <- inpa %>%
+    mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
+           year = as.numeric(lubridate::year(timestamp)))
+  return(inpa)
+}
+
+get_ornitela <- function(loginObject){
+  minDate <- "2020-09-01 00:00"
+  maxDate <- "2023-09-15 11:59"
+  ornitela <- vultureUtils::downloadVultures(loginObject = loginObject, 
+                                             removeDup = T, dfConvert = T, 
+                                             quiet = T, 
+                                             dateTimeStartUTC = minDate, 
+                                             dateTimeEndUTC = maxDate)
+  return(ornitela)
+}
+
+join_inpa_ornitela <- function(inpa, ornitela){
+  # Add the dataset names so we can keep track of where the data comes from
+  inpa_tojoin <- inpa[,names(ornitela)] %>%
+    mutate(dataset = "inpa")
+  ornitela <- ornitela %>%
+    mutate(dataset = "ornitela")
+  
+  # Join the two datasets
[…]

### 2024-04-06 scripts/dataPrep_report.R
removed obsolete

@@ -1,104 +1,0 @@
-inpa <- read_feather(here("data/dataPrep/inpa.feather"))
-inpa_nrow <- nrow(inpa)
-inpa_vultures <- unique(inpa$local_identifier)
-inpa_nvultures <- length(inpa_vultures)
-rm(inpa)
-
-ornitela <- read_feather(here("data/dataPrep/ornitela.feather"))
-ornitela_nrow <- nrow(ornitela)
-ornitela_vultures <- unique(ornitela$local_identifier)
-ornitela_nvultures <- length(ornitela_vultures)
-rm(ornitela)
-
-joined <- read_feather(here("data/dataPrep/joined.feather"))
-joined_nrow <- nrow(joined)
-joined_vultures <- unique(joined$Nili_id)
-joined_nvultures <- length(joined_vultures)
-rm(joined)
-
-load(here("data/dataPrep/removed_periods.Rda"))
-rp_nrow <- nrow(removed_periods)
-rp_vultures <- unique(removed_periods$Nili_id)
-rp_nvultures <- length(rp_vultures)
-rm(removed_periods)
-
-load(here("data/dataPrep/cleaned.Rda"))
-c_nrow <- nrow(cleaned)
-c_vultures <- unique(cleaned$Nili_id)
-c_nvultures <- length(c_vultures)
-rm(cleaned)
-
-load(here("data/dataPrep/removed_captures.Rda"))
-rc_nrow <- nrow(removed_captures)
-rc_vultures <- unique(removed_captures$Nili_id)
-rc_nvultures <- length(rc_vultures)
-rm(removed_captures)
-
-load(here("data/dataPrep/data_masked.Rda"))
-masked_nrow <- nrow(data_masked)
-masked_vultures <- unique(data_masked$Nili_id)
-masked_nvultures <- length(masked_vultures)
-rm(data_masked)
-
-# hereafter, these are per-season lists
-load(here("data/dataPrep/season_names.Rda"))
-load(here("data/dataPrep/seasons_list.Rda"))
-sl_nrow <- map_dbl(seasons_list, nrow)
-sl_vultures <- map(seasons_list, ~unique(.x$Nili_id))
-sl_nvultures <- map_dbl(sl_vultures, length)
-rm(seasons_list)
-
-load(here("data/dataPrep/removed_northern.Rda"))
-rno_nrow <- map_dbl(removed_northern, nrow)
[…]

### 2024-04-06 scripts/testscript.R

@@ -1,2 +1,0 @@
-# This is a test script!
-head(mtcars)

### 2024-04-06 scripts/04.0_mixedModels.R
removed data ingest and old response variable distribution plots

@@ -1,21 +1,13 @@
 library(car) # for p-values in mixed models?
 library(tidyverse)
 library(sf)
-library(factoextra)
-library(ggfortify)
-library(ggpmisc) # for more details on a linear regression plot
 library(lme4) # for mixed-effects models
-library(ggpubr) # for adding p-values to boxplots
-library(rstatix) # for adding p-values to boxplots
 library(easystats) # for modeling and comparisons
 library(performance) # for modeling and comparisons
 library(lmerTest) # for p-values in mixed models?
 library(glmmTMB) # more complicated than lme4; allows for beta distributions
 library(DHARMa) # for testing glmmTMB models
 # https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
-library(sjPlot)
-library(broom.mixed)
-library(jtools) # similar to sjplot, for forest plots
 library(emmeans) # estimated marginal means/trends
 library(ggeffects)
 library(gtsummary)
@@ -23,52 +15,8 @@ library(gtsummary)
 # Set ggplot theme to classic
 theme_set(theme_classic())
 
-# Load network metrics and movement variables
-load(here("data/metrics_summary.Rda"))
-metrics_summary <- metrics_summary %>% rename("seasonUnique" = "season")
-load(here("data/new_movement_vars.Rda"))
-load(here("data/dataPrep/season_names.Rda"))
-load(here("data/derived/cc.Rda"))
-load(here("data/akde/sfs_est_centroids.Rda"))
-centrs <- sfs_est_centroids %>%
-  select(seasonUnique, Nili_id, dist_szn_centr) %>%
-  group_by(seasonUnique) %>%
-  mutate(mn = mean(dist_szn_centr),
-         sd = sd(dist_szn_centr)) %>%
-  mutate(centr = (dist_szn_centr-mn)/sd,
-         centr = -1*centr) %>%
-  sf::st_drop_geometry() %>%
-  select(seasonUnique, Nili_id, centr) %>%
-  ungroup() %>%
-  distinct()
-
-# Join the two datasets (there will be 3 rows per individual per season, for the 3 different situations)
-linked <- new_movement_vars %>%
-  left_join(metrics_summary, by = c("Nili_id", "seasonUnique")) %>%
-  left_join(centrs)
-nrow(linked) == 3*nrow(new_movement_vars)
-
-# Housekeeping
-linked <- linked %>%
[…]

### 2024-04-09 docs/manuscript.qmd
calculated report stats for cleaning steps and began incorporating them into the manuscript

@@
 library(here)
+tar_config_set(store = here::here('_targets'))
 ```
@@
 ```{r}
-#| include: false
-load(here("data/dataPrep/dataPrep_report_stats.Rda"))
-dp <- dataPrep_report_stats
-cl <- which(dp$dataset == "data_cleaned")
-pctremovedcleaned <- round(100*(dp$nrow[[cl-1]] - dp$nrow[[cl]])/(dp$nrow[[cl-1]]), 2)
+tar_load(cleaned_r)
+tar_load(joined0_r)
+pctremovedcleaned <- (joined0_r[1]-cleaned_r[1])/joined0_r[1]
 ```
@@
 ```{r}
-#| include: false
-soc <- which(dp$dataset == "downsampled_10min_forSocial")
-vulturessoc <- unique(unlist(dp$vultures[[soc]]))
-nvulturessoc <- length(vulturessoc)
-nvulturessoc_szns <- dp$nvultures[[soc]]
+tar_load(downsampled_10min_forSocial_r)
+nvultures <- map_dbl(downsampled_10min_forSocial_r, "vultures")
 ```
 
-After this cleaning, the data used for social network analysis contained `r nvulturessoc` unique vultures in total (range: `r min(nvulturessoc_szns)`-`r max(nvulturessoc_szns)`; mean = `r round(mean(nvulturessoc_szns), 2)` vultures per season, sd = `r round(sd(nvulturessoc_szns), 2)` vultures per season).
+After this cleaning, the data used for social network analysis contained between `r min(nvultures)` and `r max(nvultures)` vultures per season (mean = `r mean(nvultures)`; sd = `r sd(nvultures)`).
 
@@
 ```{r}
-#| include: false
-mov <- which(dp$dataset == "downsampled_10min")
-vulturesmov <- unique(unlist(dp$vultures[[mov]]))
-nvulturesmov <- length(vulturesmov)
-nvulturesmov_szns <- dp$nvultures[[mov]]
+tar_load(downsampled_10min_r)
+nvultures <- map_dbl(downsampled_10min_r, "vultures")
 ```

### 2024-04-09 R/functions.R

@@ -1044,3 +1044,10 @@ join_movement_soc <- function(new_movement_vars, metrics_summary, centrs, season
                             type == "roosting" ~ "Ro"))
   return(linked)
 }
+
+report <- function(dataset, id){
+  n_points <- nrow(dataset)
+  n_vultures <- length(unique(dataset[[id]]))
+  vultures <- unique(dataset[[id]])
+  return(c("points" = n_points, "vultures" = n_vultures))
+}

### 2024-04-09 _targets.R

@@ -130,7 +130,16 @@ list(
   tar_target(metrics_summary, get_metrics_summary(allMetrics)),
   # Mixed models
   tar_target(centrs, get_centrs(sfs_est_centroids)),
-  tar_target(linked, join_movement_soc(new_movement_vars, metrics_summary, centrs, season_names))
+  tar_target(linked, join_movement_soc(new_movement_vars, metrics_summary, centrs, season_names)),
+  # Reporting
+  tar_target(joined0_r, report(joined0, "trackId")),
+  tar_target(cleaned_r, report(cleaned, "Nili_id")),
+  tar_target(data_masked_r, report(data_masked, "Nili_id")),
+  tar_target(seasons_list_r, map(seasons_list, ~report(.x, "Nili_id"))),
+  tar_target(removed_northern_r, map(removed_northern, ~report(.x, "Nili_id"))),
+  tar_target(removed_lfr_r, map(removed_lfr, ~report(.x, "Nili_id"))),
+  tar_target(downsampled_10min_forSocial_r, map(downsampled_10min_forSocial, ~report(.x, "Nili_id"))),
+  tar_target(downsampled_10min_r, map(downsampled_10min, ~report(.x, "Nili_id")))
 )
 
 

### 2024-04-09 scripts/04.0_mixedModels.R
updated plots with new labels, and added degree plots to manuscript

@@ -119,7 +119,7 @@ plot_d_eff_movement <- ggplot(d_eff_movement, aes(x, predicted))+
   geom_point(data = linked, aes(x = movement, y = z_deg, col = situ), alpha = 0.5)+
   geom_line(linewidth = 1, col = "black", linetype = 2)+
   scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
-  ylab("Degree (normalized)--z-score")+
+  ylab("Normalized degree (deviation from random)")+
   xlab("Movement")+
   ggtitle("")+theme_classic()+
   theme(text = element_text(size = 16))
@@ -134,7 +134,7 @@ plot_d_eff_space <- ggplot(d_eff_space, aes(x, predicted))+
   geom_point(data = linked, aes(x = space_use, y = z_deg, col = situ), alpha = 0.5)+
   geom_line(linewidth = 1, col = "black", linetype = 2)+
   scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
-  ylab("Degree (normalized)")+
+  ylab("Normalized degree (deviation from random)")+
   xlab("Space use (log-transformed)")+
   ggtitle("")+theme_classic()+
   theme(text = element_text(size = 16))
@@ -150,7 +150,7 @@ plot_d_eff_situ_space <- ggplot(d_eff_situ_space, aes(x, predicted))+
   geom_line(aes(col = group), linewidth = 1)+
   scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
   scale_fill_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
-  ylab("Degree (normalized)")+
+  ylab("Normalized degree (deviation from random)")+
   xlab("Space use (log-transformed)")+
   ggtitle("")+theme_classic()+
   theme(text = element_text(size = 16))
@@ -172,7 +172,7 @@ plot_d_emt_situ_space <- d_emt_situ_space %>%
   geom_hline(aes(yintercept = 0), linetype = 2)+
   scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
   theme(text = element_text(size = 20), legend.position = "none")+
-  ylab("Space use effect")+
[…]

### 2024-04-15 manuscript.qmd
moved .qmd docs to root

@@
 library(here)
+tar_config_set(store = here::here('_targets'))
 ```
+
+#### Social network centrality is known to be important in animals.
 
@@
 
+#### We observe that some individuals are more central than others in a social network. What causes this variation?
+
 An individual's social centrality may be determined by many factors.
 Differences in social position have been attributed to individual phenotypic characteristics, such as age, sex, or personality \[REFS\].
+
+##### Individual traits
+
+-   Multiple spatial behaviors → Social centrality [@albery2021]
+
+-   Exploration → Weak associations with more conspecifics, more movement between flocks. [@aplin2013]
+
+-   Slow-exploring → More stable relationships [@aplin2013]
+
+-   Personality → Social structure [@aplin2013]
+
+-   Stress → Weaker parental bonds, in captivity [@boogert2014]
+
+-   Stress → Foraging with more others, less choosy, more central, in captivity [@boogert2014]
+
+-   Stress → Foraging with more others, less choosy, more central, in the wild [@brandl2019]
+
+-   More exploratory → higher degree, in finches [@moyers2018]
+
+-   Quick explorers → More interactions [@moyers2018]
+
+-   Slower explorers → less central ("Yet, slower explorers overall did not travel shorter distances than faster explorers, indicating that a less central social network position was not merely a consequence of lower activity.") [@snijders2014]
+
+-   Females → higher strength, more preferred partners (only during lactation) [@wey2013]
+
+-   Females → more stable partnerships [@wey2013]
+
+#### Individuals' spatial phenotypes are important determinants of their social centrality
+
 Of course, how an animal uses space is also an important driver of how many conspecifics it interacts with, and how frequently.
@@
 In this light, the "spatial phenotype" may be considered in addition to more conventional phenotypic measures to explain why some individuals are more central than others in their social networks.
+
+-   Definition of "spatial phenotype" [@webber2023]
+
+-   Definition of "spatial behaviour" [@albery2021]
+
+-   Note that explicit spatial context is also important, and we didn't consider it. (We should have! How do we justify not doing that?)
+
+    -   An individual's observed spatial phenotype results from a combination of its internal tendency toward movement and the external environmental conditions through which it moves [(Nathan et al.
+2008)](https://www.zotero.org/google-docs/?IyWvRB)
+
+#### Social interactions occur when individuals encounter each other, in various behavioral situations.
+
+-   Co-occurrence in space and time is necessary for social interaction to occur. Therefore, more movement → more social interaction, in general.
+
+-   However, movement is also an expression of an individual's underlying tendency to move, as well as potentially an emergent consequence of things like dominance rank or social environment.
+
+-   Social centrality arises from a combination of the number of interactions (strength) and unique individuals encountered.
+An individual can achieve high social centrality by being more geographically central in the population, versus by being more versatile.
+
+-   No one has investigated the link between spatial phenotype and social centrality.
+We know that individuals vary in both spatial phenotype and social centrality, but we don't know how the one influences the other.
+
+    -   Furthermore, social networks are not monolithic.
+Because the biological mechanisms behind social network structure are different in different social situations, it stands to reason that spatial phenotype should be related to social centrality in different ways in the different social contexts.
+
+        -   For example, different components of spatial phenotype might be expected to relate differently to networks in different social situations.
+
+        -   Also, we might just see different-direction effects.
+For example, individuals that move a lot might interact more frequently in flight but interact less frequently at feeding stations (either because they spend less time feeding or because they feed at more peripheral locations).
+Therefore, it is important to consider different behavioral situations separately when teasing apart the relationship between the social and spatial phenotypes.
 
@@
 
-Eurasian griffon vultures (*Gyps fulvus*) provide a study system well suited to ascertaining the relationship between individuals' spatial phenotypes and their social centrality.
+#### We set out to uncover whether spatial phenotypes could explain observed variation in social centrality.
 
--   obligate scavengers, large soaring fliers, lots of movement
+-   Hypotheses/predictions from slides
 
--   Unusual GPS tracking coverage of the population–high proportion of individuals tracked, so we can get a population-level sense of movement behavior and social interaction patterns
-
--   Three clear social situations in which they interact, each with a different significance/relevance.
-
-    -   Communal roosts: many individuals together at once, variability in size/density of roosts.
-Used for information exchange \[REFS\] and also important for disease transmission \[or not–cf Elvira Nili\].
-Breeding/nesting also occurs at/near roost sites, on cliffs.
-
-    -   Flight: interactions in flight are important for information transfer [@harel2017], finding locations of thermal uplift [@sassi2023] and feeding sites [@spiegel2013; @cortés-avizanda2014].
-
-    -   Feeding: Communal feeding at carcasses.
-Also sometimes standing around the carcass site.
-Sometimes associate for a while before beginning to feed, e.g.
-waiting for critical mass to be reached (Orr Spiegel, pers.
-comm.) Potentially also important for disease transmission \[but not, cf.
-ref Elvira Nili\].
-
-While some griffon vultures make long-range forays \[REFS\], there is a mostly resident population in a relatively small area in southern Israel \[REF\].
-The population is small and restricted enough that individuals' home ranges are highly overlapping, with most individuals using a large fraction of available space in the region.
-This makes habitat features likely to affect most individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement [@harel2016].
-GPS fixes are taken once every 10 minutes, frequently enough to measure flight movements with good accuracy, and most individuals are tracked for several months to several years.
-Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population [@sharma2023].\
-\
-\[General introduction to the importance of time scale in social network analysis\]
-
--   [@blonder2012; @cantor2012; @hobson2013; @rasmussen2013; @holme2015; @farine2018; @fisher2021; @pasquaretta2021]
-
-Social situations occur over different time scales, and this is an important consideration when studying the causes and consequences of their structure.
-In defining co-roosting, co-flight, and co-feeding interactions based on proximity in each of these behavioral situations, we have created social networks in which edges already conceptually represent interactions over different amounts of time.
-Because individuals have one roosting location per night, edge weights in the co-roosting network are measured in units of days, while pairs of individuals can interact many times in flight or at a carcass on a given day.\
-\
-The biological processes driving the execution of these behaviors and the formation of these networks are likewise happening over different time scales.
-An individual must decide once per day which roost to go to [@harel2016].
-But flight is a continuous process driven by many decisions resulting from individual and social information [@sassi2023], and feeding events are concentrated in time on a given day.
-Feeding events are driven by the temporal rhythm of the appearance of carcasses, which occur naturally or are provisioned every \[NUMBER\] days and persist for hours to a few days (Orr Spiegel, pers.
-comm.).
-Individuals may return to the same carcass on successive days.
-While at a carcass, individuals may fight and interact many times.
-
-# Research Questions
+##### Research Questions
 
@@
 
-Q3: How does the time scale of movement predict social interactions?
+# Methods {data-link="methods"}
 
--   H3A: The relationship between movement and social centrality will be strongest within the time scales at which change occurs in the social network.
+### Study system
 
-    -   P3A: Co-flight networks will change hourly to daily, co-feeding networks will change every few days, and co-roosting networks will change over longer periods of time (weeks to months).
+This work was conducted on a population of GPS-tagged Eurasian griffon vultures in southern Israel, a study system well suited to studying the relationship between movement and sociality.
+Griffon vultures are large obligate scavengers that use soaring flight to locate food resources (carcasses) on the landscape.
+They may travel long distances without expending very much energy.
+Griffon vultures are locally endangered in Israel, where they have declined significantly since \[DATE\], and they face significant anthropogenic threats, most commonly through pesticide and NSAID poisonings, power line collisions, and shootings.
+The population in southern Israel has an unusually high GPS tracking coverage, with a high proportion of individuals (approximately 80-90%) fitted with GOS tags.
+The tags send a position reading ever ten minutes if battery condition is good, or less frequently otherwise.
+\[paper\] says that we only need \[percentage\] of individuals in order to get a good sense of the population-level social structure, and our coverage far exceeds that.
+We are able to get a very good sense of individuals' movement behavior and social interaction patterns.
 
-    -   P3B: Social position in each of the social situations will be best predicted by individuals' movement patterns aggregated over the corresponding time scale.
+Griffon vultures interact in several social situations, each with a different biological relevance.
+In particular, griffons rely on social information to locate food resources.
+Interactions in flight (i.e.
+when individuals fly within each other's perceptual distance) provide an opportunity for local enhancement of carcass resources, or following to a carcass site.
+Therefore, they are important for information transfer [@harel2017], finding locations of thermal uplift [@sassi2023], and feeding sites [@spiegel2013; @cortés-avizanda2014].
+Vultures also obtain social information about food resources at communal roosting sites.
+Roost sites can vary in size and density, with many vultures roosting together at once.
+These roosts also serve as breeding colonies during the breeding season (approximately December through May).
+Finally, vultures interact frequently at or around carcasses while eating.
+The free-living griffon vultures in the southern Israel population are provisioned at several feeding sites, where livestock carcasses are deposited irregularly to supplement the availability of naturally-occurring carcasses on the landscape.
+Vultures may co-feed, associate in close proximity near a carcass, or even fight over the food resource.
 
-# Methods
+The southern Israel griffon vulture population is mostly resident, though some individuals (especially juveniles) occasionally make long-range forays \[REF\].
+The population is small and restricted enough that individuals' home ranges are highly overlapping, with most individuals using a large fraction of available habitat in the region.
+This makes habitat features likely to affect many individuals similarly, though age has been shown to factor into flight skill and could therefore impact movement [@harel2016].
+GPS fixes are taken frequently enough to monitor flight movements with good accuracy, and most individuals are tracked for several months to several years.
+Individuals have previously been shown to vary in their social network positions in each of the three social situations, with centrality in one situation not necessarily predicting centrality in another and each situation contributing differently to the overall social structure of the population [@sharma2023].
 
-## Study system
+## Field methods
 
-To answer these questions, we analyzed data from a free-flying population of Eurasian griffon vultures (Gyps fulvus) fitted with GPS tags.
+\[Tagging methods\] For each vulture, age was determined using \[methods\]
 
@@
 
-Over the period \[date\] to \[date\], the dataset contained \[number\] of tagged vultures.
-Tags recorded GPS fixes, as well as speed and heading readings, every 10 minutes when battery conditions allowed.
-The GPS data are stored on Movebank [@kays2022] and were loaded into R using the `move` package\[ @kranstauber2023\].
+## Data preparation
 
-## Tagging and aging
+```{r}
+tar_load(cleaned_r)
+tar_load(joined0_r)
+pctremovedcleaned <- (joined0_r[1]-cleaned_r[1])/joined0_r[1]
+```
 
-\[Tagging methods\] For each vulture, age was determined using \[methods\]
+For this study, we used data between September 1, 2020 and September 15, 2023 from the \[ornitela\] and \[inpa\] datasets, which we downloaded from Movebank [@kays2022].
+For each vulture, we removed any periods of time spent in the wildlife hospital (e.g.
+while recovering from an illness, injury, or poisoning), or in captivity prior to wild release (in the case of captive-born birds).
+We cleaned the data by filtering out GPS fixes deemed unreliable (e.g.
+too few satellites for triangulation, nonsensical heading readings, impossible jumps from previous fixes), and we removed any time that vultures spent in capture cages during the tagging season in the fall.
+Altogether, approximately `r pctremovedcleaned`% of fixes were removed as compared to the raw data.
+Next, we removed any vultures that spent fewer than 30 days, in total, within a focal region encompassing the general vicinity of Israel (in order to exclude, for instance, birds that were briefly detected in the area before dispersing far away).
+For vultures detected in the area for at least 30 days during the study period, we subsequently removed any fixes for those individuals that occurred outside of the same focal region, in order to restrict our analyses to movements that occurred near the main population locus and ensure that most social interactions would be with other tagged individuals.
+(We accounted for missing time periods, including removed data, by calculating association indices in the social networks).
 
-## Data cleaning
+We divided the remaining data into seasons based on biological knowledge of vultures' breeding timelines: a breeding season (December 15 through May 14 of the following year); a summer or post-breeding season (May 15th through September 14th), and a transient or pre-breeding season (September 15th through December 14th).
+For simplicity, we refer to each breeding season by the calendar year in which it ends (e.g.
+the 2021 breeding season lasts from December 15, 2020 through May 14, 2021).
+We analyzed data from the 2020 pre-breeding/transient season through the 2023 post-breeding/summer season, for a total of nine seasons across four calendar years (three of each type of season).
 
-We subsetted the dataset to exclude periods of time that vultures spent in capture cages, in the wildlife hospital following poisonings or injuries, or in captivity prior to wild release (in the case of captive-born birds).
-We cleaned the data to remove invalid GPS fixes.
-Then we split data from each into three seasons.
-The fall, or "pre-breeding," season includes dates from 15 September through 14 December; the breeding season lasts from 15 December through 14 May of the following year; and the summer lasts from 15 May through 14 September.
-These season boundaries were chosen based on biological knowledge of the vultures' breeding timelines (see more details in the Study System section).
-Therefore, the analysis contained a total of 8 seasons, from the fall of 2020 through the 2023 breeding season (Table X).
-All of these data was reserved for use in proximity-based social network analysis.
-Since the GPS tags give few to no fixes overnight, and since reception is often very poor on the cliffs where vultures roost, we combined information from individuals' last locations of the evening and first locations of the morning to assign a roost site for each individual on each night \[more details\].
+The vulture population in Israel is geographically separated, with most individuals in the south but a small and relatively isolated population at the Hai Bar Carmel Nature Reserve and in the Golan Heights.
+While vultures occasionally move between the northern and southern population, such movements are rare, and northern vultures behave differently than southern vultures due to more consistent feeding and a more restricted range in the north.
+Therefore, for each season, we restricted our analysis to individuals whose spatial centroid (including all GPS points after data cleaning) lay south of UTM Northing 3550000 (Zone 36N), equivalent to approximately 32.086238 North latitude.
+This threshold was determined by visual examination of a histogram of centroid latitudes, which was bimodal and disjoint.
+Finally, to account for older GPS tags that were set to lower fix frequencies, we excluded any individuals from each season that never had their tags set to a 10-minute frequency during that season.
+In recent months, some tags were temporarily set to record fixes at much higher frequencies, so those individuals were downsampled to a 10-minute frequency in order to make data analysis more tractable.
+
+```{r}
+tar_load(downsampled_10min_forSocial_r)
+nvultures <- map_dbl(downsampled_10min_forSocial_r, "vultures")
+```
+
+After this cleaning, the data used for social network analysis contained between `r min(nvultures)` and `r max(nvultures)` vultures per season (mean = `r mean(nvultures)`; sd = `r sd(nvultures)`).
+
+#### Determining roost locations
+
+We used GPS fixes to determine vultures' nightly roost locations.
+The solar-powered GPS tags are set to record very few fixes overnight, and cell reception (needed for data transmission) is often very poor on the cliffs where roosting sites are located.
+Therefore, we combined information from individuals' last locations of the evening and first locations of the morning to assign a roost site for each individual on each night.
 Full details about the data cleaning and roost assignment procedure can be found in the supplementary material.
 
+#### Further cleaning for movement analysis
+
 To characterize individual movement behavior and spatial phenotype, we cleaned the data further.
-We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package suncalc [@thieurmel2022].
+We restricted movement points to the period between sunrise and sunset (calculated for each day with respect to a point at the centroid of Israel using the R package {suncalc} [@thieurmel2022].
 We restricted movement calculations to a set of focal individuals with activity regions centered in the Negev region of Israel, the site of most of our tagging effort.
@@
 Fixes were ideally collected every 10 minutes, but sometimes they were less frequent due to bad cellular reception in the cliff areas where vultures often sit or due to low battery charge.
-For each individual, only days with 10 or more GPS fixes per day were retained, unless the battery charge never dipped below 50% during that day.
-In those cases, the low number of points was attributed to the vultures spending most of their time sitting on a cliff without much movement, rather than to missed points due to low battery.
+For each individual, only days with 10 or more GPS fixes per day were retained, unless the battery charge remained above 50% for the entire day.
+In those cases, the low number of points was attributed to the vultures spending most of their time sitting on a cliff (with limited reception) without much movement, rather than to actual missing data due to low battery.
 Since these sedentary days represent an important component of vulture movement behavior, we retained this information (Orr Spiegel, pers.
@@
 
-\[Geography\]
+```{r}
+tar_load(downsampled_10min_r)
+nvultures <- map_dbl(downsampled_10min_r, "vultures")
+```
 
-For each season, we calculated movement measures only for individuals tracked for at least 30 days (after removing low-point days as described above).
-We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the `elevatr` R package [@hollister2022], at a zoom level of \[level–more details about resolution?\] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
-Finally, since some tags were occasionally set to a much higher fix rate (e.g.
-every 1 minute) for short periods of time, we downsampled the data to a 10 minute fix rate to make sure that all individuals had comparable fix rates.
+After removing days with too few fixes per day, we removed any vultures that did not have at least 30 days of valid movement data remaining in a given season, to ensure that movement estimates and averages were based on a high enough sample size of days to capture the range of the vulture's behavior.
+Since this was done on a per-season basis, individuals that were removed from one season's analysis due to not having enough valid days could still have movement measures calculated for a different season.
+Finally, we downsampled the movement data to a 10-minute interval to remove high-frequency fixes.
+Ultimately, we calculated movement measures for a total of `r nvulturesmov` **focal individuals** in total (range: `r min(nvulturesmov_szns)`-`r max(nvulturesmov_szns)`; mean = `r round(mean(nvulturesmov_szns), 2)` vultures per season, sd = `r round(sd(nvulturesmov_szns), 2)` vultures per season).
 
-## Movement calculations
+## Measuring spatial phenotype
 
-For each focal individual, we characterized a pattern-based spatial phenotype (Webber et al.
-2023) by calculating a variety of movement and space use measures over the course of each of the eight seasons.
-We calculated home range as the 95% AKDE of a vulture's GPS fixes in a season @
+We characterized the pattern-based spatial phenotype [@webber2023a] of each focal individual by measuring its overall space use, use of roost sites, and movement patterns.
+
+We calculated seasonal home ranges and core areas using autocorrelated kernel density estimates (AKDEs) via the {ctmm} R package [@fleming2023].
+We defined the home range as the 95% AKDE and the core area as the 50% AKDE.
+To measure differences in how individuals used their home ranges, we calculated the ratio of the seasonal core area size to the seasonal home range size.
 
@@
 We identified in-flight fixes (ground speed \>= 5 m/s) and calculated measures of daily flight behavior, including maximum and median daily flight altitude and daily flight time (the sum of time intervals between adjacent in-flight fixes).
+We calculated flight altitude by extracting ground elevation information from elevation rasters obtained from the {elevatr} package at a ground resolution of approximately 250 meters [@hollister2022] and subtracting those numbers from the vultures' tag-mounted altimeter readings.
 Finally, we used all fixes to calculate daily distance traveled (the sum of all distances between successive fixes); daily maximum displacement (the straight-line distance between the beginning of the daily movement track and the farthest point reached on that day); and movement tortuosity (cumulative distance to the point of maximum displacement divided by the straight-line distance to the point of maximum displacement [@benhamou2004]).
-We took seasonal means of each of these individual daily measures to get a single value per individual vulture, per season.
+We took seasonal means of each of these individual daily measures to get a single value per individual, per season.
 
@@
 
-To summarize individuals' spatial phenotype, we created three composite indices: **movement**, **space use**, and **roost diversification**.
-Each index was derived from a principal component analysis of several spatial phenotype variables, all of which were scaled and centered over the entire date range of the study before putting them into the PCA.
+```{r}
+#| include: false
+load(here("data/calcMovement/pca_space.Rda"))
+propvar_space <- as.data.frame(summary(pca_space)$importance)$PC1[2]
+
+load(here("data/calcMovement/pca_roost.Rda"))
+propvar_roost <- as.data.frame(summary(pca_roost)$importance)$PC1[2]
+
+load(here("data/calcMovement/pca_movement.Rda"))
+propvar_movement <- as.data.frame(summary(pca_movement)$importance)$PC1[2]
+```
+
+We characterized individuals' spatial phenotypes using three composite indices: movement, space use, and roost diversification.
+Each index was derived from a principal component analysis of several spatial phenotype variables, each scaled and centered over the entire date range of the study.
 
 **Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity.
+It explained `r 100*round(propvar_movement,2)`% of the variation in the PCA.
+Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall.
 **Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE.
+It explained `r 100*round(propvar_space,2)`% of the variation in the PCA.
+Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e.
+the core area made up a larger proportion of the home range).
 **Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night).
-The composite measures accounted for %, %, and % of the variation in their respective PCAs.
+It explained `r 100*round(propvar_roost,2)`% of the variation in the PCA.
+
+While the griffon vulture population in southern Israel is fairly well mixed and vultures do not have discrete, exclusive territories, there is still variability in where individuals spend their time within the landscape.
+Spatial position has the potential to directly affect how many other individuals a vulture encounters and which feeding and roost sites it frequents.
+In addition, though a high proportion of the vultures in the area are GPS-tagged, more peripheral vultures might be more likely to interact with untagged birds living farther from the population center, potentially diluting their measured measured strength or degree.
+To account for the influence of geographic centrality on social centrality, we calculated the geographic centroid of all GPS points for each individual in each season.
+Then we compared these centroids to the overall centroid of all individuals in the population in that season to obtain an index of centrality (versus peripherality).
+We incorporated spatial centrality as a covariate into our linear models.
 
@@
 comm.).
-Co-roosting Roost sites were assigned using [methods].
-Unlike flight and diurnal ground GPS fixes, vultures were only assigned one roosting location per night.
-Two vultures that roosted within the boundaries of the same roost site polygon on a given night were considered to be co-roosting.
+
+### Co-roosting
+
+Vultures were considered to be co-roosting if they roosted within 500m of each other on a given night.
+Unlike interactions in the co-feeding and co-flight networks, co-roosting interactions could occur at a maximum frequency of once per night.
+Therefore, edge weights in the co-roosting network represented the number of nights on which two individuals co-roosted divided by the number of nights when both individuals had a valid roost location (i.e.
+they theoretically could have co-roosted), regardless of proximity.
 
@@
 
-We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the lme4 and lmerTest packages [@bates2015; @kuznetsova2017].
+We modeled the relationship between the spatial phenotype variables and the network measures using linear mixed-effects regression models, fitted using the {lme4} and {lmerTest} packages [@bates2015; @kuznetsova2017].
 We fit separate models for each response variable (degree and strength).
-The predictors included were movement (PC1), exploration (PC2), situation (feeding, flight, roosting), age group (juvenile/subadult, adult), and season (breeding, summer, fall).
+The predictors included were movement (PC1), exploration (PC2), situation (feeding, flight, roosting), age group (juvenile/subadult, adult), season (breeding, summer, fall), and geographic centrality.
 To account for the different numbers of individuals in the network in each unique season, we included a random effect for the interaction of season and year, e.g.
@@
 #| include: false
+#| output: false
 source(here("scripts/04.0_mixedModels.R"))
@@
 
+## Deviations from random
+
+Individual vultures interacted with significantly more unique others over the course of a season than expected by chance (i.e.
+observed degree values compared to the distribution of degree values obtained from networks created from the wrap-around data), across all seasons and social situations.
+However, individuals' strength did not differ from random.
+
 ```{r}
 #| echo: false
+tar_load(linked)
+
+linked %>%
+  select(seasonUnique, situ, z_deg, z_str) %>%
+  ungroup() %>%
+  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
+  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
+  geom_density(aes(linetype = measure))+
+  facet_wrap(~seasonUnique, scales = "free")+
+  geom_vline(aes(xintercept = 0)) # notably, we see that the z_scores for degree are almost entirely positive, while the ones for strength are basically centered around zero, with very few exceptions.
+```
+
+## Degree
+
+```{r}
 #| label: fig-d-space
-#| fig-cap: "Relationship between normalized degree and space use"
-plot_d_eff_space
+#| fig-cap: "Relationship between movement and degree deviation"
+plot_d_eff_movement
 ```
@@
 ```{r}
-#| echo: false
-#| label: fig-d-roost
-#| fig-cap: "Relationship between normalized degree and roost diversification"
-plot_d_eff_roost
-```
-
-```{r}
-#| echo: false
 #| label: fig-d-space-situ
@@
 
+There was a significant relationship between how vultures used space and how many other individuals they encountered in flight.
+Vultures that consistently ranged widely encountered more others in flight, while those that concentrated their movement in smaller home ranges and smaller proportional core areas flew with fewer others.
+Similarly, there was a marginally significant positive relationship between space use and degree deviation in the co-feeding network as well.
+By contrast, vultures with smaller home ranges/proportionally smaller core areas had a co-roosting degree that deviated more from chance than those with larger home ranges/proportionally larger core areas.
+
 ```{r}
-#| echo: false
-#| label: fig-s-movement
-#| fig-cap: "Relationship between normalized strength and movement"
-plot_s_eff_movement
+#| label: fig-d-roost
+#| fig-cap: "Relationship between normalized degree and roost diversification"
+plot_d_eff_roost 
 ```
 
-```{r}
-#| echo: false
-#| label: fig-d-movement-situ
-#| fig-cap: "Relationship between normalized degree and movement, with separate trend lines for each social situation"
-plot_d_eff_situ_movement
-plot_d_emt_situ_movement
-```
+Vultures with greater roost diversification, had, overall, smaller degree deviations than those with smaller roost diversification, though the effect size was small.
 
-```{r}
-#| echo: false
-#| label: fig-s-roost
-#| fig-cap: "Relationship between normalized strength and roost diversification"
-plot_s_eff_roost
-```
-
-```{r}
-#| echo: false
-#| label: fig-s-space
-#| fig-cap: "Relationship between normalized strength and space use"
-plot_s_eff_situ_space
-plot_s_emt_situ_space
-```
+## Strength
 

### 2024-04-15 presentation.qmd
moved .qmd docs to root

---
title: "Movement/Social"
date: last-modified
format:
  revealjs:
    embed-resources: true
editor: visual
---

```{r}
#| include: false
#| output: false
library(tidyverse)
library(here)
library(targets)
theme_set(theme_minimal())
tar_config_set(store = here::here('_targets'))
source(here("scripts/04.0_mixedModels.R"))
```

<!-- ## Research questions (cont'd) -->

<!-- Q1: What is the relationship between spatial phenotype and social interactions? -->

<!-- -   H1A: Individuals that move more will have more, and stronger, encounters with other unique individuals. -->

<!--     -   P1A: Vultures that move more will encounter more others (higher degree centrality in the social network) and have more interactions (higher strength in the social network) than vultures who move less. -->

<!-- ## Research questions (cont'd) -->

<!-- -   H1B: Individuals that are more exploratory will have more but weaker encounters with other unique individuals. -->

<!--     -   P1B: Exploratory vultures will encounter more unique individuals (higher degree centrality), but will have fewer interactions (lower strength) compared to less exploratory vultures. -->

<!-- ## Research questions (cont'd) -->

<!-- Q2: How is the relationship between spatial phenotype and social interactions impacted by social situation and season? -->

<!-- -   H2A: Social situations in which animals move more will have a greater impact on social interactions -->

<!--     -   P2A: Social centrality in co-flight networks will be more strongly associated with spatial phenotype than in co-roosting and co-feeding behavioral situations. -->

<!-- ## Research questions (cont'd) -->

<!-- -   H2B: Seasons in which animals move more will have a stronger impact on social interactions -->

<!--     -   P2B: The relationship between movement and social position will be stronger in the non-breeding seasons (summer and fall) than during the breeding season. -->

## Examining response variable distributions

```{r}
tar_load(all_movement_vars)
tar_load(cc)
all_movement_vars %>%
  pivot_longer(cols = c(movement, space_use, roost_div), 
               names_to = "var", values_to = "measure") %>%
  mutate(season = str_extract(seasonUnique, pattern = "[a-z]+")) %>%
  ggplot(aes(x = measure, col = season, group = seasonUnique))+
  geom_density()+
  facet_wrap(~var)+
  scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor))+
  theme(legend.position = "bottom",
        axis.title = element_blank())
```

## Wrap-around randomizations (degree)

```{r}
#| fig-cap: "100 randomizations, shift +/- 1 day"
tar_load(allMetrics)
set.seed(3)
vultures <- sample(unique(allMetrics$Nili_id), 3, replace = F)
allMetrics %>%
  filter(Nili_id %in% vultures) %>%
  ggplot(aes(x = Nili_id, y = wrapped_normDegree, col = situ))+
  geom_point(aes(y = normDegree), position = position_dodge(width = 0.4))+
  geom_violin(position = position_dodge(width = 0.4))+
  facet_wrap(~season, scales = "free")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  ylab("Degree (normalized)")+
  theme(axis.title.x = element_blank(),
        legend.position = "bottom")+
  ggtitle("Degree (observed vs. random)")
```

## Wrap-around randomizations (strength)

```{r}
#| fig-cap: "100 randomizations, shift +/- 1 day"
allMetrics %>%
  filter(Nili_id %in% vultures) %>%
  ggplot(aes(x = Nili_id, y = wrapped_normStrength, col = situ))+
  geom_point(aes(y = normStrength), position = position_dodge(width = 0.4))+
  geom_violin(position = position_dodge(width = 0.4))+
  facet_wrap(~season, scales = "free")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  ylab("Strength (normalized)")+
  theme(axis.title.x = element_blank(),
        legend.position = "bottom")+
  ggtitle("Strength (observed vs. random)")
```

## Degree, but not strength, was higher than random.

```{r}
tar_load(linked)

linked %>%
  select(seasonUnique, situ, z_deg, z_str) %>%
  ungroup() %>%
  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
  geom_density()+
  facet_wrap(~measure, scales = "free")+
  geom_vline(aes(xintercept = 0), col = "red")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  theme(axis.title.y = element_blank())+
  xlab("Z-score")

linked %>%
  select(seasonUnique, situ, z_str) %>%
  ungroup() %>%
  ggplot(aes(x = z_str))
```

## Movement vs. degree deviation

```{r}
#| label: fig-d-space
#| fig-cap: "Relationship between movement and degree deviation"
plot_d_eff_movement
```

## Movement measure

```{r}
tar_load(movement_corr)
movement_corr
```

## Space use vs. degree deviation

```{r}
#| label: fig-d-space-situ
#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
plot_d_eff_situ_space
```

## Space use vs. degree deviation

```{r}
plot_d_emt_situ_space
```

## Space use measure

```{r}
tar_load(space_corr)
space_corr
```

## Roost diversification vs. degree deviation

```{r}
#| label: fig-d-roost
#| fig-cap: "Relationship between normalized degree and roost diversification"
plot_d_eff_roost 
```

## Roost diversification measure

```{r}
tar_load(roost_corr)
roost_corr
```

## Degree model summary {.smaller .scrollable}

```{r}
summary(degree_mod)
```

## The strength model did not fit well.

```{r}
simulationOutput <- DHARMa::simulateResiduals(strength_mod)
plot(simulationOutput, pch=".")
```

## Still, here is the summary as it stands {.smaller .scrollable}

```{r}
summary(strength_mod)
```

### 2024-04-15 _targets.R
added correlation plots as targets so we can pull them into the presentation

@@ -107,6 +107,10 @@ list(
   tar_target(movement, get_movement(movementBehaviorScaled)),
   tar_target(roost_behavior, get_roost_behavior(movementBehaviorScaled)),
   tar_target(new_movement_vars, get_new_movement_vars(demo, space_use, movement, roost_behavior)),
+  tar_target(all_movement_vars, get_all_movement_vars(demo, space_use, movement, roost_behavior)),
+  tar_target(space_corr, get_correlation_scatterplots(all_movement_vars, "space_use", ycols = c("homeRange_log", "coreAreaFidelity"), title = "Space use")),
+  tar_target(movement_corr, get_correlation_scatterplots(all_movement_vars, "movement", ycols = c("mnDailyMnAlt", "meanDFD", "meanDMD", "meanDDT", "mnTort"), title = "Movement")),
+  tar_target(roost_corr, get_correlation_scatterplots(all_movement_vars, "roost_div", ycols = c("propSwitch", "shannon", "uniqueRoosts"), title = "Roost diversification")),
   # Calculate social networks
   tar_target(sfdata, convertsf(downsampled_10min_forSocial)),
   tar_target(flight, get_flight(sfdata, roostPolygons)),
@@ -126,7 +130,6 @@ list(
   tar_target(with_times, prepare_times(downsampled_10min_forSocial)),
   tar_target(metrics_wrapped, get_metrics_wrapped(with_times, roosts, season_names, roostPolygons, 100, 1)),
   tar_target(allMetrics, combine_metrics(networkMetrics, metrics_wrapped)),
-  #tar_target(sample_deviations_plot, get_deviations_plot(allMetrics)),
   tar_target(metrics_summary, get_metrics_summary(allMetrics)),
   # Mixed models
   tar_target(centrs, get_centrs(sfs_est_centroids)),

### 2024-04-15 scripts/04.0_mixedModels.R
removed centrality portions

@@ -58,7 +58,7 @@ linked %>%
 # Degree ------------------------------------------------------------------
 tar_load(allMetrics)
 vultures <- sample(unique(allMetrics$Nili_id), 10)
-allMetrics %>% filter(Nili_id %in% vultures) %>% ggplot(aes(x = Nili_id, y = wrapped_normDegree, group = interaction(situ, Nili_id), col = situ))+geom_violin()+facet_wrap(~season, scales = "free")+geom_point(aes(x = Nili_id, y = normDegree)) # individuals generally *do* have degrees higher than expected by chance.
+allMetrics %>% filter(Nili_id %in% vultures) %>% ggplot(aes(x = Nili_id, y = wrapped_degree, group = interaction(situ, Nili_id), col = situ))+geom_violin()+facet_wrap(~season, scales = "free")+geom_point(aes(x = Nili_id, y = degree)) # individuals generally *do* have degrees higher than expected by chance.
 
 degree_base <- glmmTMB(z_deg ~ situ + movement + roost_div + space_use + age_group + season + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian()) 
 check_predictions(degree_base)
@@ -86,16 +86,16 @@ summary(degree_3) # season*space, movement*season, and situ*movement are all non
 check_collinearity(degree_3) # the ones with the highest collinearity have some significant effects. Maybe we can reduce the collinearity by removing non-significant ones instead? Let's start by removing season*space
 
 degree_4 <- glmmTMB(z_deg ~ situ*movement + movement*season + situ*roost_div + situ*space_use + age_group + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian())
-summary(degree_4) # 
+summary(degree_4) #
 check_collinearity(degree_4) # situ*space has high collinearity but also a significant effect. Maybe we could remove situ*movement instead...
 
 degree_5 <- glmmTMB(z_deg ~ movement*season + situ*roost_div + situ*space_use + age_group + (1|seasonUnique)+(1|Nili_id), data = linked, family = gaussian())
-summary(degree_5) # 
+summary(degree_5) #
 check_collinearity(degree_5) # it's not going to help, but we can remove movement*season last...
 
[…]

### 2024-04-15 scripts/01.2_categorizeMovement.qmd
added age group; loaded targets

@@
 library(ade4)
-load(here::here("data", "calcMovement", "movementBehavior.Rda"))
-load(here::here("data", "dataPrep", "season_names.Rda"))
-load(here::here("data", "derived", "cc.Rda"))
-
+library(here)
+tar_config_set(store = here::here('_targets'))
+tar_load(movementBehavior)
+tar_load(season_names)
+tar_load(cc)
+tar_load(downsampled_10min)
 # retrieve age information, since it wasn't carried through in the movement analysis (oops)
-load(here::here("data", "dataPrep", "downsampled_10min.Rda"))
 ages <- map(downsampled_10min, ~.x %>%
@@
   purrr::list_rbind()
-rm(downsampled_10min) # XXX need to do this in a faster way--attach age at an earlier stage probably
 
@@
          homeRange_log = log(homeRange)) %>%
-  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex), 
+  mutate(across(-c(Nili_id, seasonUnique, birth_year, sex, age_group), 
                 function(x){as.numeric(as.vector(scale(x)))}))

### 2024-04-15 R/functions.R
distinct() on movement measures to allow left joins to work

@@ -676,7 +676,7 @@ scale_movement_behavior <- function(movementBehavior){
 get_demo <- function(movementBehaviorScaled){
   demo <- movementBehaviorScaled %>%
     dplyr::select(Nili_id, seasonUnique, daysTracked, propDaysTracked, birth_year, age_group, sex)
-  return(demo)
+  return(distinct(demo))
 }
 
 get_space_use <- function(movementBehaviorScaled){
@@ -685,7 +685,7 @@ get_space_use <- function(movementBehaviorScaled){
   pca <- prcomp(space_use[,-c(1:3)])
   space_use <- space_use %>%
     mutate(space_pc1 = pca$x[,1]*-1)
-  return(space_use)
+  return(distinct(space_use))
 }
 
 get_movement <- function(movementBehaviorScaled){
@@ -694,7 +694,7 @@ get_movement <- function(movementBehaviorScaled){
     filter(!is.na(mnDailyMnAlt))
   pca <- prcomp(movement[,-c(1:2)])
   movement <- movement %>% mutate(movement_pc1 = pca$x[,1]*-1)
-  return(movement)
+  return(distinct(movement))
 }
 
 get_roost_behavior <- function(movementBehaviorScaled){
@@ -704,7 +704,7 @@ get_roost_behavior <- function(movementBehaviorScaled){
   pca <- prcomp(roost_behavior[,-c(1:2)])
   roost_behavior <- roost_behavior %>%
     mutate(roost_pc1 = pca$x[,1]*-1)
-  return(roost_behavior)
+  return(distinct(roost_behavior))
 }
 
 get_new_movement_vars <- function(demo, space_use, movement, roost_behavior){
@@ -717,6 +717,33 @@ get_new_movement_vars <- function(demo, space_use, movement, roost_behavior){
            "roost_div" = "roost_pc1") %>%
     filter(!is.na(movement))
   return(new_movement_vars)
+}
+
+get_all_movement_vars <- function(demo, space_use, movement, roost_behavior){
+  all_movement_vars <- demo %>%
+    left_join(space_use) %>%
+    left_join(movement) %>%
+    left_join(roost_behavior) %>%
+    rename("space_use" = "space_pc1",
+           "movement" = "movement_pc1",
+           "roost_div" = "roost_pc1") %>%
+    filter(!is.na(movement))
+  return(all_movement_vars)
+}
+
+get_correlation_scatterplots <- function(data, xcol, ycols, title){
+  p <- data %>%
+    pivot_longer(cols = all_of(ycols),
[…]

### 2024-04-15 R/functions.R
removed centrs from modeling prep

@@ -999,19 +999,6 @@ combine_metrics <- function(networkMetrics, metrics_wrapped){
   return(allMetrics)
 }
 
-get_deviations_plot <- function(allMetrics){
-  set.seed(3)
-  vultures <- sample(allMetrics$Nili_id, size = 15)
-  plt <- allMetrics %>% filter(season == "2023_summer", Nili_id %in% vultures) %>%
-    ggplot(aes(x = Nili_id, y = wrapped_normDegree))+
-    geom_violin()+
-    geom_point(aes(x = Nili_id, y = normDegree), col = "firebrick2", size = 2)+
-    facet_wrap(~situ, scales = "free")+
-    theme(axis.text.x = element_blank(), 
-          axis.ticks.x = element_blank())
-  return(plt)
-}
-
 get_metrics_summary <- function(allMetrics){
   metrics_summary <- allMetrics %>%
     group_by(season, situ, Nili_id) %>%
@@ -1027,23 +1014,8 @@ get_metrics_summary <- function(allMetrics){
   return(metrics_summary)
 }
 
-get_centrs <- function(sfs_est_centroids){
-  centrs <- sfs_est_centroids %>%
-    select(seasonUnique, Nili_id, dist_szn_centr) %>%
-    group_by(seasonUnique) %>%
-    mutate(mn = mean(dist_szn_centr),
-           sd = sd(dist_szn_centr)) %>%
-    mutate(centr = (dist_szn_centr-mn)/sd,
-           centr = -1*centr) %>%
-    sf::st_drop_geometry() %>%
-    select(seasonUnique, Nili_id, centr) %>%
-    ungroup() %>%
-    distinct()
-  return(centrs)
-}
-
 # Pre-modeling ------------------------------------------------------------
-join_movement_soc <- function(new_movement_vars, metrics_summary, centrs, season_names){
+join_movement_soc <- function(new_movement_vars, metrics_summary, season_names){
   oneage <- new_movement_vars %>%
     distinct() %>%
     group_by(Nili_id, seasonUnique) %>%
@@ -1052,8 +1024,7 @@ join_movement_soc <- function(new_movement_vars, metrics_summary, centrs, season
   # XXX THIS IS REALLY FRAGILE--MAKE IT MORE EXPLICIT!
   linked <- oneage %>% 
     left_join(metrics_summary, 
-              by = c("Nili_id", "seasonUnique")) %>%
-    left_join(centrs)
[…]

### 2024-04-15 _targets.R

@@ -132,8 +132,7 @@ list(
   tar_target(allMetrics, combine_metrics(networkMetrics, metrics_wrapped)),
   tar_target(metrics_summary, get_metrics_summary(allMetrics)),
   # Mixed models
-  tar_target(centrs, get_centrs(sfs_est_centroids)),
-  tar_target(linked, join_movement_soc(new_movement_vars, metrics_summary, centrs, season_names)),
+  tar_target(linked, join_movement_soc(new_movement_vars, metrics_summary, season_names)),
   # Reporting
   tar_target(joined0_r, report(joined0, "trackId")),
   tar_target(cleaned_r, report(cleaned, "Nili_id")),

### 2024-04-17 presentation.qmd
updated presentation before and after 4/17 vulture meeting

@@
 date: last-modified
+date-format: long
 format:
@@
 library(targets)
-theme_set(theme_minimal())
+library(patchwork)
+library(ggeffects)
+library(gtsummary)
+theme_set(theme_classic())
 tar_config_set(store = here::here('_targets'))
-source(here("scripts/04.0_mixedModels.R"))
+#source(here("scripts/04.0_mixedModels.R"))
+tar_load(cc)
+tar_load(linked)
+tar_load(deg_mod)
+tar_load(deg_z_mod)
+tar_load(str_mod)
+tar_load(str_z_mod)
+tar_load(situcolors)
+tar_load(seasoncolors)
+tar_load(movement_corr)
+tar_load(space_corr)
+tar_load(roost_corr)
 ```
 
-<!-- ## Research questions (cont'd) -->
+```{r}
+#| include: false
+effplot <- function(model, terms, dataset, pointX, pointY, colvar, legendTitle, values, ylab, xlab){
+  dat <- as.data.frame(ggeffect(model, terms = terms))
 
-<!-- Q1: What is the relationship between spatial phenotype and social interactions? -->
+  if(length(terms) > 1){
+    plt <- ggplot(dat, aes(x, predicted)) +
+      geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linewidth = 0.6, show.legend = F)+
+      geom_point(data = dataset %>% mutate("group" = .data[[colvar]]), 
+                 aes(.data[[pointX]], .data[[pointY]], col = group), alpha = 0.5)+
+      geom_line(aes(col = group), linewidth = 1) +
+      facet_wrap(~group, scales = "free_y")+
+      scale_color_manual(legendTitle, values = values) + 
+      scale_fill_manual(legendTitle, values = values) + 
+      ylab(ylab) + xlab(xlab)
+  }else{
+    plt <- ggplot(dat, aes(x, predicted)) +
+      geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linewidth = 0.6, show.legend = F)+
+      geom_point(data = dataset %>% mutate("group" = .data[[colvar]]), 
+                 aes(.data[[pointX]], .data[[pointY]], col = group), alpha = 0.5)+
+      geom_line(linewidth = 1) +
+      facet_wrap(~group, scales = "free_y")+
+      scale_color_manual(legendTitle, values = values) + 
+      ylab(ylab) + xlab(xlab)
+  }
+  return(plt)
+}
 
-<!-- -   H1A: Individuals that move more will have more, and stronger, encounters with other unique individuals. -->
+emtplot <- function(model, specs, var, values, ylab, xlab){
+  dat <- as.data.frame(emmeans::emtrends(model, specs, var))
+  lower <- ifelse("lower.CL" %in% names(dat), "lower.CL", "asymp.LCL")
+  upper <- ifelse("upper.CL" %in% names(dat), "upper.CL", "asymp.UCL")
+  plt <- dat %>% ggplot(aes(.data[[specs]], 
+                            .data[[paste0(var, ".", "trend")]], 
+                            col = .data[[specs]])) + 
+    geom_point(size = 6) +
+    geom_errorbar(aes(x = .data[[specs]], 
+                      ymin = .data[[lower]], 
+                      ymax = .data[[upper]]), 
+                  width = 0, linewidth = 2)+geom_hline(aes(yintercept = 0), 
+                                                       linetype = 2)+
+    scale_color_manual(values = values)+ 
+    theme(legend.position = "none") + ylab(ylab)+ xlab(xlab) + coord_flip()
+  return(plt)
+}
+```
 
-<!--     -   P1A: Vultures that move more will encounter more others (higher degree centrality in the social network) and have more interactions (higher strength in the social network) than vultures who move less.
--->
-
-<!-- ## Research questions (cont'd) -->
-
-<!-- -   H1B: Individuals that are more exploratory will have more but weaker encounters with other unique individuals. -->
-
-<!--     -   P1B: Exploratory vultures will encounter more unique individuals (higher degree centrality), but will have fewer interactions (lower strength) compared to less exploratory vultures. -->
-
-<!-- ## Research questions (cont'd) -->
-
-<!-- Q2: How is the relationship between spatial phenotype and social interactions impacted by social situation and season? -->
-
-<!-- -   H2A: Social situations in which animals move more will have a greater impact on social interactions -->
-
-<!--     -   P2A: Social centrality in co-flight networks will be more strongly associated with spatial phenotype than in co-roosting and co-feeding behavioral situations. -->
-
-<!-- ## Research questions (cont'd) -->
-
-<!-- -   H2B: Seasons in which animals move more will have a stronger impact on social interactions -->
-
-<!--     -   P2B: The relationship between movement and social position will be stronger in the non-breeding seasons (summer and fall) than during the breeding season. -->
-
-## Examining response variable distributions
+## Movement
 
 ```{r}
-tar_load(all_movement_vars)
-tar_load(cc)
-all_movement_vars %>%
+movement_corr
+```
+
+## Space use
+
+```{r}
+space_corr
+```
+
+## Roost diversification
+
+```{r}
+roost_corr
+```
+
+## Predictor variable distributions
+
+```{r}
+linked %>%
   pivot_longer(cols = c(movement, space_use, roost_div), 
@@
 
+# Predictor variable correlations
+
+## Roost \~ Movement
+
+```{r}
+linked %>%
+  select(movement, roost_div, space_use, season, seasonUnique, Nili_id) %>%
+  distinct() %>%
+  ggplot(aes(x = movement, y = roost_div, col = season, group = seasonUnique))+
+  geom_point()+
+  geom_smooth(method = "lm", alpha = 0.15)+
+  scale_color_manual("Season", values = seasoncolors)+
+  ylab("Roost diversification")+ xlab("Movement")
+```
+
+## Movement \~ Space
+
+```{r}
+linked %>%
+  select(movement, roost_div, space_use, season, seasonUnique, Nili_id) %>%
+  distinct() %>%
+  ggplot(aes(x = space_use, y = movement, col = season, group = seasonUnique))+
+  geom_point()+
+  geom_smooth(method = "lm", alpha = 0.15)+
+  scale_color_manual("Season", values = seasoncolors)+
+  ylab("Movement")+ xlab("Space use")
+```
+
+## Space \~ Roost
+
+```{r}
+linked %>%
+  select(movement, roost_div, space_use, season, seasonUnique, Nili_id) %>%
+  distinct() %>%
+  ggplot(aes(x = space_use, y = roost_div, col = season, group = seasonUnique))+
+  geom_point()+
+  geom_smooth(method = "lm", alpha = 0.15)+
+  scale_color_manual("Season", values = seasoncolors)+
+  ylab("Roost diversification")+ xlab("Space use")
+```
+
+## Response variable distributions (normalized degree)
+
+```{r}
+linked %>%
+  ggplot(aes(x = normDegree, col = situ))+
+  geom_density()+
+  facet_wrap(~seasonUnique, scales = "free_y")+
+  scale_color_manual("Situation", values = situcolors)+ylab("")
+```
+
+## Response variable distributions (normalized strength)
+
+```{r}
+linked %>%
+  ggplot(aes(x = normStrength, col = situ))+
+  geom_density()+
+  facet_wrap(~seasonUnique, scales = "free_y")+
+  scale_color_manual("Situation", values = situcolors)+ylab("")
+```
+
 ## Wrap-around randomizations (degree)
@@
   filter(Nili_id %in% vultures) %>%
-  ggplot(aes(x = Nili_id, y = wrapped_normDegree, col = situ))+
-  geom_point(aes(y = normDegree), position = position_dodge(width = 0.4))+
+  ggplot(aes(x = Nili_id, y = wrapped_degree, col = situ))+
+  geom_point(aes(y = degree), position = position_dodge(width = 0.4))+
   geom_violin(position = position_dodge(width = 0.4))+
   facet_wrap(~season, scales = "free")+
-  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
+  scale_color_manual(name = "Situation", values = situcolors)+
   theme_classic()+
-  ylab("Degree (normalized)")+
-  theme(axis.title.x = element_blank(),
-        legend.position = "bottom")+
+  ylab("Degree")+
+  theme(axis.title.x = element_blank())+
   ggtitle("Degree (observed vs. random)")
@@
   filter(Nili_id %in% vultures) %>%
-  ggplot(aes(x = Nili_id, y = wrapped_normStrength, col = situ))+
-  geom_point(aes(y = normStrength), position = position_dodge(width = 0.4))+
+  ggplot(aes(x = Nili_id, y = wrapped_strength, col = situ))+
+  geom_point(aes(y = strength), position = position_dodge(width = 0.4))+
   geom_violin(position = position_dodge(width = 0.4))+
   facet_wrap(~season, scales = "free")+
-  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
+  scale_color_manual(name = "Situation", values = situcolors)+
   theme_classic()+
-  ylab("Strength (normalized)")+
-  theme(axis.title.x = element_blank(),
-        legend.position = "bottom")+
+  ylab("Strength")+
+  theme(axis.title.x = element_blank())+
   ggtitle("Strength (observed vs. random)")
@@
 
-## Degree, but not strength, was higher than random.
+## Response variable distributions (z-scores)
 
 ```{r}
-tar_load(linked)
-
 linked %>%
@@
   ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
-  geom_density()+
-  facet_wrap(~measure, scales = "free")+
+  geom_density(aes(linetype = measure))+
+  facet_wrap(~measure*situ, scales = "free")+
   geom_vline(aes(xintercept = 0), col = "red")+
@@
   xlab("Z-score")
-
-linked %>%
-  select(seasonUnique, situ, z_str) %>%
-  ungroup() %>%
-  ggplot(aes(x = z_str))
 ```
 
-## Movement vs. degree deviation
+# Model results: Degree (z-scores, non-normalized)
+
+## Space use and situation
 
 ```{r}
-#| label: fig-d-space
-#| fig-cap: "Relationship between movement and degree deviation"
-plot_d_eff_movement
+#| fig-cap: "Degree z-scores"
+deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
+                            dataset = linked, pointX = "space_use", 
+                            pointY = "z_deg", col = "situ", 
+                            legendTitle = "Situation", values = situcolors, 
+                            ylab = "Degree (z-score)", xlab = "Space use (log)")
+
+deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
+                                var = "space_use", values = situcolors, 
+                                ylab = "Space use effect (Degree z-scores)", 
+                                xlab = "Situation")
+
+deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
 ```
 
-## Movement measure
+##
+
+![](fig/whiteboards/deg_z_space_situ_ro.jpg)
+
+##
+
+![](fig/whiteboards/deg_z_space_situ_fefl.jpg)
+
+## Roost diversification
 
 ```{r}
-tar_load(movement_corr)
-movement_corr
+#| fig-cap: "Degree z-scores"
+deg_z_roost_P <- effplot(model = deg_z_mod, terms = c("roost_div", "situ"), 
+                       dataset = linked, pointX = "roost_div", 
+                       pointY = "z_deg", col = "situ", 
+                       legendTitle = "Situation", values = situcolors, 
+                       ylab = "Degree (z-score)", xlab = "Roost diversification")
+deg_z_roost_P & theme_classic()
 ```
 
-## Space use vs. degree deviation
+##
+
+![](fig/whiteboards/deg_z_roost.jpg)
+
+# Model results: Strength (z-scores, non-normalized)
+
+## Space use and situation
 
 ```{r}
-#| label: fig-d-space-situ
-#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
-plot_d_eff_situ_space
+#| fig-cap: "Strength z-scores"
+str_z_situ_space_emt_P <- emtplot(model = str_z_mod, specs = "situ", 
+                                var = "space_use", values = situcolors, 
+                                ylab = "Space use effect (strength z-score)", 
+                                xlab = "Situation")
+
+str_z_situ_space_P <- effplot(model = str_z_mod, terms = c("space_use", "situ"), 
+                            dataset = linked, pointX = "space_use", 
+                            pointY = "z_str", col = "situ", 
+                            legendTitle = "Situation", values = situcolors, 
+                            ylab = "Strength (z-score)", xlab = "Space use (log)")
+
+str_z_situ_space_P + str_z_situ_space_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
 ```
 
-## Space use vs. degree deviation
+## Roost diversification
 
 ```{r}
-plot_d_emt_situ_space
+#| fig-cap: "Strength z-scores"
+str_z_roost_P <- effplot(model = str_z_mod, terms = c("roost_div", "situ"), 
+                         dataset = linked, pointX = "roost_div", 
+                         pointY = "z_str", col = "situ", 
+                         legendTitle = "Situation", values = situcolors, 
+                         ylab = "Strength (z-score)", xlab = "Roost diversification")
+str_z_roost_P & theme_classic()
 ```
 
-## Space use measure
+# Model results: Normalized degree
+
+## Movement and situation
 
 ```{r}
-tar_load(space_corr)
-space_corr
+#| fig-cap: "Normalized degree"
+deg_situ_movement_P <- effplot(model = deg_mod, terms = c("movement", "situ"), 
+                               dataset = linked, pointX = "movement", 
+                               pointY = "normDegree", col = "situ", 
+                               legendTitle = "Situation", values = situcolors, 
+                               ylab = "normDegree", xlab = "Movement")
+
+deg_situ_movement_emt_P <- emtplot(model = deg_mod, specs = "situ", 
+                                   var = "movement", values = situcolors, 
+                                   ylab = "Movement effect (normDegree)", 
+                                   xlab = "Situation")
+
+deg_situ_movement_P + deg_situ_movement_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
 ```
 
-## Roost diversification vs. degree deviation
+## Movement and season
 
 ```{r}
-#| label: fig-d-roost
-#| fig-cap: "Relationship between normalized degree and roost diversification"
-plot_d_eff_roost 
+#| fig-cap: "Normalized degree"
+deg_season_movement_P <- effplot(model = deg_mod, terms = c("movement", "season"), 
+                                 dataset = linked, pointX = "movement", 
+                                 pointY = "normDegree", col = "season", 
+                                 legendTitle = "Season", values = seasoncolors, 
+                                 ylab = "normDegree", xlab = "Movement")
+
+deg_season_movement_emt_P <- emtplot(model = deg_mod, specs = "season", 
+                                     var = "movement", values = seasoncolors, 
+                                     ylab = "Movement effect (normDegree)", 
+                                     xlab = "Season")
+
+deg_season_movement_P + deg_season_movement_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
 ```
 
-## Roost diversification measure
+## Space use and situation
 
 ```{r}
-tar_load(roost_corr)
-roost_corr
+#| fig-cap: "Normalized degree"
+
+deg_situ_space_P <- effplot(model = deg_mod, terms = c("space_use", "situ"), 
+                            dataset = linked, pointX = "space_use", 
+                            pointY = "normDegree", col = "situ", 
+                            legendTitle = "Situation", values = situcolors, 
+                            ylab = "normDegree", xlab = "Space use (log)")
+
+deg_situ_space_emt_P <- emtplot(model = deg_mod, specs = "situ", 
+                                var = "space_use", values = situcolors, 
+                                ylab = "Space use effect (normDegree)", 
+                                xlab = "Situation")
+
+deg_situ_space_P + deg_situ_space_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
 ```
 
-## Degree model summary {.smaller .scrollable}
+## Roost diversification
 
 ```{r}
-summary(degree_mod)
+#| fig-cap: "Normalized degree"
+deg_roost_P <- effplot(model = deg_mod, terms = c("roost_div", "situ"), 
+                               dataset = linked, pointX = "roost_div", 
+                               pointY = "normDegree", col = "situ", 
+                               legendTitle = "Situation", values = situcolors, 
+                               ylab = "normDegree", xlab = "Roost diversification")
+deg_roost_P & theme_classic()
 ```
 
-## The strength model did not fit well.
+# Model results: Normalized strength
+
+## Roost diversification and situation
 
 ```{r}
-simulationOutput <- DHARMa::simulateResiduals(strength_mod)
-plot(simulationOutput, pch=".")
+#| fig-cap: "Normalized strength"
+str_situ_roost_P <- effplot(model = str_mod, terms = c("roost_div", "situ"), 
+                               dataset = linked, pointX = "roost_div", 
+                               pointY = "normStrength", col = "situ", 
+                               legendTitle = "Situation", values = situcolors, 
+                               ylab = "normStrength", xlab = "Roost diversification")
+
+str_situ_roost_emt_P <- emtplot(model = str_mod, specs = "situ", 
+                                   var = "roost_div", values = situcolors, 
+                                   ylab = "Roost diversification effect (normStrength)", 
+                                   xlab = "Situation")
+str_situ_roost_P + str_situ_roost_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
 ```
 
-## Still, here is the summary as it stands {.smaller .scrollable}
+## Takeaways
+
+1.  Movement relates to strength/degree, but much of this is incidental--effect drops away after randomizing
+2.  Differences in excess/preferential sociality relate mostly to space use
+
+# Model summaries
+
+## Normalized (not z-scores) {.scrollable}
 
 ```{r}
-summary(strength_mod)
+tbl_deg_mod <- tbl_regression(deg_mod)
+tbl_str_mod <- tbl_regression(str_mod)
+
+merged <- tbl_merge(
+  tbls <- list(tbl_deg_mod, tbl_str_mod),
+  tab_spanner = c("**Degree (normalized)**", "**Strength (normalized)**")
+)
+merged
+
+# save the table (requires the `gt` package)
+#gt::gtsave(as_gt(merged), file = file.path("fig", "mm_summary.png"))
 ```
+
+## Z-scores {.scrollable}
+
+```{r}
+tbl_deg_z_mod <- tbl_regression(deg_z_mod)
+tbl_str_z_mod <- tbl_regression(str_z_mod)
+
+merged_z <- tbl_merge(
+  tbls <- list(tbl_deg_z_mod, tbl_str_z_mod),
+  tab_spanner = c("**Degree (z-score)**", "**Strength (z-score)**")
+)
+merged_z
+
+# save the table (requires the `gt` package)
+#gt::gtsave(as_gt(merged_z), file = file.path("fig", "mm_summary_z.png"))
+```

### 2024-04-17 R/functions.R

@@ -21,6 +21,16 @@ get_cc <- function(){
   return(cc)
 }
 
+get_situcolors <- function(cc){
+  situcolors <- c(cc$feedingColor, cc$flightColor, cc$roostingColor)
+  return(situcolors)
+}
+
+get_seasoncolors <- function(cc){
+  seasoncolors <- c(cc$breedingColor, cc$summerColor, cc$fallColor)
+  return(seasoncolors)
+}
+
 # Data prep ---------------------------------------------------------------
 get_loginObject <- function(pw){
   load(pw)
@@ -1043,6 +1053,22 @@ join_movement_soc <- function(new_movement_vars, metrics_summary, season_names){
     mutate(situ = case_when(type == "flight" ~ "Fl", 
                             type == "feeding" ~ "Fe",
                             type == "roosting" ~ "Ro"))
+  
+  # Let's examine zeroes for the social network measures. I know that when we calculate the social networks, we had a lot of zeroes for both degree and strength. But most of the individuals that didn't have network connections probably aren't our focal individuals for the movement measures.
+  # linked %>% filter(z_deg == 0 | z_str == 0) # nobody
+  # linked %>% filter(degree == 0, strength == 0) # just one individual in one season
+  # linked %>% filter(is.na(z_deg) | is.na(z_str)) # just one individual in one season
+  # linked %>% filter(is.na(z_deg) | is.na(z_str)) # just one individual in one season
+  
+  # Let's remove that individual in case she poses a problem for modeling
+  linked <- linked %>%
+    filter(!is.na(z_deg), !is.na(z_str))
+  #nrow(linked)
+  
+  #linked %>% filter(is.infinite(z_deg) | is.infinite(z_str)) # likewise, removing the infinite individual
+  linked <- linked %>%
+    filter(!is.infinite(z_deg), !is.infinite(z_str))
+  
   return(linked)
 }
 
@@ -1051,4 +1077,23 @@ report <- function(dataset, id){
   n_vultures <- length(unique(dataset[[id]]))
   vultures <- unique(dataset[[id]])
   return(c("points" = n_points, "vultures" = n_vultures))
+}
+
+# Models
+# Code for these is taken over from 04.0_mixedModels.R once I'm done testing them.
+get_deg_mod <- function(linked){
[…]

### 2024-04-17 _targets.R

@@ -52,6 +52,8 @@ lapply(list.files("R", full.names = TRUE), source) # source all scripts in the R
 list(
   # Prepare data
   tar_target(cc, get_cc()),
+  tar_target(situcolors, get_situcolors(cc)), #XXX
+  tar_target(seasoncolors, get_seasoncolors(cc)), #XXX
   tar_target(pw, "movebankCredentials/pw.Rda", format = "file"),
   tar_target(loginObject, get_loginObject(pw)),
   tar_target(inpa, get_inpa(loginObject)),
@@ -131,8 +133,6 @@ list(
   tar_target(metrics_wrapped, get_metrics_wrapped(with_times, roosts, season_names, roostPolygons, 100, 1)),
   tar_target(allMetrics, combine_metrics(networkMetrics, metrics_wrapped)),
   tar_target(metrics_summary, get_metrics_summary(allMetrics)),
-  # Mixed models
-  tar_target(linked, join_movement_soc(new_movement_vars, metrics_summary, season_names)),
   # Reporting
   tar_target(joined0_r, report(joined0, "trackId")),
   tar_target(cleaned_r, report(cleaned, "Nili_id")),
@@ -141,7 +141,13 @@ list(
   tar_target(removed_northern_r, map(removed_northern, ~report(.x, "Nili_id"))),
   tar_target(removed_lfr_r, map(removed_lfr, ~report(.x, "Nili_id"))),
   tar_target(downsampled_10min_forSocial_r, map(downsampled_10min_forSocial, ~report(.x, "Nili_id"))),
-  tar_target(downsampled_10min_r, map(downsampled_10min, ~report(.x, "Nili_id")))
+  tar_target(downsampled_10min_r, map(downsampled_10min, ~report(.x, "Nili_id"))),
+  # Mixed models
+  tar_target(linked, join_movement_soc(new_movement_vars, metrics_summary, season_names)),
+  tar_target(deg_mod, get_deg_mod(linked)),
+  tar_target(deg_z_mod, get_deg_z_mod(linked)),
+  tar_target(str_mod, get_str_mod(linked)),
+  tar_target(str_z_mod, get_str_z_mod(linked))
 )
 
 

### 2024-04-17 scripts/04.0_mixedModels.R

@@ -12,6 +12,7 @@ library(emmeans) # estimated marginal means/trends
 library(ggeffects)
 library(gtsummary)
 library(here)
+# THIS SCRIPT IS FOR TESTING OUT MODELS. FINAL MODEL CODE GOES OVER INTO FUNCTIONS.R TO BE INCORPORATED INTO THE TARGETS PIPELINE.
 
 # Set ggplot theme to classic
 theme_set(theme_classic())
@@ -25,32 +26,32 @@ seasoncolors <- c(cc$breedingColor, cc$summerColor, cc$fallColor)
 # Examine response variable distributions ---------------------------------
 
 # Normalized degree and strength ------------------------------------------
-linked %>%
-  ggplot(aes(x = degree, col = season, group = seasonUnique))+
-  geom_density()+
-  facet_wrap(~type, scales = "free") # very distinct distributions, but since we will have a random effect for seasonUnique, that should cover it... 
-
-linked %>%
-  ggplot(aes(x = strength, col = season, group = seasonUnique))+
-  geom_density()+
-  facet_wrap(~type, scales = "free") 
-
-# Degree and strength z-scores (non-normalized) ---------------------------
-linked %>%
-  ggplot(aes(x = z_deg, col = season, group = seasonUnique))+
-  geom_density()+
-  facet_wrap(~type, scales = "free") # these are weird distributions. I wonder what model family we should use here...
-
-linked %>%
-  ggplot(aes(x = z_str, col = season, group = seasonUnique))+
-  geom_density()+
-  facet_wrap(~type, scales = "free")
+# linked %>%
+#   ggplot(aes(x = degree, col = season, group = seasonUnique))+
+#   geom_density()+
+#   facet_wrap(~type, scales = "free") # very distinct distributions, but since we will have a random effect for seasonUnique, that should cover it... 
+# 
+# linked %>%
+#   ggplot(aes(x = strength, col = season, group = seasonUnique))+
+#   geom_density()+
+#   facet_wrap(~type, scales = "free") 
+# 
+# # Degree and strength z-scores (non-normalized) ---------------------------
+# linked %>%
+#   ggplot(aes(x = z_deg, col = season, group = seasonUnique))+
+#   geom_density()+
+#   facet_wrap(~type, scales = "free") # these are weird distributions. I wonder what model family we should use here...
+# 
+# linked %>%
+#   ggplot(aes(x = z_str, col = season, group = seasonUnique))+
+#   geom_density()+
[…]

### 2024-04-17 scripts/05.0_modelPlots.R

@@ -1,0 +1,170 @@
+library(tidyverse)
+library(performance) # for modeling and comparisons
+library(emmeans) # estimated marginal means/trends
+library(ggeffects)
+library(here)
+
+situcolors <- c(cc$feedingColor, cc$flightColor, cc$roostingColor)
+seasoncolors <- c(cc$breedingColor, cc$summerColor, cc$fallColor)
+
+# Set ggplot theme to classic
+theme_set(theme_classic())
+
+tar_load(linked)
+
+load(here("data/deg_mod.Rda"))
+load(here("data/deg_z_mod.Rda"))
+load(here("data/str_mod.Rda"))
+load(here("data/str_z_mod.Rda"))
+
+# Degree models to plot -----------------------------------------------------
+### Without randomizations:
+summary(deg_mod) # situ*space, movement*season, situ*movement, roost_div
+### With randomizations:
+summary(deg_z_mod) # situ*space, roost_div
+
+# Strength models to plot ---------------------------------------------------
+### Without randomizations:
+summary(str_mod) # situ*roost, roost_div
+### With randomizations:
+summary(str_z_mod) # situ*space, roost_div
+
+# Define plotting functions -----------------------------------------------
+effplot <- function(model, terms, dataset, pointX, pointY, colvar, legendTitle, values, ylab, xlab){
+  dat <- as.data.frame(ggeffect(model, terms = terms))
+  
+  if(length(terms) > 1){
+    plt <- ggplot(dat, aes(x, predicted)) +
+      geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linewidth = 0.6, show.legend = F)+
+      geom_point(data = dataset, aes(.data[[pointX]], .data[[pointY]], col = .data[[colvar]]), alpha = 0.5)+
+      geom_line(aes(col = group), linewidth = 1) +
+      scale_color_manual(legendTitle, values = values) + 
+      scale_fill_manual(legendTitle, values = values) + 
+      ylab(ylab) + xlab(xlab)
+  }else{
+    plt <- ggplot(dat, aes(x, predicted)) +
+      geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linewidth = 0.6, show.legend = F)+
[…]

### 2024-04-17 manuscript.rmd
switched to rmd so I can try using trackdown/google docs

--- 2024-04-09 docs/manuscript.qmd
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
 # Introduction
 
-```{r}
-#| include: false
+```{r include = FALSE}
 library(tidyverse)
 library(here)
+library(targets)
+library(patchwork)
+library(ggeffects)
+library(gtsummary)
+theme_set(theme_classic())
 tar_config_set(store = here::here('_targets'))
+#source(here("scripts/04.0_mixedModels.R"))
+tar_load(cleaned_r)
+tar_load(joined0_r)
+tar_load(downsampled_10min_forSocial_r)
+tar_load(downsampled_10min_r)
+tar_load(cc)
+tar_load(linked)
+tar_load(deg_mod)
+tar_load(deg_z_mod)
+tar_load(str_mod)
+tar_load(str_z_mod)
+tar_load(situcolors)
+tar_load(seasoncolors)
+tar_load(movement_corr)
+tar_load(space_corr)
+tar_load(roost_corr)
 ```
@@
 ```{r}
-tar_load(cleaned_r)
-tar_load(joined0_r)
 pctremovedcleaned <- (joined0_r[1]-cleaned_r[1])/joined0_r[1]
@@
 ```{r}
-tar_load(downsampled_10min_forSocial_r)
 nvultures <- map_dbl(downsampled_10min_forSocial_r, "vultures")
@@
 ```{r}
-tar_load(downsampled_10min_r)
-nvultures <- map_dbl(downsampled_10min_r, "vultures")
+nvulturesmov <- map_dbl(downsampled_10min_r, "vultures")
 ```
@@
 Finally, we downsampled the movement data to a 10-minute interval to remove high-frequency fixes.
-Ultimately, we calculated movement measures for a total of `r nvulturesmov` **focal individuals** in total (range: `r min(nvulturesmov_szns)`-`r max(nvulturesmov_szns)`; mean = `r round(mean(nvulturesmov_szns), 2)` vultures per season, sd = `r round(sd(nvulturesmov_szns), 2)` vultures per season).
+The data used for movement analysis contained between `r min(nvulturesmov)` and `r max(nvulturesmov)` vultures per season (mean = `r mean(nvulturesmov)`; sd = `r sd(nvulturesmov)`).
 
@@
 
-```{r}
-#| include: false
-load(here("data/calcMovement/pca_space.Rda"))
-propvar_space <- as.data.frame(summary(pca_space)$importance)$PC1[2]
-
-load(here("data/calcMovement/pca_roost.Rda"))
-propvar_roost <- as.data.frame(summary(pca_roost)$importance)$PC1[2]
-
-load(here("data/calcMovement/pca_movement.Rda"))
-propvar_movement <- as.data.frame(summary(pca_movement)$importance)$PC1[2]
+```{r include = FALSE}
+# INSERT VARIABLE CONTRIBUTION PLOTS HERE
 ```
@@
 **Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity.
-It explained `r 100*round(propvar_movement,2)`% of the variation in the PCA.
+It explained [PERCENT]% of the variation in the PCA.
 Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall.
 **Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE.
-It explained `r 100*round(propvar_space,2)`% of the variation in the PCA.
+It explained [PERCENT]% of the variation in the PCA.
 Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e.
@@
 **Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night).
-It explained `r 100*round(propvar_roost,2)`% of the variation in the PCA.
+It explained [PERCENT]% of the variation in the PCA.
 
@@
 
-```{r}
-#| include: false
-#| output: false
-source(here("scripts/04.0_mixedModels.R"))
+## Social preference (deviations from random sociality)
+
+Over the course of a season, individual vultures interacted with significantly more conspecifics, and had significantly more interactions, than expected by chance (i.e.
+observed degree and strength values were higher than those derived from the randomized datasets).
+The magnitude of the deviations from random varied by social situation.
+
+```{r message = FALSE}
+linked %>%
+  select(seasonUnique, situ, z_deg, z_str) %>%
+  ungroup() %>%
+  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
+  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
+  geom_density(aes(linetype = measure))+
+  facet_wrap(~measure*situ, scales = "free")+
+  geom_vline(aes(xintercept = 0), col = "red")+
+  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
+  theme_classic()+
+  theme(axis.title.y = element_blank())+
+  xlab("Z-score")
 ```
@@
 
-```{r}
-#| eval: false
-#| output: false
-#| label: fig-d-space
-#| fig-cap: "Relationship between normalized degree and space use"
-# NOT SHOW
-plot_d_eff_space # null--no overall relationship between space use and degree
-```
-
-```{r}
-#| label: fig-d-space-situ
-#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
-plot_d_eff_situ_space
-plot_d_emt_situ_space
-```
-
-There was a significant relationship between how vultures used space and how many other individuals they encountered in flight.
-Vultures that consistently ranged widely encountered more others in flight, while those that concentrated their movement in smaller home ranges and smaller proportional core areas flew with fewer others.
-There was a slight, albeit non-statistically-significant, trend toward lower degree in the co-feeding and co-roosting networks for vultures with larger ranges.
-
-```{r}
-#| eval: false
-#| output: false
-#| label: fig-d-roost
-#| fig-cap: "Relationship between normalized degree and roost diversification"
-# NOT SHOW
-plot_d_eff_roost # null--no overall relationship between roost diversification and degree
-```
-
-```{r}
-#| eval: false
-#| output: false
-#| label: fig-d-movement-situ
-#| fig-cap: "Relationship between normalized degree and movement, with separate trend lines for each social situation"
-# NOT SHOW
-plot_d_eff_situ_movement
-plot_d_emt_situ_movement # going to show the plot by season instead of this one
-```
-
-```{r}
-#| label: fig-d-movement-situ-season
-#| fig-cap: "Relationship between normalized degree and movement, for each season and social situation."
-plot_d_emt_situ_movement_season
-```
-
-Movement behavior affected degree in all three social situations.
-Vultures with a higher movement score co-roosted with fewer conspecifics only in the transient, pre-breeding season; otherwise movement did not affect their co-roosting degree.
-In the breeding and post-breeding seasons, vultures that moved more encountered more conspecifics in flight and while feeding, but this relationship did not hold in the pre-breeding season.
-
 ## Strength
 
-```{r}
-#| eval: false
-#| output: false
-#| label: fig-s-movement
-#| fig-cap: "Relationship between normalized strength and movement"
-plot_s_eff_movement # null
-```
-
-There was no relationship between movement and strength, either overall or in any season or situation.
-
-```{r}
-#| eval: false
-#| label: fig-s-roost
-#| fig-cap: "Relationship between normalized strength and roost diversification"
-plot_s_eff_roost
-```
-
-However, individuals with more diversified roosts had lower strength (i.e. they interacted less) overall, considering seasons and situations together.
-
-```{r}
-#| eval: false
-#| label: fig-s-space
-#| fig-cap: "Relationship between normalized strength and space use"
-plot_s_eff_situ_space
-plot_s_emt_situ_space
-```
-
-Vultures with greater space use had significantly more co-flight interactions (higher strength in the co-flight network).
-They also tended to have a higher co-roosting strength, though the trend was not statistically significant.
-There was no relationship between seasonal space use and co-feeding strength.
-
 # Discussion

### 2024-04-17 manuscript.rmd
updated with more of the results

@@
 tar_load(roost_corr)
+
+effplot <- function(model, terms, dataset, pointX, pointY, colvar, legendTitle, values, ylab, xlab){
+  dat <- as.data.frame(ggeffect(model, terms = terms))
+
+  if(length(terms) > 1){
+    plt <- ggplot(dat, aes(x, predicted)) +
+      geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linewidth = 0.6, show.legend = F)+
+      geom_point(data = dataset %>% mutate("group" = .data[[colvar]]), 
+                 aes(.data[[pointX]], .data[[pointY]], col = group), alpha = 0.5)+
+      geom_line(aes(col = group), linewidth = 1) +
+      facet_wrap(~group, scales = "free_y")+
+      scale_color_manual(legendTitle, values = values) + 
+      scale_fill_manual(legendTitle, values = values) + 
+      ylab(ylab) + xlab(xlab)
+  }else{
+    plt <- ggplot(dat, aes(x, predicted)) +
+      geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linewidth = 0.6, show.legend = F)+
+      geom_point(data = dataset %>% mutate("group" = .data[[colvar]]), 
+                 aes(.data[[pointX]], .data[[pointY]], col = group), alpha = 0.5)+
+      geom_line(linewidth = 1) +
+      facet_wrap(~group, scales = "free_y")+
+      scale_color_manual(legendTitle, values = values) + 
+      ylab(ylab) + xlab(xlab)
+  }
+  return(plt)
+}
+
+emtplot <- function(model, specs, var, values, ylab, xlab){
+  dat <- as.data.frame(emmeans::emtrends(model, specs, var))
+  lower <- ifelse("lower.CL" %in% names(dat), "lower.CL", "asymp.LCL")
+  upper <- ifelse("upper.CL" %in% names(dat), "upper.CL", "asymp.UCL")
+  plt <- dat %>% ggplot(aes(.data[[specs]], 
+                            .data[[paste0(var, ".", "trend")]], 
+                            col = .data[[specs]])) + 
+    geom_point(size = 6) +
+    geom_errorbar(aes(x = .data[[specs]], 
+                      ymin = .data[[lower]], 
+                      ymax = .data[[upper]]), 
+                  width = 0, linewidth = 2)+geom_hline(aes(yintercept = 0), 
+                                                       linetype = 2)+
+    scale_color_manual(values = values)+ 
+    theme(legend.position = "none") + ylab(ylab)+ xlab(xlab) + coord_flip()
+  return(plt)
+}
 ```
@@
 
-```{r message = FALSE}
+```{r message = FALSE, out.width = "100%"}
 linked %>%
@@
 
+Preferential degree (i.e.
+degree z-score) was significantly associated with space use and roost diversification, but not with movement.
+Vultures that used more space (i.e.
+those with larger home ranges and larger proportional core areas) had a co-flight degree that deviated more from random (i.e.
+more preferential degree) than those with smaller space use.
+By contrast, co-roosting degree deviated less, not more, from random for wide-ranging vultures.
+[FIGURE] (Co-roosting degree showed a similar trend to co-flight, though it was only marginally significant).
+
+```{r fig.cap = "Degree z-scores", out.width = "100%"}
+deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
+                            dataset = linked, pointX = "space_use", 
+                            pointY = "z_deg", col = "situ", 
+                            legendTitle = "Situation", values = situcolors, 
+                            ylab = "Degree (z-score)", xlab = "Space use (log)")
+
+deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
+                                var = "space_use", values = situcolors, 
+                                ylab = "Space use effect (Degree z-scores)", 
+                                xlab = "Situation")
+
+deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(4, 1), guides = "collect") & theme_classic()
+```
+
+Meanwhile, roost diversification negatively predicted preferential degree in all social situations, with no significant difference in effect magnitude between situations.
+Vultures with more unique roosts, more roost switching, and/or a higher diversity of roosts differed less from the degree expected by chance than those with lower roost diversification.
+
+```{r}
+deg_z_roost_P <- effplot(model = deg_z_mod, terms = c("roost_div", "situ"), 
+                       dataset = linked, pointX = "roost_div", 
+                       pointY = "z_deg", col = "situ", 
+                       legendTitle = "Situation", values = situcolors, 
+                       ylab = "Degree (z-score)", xlab = "Roost diversification")
+deg_z_roost_P & theme_classic()
+```
+
 ## Strength
 
+The relationships were very similar for strength.
+Preferential strength in the co-flight and co-feeding networks was positively associated with space use, with wider-ranging vultures' strength deviating the most from random.
+Meanwhile, the opposite was once again true for the co-roosting network.
+Vultures with smaller ranges deviated singificantly more from random than those with larger ranges.
+
 # Discussion

### 2024-05-06 manuscript.rmd
figures and further exploration for manuscript

@@
 
-# Introduction
+Introduction
 
@@
 library(gtsummary)
-theme_set(theme_classic())
+library(extrafont)
+loadfonts(device = "win")
+theme_set(theme_classic() + theme(text = element_text(family="Verdana")))
 tar_config_set(store = here::here('_targets'))
@@
 
-#### Social network centrality is known to be important in animals.
+Social network centrality is known to be important in animals.
 
-The social structure of animal groups has consequences for disease transmission, information sharing, and population dynamics.
+The social structure of animal groups \[a\]has consequences for disease transmission, information sharing, and population dynamics.
 At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration [@sah2017; @silk2017], govern the flow of information about food resources \[REF\], and contribute to population persistence through Allee effects \[REF\].
@@
 
-#### We observe that some individuals are more central than others in a social network. What causes this variation?
+We observe that some individuals are more central than others in a social network. What causes this variation?
 
@@
 
-##### Individual traits
+Individual traits
 
@@
 
-#### Individuals' spatial phenotypes are important determinants of their social centrality
+Individuals' spatial phenotypes are important determinants of their social centrality
 
@@
 
--   Note that explicit spatial context is also important, and we didn't consider it. (We should have! How do we justify not doing that?)
-
     -   An individual's observed spatial phenotype results from a combination of its internal tendency toward movement and the external environmental conditions through which it moves [(Nathan et al.
@@
 
-#### Social interactions occur when individuals encounter each other, in various behavioral situations.
+Social interactions occur when individuals encounter each other, in various behavioral situations.
 
@@
 
-#### We set out to uncover whether spatial phenotypes could explain observed variation in social centrality.
+We set out to uncover whether spatial phenotypes could explain observed variation in social centrality.
 
@@
 
-##### Research Questions
+Research Questions
 
@@
 
-# Methods {data-link="methods"}
+Methods {data-link="methods"}
 
-### Study system
+Study system
 
@@
 
-## Field methods
+Field methods
 
@@
 
-## Data preparation
+Data preparation
 
@@
 
-#### Determining roost locations
+Determining roost locations
 
@@
 
-#### Further cleaning for movement analysis
+Further cleaning for movement analysis
 
@@
 
-## Measuring spatial phenotype
+Measuring spatial phenotype
 
@@
 
-## Summarizing movement measures
+Summarizing movement measures
 
-```{r include = FALSE}
-# INSERT VARIABLE CONTRIBUTION PLOTS HERE
+```{r eval = FALSE}
+ INSERT VARIABLE CONTRIBUTION PLOTS HERE
 ```
@@
 
-**Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity.
-It explained [PERCENT]% of the variation in the PCA.
+Movement was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity.
+It explained \[PERCENT\]% of the variation in the PCA.
 Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall.
-**Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE.
-It explained [PERCENT]% of the variation in the PCA.
+Space use was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE.
+It explained \[PERCENT\]% of the variation in the PCA.
 Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e.
 the core area made up a larger proportion of the home range).
-**Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night).
-It explained [PERCENT]% of the variation in the PCA.
+Roost diversification was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night).
+It explained \[PERCENT\]% of the variation in the PCA.
 
@@
 
-## Construction of proximity-based social networks
+Construction of proximity-based social networks
 
@@
 
-### Co-flight
+Co-flight
 
@@
 
-### Co-feeding
+Co-feeding
 
@@
 
-### Co-roosting
+Co-roosting
 
@@
 
-## Calculation of social network measures
+Calculation of social network measures
 
@@
 
-## Comparison to a reference model
+Comparison to a reference model
 
@@
 
-## Model fitting and selection
+Model fitting and selection
 
@@
 
-# Results
+Results
 
-## Social preference (deviations from random sociality)
+Social preference (deviations from random sociality)
 
@@
   scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
-  theme_classic()+
   theme(axis.title.y = element_blank())+
@@
 
-## Degree
+Degree
 
@@
 By contrast, co-roosting degree deviated less, not more, from random for wide-ranging vultures.
-[FIGURE] (Co-roosting degree showed a similar trend to co-flight, though it was only marginally significant).
+\[FIGURE\] (Co-roosting degree showed a similar trend to co-flight, though it was only marginally significant).
 
 ```{r fig.cap = "Degree z-scores", out.width = "100%"}
-deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
-                            dataset = linked, pointX = "space_use", 
-                            pointY = "z_deg", col = "situ", 
-                            legendTitle = "Situation", values = situcolors, 
-                            ylab = "Degree (z-score)", xlab = "Space use (log)")
+# deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
+#                             dataset = linked, pointX = "space_use", 
+#                             pointY = "z_deg", col = "situ", 
+#                             legendTitle = "Situation", values = situcolors, 
+#                             ylab = "Degree (z-score)", xlab = "Space use (log)")
+# 
+# deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
+#                                 var = "space_use", values = situcolors, 
+#                                 ylab = "Space use effect (Degree z-scores)", 
+#                                 xlab = "Situation")
+# 
+# deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(4, 1), guides = "collect") & theme_classic()
+pts <- linked %>% mutate(group = case_when(situ == "Fe" ~ "Feeding",
+                                           situ == "Fl" ~ "Flight",
+                                           situ == "Ro" ~ "Roosting"))
 
-deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
-                                var = "space_use", values = situcolors, 
-                                ylab = "Space use effect (Degree z-scores)", 
-                                xlab = "Situation")
+dataset <- as.data.frame(ggeffect(model = deg_z_mod, terms = c("space_use", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
 
-deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(4, 1), guides = "collect") & theme_classic()
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = space_use, y = z_deg, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (z-score)")+ xlab("Space use (log)")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_z_mod, specs = "situ", var = "space_use")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = space_use.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Space use \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/sp_situ.png"), width = 12, height = 6)
 ```
@@
 ```{r}
-deg_z_roost_P <- effplot(model = deg_z_mod, terms = c("roost_div", "situ"), 
-                       dataset = linked, pointX = "roost_div", 
-                       pointY = "z_deg", col = "situ", 
-                       legendTitle = "Situation", values = situcolors, 
-                       ylab = "Degree (z-score)", xlab = "Roost diversification")
-deg_z_roost_P & theme_classic()
+dataset <- as.data.frame(ggeffect(model = deg_z_mod, terms = c("roost_div", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = space_use, y = z_deg, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (z-score)")+ xlab("Roost diversification")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_z_mod, specs = "situ", var = "roost_div")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = roost_div.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Roost diversification \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/rd_situ.png"), width = 12, height = 6)
 ```
 
-## Strength
+Strength
 
@@
 Meanwhile, the opposite was once again true for the co-roosting network.
-Vultures with smaller ranges deviated singificantly more from random than those with larger ranges.
+Vultures with smaller ranges deviated significantly more from random than those with larger ranges.
 
-# Discussion
+```{r}
+dataset <- as.data.frame(ggeffect(model = str_z_mod, terms = c("space_use", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = space_use, y = z_str, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Strength (z-score)")+ xlab("Space use (log)")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(str_z_mod, specs = "situ", var = "space_use")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = space_use.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Space use \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/strength/sp_situ.png"), width = 12, height = 6)
+```
+
+Similarly, roost diversification negatively predicted preferential strength \[...\]
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = str_z_mod, terms = c("roost_div", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = space_use, y = z_str, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Strength (z-score)")+ xlab("Roost diversification")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(str_z_mod, specs = "situ", var = "roost_div")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = roost_div.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Roost diversification \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/strength/rd_situ.png"), width = 12, height = 6)
+```
+
+## How much does z-score vary by individual across situations and seasons?
+
+```{r}
+ll <- linked %>%
+  select(Nili_id, seasonUnique, season, year, situ, age_group, sex, z_deg, z_str)
+
+pltfun <- function(sit){
+  dat <- ll %>%
+    filter(situ == sit) %>%
+    group_by(Nili_id) %>%
+    mutate(mn = mean(z_deg)) %>%
+    ungroup()
+
+  highlight <- dat %>% group_by(Nili_id) %>%
+    filter(n() > 4) %>%
+    ungroup() %>%
+    filter(mn %in% c(min(mn), max(mn), median(mn))) %>%
+    pull(Nili_id) %>% unique()
+
+  dat <- dat %>%
+    mutate(hl = ifelse(Nili_id %in% highlight, T, F))
+
+  plt <- dat %>%
+    ggplot(aes(x = seasonUnique, y = z_deg, col = mn))+
+    geom_jitter(pch = 1, width = 0.1, alpha = 0.2)+
+    geom_line(aes(group = Nili_id, alpha = hl, linewidth = hl))+
+    scale_color_viridis_c()+
+    scale_alpha_manual(values = c(0.2, 1))+
+    scale_linewidth_manual(values = c(0.2, 1))
+  return(plt)
+}
+
+```
+
+Looks like there might be a little bit of repeatability within an individual, but there's still a lot of variation as well.
+
+What about correlation between z-score in one social situation and z-score in the others?
+
+```{r}
+llw <- ll %>%
+  select(Nili_id, seasonUnique, situ, z_deg) %>%
+  pivot_wider(id_cols = c("Nili_id", "seasonUnique"), names_from = "situ", values_from = "z_deg")
+
+llw %>%
+  ggplot(aes(x = Fe, y = Fl, col = seasonUnique))+geom_point(alpha = 0.5, pch = 1)+geom_smooth(method = "lm", se = F, linewidth = 0.5) # there is generally a positive correlation between feeding and flight
+
+llw %>%
+  ggplot(aes(x = Fl, y = Ro, col = seasonUnique))+geom_point(alpha = 0.5, pch = 1)+geom_smooth(method = "lm", se = F, linewidth = 0.5) # generally a negative correlation between flight and roosting (makes sense, given the opposite trends)
+
+llw %>%
+  ggplot(aes(x = Ro, y = Fe, col = seasonUnique))+geom_point(alpha = 0.5, pch = 1)+geom_smooth(method = "lm", se = F, linewidth = 0.5) # Also a bit of a negative relationship here.
+```

### 2024-06-10 forElvira_wrapAroundExample.R
wrap around example for Elvira

@@ -1,0 +1,113 @@
+# Steps to perform a wrap-around randomization
+# 1. Extract an explicit "time" column from your timestamps. Usually I use substr() or a similar function. Make sure you also have an explicit "date" column--i.e. you will need to have the full timestamp separated into separate date and time columns. 
+# 2. Define number of repetitions `n` (e.g. 100) and a shiftMax (the maximum number of days in each direction that the dataset can be shifted; minimum 1)
+# 3. Using a for loop or a map() statement, apply rotate_data_table() `n` times to your dataset. Save the repetitions.
+# 4. For each of the repetitions, create a social network using the get_*_edges() functions in vultureUtils.
+
+# rotate_data_table function
+# Requires that you separate the dates and times into separate columns beforehand.
+rotate_data_table <- function(dataset, shiftMax, idCol = "indiv", dateCol = "date", timeCol = "time"){
+  indivList <- dataset %>%
+    group_by(.data[[idCol]]) %>%
+    group_split(.keep = T)
+  joined <- vector(mode = "list", length = length(indivList))
+  for(indiv in 1:length(indivList)){
+    x <- indivList[[indiv]]
+    shift <- sample(-(shiftMax):shiftMax, size = 1)
+    #cat(shift, "\n")
+    # get all unique days that show up
+    days <- sort(unique(x[[dateCol]]))
+    
+    # get min and max dates to shift around (the "poles" of the conveyor)
+    selfMinDate <- min(days, na.rm = T)
+    selfMaxDate <- max(days, na.rm = T)
+    
+    # create a total sequence of dates to select from
+    daysFilled <- seq(lubridate::ymd(selfMinDate), lubridate::ymd(selfMaxDate), by = "day")
+    # converting to numbers so we can use %%--which dates are the ones we started with?
+    vec <- which(daysFilled %in% days)
+    shiftedvec <- vec + shift # shift
+    new <- (shiftedvec - min(vec)) %% (max(vec)-min(vec)+1)+1 # new dates as numbers
+    shiftedDates <- daysFilled[new] # select those dates from the possibilities
+    
+    # Make a data frame to hold the old and new dates
+    daysDF <- bind_cols({{dateCol}} := days, 
+                        "newDate" = shiftedDates,
+                        shift = shift)
+    nw <- left_join(x, daysDF, by = dateCol)
+    
+    if(!is.null(timeCol)){
[…]

### 2024-06-10 manuscript.rmd
more figures in ms

@@
 plt <- a + b + plot_layout(widths = c(4, 1))
-ggsave(plot = plt, filename = here("fig/figures/degree/sp_situ.png"), width = 12, height = 6)
+ggsave(plot = plt, filename = here("fig/figures/degree/deg_sp_situ.png"), width = 12, height = 6)
 ```
@@
 plt <- a + b + plot_layout(widths = c(4, 1))
-ggsave(plot = plt, filename = here("fig/figures/degree/rd_situ.png"), width = 12, height = 6)
+ggsave(plot = plt, filename = here("fig/figures/degree/deg_rd_situ.png"), width = 12, height = 6)
 ```
@@
 plt <- a + b + plot_layout(widths = c(4, 1))
-ggsave(plot = plt, filename = here("fig/figures/strength/sp_situ.png"), width = 12, height = 6)
+ggsave(plot = plt, filename = here("fig/figures/strength/str_sp_situ.png"), width = 12, height = 6)
 ```
@@
 plt <- a + b + plot_layout(widths = c(4, 1))
-ggsave(plot = plt, filename = here("fig/figures/strength/rd_situ.png"), width = 12, height = 6)
+ggsave(plot = plt, filename = here("fig/figures/strength/str_rd_situ.png"), width = 12, height = 6)
+```
+
+## Movement effects (non-significant)
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = deg_z_mod, terms = c("movement", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = movement, y = z_deg, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (z-score)")+ xlab("Movement")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_z_mod, specs = "situ", var = "movement")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = movement.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Movement \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/deg_mv_situ.png"), width = 12, height = 6)
+```
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = str_z_mod, terms = c("movement", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = movement, y = z_str, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Strength (z-score)")+ xlab("Movement")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(str_z_mod, specs = "situ", var = "movement")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = movement.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Movement \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/strength/str_mv_situ.png"), width = 12, height = 6)
 ```
@@
 ```
+
+## Plots for models before accounting for explicit space use
+
+```{r}
+summary(deg_mod) # space*situ, movement*season, movement*situ, roost_div
+summary(str_mod) # roost_div*situ only
+```
+
+Degree: space\*situ
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = deg_mod, terms = c("space_use", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = space_use, y = normDegree, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (normalized)")+ xlab("Space use")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_mod, specs = "situ", var = "space_use")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = space_use.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Space use \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/nonpermuted/np_deg_sp_situ.png"), width = 12, height = 6)
+```
+
+Degree: movement\*season
+
+```{r}
+pts_szn <- linked %>% mutate(group = season)
+dataset <- as.data.frame(ggeffect(model = deg_mod, terms = c("movement", "season")))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts_szn,
+             aes(x = movement, y = normDegree, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Season", values = seasoncolors)+
+  scale_fill_manual(name = "Season", values = seasoncolors)+
+  ylab("Degree (normalized)")+ xlab("Movement")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_mod, specs = "season", var = "movement"))
+
+b <- dat %>%
+  ggplot(aes(x = season, y = movement.trend, col = season))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = season, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = seasoncolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Movement \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/nonpermuted/np_deg_mv_season.png"), width = 12, height = 6)
+```
+
+Degree: movement\*situ
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = deg_mod, terms = c("movement", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = movement, y = normDegree, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (normalized)")+ xlab("Movement")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_mod, specs = "situ", var = "movement")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = movement.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Movement \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/nonpermuted/np_deg_mv_situ.png"), width = 12, height = 6)
+```
+
+Degree: roost_div(\*situ)
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = deg_mod, terms = c("roost_div", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = roost_div, y = normDegree, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (normalized)")+ xlab("Roost diversification")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(deg_mod, specs = "situ", var = "roost_div")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = roost_div.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Roost diversification \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/degree/nonpermuted/np_deg_rd_situ.png"), width = 12, height = 6)
+```
+
+Strength: roost_div\*situ
+
+```{r}
+dataset <- as.data.frame(ggeffect(model = str_mod, terms = c("roost_div", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
+                                                                                                                   group == "Fl" ~ "Flight",
+                                                                                                                   group == "Ro" ~ "Roosting"))
+
+a <- ggplot(data = dataset, 
+       aes(x, predicted))+
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
+              alpha = 0.2, linewidth = 0.6, show.legend = F)+
+  geom_point(data = pts,
+             aes(x = roost_div, y = normStrength, col = group),
+             pch = 1)+
+  geom_line(aes(col = group), linewidth = 1)+
+  facet_wrap(~group, scales = "free_y")+
+  scale_color_manual(name = "Situation", values = situcolors)+
+  scale_fill_manual(name = "Situation", values = situcolors)+
+  ylab("Degree (normalized)")+ xlab("Roost diversification")+
+  theme(legend.position = "none",
+        text = element_text(size = 16))
+
+dat <- as.data.frame(emmeans::emtrends(str_mod, specs = "situ", var = "roost_div")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
+                                                                                                                  situ == "Fl" ~ "Flight",
+                                                                                                                  situ == "Ro" ~ "Roosting"))
+
+b <- dat %>%
+  ggplot(aes(x = situ, y = roost_div.trend, col = situ))+
+  geom_point(size = 6)+
+  geom_errorbar(aes(x = situ, ymin = asymp.LCL, ymax = asymp.UCL), 
+                width = 0, linewidth = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2)+
+  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
+  coord_flip()+ ylab("Roost diversification \neffect")+
+  theme(legend.position = "none", 
+        text = element_text(size = 16),
+        axis.title.y = element_blank())
+
+plt <- a + b + plot_layout(widths = c(4, 1))
+ggsave(plot = plt, filename = here("fig/figures/strength/nonpermuted/np_deg_rd_situ.png"), width = 12, height = 6)
+```

### 2025-02-04 scripts/tables.R
updated sup mat table

@@ -17,7 +17,7 @@ all <- data.frame(seasonUnique = season_names,
   mutate(year = str_extract(seasonUnique, "[0-9]+"),
          season = str_extract(seasonUnique, "[a-z]+")) %>%
   select(-seasonUnique) %>%
-  mutate(type = "all")# there are some differences--and I think that's because of how I factored out the individuals that were never observed participating in that behavior, before calculating interactions.
+  mutate(type = "social network")# there are some differences--and I think that's because of how I factored out the individuals that were never observed participating in that behavior, before calculating interactions.
 
 tar_load(linked)
 focal <- linked %>%
@@ -28,27 +28,47 @@ focal <- linked %>%
          season = str_extract(seasonUnique, "[a-z]+")) %>%
   ungroup() %>%
   select(-seasonUnique) %>%
-  mutate(type = "focal")
+  mutate(type = "space use")
 
 indivs <- bind_rows(all, focal) %>%
   relocate(year, season) %>%
   arrange(year, season) %>%
   pivot_longer(c("flight", "feeding", "roosting"), names_to = "situ", values_to = "n") %>%
-  pivot_wider(names_from = "type", values_from = "n") %>%
-  mutate(proportion = round(focal/all, 2))
+  pivot_wider(names_from = "type", values_from = "n")
 
-indivs %>%
-  summarize(mn_all = mean(all),
-            sd_all = sd(all),
-            mn_focal = mean(focal),
-            sd_focal = sd(focal),
-            mn_prop = mean(proportion),
-            sd_prop = sd(proportion))
+# For comparison: how many tagged birds in the entire uncleaned da --------
+tar_load(joined0)# to get total numbers of tagged birds during the same time periods
+with_seasons <- joined0 %>% mutate(month = lubridate::month(timestamp),
+                                   day = lubridate::day(timestamp),
+                                   year = lubridate::year(timestamp)) %>%
+  mutate(season = case_when(((month == 12 & day >= 15) | 
+                               (month %in% 1:4) | 
+                               (month == 5 & day < 15)) ~ "breeding",
+                            ((month == 5 & day >= 15) | 
+                               (month %in% 6:8) | 
+                               (month == 9 & day < 15)) ~ "summer",
+                            .default = "fall")) %>%
[…]

### 2025-02-04 scripts/06.0_figures.R
updated map figure

@@ -421,13 +421,15 @@ dat <- st_drop_geometry(r) %>% select("long" = location_long, "lat" = location_l
   bind_rows(st_drop_geometry(fee) %>% select("long" = interactionLong, "lat" = interactionLat) %>%
               mutate(type = "Co-feeding\ninteractions\n"))
 mymap <- mp+
-  geom_point(data = dat, aes(x = long, y = lat, col = type, shape = type), alpha = 0.4)+
+  geom_point(data = dat, aes(x = long, y = lat, col = type, shape = type),
+             alpha = 0.9, size = 2)+
   scale_color_manual(values = situcolors)+
   scale_shape_manual(values = c(4, 1, 2))+
-  guides(shape = guide_legend(override.aes = list(alpha = 1, size = 3)))+
+  guides(shape = guide_legend(override.aes = list(alpha = 1, size = 5)))+
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         legend.title = element_blank(),
-        text = element_text(family = "Verdana", size = 14),
+        text = element_text(size = 18),
         legend.position = "left")
+mymap
 ggsave(mymap, filename = here("fig/mymap.png"), width = 7, height = 6)

### 2025-02-04 scripts/tables.R
changed from "tracked" to "tagged"

@@ -54,21 +54,21 @@ with_seasons <- joined0 %>% mutate(month = lubridate::month(timestamp),
                                   .default = paste(as.character(year), season, sep = "_"))) %>%
   mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding", "2023_summer")),
          season = factor(season, levels = c("breeding", "summer", "fall")))
-all_tracked <- with_seasons %>%
+all_tagged <- with_seasons %>%
   select(local_identifier, month, day, year, season, seasonUnique) %>%
   group_by(year, season) %>%
-  summarize(`all tracked` = length(unique(local_identifier)))
-a_t <- bind_rows(all_tracked %>% mutate(situ = "flight"),
-                 all_tracked %>% mutate(situ = "feeding"),
-                 all_tracked %>% mutate(situ = "roosting")) %>%
+  summarize(`all tagged` = length(unique(local_identifier)))
+a_t <- bind_rows(all_tagged %>% mutate(situ = "flight"),
+                 all_tagged %>% mutate(situ = "feeding"),
+                 all_tagged %>% mutate(situ = "roosting")) %>%
   mutate(year = as.character(year))
 
 indivs <- left_join(indivs, a_t)
 
 
 # Make a nice table -------------------------------------------------------
-min <- min(c(indivs$`space use`, indivs$`social network`, indivs$`all tracked`))
-max <- max(c(indivs$`space use`, indivs$`social network`, indivs$`all tracked`))
+min <- min(c(indivs$`space use`, indivs$`social network`, indivs$`all tagged`))
+max <- max(c(indivs$`space use`, indivs$`social network`, indivs$`all tagged`))
 tar_load(season_names)
 
 tab <- indivs %>%
@@ -76,7 +76,7 @@ tab <- indivs %>%
   mutate(yrsz = factor(yrsz, levels = str_replace(season_names, "_", " "))) %>%
   arrange(yrsz) %>%
   select(-c("year", "season")) %>%
-  pivot_longer(cols = c("social network", "space use", "all tracked"), names_to = "type", values_to = "count") %>%
+  pivot_longer(cols = c("social network", "space use", "all tagged"), names_to = "type", values_to = "count") %>%
   pivot_wider(names_from = "yrsz", values_from = "count") %>%
   arrange(situ, type) %>%
   rename("Situation" = situ,

### 2025-02-04 scripts/00.0_dataPrep.qmd
tidying up

@@
 ```{r load-packages}
+# First install the package from GitHub (do this in the console, not the script)
+# remotes::install_github("kaijagahm/vultureUtils")
+# devtools::install_github("kaijagahm/vultureUtils")
 library(vultureUtils)
@@
 library(tidyverse)
-library(move)
+library(move) # the R package that allows you to download Movebank data through R instead of manually
 library(feather)
@@
 carmel <- read.csv(here("data/raw/all_captures_carmel_2010-2021.csv"))
-removed_captures <- removeCaptures(data = cleaned, 
+removed_captures <- vultureUtils::removeCaptures(data = cleaned, 
                                        captureSites = capture_sites, 
@@
 
-mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
-         sex = case_when(sex == "f/m" ~ "u",
-                         TRUE ~ sex)) %>%
-
 # tidy up
@@
 ```{r region-masking}
-mask <- sf::st_read(here("data/raw/CutOffRegion.kml"))
+mask <- sf::st_read(here("data/CutOffRegion.kml"))
 
